### 2023-01-24 reports/01_predators.qmd
Predator counts

---
title: "Predator summary"
format:
  html:
    code-fold: true
editor: visual
---

```{r}
#| label: setup
#| include: false

library(targets)
library(tidyverse)

tar_load("predators_agg", store = here::here("_targets"))
```

# South Georgia winter predators

## Counts

The six most abundant species.

```{r}

top_six <- predators_agg %>% 
  group_by(species) %>% 
  summarize(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>%
  slice(1:6)

top_six
```

```{r}

top_six_plot_data <- predators_agg %>% 
  semi_join(top_six, by = "species") %>% 
  group_by(year, species) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(species = fct_reorder(species, total, .fun = sum),
         year = factor(year, levels = 2016:2012))
year_labels <- top_six_plot_data %>% 
  filter(species == "SNPT") %>% 
  arrange(desc(year)) %>%                   # Make the labels show up at
  mutate(total = cumsum(total) - total / 2) # correct y-axis position

ggplot(top_six_plot_data, aes(species, total, group = year)) +
  geom_col(color = "black", fill = "grey80") +
  geom_text(aes(label = year), year_labels) +
  labs(y = "Total species count") +
  theme_classic() +
  theme(axis.title.x = element_blank())
```

### 2023-01-24 R/predators.R

@@ -1,0 +1,7 @@
+aggregate_predators <- function(predators) {
+  predators %>%
+    filter(species != "NULL") %>%
+    mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
+    group_by(year, interval, species, lon_mean, lat_mean) %>%
+    summarize(count = sum(count_species), .groups = "drop")
+}

### 2023-01-24 _targets.R

@@ -1,0 +1,26 @@
+library(here)
+library(targets)
+library(tarchetypes)
+library(tidyverse)
+
+# Set target options:
+tar_option_set(
+  packages = "tidyverse",
+  format = "rds"
+)
+
+# Source scripts in R/
+tar_source()
+
+# Pipeline
+predators <- read_csv("data/WAMLR/WAMLR-2012-2016-Underway-Predator-Sightings.csv")
+list(
+  tar_target(
+    predators_file,
+    here("data", "WAMLR", "WAMLR-2012-2016-Underway-Predator-Sightings.csv"),
+    format = "file"
+  ),
+  tar_target(predators, readr::read_csv(predators_file)),
+  tar_target(predators_agg, aggregate_predators(predators)),
+  tar_quarto(predator_summary, here("reports", "01_predators.qmd"))
+)

### 2023-01-25 reports/01_predators.qmd
Begin predator sighting map

@@
 
-tar_load("predators_agg", store = here::here("_targets"))
+tar_load(c("predators_agg", "sg_sf"), store = here::here("_targets"))
+source(here::here("R", "sp.R"))
 ```
 
-# South Georgia winter predators
+## South Georgia winter predators
 
-## Counts
+### Counts
 
@@
 ```
+
+### Maps
+
+```{r}
+#| label: predator_map
+
+predators_sf <- predators_agg %>% 
+  semi_join(top_six, by = "species") %>% 
+  filter(year == 2016) %>% 
+  sf::st_as_sf(coords = c("lon_mean", "lat_mean"), 
+               crs = sp::CRS("+proj=longlat +datum=WGS84"))
+
+map_lim <- sf::st_bbox(predators_sf) %>% 
+  sf::st_as_sfc(crs = "EPSG:4326") %>% 
+  sf::st_transform(sg_proj()) %>% 
+  sf::st_bbox()
+
+ggplot() +
+  geom_sf(data = sg_sf) +
+  geom_sf(aes(color = count), data = predators_sf) +
+  scale_color_gradient(low = "lightblue", high = "darkred",
+                       limits = c(0, 1000)) +
+  facet_wrap(~species) +
+  coord_sf(xlim = map_lim[c("xmin", "xmax")] + c(-1e5, 1e5),
+           ylim = map_lim[c("ymin", "ymax")] + c(-1e5, 1e5),
+           expand = FALSE,
+           crs = sg_proj()) +
+  theme_minimal()
+
+```

### 2023-01-25 R/sp.R

@@ -1,0 +1,24 @@
+# Use WGS 84 / South Georgia Lambert for mapping
+sg_proj <- function(){
+  sf::st_crs("EPSG:3762")
+}
+
+download_ne <- function(ne_dir) {
+  rnaturalearth::ne_download(scale = "large",
+                             type = "land",
+                             category = "physical",
+                             destdir = ne_dir,
+                             returnclass = "sf")
+}
+
+create_sg_sf <- function(
+    ne_dir,
+    limits = c(xmin = -67, xmax = -50, ymin = -68, ymax = -55)
+) {
+  land <- rnaturalearth::ne_load(scale = "large",
+                                 type = "land",
+                                 category = "physical",
+                                 destdir = ne_dir,
+                                 returnclass = "sf")
+  sg_sf <- sf::st_crop(land, sf::st_bbox(limits))
+}

### 2023-01-25 _targets.R

@@ -13,8 +13,22 @@ tar_option_set(
 tar_source()
 
 # Pipeline
-predators <- read_csv("data/WAMLR/WAMLR-2012-2016-Underway-Predator-Sightings.csv")
 list(
+  # Spatial data
+  tar_target(
+    ne_dir,
+    here("data", "geospatial", "ne"),
+    format = "file"
+  ),
+  tar_target(
+    ne_data,
+    download_ne(ne_dir)
+  ),
+  tar_target(
+    sg_sf,
+    create_sg_sf(ne_dir)
+  ),
+  # Predators
   tar_target(
     predators_file,
     here("data", "WAMLR", "WAMLR-2012-2016-Underway-Predator-Sightings.csv"),

### 2023-01-25 reports/01_predators.qmd
rasterized predator counts

--- 2023-01-24 reports/01_predators.qmd
@@
 ```
+
+### Maps
+
+```{r}
+#| label: snpt_map
+#| warning: false
+
+snpt_sf <- predators_agg %>% 
+  filter(species == "SNPT",
+         year == 2016) %>% 
+  sf::st_as_sf(coords = c("lon_mean", "lat_mean"), 
+               crs = sp::CRS("+proj=longlat +datum=WGS84"))
+
+snpt_stars <- rasterize_counts(snpt_sf, res = 1e4, layer_name = "SNPT")
+
+map_lim <- sf::st_bbox(snpt_sf) %>% 
+  sf::st_as_sfc(crs = "EPSG:4326") %>% 
+  sf::st_transform(sg_proj()) %>% 
+  sf::st_bbox()
+
+ggplot() +
+  geom_sf(data = sg_sf) +
+  stars::geom_stars(aes(fill = SNPT), data = snpt_stars) +
+  scale_fill_gradient(low = "lightblue", high = "darkred",
+                      na.value = NA) +
+  expand_limits(fill = 0) +
+  coord_sf(xlim = map_lim[c("xmin", "xmax")],
+           ylim = map_lim[c("ymin", "ymax")],
+           expand = FALSE,
+           crs = sg_proj()) +
+  theme_minimal() +
+  theme(axis.title = element_blank())
+
+```

### 2023-01-25 R/sp.R

@@ -1,6 +1,10 @@
 # Use WGS 84 / South Georgia Lambert for mapping
-sg_proj <- function(){
+sg_proj <- function() {
   sf::st_crs("EPSG:3762")
+}
+
+sg_lims <- function() {
+  c(xmin = -67, xmax = -50, ymin = -68, ymax = -55)
 }
 
 download_ne <- function(ne_dir) {
@@ -11,10 +15,7 @@ download_ne <- function(ne_dir) {
                              returnclass = "sf")
 }
 
-create_sg_sf <- function(
-    ne_dir,
-    limits = c(xmin = -67, xmax = -50, ymin = -68, ymax = -55)
-) {
+create_sg_sf <- function(ne_dir, limits = sg_lims()) {
   land <- rnaturalearth::ne_load(scale = "large",
                                  type = "land",
                                  category = "physical",
@@ -22,3 +23,32 @@ create_sg_sf <- function(
                                  returnclass = "sf")
   sg_sf <- sf::st_crop(land, sf::st_bbox(limits))
 }
+
+sg_raster_template <- function(limits = sg_lims(), res = 5e4) {
+  lims_prj <- sf::st_bbox(limits, crs = "EPSG:4326") %>%
+    sf::st_as_sfc() %>%
+    sf::st_transform(sg_proj()) %>%
+    sf::st_bbox()
+  raster::raster(xmn = lims_prj["xmin"],
+                 xmx = lims_prj["xmax"],
+                 ymn = lims_prj["ymin"],
+                 ymx = lims_prj["ymax"],
+                 crs = sg_proj()[1]$input,
+                 resolution = res)
+}
+
+rasterize_counts <- function(counts_sf,
+                             limits = sg_lims(),
+                             res = 5e4,
+                             layer_name = "counts") {
+  raster::rasterize(
+    sf::as_Spatial(sf::st_transform(counts_sf, sg_proj())),
+    sg_raster_template(limits, res),
+    field = "count",
+    fun = "sum",
+    na.rm = TRUE
+  ) %>%
+    stars::st_as_stars() %>%
+    setNames(layer_name)
+}
+

### 2023-01-25 _targets.R
Use Quarto project to fix output directory

@@ -36,5 +36,6 @@ list(
   ),
   tar_target(predators, readr::read_csv(predators_file)),
   tar_target(predators_agg, aggregate_predators(predators)),
-  tar_quarto(predator_summary, here("reports", "01_predators.qmd"))
+  # Reports
+  tar_quarto(reports, here("reports"))
 )

### 2023-02-09 reports/02_stations.qmd
Start looking at zooplankton

--- 2023-01-25 reports/01_predators.qmd
@@
 ---
-title: "Predator summary"
+title: "Station summary"
 format:
@@
 editor: visual
+output-dir: "docs"
 ---
@@
 
-tar_load(c("predators_agg", "sg_sf"), store = here::here("_targets"))
-source(here::here("R", "sp.R"))
+tar_load(c("zoop"), store = here::here("_targets"))
+# source(here::here("R", "sp.R"))
 ```
 
-## South Georgia winter predators
+## Station maps
 
-### Counts
-
-The six most abundant species.
+By year.
 

### 2023-02-09 _targets.R

@@ -36,6 +36,13 @@ list(
   ),
   tar_target(predators, readr::read_csv(predators_file)),
   tar_target(predators_agg, aggregate_predators(predators)),
+  # Zooplankton
+  tar_target(
+    zoop_file,
+    here("data", "WAMLR", "WAMLR-2012-2016-Zooplnkton-Physics.xlsx"),
+    format = "file"
+  ),
+  tar_target(zoop, read_zoop(zoop_file)),
   # Reports
   tar_quarto(reports, here("reports"))
 )

### 2023-02-09 reports/01_predators.qmd
Use Quarto project to fix output directory
Start on a basemap function

@@
          year == 2016) %>% 
-  sf::st_as_sf(coords = c("lon_mean", "lat_mean"), 
-               crs = sp::CRS("+proj=longlat +datum=WGS84"))
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
 

### 2023-02-09 R/sp.R

@@ -52,3 +52,24 @@ rasterize_counts <- function(counts_sf,
     setNames(layer_name)
 }
 
+latlon_to_sf <- function(df, coords = c("x", "y")) {
+  sf::st_as_sf(df,
+               coords = coords,
+               crs = sp::CRS("+proj=longlat +datum=WGS84"))
+}
+
+sg_basemap <- function() {
+  map_lim <- sf::st_bbox(sg_lims(), crs = "EPSG:4326") %>%
+    sf::st_as_sfc() %>%
+    sf::st_transform(sg_proj()) %>%
+    sf::st_bbox()
+
+  ggplot() +
+    geom_sf(data = tar_read("sg_sf", store = here::here("_targets"))) +
+    coord_sf(xlim = map_lim[c("xmin", "xmax")],
+             ylim = map_lim[c("ymin", "ymax")],
+             expand = FALSE,
+             crs = sg_proj()) +
+    theme_minimal() +
+    theme(axis.title = element_blank())
+}

### 2023-02-09 reports/01_predators.qmd
Change over from SG tto ANT terminology

@@
 
-tar_load(c("predators_agg", "sg_sf"), store = here::here("_targets"))
+tar_load(c("predators_agg", "ant_sf"), store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 
-## South Georgia winter predators
+## Antarctic Peninsula winter predators
 
@@
   sf::st_as_sfc(crs = "EPSG:4326") %>% 
-  sf::st_transform(sg_proj()) %>% 
+  sf::st_transform(ant_proj()) %>% 
   sf::st_bbox()
@@
 ggplot() +
-  geom_sf(data = sg_sf) +
+  geom_sf(data = ant_sf) +
   stars::geom_stars(aes(fill = SNPT), data = snpt_stars) +
@@
            expand = FALSE,
-           crs = sg_proj()) +
+           crs = ant_proj()) +
   theme_minimal() +

### 2023-02-09 reports/02_stations.qmd
Change over from SG tto ANT terminology

@@
 ## Station maps
-
-By year.
-
-```{r}
-
-top_six <- predators_agg %>% 
-  group_by(species) %>% 
-  summarize(total_count = sum(count)) %>% 
-  arrange(desc(total_count)) %>%
-  slice(1:6)
-
-top_six
-```
-
-```{r}
-
-top_six_plot_data <- predators_agg %>% 
-  semi_join(top_six, by = "species") %>% 
-  group_by(year, species) %>% 
-  summarize(total = sum(count), .groups = "drop") %>% 
-  mutate(species = fct_reorder(species, total, .fun = sum),
-         year = factor(year, levels = 2016:2012))
-year_labels <- top_six_plot_data %>% 
-  filter(species == "SNPT") %>% 
-  arrange(desc(year)) %>%                   # Make the labels show up at
-  mutate(total = cumsum(total) - total / 2) # correct y-axis position
-
-ggplot(top_six_plot_data, aes(species, total, group = year)) +
-  geom_col(color = "black", fill = "grey80") +
-  geom_text(aes(label = year), year_labels) +
-  labs(y = "Total species count") +
-  theme_classic() +
-  theme(axis.title.x = element_blank())
-```
-
-### Maps
-
-```{r}
-#| label: snpt_map
-#| warning: false
-
-snpt_sf <- predators_agg %>% 
-  filter(species == "SNPT",
-         year == 2016) %>% 
-  sf::st_as_sf(coords = c("lon_mean", "lat_mean"), 
-               crs = sp::CRS("+proj=longlat +datum=WGS84"))
-
-snpt_stars <- rasterize_counts(snpt_sf, res = 1e4, layer_name = "SNPT")
-
-map_lim <- sf::st_bbox(snpt_sf) %>% 
-  sf::st_as_sfc(crs = "EPSG:4326") %>% 
-  sf::st_transform(sg_proj()) %>% 
-  sf::st_bbox()
-
-ggplot() +
-  geom_sf(data = sg_sf) +
-  stars::geom_stars(aes(fill = SNPT), data = snpt_stars) +
-  scale_fill_gradient(low = "lightblue", high = "darkred",
-                      na.value = NA) +
-  expand_limits(fill = 0) +
-  coord_sf(xlim = map_lim[c("xmin", "xmax")],
-           ylim = map_lim[c("ymin", "ymax")],
-           expand = FALSE,
-           crs = sg_proj()) +
-  theme_minimal() +
-  theme(axis.title = element_blank())
-
-```

### 2023-02-09 R/sp.R

@@ -1,9 +1,9 @@
-# Use WGS 84 / South Georgia Lambert for mapping
-sg_proj <- function() {
-  sf::st_crs("EPSG:3762")
+# Use WGS 84 / Antarctic Polar Stereographic for mapping
+ant_proj <- function() {
+  sf::st_crs("EPSG:3031")
 }
 
-sg_lims <- function() {
+ant_lims <- function() {
   c(xmin = -67, xmax = -50, ymin = -68, ymax = -55)
 }
 
@@ -15,35 +15,35 @@ download_ne <- function(ne_dir) {
                              returnclass = "sf")
 }
 
-create_sg_sf <- function(ne_dir, limits = sg_lims()) {
+create_ant_sf <- function(ne_dir, limits = ant_lims()) {
   land <- rnaturalearth::ne_load(scale = "large",
                                  type = "land",
                                  category = "physical",
                                  destdir = ne_dir,
                                  returnclass = "sf")
-  sg_sf <- sf::st_crop(land, sf::st_bbox(limits))
+  sf::st_crop(land, sf::st_bbox(limits))
 }
 
-sg_raster_template <- function(limits = sg_lims(), res = 5e4) {
+ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
   lims_prj <- sf::st_bbox(limits, crs = "EPSG:4326") %>%
     sf::st_as_sfc() %>%
-    sf::st_transform(sg_proj()) %>%
+    sf::st_transform(ant_proj()) %>%
     sf::st_bbox()
   raster::raster(xmn = lims_prj["xmin"],
                  xmx = lims_prj["xmax"],
                  ymn = lims_prj["ymin"],
                  ymx = lims_prj["ymax"],
-                 crs = sg_proj()[1]$input,
+                 crs = ant_proj()[1]$input,
                  resolution = res)
 }
 
 rasterize_counts <- function(counts_sf,
-                             limits = sg_lims(),
+                             limits = ant_lims(),
                              res = 5e4,
                              layer_name = "counts") {
   raster::rasterize(
-    sf::as_Spatial(sf::st_transform(counts_sf, sg_proj())),
-    sg_raster_template(limits, res),
+    sf::as_Spatial(sf::st_transform(counts_sf, ant_proj())),
+    ant_raster_template(limits, res),
     field = "count",
[…]

### 2023-02-09 _targets.R

@@ -25,8 +25,8 @@ list(
     download_ne(ne_dir)
   ),
   tar_target(
-    sg_sf,
-    create_sg_sf(ne_dir)
+    ant_sf,
+    create_ant_sf(ne_dir)
   ),
   # Predators
   tar_target(

### 2023-02-09 reports/01_predators.qmd
Fixed projection and finished implementing basemap

@@
 map_lim <- sf::st_bbox(snpt_sf) %>% 
-  sf::st_as_sfc(crs = "EPSG:4326") %>% 
-  sf::st_transform(ant_proj()) %>% 
-  sf::st_bbox()
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
 
-ggplot() +
-  geom_sf(data = ant_sf) +
+ant_basemap(map_lim) +
   stars::geom_stars(aes(fill = SNPT), data = snpt_stars) +
@@
                       na.value = NA) +
-  expand_limits(fill = 0) +
-  coord_sf(xlim = map_lim[c("xmin", "xmax")],
-           ylim = map_lim[c("ymin", "ymax")],
-           expand = FALSE,
-           crs = ant_proj()) +
-  theme_minimal() +
-  theme(axis.title = element_blank())
+  expand_limits(fill = 0)
 

### 2023-02-09 R/sp.R

@@ -1,6 +1,11 @@
-# Use WGS 84 / Antarctic Polar Stereographic for mapping
+# Use WGS 84 / Antarctic Polar Stereographic for mapping, recentered on ant_lims()
 ant_proj <- function() {
-  sf::st_crs("EPSG:3031")
+  lat_ts <- (ant_lims()["ymin"] + ant_lims()["ymax"]) / 2
+  lon_0 <- (ant_lims()["xmin"] + ant_lims()["xmax"]) / 2
+  glue::glue(
+    "+proj=stere +lat_0=-90 +lat_ts={lat_ts} +lon_0={lon_0} +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
+  ) %>%
+  sf::st_crs()
 }
 
 ant_lims <- function() {
@@ -26,9 +31,7 @@ create_ant_sf <- function(ne_dir, limits = ant_lims()) {
 
 ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
   lims_prj <- sf::st_bbox(limits, crs = "EPSG:4326") %>%
-    sf::st_as_sfc() %>%
-    sf::st_transform(ant_proj()) %>%
-    sf::st_bbox()
+    project_bbox()
   raster::raster(xmn = lims_prj["xmin"],
                  xmx = lims_prj["xmax"],
                  ymn = lims_prj["ymin"],
@@ -58,11 +61,11 @@ latlon_to_sf <- function(df, coords = c("x", "y")) {
                crs = sp::CRS("+proj=longlat +datum=WGS84"))
 }
 
-ant_basemap <- function() {
-  map_lim <- sf::st_bbox(ant_lims(), crs = "EPSG:4326") %>%
-    sf::st_as_sfc() %>%
-    sf::st_transform(ant_proj()) %>%
-    sf::st_bbox()
+ant_basemap <- function(map_lim = NULL) {
+  if (is.null(map_lim)) {
+    map_lim <- sf::st_bbox(ant_lims(), crs = "EPSG:4326") %>%
+      project_bbox()
+  }
 
   ggplot() +
     geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
@@ -72,4 +75,24 @@ ant_basemap <- function() {
              crs = ant_proj()) +
     theme_minimal() +
     theme(axis.title = element_blank())
+}
+
[…]

### 2023-02-09 reports/01_predators.qmd
Rasterize counts by year and species

@@
 
-The six most abundant species.
+The five most abundant species.
 
@@
 
-top_six <- predators_agg %>% 
+top_five <- predators_agg %>% 
   group_by(species) %>% 
@@
   arrange(desc(total_count)) %>%
-  slice(1:6)
+  slice(1:5)
 
-top_six
+top_five
 ```
@@
 
-top_six_plot_data <- predators_agg %>% 
-  semi_join(top_six, by = "species") %>% 
+top_five_plot_data <- predators_agg %>% 
+  semi_join(top_five, by = "species") %>% 
   group_by(year, species) %>% 
@@
          year = factor(year, levels = 2016:2012))
-year_labels <- top_six_plot_data %>% 
+year_labels <- top_five_plot_data %>% 
   filter(species == "SNPT") %>% 
@@
 
-ggplot(top_six_plot_data, aes(species, total, group = year)) +
+ggplot(top_five_plot_data, aes(species, total, group = year)) +
   geom_col(color = "black", fill = "grey80") +
@@
 ```{r}
-#| label: snpt_map
+#| label: topfive_map
 #| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-width: 6
+#| fig-dpi: 600
 
-snpt_sf <- predators_agg %>% 
-  filter(species == "SNPT",
-         year == 2016) %>% 
+topfive_sf <- predators_agg %>% 
+  semi_join(top_five, by = "species") %>% 
   latlon_to_sf(coords = c("lon_mean", "lat_mean"))
 
-snpt_stars <- rasterize_counts(snpt_sf, res = 1e4, layer_name = "SNPT")
+topfive_stars <- rasterize_counts(topfive_sf, res = 5e4)
 
-map_lim <- sf::st_bbox(snpt_sf) %>% 
+map_lim <- sf::st_bbox(topfive_sf) %>% 
   project_bbox() %>% 
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = SNPT), data = snpt_stars) +
+  stars::geom_stars(aes(fill = count), data = topfive_stars) +
   scale_fill_gradient(low = "lightblue", high = "darkred",
+                      trans = scales::pseudo_log_trans(),
+                      breaks = c(0, 25, 100, 400, 1000, 4000),
                       na.value = NA) +
-  expand_limits(fill = 0)
-
+  scale_x_continuous(breaks = c(-65, -60, -55)) +
+  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
+  facet_grid(species ~ year) +
+  expand_limits(fill = 0) +
+  theme(axis.text.x = element_text(size = 7))
 ```

### 2023-02-09 R/sp.R

@@ -40,19 +40,39 @@ ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
                  resolution = res)
 }
 
+# By species and year
 rasterize_counts <- function(counts_sf,
+                             count_col = "count",
+                             species_col = "species",
+                             year_col = "year",
+                             layer_name = count_col,
                              limits = ant_lims(),
-                             res = 5e4,
-                             layer_name = "counts") {
-  raster::rasterize(
-    sf::as_Spatial(sf::st_transform(counts_sf, ant_proj())),
-    ant_raster_template(limits, res),
-    field = "count",
-    fun = "sum",
-    na.rm = TRUE
+                             res = 5e4) {
+
+  # Group counts, call rasterize on each group, convert back to df
+  rasterized_df <- counts_sf %>%
+    group_by(.data[[species_col]], .data[[year_col]]) %>%
+    group_modify(
+      function(rows, key) {
+        result <- raster::rasterize(
+          sf::as_Spatial(sf::st_transform(rows, ant_proj())),
+          ant_raster_template(limits, res),
+          field = count_col,
+          fun = "sum",
+          na.rm = TRUE
+        ) %>%
+          raster::as.data.frame(xy = TRUE) %>%
+          rename(!!layer_name := layer)
+      }
+    )
+
+  # Convert df to stars and set projection
+  stars::st_as_stars(
+    rasterized_df,
+    dims = 1:4,
+    xy = 3:4
   ) %>%
-    stars::st_as_stars() %>%
-    setNames(layer_name)
+    sf::st_set_crs(ant_proj())
 }
 
 latlon_to_sf <- function(df, coords = c("x", "y")) {

### 2023-02-09 reports/01_predators.qmd
Create species-of-interest maps

@@
 
+Distribution maps of the five most abundant species in each survey year, aggregated to 50km bins.
+
 ```{r}
@@
 ```
+
+Distribution maps of five species of interest in each survey year, also aggregated to 50km bins.
+
+```{r}
+#| label: soi_map
+#| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-width: 6
+#| fig-dpi: 600
+
+soi <- c("GEPN", "BLPT", "ANTE", "LESE", "KIWH")
+
+soi_sf <- predators_agg %>% 
+  filter(species %in% soi) %>% 
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+
+soi_stars <- rasterize_counts(soi_sf, res = 5e4)
+
+map_lim <- sf::st_bbox(soi_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  stars::geom_stars(aes(fill = count), data = soi_stars) +
+  scale_fill_gradient(low = "lightblue", high = "darkred",
+                      trans = scales::pseudo_log_trans(),
+                      breaks = c(0, 25, 100, 400, 1000, 4000),
+                      na.value = NA) +
+  scale_x_continuous(breaks = c(-65, -60, -55)) +
+  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
+  facet_grid(species ~ year) +
+  expand_limits(fill = 0) +
+  theme(axis.text.x = element_text(size = 7))
+```

### 2023-02-09 reports/02_stations.qmd
Create a station map color-coded by time-of-day

@@
 
-tar_load(c("zoop"), store = here::here("_targets"))
-# source(here::here("R", "sp.R"))
+tar_load(c("zoop", "zoop_long", "ant_sf"), store = here::here("_targets"))
+source(here::here("R", "sp.R"))
+source(here::here("R", "zoop.R"))
 ```
@@
 ## Station maps
+
+```{r}
+#| label: station_maps
+
+zoop_sf <- latlon_to_sf(zoop, 
+                        coords = c("dec.longitude", "dec.latitude")) %>% 
+  sf::st_transform(ant_proj())
+
+map_lim <- sf::st_bbox(zoop_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = time.of.day), zoop_sf) +
+  facet_wrap(~ Year) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  scale_color_manual(values = tod_pal()) +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
+```

### 2023-02-09 R/sp.R

@@ -82,19 +82,21 @@ latlon_to_sf <- function(df, coords = c("x", "y")) {
 }
 
 ant_basemap <- function(map_lim = NULL) {
+  ggplot() +
+    geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+    theme_minimal() +
+    theme(axis.title = element_blank())
+}
+
+coord_ant <- function(map_lim = NULL) {
   if (is.null(map_lim)) {
     map_lim <- sf::st_bbox(ant_lims(), crs = "EPSG:4326") %>%
       project_bbox()
   }
-
-  ggplot() +
-    geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
-    coord_sf(xlim = map_lim[c("xmin", "xmax")],
-             ylim = map_lim[c("ymin", "ymax")],
-             expand = FALSE,
-             crs = ant_proj()) +
-    theme_minimal() +
-    theme(axis.title = element_blank())
+  coord_sf(xlim = map_lim[c("xmin", "xmax")],
+           ylim = map_lim[c("ymin", "ymax")],
+           expand = FALSE,
+           crs = ant_proj())
 }
 
 project_bbox <- function(x) {

### 2023-02-09 R/zoop.R

@@ -2,5 +2,13 @@ read_zoop <- function(zoop_path) {
   readxl::read_excel(zoop_path, sheet = 1) %>%
     mutate(time.of.day = factor(time.of.day,
                                 levels = c("D", "T", "N"),
-                                labels = c("Day", "Twilight", "Night")))
+                                labels = c("Day", "Twilight", "Night"))) %>%
+    select(-57) %>%
+    filter(leg == "W")
+}
+
+tod_pal <- function() {
+  c(Day = "gold",
+    Twilight = "violet",
+    Night = "darkblue")
 }

### 2023-02-09 _targets.R

@@ -43,6 +43,10 @@ list(
     format = "file"
   ),
   tar_target(zoop, read_zoop(zoop_file)),
+  tar_target(
+    zoop_long,
+    pivot_longer(zoop, -(1:53), names_to = "taxa", values_to = "abundance")
+  ),
   # Reports
   tar_quarto(reports, here("reports"))
 )

### 2023-02-10 reports/02_stations.qmd
Add buffers to station map

--- 2023-02-09 reports/02_stations.qmd
@@
 
-tar_load(c("zoop"), store = here::here("_targets"))
-# source(here::here("R", "sp.R"))
+tar_load(c("zoop", "ant_sf"), store = here::here("_targets"))
+source(here::here("R", "sp.R"))
+source(here::here("R", "zoop.R"))
 ```
@@
 ## Station maps
+
+Sampling stations across survey years, color-coded by time-of-day. Grey circles indicate 10k, 15k, and 20k buffers.
+
+```{r}
+#| label: station_maps
+
+zoop_sf <- latlon_to_sf(zoop, 
+                        coords = c("dec.longitude", "dec.latitude")) %>% 
+  sf::st_transform(ant_proj())
+
+zoop_buffers_10k <- sf::st_buffer(zoop_sf, 5e3)
+zoop_buffers_15k <- sf::st_buffer(zoop_sf, 10e3)
+zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)
+
+map_lim <- sf::st_bbox(zoop_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
+  geom_sf(data = zoop_buffers_10k, color = "lightgrey", fill = NA) +
+  geom_sf(data = zoop_buffers_15k, color = "lightgrey", fill = NA) +
+  geom_sf(data = zoop_buffers_20k, color = "lightgrey", fill = NA) +
+  facet_wrap(~ Year) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  scale_color_manual(values = tod_pal()) +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
+```

### 2023-02-10 reports/01_predators.qmd
Fix predator maps

--- 2023-02-09 reports/01_predators.qmd
@@
 
+Distribution maps of the five most abundant species in each survey year, aggregated to 25km bins.
+
 ```{r}
@@
 
-topfive_stars <- rasterize_counts(topfive_sf, res = 5e4)
+topfive_stars <- rasterize_counts(topfive_sf, res = 2.5e4)
 
-map_lim <- sf::st_bbox(topfive_sf) %>% 
+map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
+                       crs = "EPSG:4326") %>% 
   project_bbox() %>% 
-  expand_bbox(factor = 1.2)
+  expand_bbox(factor = 1.1)
 
@@
   scale_x_continuous(breaks = c(-65, -60, -55)) +
-  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
+  scale_y_continuous(breaks = c(-66, -64, -62, -60)) +
+  coord_ant(map_lim) +
   facet_grid(species ~ year) +
@@
 ```
+
+Distribution maps of five species of interest in each survey year, also aggregated to 25km bins.
+
+```{r}
+#| label: soi_map
+#| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-width: 6
+#| fig-dpi: 600
+
+soi <- c("GEPN", "BLPT", "ANTE", "LESE", "KIWH")
+
+soi_sf <- predators_agg %>% 
+  filter(species %in% soi) %>% 
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+
+soi_stars <- rasterize_counts(soi_sf, res = 2.5e4)
+
+map_lim <- sf::st_bbox(soi_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  stars::geom_stars(aes(fill = count), data = soi_stars) +
+  scale_fill_gradient(low = "lightblue", high = "darkred",
+                      trans = scales::pseudo_log_trans(),
+                      breaks = c(0, 25, 100, 400, 1000, 4000),
+                      na.value = NA) +
+  scale_x_continuous(breaks = c(-65, -60, -55)) +
+  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
+  facet_grid(species ~ year) +
+  expand_limits(fill = 0) +
+  coord_ant(map_lim) +
+  theme(axis.text.x = element_text(size = 7))
+```

### 2023-02-10 reports/02_stations.qmd
Create combined station-predator report and clean up documents

@@
   scale_color_manual(values = tod_pal()) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
   coord_ant(map_lim) +

### 2023-02-10 reports/03_combined.qmd
Create combined station-predator report and clean up documents

---
title: "Combined stations and predators"
format:
  html:
    code-fold: true
editor: visual
output-dir: "docs"
---

```{r}
#| label: setup
#| include: false

library(targets)
library(tidyverse)

tar_load(c("predators_agg", "zoop", "ant_sf"), 
         store = here::here("_targets"))
source(here::here("R", "sp.R"))
source(here::here("R", "zoop.R"))
```

Station locations on predator (five most abundant) count maps. Spatially restricted to station transects.

```{r}
#| label: combined_map
#| warning: false
#| column: screen
#| out-width: 100%
#| fig-width: 6
#| fig-dpi: 600

top_five <- predators_agg %>% 
  group_by(species) %>% 
  summarize(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>%
  slice(1:5)

topfive_sf <- predators_agg %>% 
  semi_join(top_five, by = "species") %>% 
  latlon_to_sf(coords = c("lon_mean", "lat_mean"))

topfive_stars <- rasterize_counts(topfive_sf, res = 2.5e4)

zoop_sf <- latlon_to_sf(zoop, 
                        coords = c("dec.longitude", "dec.latitude")) %>% 
  sf::st_transform(ant_proj()) %>% 
  rename(year = Year)
zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)

map_lim <- sf::st_bbox(zoop_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

ant_basemap(map_lim) +
  stars::geom_stars(aes(fill = count), data = topfive_stars) +
  geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
  scale_x_continuous(breaks = c(-65, -60, -55)) +
  scale_y_continuous(breaks = c(-66, -64, -62, -60)) +
  scale_fill_gradient(low = "lightblue", high = "darkred",
                      trans = scales::pseudo_log_trans(),
                      breaks = c(0, 25, 100, 400, 1000, 4000),
                      na.value = NA) +
  scale_color_manual(values = tod_pal()) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  coord_ant(map_lim) +
  facet_grid(species ~ year) +
  expand_limits(fill = 0) +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom",
        legend.title = element_blank())
```

As above, but for species of interest (Gentoo Penguins, Blue Petrels, Antarctic Terns, Leopard Seals, and Killer Whales).

```{r}
#| label: combined_map_soi
#| warning: false
#| column: screen
#| out-width: 100%
#| fig-width: 6
#| fig-dpi: 600

soi <- c("GEPN", "BLPT", "ANTE", "LESE", "KIWH")

soi_sf <- predators_agg %>% 
  filter(species %in% soi) %>% 
  latlon_to_sf(coords = c("lon_mean", "lat_mean"))

soi_stars <- rasterize_counts(soi_sf, res = 2.5e4)

ant_basemap(map_lim) +
  stars::geom_stars(aes(fill = count), data = soi_stars) +
  geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
  scale_x_continuous(breaks = c(-65, -60, -55)) +
  scale_y_continuous(breaks = c(-66, -64, -62, -60)) +
  scale_fill_gradient(low = "lightblue", high = "darkred",
                      trans = scales::pseudo_log_trans(),
                      breaks = c(0, 25, 100, 400, 1000, 4000),
                      na.value = NA) +
  scale_color_manual(values = tod_pal()) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  coord_ant(map_lim) +
  facet_grid(species ~ year) +
  expand_limits(fill = 0) +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom",
        legend.title = element_blank())
```

### 2023-02-14 reports/04_pred_clus.qmd
Perform clustering analysis

---
title: "Predator clusters"
format:
  html:
    code-fold: true
editor: visual
output-dir: "docs"
---

```{r}
#| label: setup
#| include: false

library(ggtree)
library(sf)
library(targets)
library(tidyverse)

tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
           "zoop_sf"), 
         store = here::here("_targets"))
source(here::here("R", "sp.R"))
source(here::here("R", "zoop.R"))
```

Following the methods of Dietrich et al. (2021), clustering analysis is restricted to species sighted near 5% or more of stations. This avoids granting undue leverage to rare species. `r n_distinct(predators_abundant$species)` species retained.

```{r}
#| label: abundant_predators

total_stations <- n_distinct(zoop_sf$amlr.station)
predators_stations %>%
  group_by(amlr.station, species) %>% 
  summarize(count = sum(count), .groups = "drop") %>%
  group_by(species) %>%
  summarize(stations_present = n(),
            station_frac = stations_present / total_stations,
            .groups = "drop") %>%
  arrange(desc(station_frac))

```

Hierarchical clustering of stations by predator community.

```{r}
#| label: predator_clusters
#| warning: false
#| column: screen
#| out-width: 100%
#| fig-width: 6
#| fig-dpi: 600

clus <- cutree(sightings_clust, 4)
g <- split(names(clus), clus)
p <- ggtree(sightings_clust)
station_clades <- sapply(g, function(n) MRCA(p, n))
p <- groupClade(p, station_clades, group_name = "subtree") + 
  aes(color = subtree)

p +
  layout_dendrogram() +
  theme_dendrogram()

```

### 2023-02-14 R/cluster.R

@@ -1,0 +1,75 @@
+# From Dietrich et al. 2021:
+# Abundance was log-transformed (log10(x + 1)) to minimize the influence of
+# highly abundant taxa in subsequent multivariate analysis. To minimize the
+# influence of rare taxa, taxa were excluded if they did not occur in at least
+# 5% of the net hauls. [predators sighted near >=5% of stations]
+# ...
+# A Bray-Curtis dissimilarity matrix based on abundance
+# ...
+# Separate hierarchical agglomerative cluster analyses of taxa abundance were
+# performed with the R package cluster
+# ...
+# Ward’s minimum variance was selected as the linkage method due to having the
+# highest agglomerative coefficient
+# ...
+# Indicator values were calculated for each taxa cluster combination based on
+# formulas in Dufrene and Leg- endre (1997) and implemented in the R package
+# labdsv
+
+assign_sightings <- function(sightings, stations, max_dist_km) {
+  stopifnot(inherits(sightings, "sf"),
+            inherits(stations, "sf"))
+
+  # Assign sightings to stations within year
+  assign_group <- function(sightings_group, sightings_key) {
+    # Note difference in column title Year vs year
+    stations_group <- stations[stations$Year == sightings_key$year, ]
+    # Find nearest station within year
+    nearest_station <- sf::st_join(sightings_group,
+                                   select(stations_group, amlr.station),
+                                   join = sf::st_nearest_feature)
+    # Get station coordinates
+    station_coords <- stations_group %>%
+      slice(map_int(nearest_station$amlr.station,
+                    ~ which(stations_group$amlr.station == .x)))
+    # Calculate distance from predator sighting to nearest station
+    nearest_station %>%
+      mutate(dist_to_station = sf::st_distance(nearest_station,
+                                               station_coords,
+                                               by_element = TRUE))
+  }
+
+  # browser()
+  # Assign sightings across years
+  sightings %>%
+    group_by(year) %>%
+    group_modify(assign_group) %>%
+    ungroup() %>%
+    filter(dist_to_station <= units::as_units(max_dist_km, "km"))
+}
+
+filter_sightings <- function(sightings, stations, station_thr) {
+  total_stations <- n_distinct(stations$amlr.station)
+  sightings_agg <- sightings %>%
+    group_by(amlr.station, species) %>%
+    summarize(count = sum(count), .groups = "drop")
+  retain <- sightings_agg %>%
[…]

### 2023-02-14 _targets.R

@@ -36,6 +36,8 @@ list(
   ),
   tar_target(predators, readr::read_csv(predators_file)),
   tar_target(predators_agg, aggregate_predators(predators)),
+  tar_target(predators_sf,
+             latlon_to_sf(predators_agg, coords = c("lon_mean", "lat_mean"))),
   # Zooplankton
   tar_target(
     zoop_file,
@@ -43,10 +45,22 @@ list(
     format = "file"
   ),
   tar_target(zoop, read_zoop(zoop_file)),
+  tar_target(zoop_sf,
+             latlon_to_sf(zoop, coords = c("dec.longitude", "dec.latitude"))),
   tar_target(
     zoop_long,
     pivot_longer(zoop, -(1:53), names_to = "taxa", values_to = "abundance")
   ),
+  # Clustering
+  tar_target(predators_stations,
+             assign_sightings(predators_sf, zoop_sf, max_dist_km = 15)),
+  tar_target(
+    predators_abundant,
+    filter_sightings(predators_stations, zoop_sf, station_thr = 0.05)
+  ),
+  tar_target(sightings_mtx, sightings_to_matrix(predators_abundant)),
+  tar_target(sightings_dist, vegan::vegdist(sightings_mtx, method = "bray")),
+  tar_target(sightings_clust, hclust(sightings_dist, method = "ward.D2")),
   # Reports
   tar_quarto(reports, here("reports"))
 )

### 2023-02-14 reports/04_pred_clus.qmd
Clean up cluster report

@@
 This avoids granting undue leverage to rare species.
-`r n_distinct(predators_abundant$species)` species retained.
+`r n_distinct(predators_abundant$species)` species retained (SNPT to MIWH).
 
@@
             .groups = "drop") %>%
-  arrange(desc(station_frac))
+  arrange(desc(station_frac)) %>% 
+  knitr::kable()
 
@@
 #| label: predator_clusters
-#| warning: false
-#| column: screen
-#| out-width: 100%
-#| fig-width: 6
-#| fig-dpi: 600
 

### 2023-02-15 reports/02_stations.qmd
Add zoop/physics maps

@@
 
-tar_load(c("zoop", "ant_sf"), store = here::here("_targets"))
+tar_load(c("zoop_sf", "ant_sf"), store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 #| label: station_maps
-
-zoop_sf <- latlon_to_sf(zoop, 
-                        coords = c("dec.longitude", "dec.latitude")) %>% 
-  sf::st_transform(ant_proj())
 
@@
 ```
+
+## Zooplankton community/physics maps
+
+```{r}
+#| label: zoop_physics_maps
+
+zoop_physics_vars <- c("Winter Cluster factor",
+                       "Net.ice.type",
+                       "Net.ice.percent",
+                       "zuml_m",
+                       "avg.Dens.Diff",
+                       "avg.salinity",
+                       "avg.temp",
+                       "Water.Zone",
+                       "Integ.chla.zuml",
+                       "Integ.phae.zuml",
+                       "Integ.chla.100m",
+                       "Integ.phae.100m")
+
+station_map <- function(var) {
+  ant_basemap(map_lim) +
+    geom_sf(aes(color = .data[[var]]), zoop_sf, size = 2, shape = 15) +
+    facet_wrap(~ Year) +
+    scale_x_continuous(breaks = c(-60, -55)) +
+    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+    coord_ant(map_lim) +
+    theme(legend.position = "bottom")
+}
+
+walk(zoop_physics_vars, compose(print, station_map))
+
+```

### 2023-02-15 R/zoop.R

@@ -2,7 +2,12 @@ read_zoop <- function(zoop_path) {
   readxl::read_excel(zoop_path, sheet = 1) %>%
     mutate(time.of.day = factor(time.of.day,
                                 levels = c("D", "T", "N"),
-                                labels = c("Day", "Twilight", "Night"))) %>%
+                                labels = c("Day", "Twilight", "Night")),
+           across(c(zuml_m,
+                    starts_with("avg"),
+                    starts_with("Integ"),
+                    Net.ice.percent),
+                  parse_number)) %>%
     select(-57) %>%
     filter(leg == "W")
 }

### 2023-02-16 reports/02_stations.qmd
Fix index.md
add zooplankton maps

@@
     code-fold: true
+    toc: true
 editor: visual
@@
 
-## Zooplankton community/physics maps
+## Zooplankton maps
 
 ```{r}
-#| label: zoop_physics_maps
+#| label: station_map_fn
 
-zoop_physics_vars <- c("Winter Cluster factor",
-                       "Net.ice.type",
+station_map <- function(var) {
+  ant_basemap(map_lim) +
+    geom_sf(aes(color = .data[[var]]), zoop_sf, size = 2, shape = 15) +
+    facet_wrap(~ Year) +
+    scale_x_continuous(breaks = c(-60, -55)) +
+    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+    coord_ant(map_lim) +
+    theme(legend.position = "bottom")
+}
+
+```
+
+```{r}
+#| label: zoop_maps
+
+zoop_vars <- c("Winter Cluster factor",
+               "NMDS_x",
+               "NMDS_y",
+               "NMDS_z",
+               "Amphipoda_Primno_N.A",
+               "Amphipoda_Themisto_N.A",
+               "Chaetognatha_N.A",
+               "Copepoda_Cacutus_N.A",
+               "Copepoda_Cprop_N.A",
+               "Copepoda_Heterorhabdus_N.A",
+               "Copepoda_Metridia_N.A",
+               "Copepoda_Paraeuchaeta_N.A",
+               "Copepoda_Pleuromama_N.A",
+               "Copepoda_Rhincalanus_N.A",
+               "Euphausiidae_Ecrystall_Adult",
+               "Euphausiidae_Efrigida_Adult",
+               "Euphausiidae_Efrigida_Larvae",
+               "Euphausiidae_Esuperba_Adult",
+               "Euphausiidae_Etricantha_Adult",
+               "Euphausiidae_Tmac_Adult",
+               "Euphausiidae_Tmac_Larvae",
+               "Gastropoda_Clione_N.A",
+               "Gastropoda_Limacina_N.A",
+               "Ostracoda_N.A",
+               "Radiolaria_N.A",
+               "Siphonophora_N.A",
+               "Tunicata_Salpa_N.A",
+               "Euphausiidae_Esuperba_F_Tot",
+               "Pisces_Paralepididae_NotolepisSpp_Larvae_Tot",
+               "Pisces_Myctophidae_Adult_Tot",
+               "Pisces_Myctophidae_Larvae_Tot",
+               "Pisces_Other_Larvae_Tot",
+               "Pisces_Chaenichthyidae_Larvae_Tot",
+               "Pisces_Nototheniidae_Larvae_Tot")
+
+walk(zoop_vars, compose(print, station_map))
+
+```
+
+## Physics maps
+
+```{r}
+#| label: physics_maps
+
+physics_vars <- c("Net.ice.type",
                   "Net.ice.percent",
@@
 
-station_map <- function(var) {
-  ant_basemap(map_lim) +
-    geom_sf(aes(color = .data[[var]]), zoop_sf, size = 2, shape = 15) +
-    facet_wrap(~ Year) +
-    scale_x_continuous(breaks = c(-60, -55)) +
-    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-    coord_ant(map_lim) +
-    theme(legend.position = "bottom")
-}
-
-walk(zoop_physics_vars, compose(print, station_map))
+walk(physics_vars, compose(print, station_map))
 

### 2023-02-21 reports/02_stations.qmd
fix zoop station maps (pseudo-log abundance, better gradient)

--- 2023-02-10 reports/02_stations.qmd
@@
 ```
+
+## Zooplankton maps
+
+```{r}
+#| label: station_map_fn
+
+station_map <- function(var, pseudolog = FALSE) {
+  zoop_trans <- zoop_sf
+  if (pseudolog) {
+    old_var <- var 
+    var <- str_glue("log({var}+1)")
+    zoop_trans[[var]] <- log(zoop_trans[[old_var]] + 1)
+  }
+  ant_basemap(map_lim) +
+    geom_sf(aes(color = .data[[var]]), zoop_trans, size = 2, shape = 15) +
+    facet_wrap(~ Year) +
+    scale_x_continuous(breaks = c(-60, -55)) +
+    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+    { if (is.numeric(zoop_trans[[var]])) 
+      scale_color_distiller(palette = "PuRd", direction = 1) } +
+    coord_ant(map_lim) +
+    theme(legend.position = "bottom")
+}
+
+```
+
+```{r}
+#| label: zoop_maps
+
+zoop_vars_raw <- c("Winter Cluster factor",
+                   "NMDS_x",
+                   "NMDS_y",
+                   "NMDS_z")
+zoop_vars_pseudolog <- c("Amphipoda_Primno_N.A",
+                         "Amphipoda_Themisto_N.A",
+                         "Chaetognatha_N.A",
+                         "Copepoda_Cacutus_N.A",
+                         "Copepoda_Cprop_N.A",
+                         "Copepoda_Heterorhabdus_N.A",
+                         "Copepoda_Metridia_N.A",
+                         "Copepoda_Paraeuchaeta_N.A",
+                         "Copepoda_Pleuromama_N.A",
+                         "Copepoda_Rhincalanus_N.A",
+                         "Euphausiidae_Ecrystall_Adult",
+                         "Euphausiidae_Efrigida_Adult",
+                         "Euphausiidae_Efrigida_Larvae",
+                         "Euphausiidae_Esuperba_Adult",
+                         "Euphausiidae_Etricantha_Adult",
+                         "Euphausiidae_Tmac_Adult",
+                         "Euphausiidae_Tmac_Larvae",
+                         "Gastropoda_Clione_N.A",
+                         "Gastropoda_Limacina_N.A",
+                         "Ostracoda_N.A",
+                         "Radiolaria_N.A",
+                         "Siphonophora_N.A",
+                         "Tunicata_Salpa_N.A",
+                         "Euphausiidae_Esuperba_F_Tot",
+                         "Pisces_Paralepididae_NotolepisSpp_Larvae_Tot",
+                         "Pisces_Myctophidae_Adult_Tot",
+                         "Pisces_Myctophidae_Larvae_Tot",
+                         "Pisces_Other_Larvae_Tot",
+                         "Pisces_Chaenichthyidae_Larvae_Tot",
+                         "Pisces_Nototheniidae_Larvae_Tot")
+
+walk(zoop_vars_raw, compose(print, station_map))
+walk(zoop_vars_pseudolog, 
+     compose(print, partial(station_map, pseudolog = TRUE)))
+
+```
+
+## Physics maps
+
+```{r}
+#| label: physics_maps
+
+physics_vars <- c("Net.ice.type",
+                  "Net.ice.percent",
+                  "zuml_m",
+                  "avg.Dens.Diff",
+                  "avg.salinity",
+                  "avg.temp",
+                  "Integ.chla.zuml",
+                  "Integ.phae.zuml",
+                  "Integ.chla.100m",
+                  "Integ.phae.100m")
+
+walk(physics_vars, compose(print, station_map))
+
+```

### 2023-02-22 reports/01_predators.qmd
Add effort map

@@
 
+library(sf)
 library(targets)
@@
 
-tar_load(c("predators_agg", "ant_sf"), store = here::here("_targets"))
+tar_load(c("predators_agg", "ant_sf", "effort_sf"), 
+         store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 
+Survey effort. 21 out of the 19404 intervals were recorded as nmi \> 1000 (as opposed to \~1 nmi for the rest). Those intervals excluded from the survey effort.
+
+```{r}
+#| label: survey_effort
+#| warning: false
+
+effort_stars <- effort_sf %>% 
+  filter(nmi < 1000) %>% 
+  rasterize_counts(count_col = "nmi",
+                   res = 2.5e4)
+map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
+                       crs = "EPSG:4326") %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+ant_basemap(map_lim) + 
+  stars::geom_stars(aes(fill = nmi), data = effort_stars) +
+  scale_fill_distiller(palette = "PuRd", 
+                       direction = 1, 
+                       na.value = NA) +
+  scale_x_continuous(breaks = seq(-65, -55, by = 5)) +
+  scale_y_continuous(breaks = seq(-65, -58, by = 2)) +
+  coord_ant(map_lim) +
+  facet_wrap(vars(year)) +
+  expand_limits(fill = 0)
+
+```
+
 Distribution maps of the five most abundant species in each survey year, aggregated to 25km bins.
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = count), data = topfive_stars) +
-  scale_fill_gradient(low = "lightblue", high = "darkred",
-                      trans = scales::pseudo_log_trans(),
-                      breaks = c(0, 25, 100, 400, 1000, 4000),
+  stars::geom_stars(aes(fill = log(count + 1)), data = topfive_stars) +
+  scale_fill_distiller(palette = "PuRd",
+                       direction = 1,
                        na.value = NA) +
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = count), data = soi_stars) +
-  scale_fill_gradient(low = "lightblue", high = "darkred",
-                      trans = scales::pseudo_log_trans(),
-                      breaks = c(0, 25, 100, 400, 1000, 4000),
+  stars::geom_stars(aes(fill = log(count + 1)), data = soi_stars) +
+  scale_fill_distiller(palette = "PuRd",
+                       direction = 1,
                        na.value = NA) +

### 2023-02-22 R/sp.R

@@ -48,7 +48,8 @@ rasterize_counts <- function(counts_sf,
                              layer_name = count_col,
                              limits = ant_lims(),
                              res = 5e4) {
-
+  stopifnot(inherits(counts_sf, "sf"),
+            all(c(count_col, species_col, year_col) %in% colnames(counts_sf)))
   # Group counts, call rasterize on each group, convert back to df
   rasterized_df <- counts_sf %>%
     group_by(.data[[species_col]], .data[[year_col]]) %>%

### 2023-02-22 reports/04_pred_clus.qmd
Clean up predator table in cluster report

@@
 This avoids granting undue leverage to rare species.
-`r n_distinct(predators_abundant$species)` species retained (SNPT to MIWH).
+`r n_distinct(predators_abundant$species)` species retained (i.e., species above red line).
 
@@
 total_stations <- n_distinct(zoop_sf$amlr.station)
-predators_stations %>%
+pred_frac <- predators_stations %>%
   group_by(amlr.station, species) %>% 
@@
             .groups = "drop") %>%
-  arrange(desc(station_frac)) %>% 
-  knitr::kable()
+  arrange(desc(station_frac))
+cutoff_row <- sum(pred_frac$station_frac >= 0.05)
+
+frac_fn <- scales::label_percent(accuracy = 0.1)
+pred_frac %>% 
+  set_names(c("Species", "Stations", "Fraction")) %>% 
+  mutate(Fraction = frac_fn(Fraction)) %>% 
+  knitr::kable(align = "lrr") %>% 
+  kableExtra::kable_styling("striped",
+                            full_width = FALSE) %>% 
+  kableExtra::row_spec(cutoff_row, 
+                       extra_css = "border-bottom: 2px dashed red")
 

### 2023-02-22 reports/04_pred_clus.qmd
Cluster comparison

@@
 
+library(ggmosaic)
 library(ggtree)
@@
 tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
-           "zoop_sf"), 
+           "zoop"), 
          store = here::here("_targets"))
@@
 
-total_stations <- n_distinct(zoop_sf$amlr.station)
+total_stations <- n_distinct(zoop$amlr.station)
 pred_frac <- predators_stations %>%
@@
 Hierarchical clustering of stations by predator community.
+Chose 5 clusters a priori to match Dietrich et al.
+(2021).
+Note: `r nrow(zoop)` stations were included in the zooplankton clustering, but only `r length(sightings_clust$labels)` stations were used for predator clustering.
 
 ```{r}
-#| label: predator_clusters
+#| label: predator_dendro
 
-clus <- cutree(sightings_clust, 4)
-g <- split(names(clus), clus)
-p <- ggtree(sightings_clust)
-station_clades <- sapply(g, function(n) MRCA(p, n))
-p <- groupClade(p, station_clades, group_name = "subtree") + 
-  aes(color = subtree)
+# Cut predators into five clusters
+clust <- cutree(sightings_clust, 5)
+g <- split(names(clust), clust)
+p <- ggtree(sightings_clust, hang = -1)
+pred_clust_mrca <- sapply(g, function(n) MRCA(p, n))
 
-p +
+p %>% 
+  groupClade(pred_clust_mrca, group_name = "Predator cluster") + 
+  aes(color = `Predator cluster`) +
   layout_dendrogram() +
   theme_dendrogram()
+```
+
+How do the predator clusters compare to the zooplankton clusters?
+
+```{r}
+#| label: predator_mosaic
+#| warning: false
+
+# Join with zooplankton clusters
+both_clust <- zoop %>% 
+  transmute(amlr.station, 
+            zoop_clust = factor(`Winter Cluster factor`)) %>% 
+  right_join(tibble(amlr.station = names(clust),
+                    pred_clust = factor(clust)),
+             by = "amlr.station")
+
+# Mosaic plot
+cluster_pal = c(`1` = "#96B687",
+                `2a` = "#9CB2C8",
+                `2b` = "#CAB3CC",
+                `3a` = "#D29A97",
+                `3b` = "#E8AD5C")
+ggplot(both_clust) +
+  geom_mosaic(aes(x = product(zoop_clust, pred_clust), 
+                  fill = zoop_clust)) +
+  geom_mosaic_text(aes(x = product(zoop_clust, pred_clust),
+                       label = after_stat(.wt)),
+                   as.label = TRUE) +
+  scale_fill_manual(values = cluster_pal) +
+  labs(x = "Predator cluster",
+       y = "Zooplankton cluster",
+       fill = "Zooplankton cluster") +
+  theme_classic()
+```
+
+The above mosaic plot shows the distribution of stations by predator and zooplankton stations.
+For example, the bottom left rectangle indicates 9 stations were assigned to zooplankton cluster 1 and predator cluster 1.
+Rectangles color-coded by zooplankton cluster using the color palette from Dietrich et al.
+2021 Fig 3a.
+
+Upon visual inspection of the mosaic plot, I don't perceive any clear correlations between the zooplankton and predator cluster assignments.
+I tested the strength of the relationship between the two cluster assignments through simulation of the Rand index (RI).
+RI is a measure of similarity between two clusterings ranging from 0 (no agreement) to 1 (total agreement).
+The histogram below shows the distribution of RI values when the cluster assignments were randomly shuffled 100,000 times.
+The vertical red line is the observed RI value.
+
+```{r}
+#| label: randindex
+
+set.seed(1311)
+
+expected_ri <- replicate(
+  100000,
+  aricode::RI(
+    sample(both_clust$zoop_clust, nrow(both_clust)),
+    sample(both_clust$pred_clust, nrow(both_clust))
+  )
+)
+
+observed_ri <- aricode::RI(both_clust$zoop_clust, both_clust$pred_clust)
+
+ri_pval <- sum(expected_ri > observed_ri) / 100000
+
+ggplot(tibble(`Rand index` = expected_ri), aes(`Rand index`)) +
+  geom_histogram(binwidth = 0.001) +
+  geom_vline(xintercept = observed_ri, color = "red") +
+  theme_classic()
 
 ```
+
+Based on this simulation, the correlation between the zooplankton and predator assignments *is* significant (p = `r as.character(ri_pval)`).
+But I haven't figured out how to assess effect size yet, which I imagine is small.

### 2023-02-23 reports/04_pred_clus.qmd
Indicator species and predator cluster maps

@@
 tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
-           "zoop"), 
+           "zoop_sf"), 
          store = here::here("_targets"))
@@
 ```
+
+## Predator clusters
 
@@
 
-total_stations <- n_distinct(zoop$amlr.station)
+total_stations <- n_distinct(zoop_sf$amlr.station)
 pred_frac <- predators_stations %>%
@@
 (2021).
-Note: `r nrow(zoop)` stations were included in the zooplankton clustering, but only `r length(sightings_clust$labels)` stations were used for predator clustering.
+Note: `r nrow(zoop_sf)` stations were included in the zooplankton clustering, but only `r length(sightings_clust$labels)` stations were used for predator clustering.
 
@@
 # Join with zooplankton clusters
-both_clust <- zoop %>% 
+both_clust <- zoop_sf %>% 
   transmute(amlr.station, 
@@
 
-The above mosaic plot shows the distribution of stations by predator and zooplankton stations.
+The above mosaic plot shows the distribution of stations by predator and zooplankton clusters.
 For example, the bottom left rectangle indicates 9 stations were assigned to zooplankton cluster 1 and predator cluster 1.
@@
 But I haven't figured out how to assess effect size yet, which I imagine is small.
+
+## Indicator species
+
+Which species best represent each cluster? Using Dufrene-Legendre indicator species analysis. Apply indicator value cutoff at 25%.
+
+```{r}
+#| label: indicator_species
+
+species_wide <- pivot_wider(predators_abundant,
+                            names_from = species,
+                            values_from = count,
+                            values_fill = 0) %>% 
+  column_to_rownames("amlr.station")
+cluster_indval <- labdsv::indval(species_wide, clust)
+indval_mtx <- cluster_indval$indval
+indval_fmt <- apply(indval_mtx, 2, function(x) {
+  ord <- order(x, decreasing = TRUE)
+  result <- ifelse(x >= 0.25, 
+         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
+         NA) %>% 
+    `[`(ord) %>% 
+    na.omit() %>% 
+    paste(collapse = "<br>")
+  ifelse(result == "", "None", result)
+})
+as_tibble(t(indval_fmt)) %>% 
+  knitr::kable(format = "html", escape = FALSE) %>% 
+  kableExtra::kable_styling()
+
+```
+
+According to the 25% cutoff, clusters 1 and 4 lack indicator species.
+In agreement, every species' highest indicator value falls in clusters 2, 3, or 5.
+This suggests to me 5 clusters is a poor choice for the predators and we should use fewer.
+I'll calculate the gap statistic for the hierarchical cluster analysis to quantitatively choose a data-informed number of clusters and repeat this entire analysis.
+For now, continuing with 5 clusters.
+
+I don't know enough about the Antarctic Peninsula megafauna to say whether the indicator species cluster together logically.
+The fulmarine/*Daption* petrels are split across Clusters 2 and 3.
+Cluster 2 lumps Cape Petrel and Antarctic Fulmar with the prion-like Blue Petrel.
+Cluster 3 combines the fulmarine Antarctic and Snow Petrels with the Kerguelen Petrel and fur seals.
+Cluster 5 seems to have the most ice-associated species (phocids and Adelie penguins).
+
+## Cluster maps
+
+What geospatial patterns do we see in the predator clusters?
+Cluster 5 is mostly found in the Bransfield Strait, which makes sense for the Adelie Penguin-crabeater seal-leopard seal complex.
+Cluster 2 (Antarctic Fulmar, Cape Petrel, and Blue Petrel) tends to show up oceanward of the South Shetland Islands.
+Cluster 3 (fur seal, snow petrel, Kerguelen petrel, and Antarctic Petrel) seems the most variable; found both inside and outside the Strait depending on the year.
+
+```{r}
+#| label: cluster_maps
+
+pred_clust_sf <- tibble(amlr.station = names(clust),
+                        pred_clust = factor(clust)) %>% 
+  left_join(select(zoop_sf,
+                   amlr.station,
+                   geometry,
+                   Year),
+            by = "amlr.station") %>% 
+  st_as_sf() %>% 
+  st_transform(ant_proj())
+
+map_lim <- sf::st_bbox(pred_clust_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
+  facet_wrap(~ Year) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom")
+```

### 2023-03-13 README.Rmd
Add README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# antwincomm

<!-- badges: start -->

<!-- badges: end -->

Characterizing the marine winter predator community (seabirds and marine mammals) around the northern Antarctic Peninsula (including the Bransfield Strait and South Shetland Islands) using observations collected during the 2012-2016 NOAA/NSF austral winter cruises on the RV Ice Breaker Nathaniel B. Palmer.

*Work in progress!!*

## Code

All code under development for **antwincomm** can be found in this repository. The analysis pipeline [uses `targets`](https://books.ropensci.org/targets/) to manage workflow caching and dependencies. Better documentation to follow (promise). This repository will be archived on Zenodo or somewhere similar at publication.

## Data

The **antwincomm** data are too large for GitHub, so they're stored in the [antwincomm Open Science Framework project](https://osf.io/hwnvy/). Currently private, will be made public no later than time of publication. For access to the OSF project, please contact Max Czapanskiy (max [dot] czapanskiy [at] noaa [dot] gov).

For the code to work, you must download data/WAMLR from the OSF project into data/WAMLR in your local copy of the repository. You can do that manually or by running the following lines of code. Until the project is made public, though, you'll first have to [authenticate `osfr`](https://docs.ropensci.org/osfr/articles/auth.html).

```{r eval=FALSE}
dir.create(here::here("data"))
osf_wamlr <- osfr::osf_retrieve_file("6407bf6228e5c503cc9376b3")
osfr::osf_download(osf_wamlr, path = here::here("data"), recurse = TRUE)
```

## Outputs

See the [project website](https://flukeandfeather.github/io/antwincomm) for a list of reports documenting the analysis pipeline.

### 2023-03-13 R/wamlr.R

@@ -1,21 +1,6 @@
-download_wamlr <- function(wamlr_dir) {
-  # Don't proceed if the WAMLR data directory has files in it
-  if (length(dir(wamlr_dir)) > 0)
-    return()
-
-  # We need the OSF_PAT to proceed
-  if (Sys.getenv("OSF_PAT") == "")
-    stop("No OSF_PAT environment variable defined. Create an OSF personal access token and define OSF_PAT in .Renviron (see https://docs.ropensci.org/osfr/articles/auth.html)")
-
-  # Retreive the WAMLR node
-  antwincomm_prj <- osfr::osf_retrieve_node("https://osf.io/hwnvy/")
-  antwincomm_data <- filter(osfr::osf_ls_files(antwincomm_prj),
-                            name == "data")
-  antwincomm_wamlr <- filter(osfr::osf_ls_files(antwincomm_data),
-                             name == "WAMLR")
-
-  # Download data
-  osfr::osf_download(antwincomm_wamlr,
-                     path = dirname(wamlr_dir), # So we don't get WAMLR/WAMLR/
-                     recurse = TRUE)
+check_wamlr <- function(wamlr_path) {
+  if (!dir.exists(path))
+    stop(str_glue("WAMLR directory not found at {wamlr_path}. Follow the ",
+                  "instructions in the README for downloading these data."))
+  wamlr_path
 }

### 2023-03-13 _targets.R

@@ -17,16 +17,7 @@ list(
   # Download WAMLR data from OSF (if necessary)
   tar_target(
     wamlr_dir,
-    {
-      path <- here("data", "WAMLR")
-      if (!dir.exists(path)) dir.create(path, recursive = TRUE)
-      path
-    },
-    format = "file"
-  ),
-  tar_target(
-    wamlr_data,
-    download_wamlr(wamlr_dir)
+    check_wamlr(here("data", "WAMLR"))
   ),
   # Spatial data
   tar_target(
@@ -45,11 +36,7 @@ list(
   # Predators
   tar_target(
     predators_file,
-    {
-      # Create a dependency on wamlr_data so download_wamlr() happens first
-      wamlr_data
-      here(wamlr_dir, "WAMLR-2012-2016-Underway-Predator-Sightings.csv")
-    },
+    here(wamlr_dir, "WAMLR-2012-2016-Underway-Predator-Sightings.csv"),
     format = "file"
   ),
   tar_target(predators, readr::read_csv(predators_file)),
@@ -68,11 +55,7 @@ list(
   # Zooplankton
   tar_target(
     zoop_file,
-    {
-      # Create a dependency on wamlr_data so download_wamlr() happens first
-      wamlr_data
-      here(wamlr_dir, "WAMLR-2012-2016-Zooplnkton-Physics.xlsx")
-    },
+    here(wamlr_dir, "WAMLR-2012-2016-Zooplnkton-Physics.xlsx"),
     format = "file"
   ),
   tar_target(zoop, read_zoop(zoop_file)),

### 2023-03-14 reports/04_pred_clus.qmd
Select optimal number of clusters

@@
 output-dir: "docs"
+bibliography: references.bib
 ---
@@
 tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
-           "zoop_sf"), 
+           "sightings_mtx", "sightings_dist", "zoop_sf"), 
          store = here::here("_targets"))
@@
 ```
+
+## Optimal number of clusters
+
+We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters, $k$.
+Using the "first SE max" rule, $k$ is defined as the smallest number of clusters where the gap statistic is greater than the first local maximum less its standard error.
+In the following figure, the first local maximum of the gap statistic occurs at four clusters.
+The threshold (local maximum less its standard error) is represented by the red dashed line.
+The smallest number of clusters with a gap statistic greater than that threshold is $k=3$, indicated by the red point and standard error lines.
+
+```{r}
+cluster_fn <- function(x, k) {
+  bray_dist <- function(x) {
+    vegan::vegdist(x, method = "bray")
+  }
+  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
+}
+sightings_gap <- suppressWarnings(
+  cluster::clusGap(sightings_mtx, 
+                   FUN = cluster_fn,
+                   K.max = 10, 
+                   B = 100,
+                   verbose = FALSE)
+)
+sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
+  mutate(
+    k = seq_along(logW),
+    is_optimal = k == cluster::maxSE(gap, SE.sim, method = "firstSEmax")
+  )
+firstmaxk <- with(sightings_gap_df,
+                  cluster::maxSE(gap, SE.sim, method = "firstmax"))
+gap_thr <- with(sightings_gap_df, gap[firstmaxk] - SE.sim[firstmaxk])
+ggplot(sightings_gap_df, aes(x = k, y = gap)) +
+  geom_line() +
+  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
+                      color = is_optimal)) +
+  geom_hline(yintercept = gap_thr, color = "red", linetype = "dashed") +
+  scale_color_manual(values = c(`TRUE` = "red",
+                                `FALSE` = "black")) +
+  theme_classic() +
+  theme(legend.position = "none")
+```

### 2023-03-14 R/cluster.R

@@ -39,7 +39,6 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
                                                by_element = TRUE))
   }
 
-  # browser()
   # Assign sightings across years
   sightings %>%
     group_by(year) %>%

### 2023-03-14 reports/04_pred_clus.qmd
Repeat analysis with 3 clusters

@@
 ```
+
+## 3-cluster analysis
+
+```{r}
+#| label: predator_dendro2
+
+# Cut predators into 3 clusters
+clust <- cutree(sightings_clust, 3)
+g <- split(names(clust), clust)
+p <- ggtree(sightings_clust, hang = -1)
+pred_clust_mrca <- sapply(g, function(n) MRCA(p, n))
+
+p %>% 
+  groupClade(pred_clust_mrca, group_name = "Predator cluster") + 
+  aes(color = `Predator cluster`) +
+  layout_dendrogram() +
+  theme_dendrogram()
+```
+
+```{r}
+#| label: predator_mosaic2
+#| warning: false
+
+# Join with zooplankton clusters
+both_clust <- zoop_sf %>% 
+  transmute(amlr.station, 
+            zoop_clust = factor(`Winter Cluster factor`)) %>% 
+  right_join(tibble(amlr.station = names(clust),
+                    pred_clust = factor(clust)),
+             by = "amlr.station")
+
+# Mosaic plot
+cluster_pal = c(`1` = "#96B687",
+                `2a` = "#9CB2C8",
+                `2b` = "#CAB3CC",
+                `3a` = "#D29A97",
+                `3b` = "#E8AD5C")
+ggplot(both_clust) +
+  geom_mosaic(aes(x = product(zoop_clust, pred_clust), 
+                  fill = zoop_clust)) +
+  geom_mosaic_text(aes(x = product(zoop_clust, pred_clust),
+                       label = after_stat(.wt)),
+                   as.label = TRUE) +
+  scale_fill_manual(values = cluster_pal) +
+  labs(x = "Predator cluster",
+       y = "Zooplankton cluster",
+       fill = "Zooplankton cluster") +
+  theme_classic()
+```
+
+```{r}
+#| label: indicator_species2
+
+species_wide <- pivot_wider(predators_abundant,
+                            names_from = species,
+                            values_from = count,
+                            values_fill = 0) %>% 
+  column_to_rownames("amlr.station")
+cluster_indval <- labdsv::indval(species_wide, clust)
+indval_mtx <- cluster_indval$indval
+indval_fmt <- apply(indval_mtx, 2, function(x) {
+  ord <- order(x, decreasing = TRUE)
+  result <- ifelse(x >= 0.25, 
+         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
+         NA) %>% 
+    `[`(ord) %>% 
+    na.omit() %>% 
+    paste(collapse = "<br>")
+  ifelse(result == "", "None", result)
+})
+as_tibble(t(indval_fmt)) %>% 
+  knitr::kable(format = "html", escape = FALSE) %>% 
+  kableExtra::kable_styling()
+```
+
+```{r}
+#| label: cluster_maps2
+
+pred_clust_sf <- tibble(amlr.station = names(clust),
+                        pred_clust = factor(clust)) %>% 
+  left_join(select(zoop_sf,
+                   amlr.station,
+                   geometry,
+                   Year),
+            by = "amlr.station") %>% 
+  st_as_sf() %>% 
+  st_transform(ant_proj())
+
+map_lim <- sf::st_bbox(pred_clust_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
+  facet_wrap(~ Year) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom")
+```

### 2023-03-14 reports/04_pred_clus.qmd
Reduce predator clusters

--- 2023-02-23 reports/04_pred_clus.qmd
@@
 library(tidyverse)
+library(vegan)
 
 tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
-           "zoop_sf"), 
+           "sightings_mtx", "sightings_dist", "zoop_sf"), 
          store = here::here("_targets"))
@@
 ```
+
+## Optimal number of clusters
+
+We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters ($k$).
+The gap statistic ($f$) is a goodness of clustering measure.
+The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s_{k+1}$ where $s$ is the standard error of $f$.
+According to this rule, the predators are best described by two clusters.
+
+```{r}
+cluster_fn <- function(x, k) {
+  bray_dist <- function(x) {
+    vegdist(x, method = "bray")
+  }
+  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
+}
+sightings_gap <- suppressWarnings(
+  cluster::clusGap(sightings_mtx, 
+                   FUN = cluster_fn,
+                   K.max = 10, 
+                   B = 500,
+                   verbose = FALSE)
+)
+sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
+  mutate(
+    k = seq_along(logW),
+    is_optimal = k == cluster::maxSE(gap, SE.sim, method = "Tibs2001SEmax")
+  )
+firstmaxk <- with(sightings_gap_df,
+                  cluster::maxSE(gap, SE.sim, method = "firstmax"))
+gap_thr <- with(sightings_gap_df, gap[firstmaxk] - SE.sim[firstmaxk])
+ggplot(sightings_gap_df, aes(x = k, y = gap)) +
+  geom_line() +
+  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
+                      color = is_optimal)) +
+  geom_hline(yintercept = gap_thr, color = "red", linetype = "dashed") +
+  scale_color_manual(values = c(`TRUE` = "red",
+                                `FALSE` = "black")) +
+  theme_classic() +
+  theme(legend.position = "none")
+```
+
+## 2-cluster analysis
+
+Repeat the indicator species analysis and maps with two clusters.
+
+```{r}
+#| label: predator_dendro2
+
+# Cut predators into 2 clusters
+clust <- cutree(sightings_clust, 2)
+g <- split(names(clust), clust)
+p <- ggtree(sightings_clust, hang = -1)
+pred_clust_mrca <- sapply(g, function(n) MRCA(p, n))
+
+p %>% 
+  groupClade(pred_clust_mrca, group_name = "Predator cluster") + 
+  aes(color = `Predator cluster`) +
+  layout_dendrogram() +
+  theme_dendrogram()
+```
+
+```{r}
+#| label: indicator_species2
+
+species_wide <- pivot_wider(predators_abundant,
+                            names_from = species,
+                            values_from = count,
+                            values_fill = 0) %>% 
+  column_to_rownames("amlr.station")
+cluster_indval <- labdsv::indval(species_wide, clust)
+indval_mtx <- cluster_indval$indval
+indval_fmt <- apply(indval_mtx, 2, function(x) {
+  ord <- order(x, decreasing = TRUE)
+  result <- ifelse(x >= 0.25, 
+         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
+         NA) %>% 
+    `[`(ord) %>% 
+    na.omit() %>% 
+    paste(collapse = "<br>")
+  ifelse(result == "", "None", result)
+})
+as_tibble(t(indval_fmt)) %>% 
+  knitr::kable(format = "html", escape = FALSE) %>% 
+  kableExtra::kable_styling()
+```
+
+```{r}
+#| label: cluster_maps2
+
+pred_clust_sf <- tibble(amlr.station = names(clust),
+                        pred_clust = factor(clust)) %>% 
+  left_join(select(zoop_sf,
+                   amlr.station,
+                   geometry,
+                   Year),
+            by = "amlr.station") %>% 
+  st_as_sf() %>% 
+  st_transform(ant_proj())
+
+map_lim <- sf::st_bbox(pred_clust_sf) %>% 
+  project_bbox() %>% 
+  expand_bbox(factor = 1.2)
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
+  facet_wrap(~ Year) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom")
+```
+
+## NMDS
+
+```{r}
+#| label: nmds
+
+sightings_nmds <- metaMDS(
+  sightings_mtx, 
+  distance = "bray",
+  k = 3,
+  trymax = 100
+)
+
+stressplot(sightings_nmds)
+
+```
+
+```{r}
+#| label: loading
+
+nmds_df <- sightings_nmds$points %>% 
+  as_tibble(rownames = "amlr.station") %>% 
+  rename_with(~ paste0("N", .x), starts_with("MDS")) %>% 
+  right_join(tibble(amlr.station = names(clust),
+                    pred_clust = factor(clust)),
+             by = "amlr.station")
+
+ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust)) +
+  theme_classic()
+
+ggplot(nmds_df, aes(NMDS1, NMDS3)) +
+  geom_point(aes(color = pred_clust)) +
+  theme_classic()
+
+ggplot(nmds_df, aes(NMDS2, NMDS3)) +
+  geom_point(aes(color = pred_clust)) +
+  theme_classic()
+```

### 2023-03-14 reports/04_pred_clus.qmd
Details for optimal cluster number

@@
 The gap statistic ($f$) is a goodness of clustering measure.
-The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s_{k+1}$ where $s$ is the standard error of $f$.
+The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
 According to this rule, the predators are best described by two clusters.
+However, $f(2)$ is only slightly greater than $f(3)-s(3)$ and, because $s$ is estimated by bootstrapping, $k=3$ may be selected sometimes depending on stochasticity.
 
@@
 }
+set.seed(931)
 sightings_gap <- suppressWarnings(
@@
   )
-firstmaxk <- with(sightings_gap_df,
-                  cluster::maxSE(gap, SE.sim, method = "firstmax"))
-gap_thr <- with(sightings_gap_df, gap[firstmaxk] - SE.sim[firstmaxk])
 ggplot(sightings_gap_df, aes(x = k, y = gap)) +
@@
                       color = is_optimal)) +
-  geom_hline(yintercept = gap_thr, color = "red", linetype = "dashed") +
   scale_color_manual(values = c(`TRUE` = "red",

### 2023-03-22 reports/04_pred_clus.qmd
NMDS and ice categories

@@
 
+library(cluster)
 library(ggmosaic)
@@
 
-tar_load(c("predators_stations", "predators_abundant", "sightings_clust",
-           "sightings_mtx", "sightings_dist", "zoop_sf"), 
+tar_load(c("predators", "predators_stations", 
+           "predators_abundant", "sightings_clust", "sightings_mtx", 
+           "sightings_dist", "zoop_sf"), 
          store = here::here("_targets"))
@@
 The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
-According to this rule, the predators are best described by two clusters.
-However, $f(2)$ is only slightly greater than $f(3)-s(3)$ and, because $s$ is estimated by bootstrapping, $k=3$ may be selected sometimes depending on stochasticity.
+According to this rule, the predators are best described by two clusters (indicated in blue).
+However, $f(2)$ is only slightly greater than $f(3)-s(3)$ and, because $s$ is estimated by bootstrapping, sometimes re-running the gap statistic estimate yields an optimum at $k=3$ instead.
+An alternative rule suggests choosing the smallest $k$ such $f(k) \geq f_{max} - s_{max}$, where $f_{max}$ and $s_{max}$ are $f$ and $s$ at the first local maximum, respectively.
+In the figure below, $f_{max} - s_{max}$ is represented by the dashed red line, and the smallest $k$ greater than that threshold is $k=3$ (red).
 
@@
 sightings_gap <- suppressWarnings(
-  cluster::clusGap(sightings_mtx, 
+  clusGap(sightings_mtx, 
                    FUN = cluster_fn,
@@
     k = seq_along(logW),
-    is_optimal = k == cluster::maxSE(gap, SE.sim, method = "Tibs2001SEmax")
+    is_optimum = case_when(
+      k == maxSE(gap, SE.sim, method = "Tibs2001SEmax") ~ "tibsSEmax",
+      k == maxSE(gap, SE.sim, method = "firstSEmax") ~ "firstSEmax",
+      TRUE ~ "none"
+  ))
+kfirstmax <- with(
+  sightings_gap_df,
+  maxSE(gap, SE.sim, method = "firstmax")
 )
+ktibs <- with(sightings_gap_df, k[is_optimum == "tibsSEmax"])
+firstSEmax_thr <- with(sightings_gap_df, gap[kfirstmax] - SE.sim[kfirstmax])
+tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
 ggplot(sightings_gap_df, aes(x = k, y = gap)) +
@@
   geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
-                      color = is_optimal)) +
-  scale_color_manual(values = c(`TRUE` = "red",
-                                `FALSE` = "black")) +
+                      color = is_optimum)) +
+  geom_segment(x = 3, xend = 4, 
+               y = firstSEmax_thr, yend = firstSEmax_thr,
+               color = "red",
+               linetype = "dashed") +
+  geom_segment(x = ktibs, xend = ktibs + 1, 
+               y = tibs_thr, yend = tibs_thr,
+               color = "blue",
+               linetype = "dashed") +
+  scale_color_manual(values = c(tibsSEmax = "blue",
+                                firstSEmax = "red",
+                                none = "black")) +
   theme_classic() +
@@
 
-## 2-cluster analysis
+## 3-cluster analysis
 
-Repeat the indicator species analysis and maps with two clusters.
+Repeat the indicator species analysis and maps with three clusters.
 
@@
 
-# Cut predators into 2 clusters
-clust <- cutree(sightings_clust, 2)
+# Cut predators into 3 clusters
+clust <- cutree(sightings_clust, 3)
 g <- split(names(clust), clust)
@@
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  scale_color_brewer(palette = "Dark2") +
   coord_ant(map_lim) +
@@
 
+How many NMDS axes should we use? Scree plot
+
 ```{r}
-#| label: nmds
+#| label: nmds-scree
+
+nmds_stress <- tibble(
+  k = 1:6,
+  stress = map_dbl(k, function(k) {
+    metaMDS(
+      sightings_mtx, 
+      distance = "bray",
+      k = k,
+      trymax = 100,
+      trace = FALSE
+    )$stress
+  })
+)
+
+ggplot(nmds_stress, aes(k, stress)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = c(0.2, 0.1, 0.05), linetype = "dotted") +
+  scale_x_continuous(breaks = nmds_stress$k) +
+  expand_limits(y = 0) +
+  theme_classic()
+
+```
+
+3 axes for NMDS looks acceptable. How does ordination distance compare to dissimilarity (Shepard diagram)?
+
+```{r}
+#| label: nmds-shepard
 
@@
   k = 3,
-  trymax = 100
+  trymax = 100,
+  trace = FALSE
 )
 
-stressplot(sightings_nmds)
+nmds_isoreg <- isoreg(sightings_nmds$diss, sightings_nmds$dist)
+
+nmds_shepard <- tibble(
+  diss = sightings_nmds$diss,
+  dist = sightings_nmds$dist,
+  mono = nmds_isoreg$yf
+)
+
+nmds_spear <- cor(nmds_shepard$diss,
+                  nmds_shepard$dist,
+                  method = "spearman")
+nmds_label <- str_glue(
+  "Non-metric fit, R^2 = {round(1 - sightings_nmds$stress^2, 3)}\n",
+  "Linear fit, R^2 = {round(nmds_spear^2, 3)}\n",
+  "k = {sightings_nmds$ndim}\n",
+  "Stress = {round(sightings_nmds$stress, 3)}"
+)
+
+ggplot(nmds_shepard, aes(diss, dist)) +
+  geom_point(size = 0.75, alpha = 0.1) +
+  geom_line(aes(y = mono), color = "red") +
+  annotate("text", x = 0.05, y = 2.9, label = nmds_label,
+           hjust = 0, vjust = 1) +
+  expand_limits(x = 0, y = 0) +
+  labs(x = "Observed dissimilarity",
+       y = "Ordination distance") +
+  theme_classic()
 
 ```
+
+NMDS plots, color-coded by predator cluster.
 
@@
   geom_point(aes(color = pred_clust)) +
+  scale_color_brewer(palette = "Dark2") +
   theme_classic()
@@
   geom_point(aes(color = pred_clust)) +
+  scale_color_brewer(palette = "Dark2") +
   theme_classic()
 ```
+
+## \[Insert Frozen or Vanilla Ice pun here\]
+
+Get the modal ice type (multi-year, first-year, thin, or open) for each station from the along-track observations.
+Seems weird that no stations shook out as multi-year ice, right?
+This figure shows how frequent the modal ice type was at each station.
+The two grey lines represent 1:1 and 2:1.
+Stations falling on the upper line had *uniform* ice category observations i.e., every interval within 15 km had the same ice category.
+Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station.
+Most points on the figure include multiple stations, indicated by the color.
+For example, in the THIN ice category (bottom), 8 stations (red point) had 4 intervals with ice category observations where all 4 were THIN (x,y = 4,4).
+
+```{r}
+#| label: ice
+
+# Most frequent value i.e., mode
+# Note: returns a character
+mfv <- function(x) {
+  x_tbl <- table(x)
+  names(x_tbl)[which.max(x_tbl)]
+}
+underway_ice <- predators_stations %>% 
+  left_join(transmute(predators,
+                      year = as.numeric(substr(cruise, 5, 8)),
+                      interval,
+                      ice_type_desc),
+            by = c("year", "interval"),
+            multiple = "all") %>% 
+  filter(ice_type_desc != "NULL") %>% 
+  left_join(read_csv(here::here("data", "ICECAT.csv"),
+                     show_col_types = FALSE),
+            by = "ice_type_desc") %>% 
+  # Ice observations were unique to intervals, but intervals are
+  # repeated if there were multiple predator sightings
+  group_by(amlr.station,
+           interval) %>% 
+  summarize(interval_ice_cat = first(ice_cat),
+            .groups = "drop_last") %>% 
+  summarize(station_ice_cat = mfv(interval_ice_cat),
+            n_obs = n(),
+            n_mode = sum(interval_ice_cat == station_ice_cat))
+
+underway_ice %>% 
+  group_by(n_obs, n_mode, station_ice_cat) %>% 
+  summarize(Stations = n(), .groups = "drop") %>% 
+  ggplot(aes(n_obs, n_mode, color = Stations)) +
+  geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
+  geom_point() +
+  labs(x = "Count of intervals with ice category observation",
+       y = "Count of intervals with most frequent ice category") +
+  scale_color_fermenter(palette = "RdYlBu",
+                        breaks = c(1, 2, 3, 5, 8)) +
+  coord_fixed() +
+  facet_wrap(~ station_ice_cat, ncol = 2) +
+  theme_classic()
+
+```

### 2023-03-23 reports/03_combined.qmd
Move ice summary to combined report

@@
 
-tar_load(c("predators_agg", "zoop", "ant_sf"), 
+tar_load(c("predators_agg", "predators_stations", "zoop", "ant_sf"), 
          store = here::here("_targets"))
@@
 ```
+
+## Sightings and stations
 
@@
 ```
+
+## Aggregated ice observations
+
+Each station was associated with the modal ice type (multi-year, first-year, thin, or open) from the underway observations, which were only associated with predator sightings.
+**Seems weird that no stations shook out as multi-year ice, right?** This figure shows how frequent the modal ice type was at each station.
+The two grey lines represent 1:1 and 2:1.
+Stations falling on the upper line had *uniform* ice category observations i.e., every interval within 15 km had the same ice category.
+Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station.
+Most points on the figure include multiple stations, indicated by the color.
+For example, in the THIN ice category (bottom), 8 stations (red point) had 4 intervals with ice category observations where all 4 were THIN (x,y = 4,4).
+
+```{r}
+#| label: underway-ice
+
+predators_stations %>% 
+  mutate(n_obs = ice_intervals,
+         n_mode = round(ice_cat_freq * n_obs, digits = 0)) %>% 
+  group_by(n_obs, n_mode, ice_cat) %>% 
+  summarize(Stations = n(), .groups = "drop") %>% 
+  ggplot(aes(n_obs, n_mode, color = Stations)) +
+  geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
+  geom_point() +
+  labs(x = "Count of intervals with ice category observation",
+       y = "Count of intervals with most frequent ice category") +
+  scale_color_fermenter(palette = "RdYlBu",
+                        breaks = c(1, 2, 3, 5, 8)) +
+  coord_fixed() +
+  facet_wrap(~ ice_cat, ncol = 2) +
+  theme_classic()
+```

### 2023-03-23 reports/04_pred_clus.qmd
Move ice summary to combined report

@@
 
-tar_load(c("predators", "predators_stations", 
-           "predators_abundant", "sightings_clust", "sightings_mtx", 
-           "sightings_dist", "zoop_sf"), 
+tar_load(c("predators_stations", "predators_abundant", 
+           "sightings_clust", "sightings_mtx", "sightings_dist", 
+           "zoop_sf"), 
          store = here::here("_targets"))
@@
 ```
-
-## \[Insert Frozen or Vanilla Ice pun here\]
-
-Get the modal ice type (multi-year, first-year, thin, or open) for each station from the along-track observations.
-Seems weird that no stations shook out as multi-year ice, right?
-This figure shows how frequent the modal ice type was at each station.
-The two grey lines represent 1:1 and 2:1.
-Stations falling on the upper line had *uniform* ice category observations i.e., every interval within 15 km had the same ice category.
-Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station.
-Most points on the figure include multiple stations, indicated by the color.
-For example, in the THIN ice category (bottom), 8 stations (red point) had 4 intervals with ice category observations where all 4 were THIN (x,y = 4,4).
-
-```{r}
-#| label: ice
-
-# Most frequent value i.e., mode
-# Note: returns a character
-mfv <- function(x) {
-  x_tbl <- table(x)
-  names(x_tbl)[which.max(x_tbl)]
-}
-underway_ice <- predators_stations %>% 
-  left_join(transmute(predators,
-                      year = as.numeric(substr(cruise, 5, 8)),
-                      interval,
-                      ice_type_desc),
-            by = c("year", "interval"),
-            multiple = "all") %>% 
-  filter(ice_type_desc != "NULL") %>% 
-  left_join(read_csv(here::here("data", "ICECAT.csv"),
-                     show_col_types = FALSE),
-            by = "ice_type_desc") %>% 
-  # Ice observations were unique to intervals, but intervals are
-  # repeated if there were multiple predator sightings
-  group_by(amlr.station,
-           interval) %>% 
-  summarize(interval_ice_cat = first(ice_cat),
-            .groups = "drop_last") %>% 
-  summarize(station_ice_cat = mfv(interval_ice_cat),
-            n_obs = n(),
-            n_mode = sum(interval_ice_cat == station_ice_cat))
-
-underway_ice %>% 
-  group_by(n_obs, n_mode, station_ice_cat) %>% 
-  summarize(Stations = n(), .groups = "drop") %>% 
-  ggplot(aes(n_obs, n_mode, color = Stations)) +
-  geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
-  geom_point() +
-  labs(x = "Count of intervals with ice category observation",
-       y = "Count of intervals with most frequent ice category") +
-  scale_color_fermenter(palette = "RdYlBu",
-                        breaks = c(1, 2, 3, 5, 8)) +
-  coord_fixed() +
-  facet_wrap(~ station_ice_cat, ncol = 2) +
-  theme_classic()
-
-```

### 2023-03-26 reports/01_predators.qmd
Clean up ice and reports

@@
                    res = 2.5e4)
-map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
+map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
                        crs = "EPSG:4326") %>% 
@@
 
-map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
+map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
                        crs = "EPSG:4326") %>% 
@@
 
-map_lim <- sf::st_bbox(soi_sf) %>% 
+map_lim <- st_bbox(soi_sf) %>% 
   project_bbox() %>% 

### 2023-03-26 reports/02_stations.qmd
Clean up ice and reports

@@
 
-zoop_buffers_10k <- sf::st_buffer(zoop_sf, 5e3)
-zoop_buffers_15k <- sf::st_buffer(zoop_sf, 10e3)
-zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)
+zoop_buffers_10k <- st_buffer(zoop_sf, 5e3)
+zoop_buffers_15k <- st_buffer(zoop_sf, 10e3)
+zoop_buffers_20k <- st_buffer(zoop_sf, 20e3)
 
-map_lim <- sf::st_bbox(zoop_sf) %>% 
+map_lim <- st_bbox(zoop_sf) %>% 
   project_bbox() %>% 

### 2023-03-26 reports/03_combined.qmd
Clean up ice and reports

@@
                         coords = c("dec.longitude", "dec.latitude")) %>% 
-  sf::st_transform(ant_proj()) %>% 
+  st_transform(ant_proj()) %>% 
   rename(year = Year)
-zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)
+zoop_buffers_20k <- st_buffer(zoop_sf, 20e3)
 
-map_lim <- sf::st_bbox(zoop_sf) %>% 
+map_lim <- st_bbox(zoop_sf) %>% 
   project_bbox() %>% 
@@
 ```
+
+```{r}
+#| label: ice-map
+#| column: screen
+#| out-width: 100%
+
+ant_basemap(map_lim) +
+  geom_sf(aes(color = ice_cat), 
+          predators_stations,
+          size = 0.5) +
+  facet_wrap(~ year) +
+  theme(legend.position = "bottom")
+
+```

### 2023-03-26 reports/04_pred_clus.qmd
Clean up ice and reports

@@
 pred_frac <- predators_stations %>%
+  as_tibble() %>% 
   group_by(amlr.station, species) %>% 
@@
 
-species_wide <- pivot_wider(predators_abundant,
-                            names_from = species,
+species_wide <- predators_abundant %>% 
+  pivot_wider(names_from = species,
               values_from = count,
@@
 
-map_lim <- sf::st_bbox(pred_clust_sf) %>% 
+map_lim <- st_bbox(pred_clust_sf) %>% 
   project_bbox() %>% 
@@
 ```{r}
+#| label: optimal-clusters
+
 cluster_fn <- function(x, k) {

### 2023-03-26 R/cluster.R

@@ -25,16 +25,16 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     # Note difference in column title Year vs year
     stations_group <- stations[stations$Year == sightings_key$year, ]
     # Find nearest station within year
-    nearest_station <- sf::st_join(sightings_group,
+    nearest_station <- st_join(sightings_group,
                                    select(stations_group, amlr.station),
-                                   join = sf::st_nearest_feature)
+                                   join = st_nearest_feature)
     # Get station coordinates
     station_coords <- stations_group %>%
       slice(map_int(nearest_station$amlr.station,
                     ~ which(stations_group$amlr.station == .x)))
     # Calculate distance from predator sighting to nearest station
     nearest_station %>%
-      mutate(dist_to_station = sf::st_distance(nearest_station,
+      mutate(dist_to_station = st_distance(nearest_station,
                                                station_coords,
                                                by_element = TRUE))
   }
@@ -44,12 +44,14 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     group_by(year) %>%
     group_modify(assign_group) %>%
     ungroup() %>%
-    filter(dist_to_station <= units::as_units(max_dist_km, "km"))
+    filter(dist_to_station <= units::as_units(max_dist_km, "km")) %>%
+    st_as_sf()
 }
 
 filter_sightings <- function(sightings, stations, station_thr) {
   total_stations <- n_distinct(stations$amlr.station)
   sightings_agg <- sightings %>%
+    as_tibble() %>%
     group_by(amlr.station, species) %>%
     summarize(count = sum(count), .groups = "drop")
   retain <- sightings_agg %>%
@@ -63,6 +65,7 @@ filter_sightings <- function(sightings, stations, station_thr) {
 
 sightings_to_matrix <- function(sightings) {
   sightings_wide <- sightings %>%
+    as_tibble() %>%
     pivot_wider(id_cols = amlr.station,
                 names_from = species,
                 values_from = count,

### 2023-03-26 R/ice.R

@@ -12,7 +12,7 @@ mfv <- function(x) {
 aggregate_ice <- function(stations, underway, ice_cat) {
   # Aggregates underway ice observations by nearest AMLR station
   # Uses mode as aggregation function i.e. mfv()
-  underway_ice <- stations %>%
+  underway_ice <- as_tibble(stations) %>%
     left_join(transmute(underway,
                         year = as.numeric(substr(cruise, 5, 8)),
                         interval,
@@ -20,9 +20,7 @@ aggregate_ice <- function(stations, underway, ice_cat) {
               by = c("year", "interval"),
               multiple = "all") %>%
     filter(ice_type_desc != "NULL") %>%
-    left_join(read_csv(here::here("data", "ICECAT.csv"),
-                       show_col_types = FALSE),
-              by = "ice_type_desc") %>%
+    left_join(ice_cat, by = "ice_type_desc") %>%
     # Ice observations were unique to intervals, but intervals are
     # repeated if there were multiple predator sightings
     group_by(amlr.station,

### 2023-03-26 R/sp.R

@@ -5,7 +5,7 @@ ant_proj <- function() {
   glue::glue(
     "+proj=stere +lat_0=-90 +lat_ts={lat_ts} +lon_0={lon_0} +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
   ) %>%
-  sf::st_crs()
+  st_crs()
 }
 
 ant_lims <- function() {
@@ -27,11 +27,11 @@ create_ant_sf <- function(ne_dir, limits = ant_lims()) {
                                  category = "physical",
                                  destdir = ne_dir,
                                  returnclass = "sf")
-  sf::st_crop(land, sf::st_bbox(limits))
+  st_crop(land, st_bbox(limits))
 }
 
 ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
-  lims_prj <- sf::st_bbox(limits, crs = "EPSG:4326") %>%
+  lims_prj <- st_bbox(limits, crs = "EPSG:4326") %>%
     project_bbox()
   raster::raster(xmn = lims_prj["xmin"],
                  xmx = lims_prj["xmax"],
@@ -57,7 +57,7 @@ rasterize_counts <- function(counts_sf,
     group_modify(
       function(rows, key) {
         result <- raster::rasterize(
-          sf::as_Spatial(sf::st_transform(rows, ant_proj())),
+          as_Spatial(st_transform(rows, ant_proj())),
           ant_raster_template(limits, res),
           field = count_col,
           fun = "sum",
@@ -74,11 +74,11 @@ rasterize_counts <- function(counts_sf,
     dims = 1:4,
     xy = 3:4
   ) %>%
-    sf::st_set_crs(ant_proj())
+    st_set_crs(ant_proj())
 }
 
 latlon_to_sf <- function(df, coords = c("x", "y")) {
-  sf::st_as_sf(df,
+  st_as_sf(df,
                coords = coords,
                crs = sp::CRS("+proj=longlat +datum=WGS84"))
 }
@@ -92,7 +92,7 @@ ant_basemap <- function(map_lim = NULL) {
 
 coord_ant <- function(map_lim = NULL) {
   if (is.null(map_lim)) {
-    map_lim <- sf::st_bbox(ant_lims(), crs = "EPSG:4326") %>%
+    map_lim <- st_bbox(ant_lims(), crs = "EPSG:4326") %>%
[…]

### 2023-05-18 reports/01_predators.qmd
nmds plots finished

@@
 
-The five most abundant species.
+The five most abundant species (in ind nmi^-1).
 
@@
   group_by(species) %>% 
-  summarize(total_count = sum(count)) %>% 
+  summarize(total_count = sum(count_norm)) %>% 
   arrange(desc(total_count)) %>%
@@
   group_by(year, species) %>% 
-  summarize(total = sum(count), .groups = "drop") %>% 
+  summarize(total = sum(count_norm), .groups = "drop") %>% 
   mutate(species = fct_reorder(species, total, .fun = sum),
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = log(count + 1)), data = topfive_stars) +
+  stars::geom_stars(aes(fill = log(count_norm + 1)), 
+                    data = topfive_stars) +
   scale_fill_distiller(palette = "PuRd",
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = log(count + 1)), data = soi_stars) +
+  stars::geom_stars(aes(fill = log(count_norm + 1)), data = soi_stars) +
   scale_fill_distiller(palette = "PuRd",

### 2023-05-18 reports/03_combined.qmd
nmds plots finished

@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = count), data = topfive_stars) +
+  stars::geom_stars(aes(fill = count_norm), data = topfive_stars) +
   geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
@@
 ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = count), data = soi_stars) +
+  stars::geom_stars(aes(fill = count_norm), data = soi_stars) +
   geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
@@
   scale_color_fermenter(palette = "RdYlBu",
-                        breaks = c(1, 2, 3, 5, 8)) +
+                        breaks = c(10, 20, 100, 200)) +
   coord_fixed() +
@@
           size = 0.5) +
+  scale_color_brewer("Ice category", palette = "YlGnBu") +
+  guides(color = guide_legend(override.aes = list(size = 5))) +
   facet_wrap(~ year) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "left")
 

### 2023-05-18 reports/04_pred_clus.qmd
nmds plots finished

@@
   group_by(amlr.station, species) %>% 
-  summarize(count = sum(count), .groups = "drop") %>%
+  summarize(count_norm = sum(count_norm), .groups = "drop") %>%
   group_by(species) %>%
@@
 
+## Optimal number of clusters
+
+We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters ($k$).
+The gap statistic ($f$) is a goodness of clustering measure.
+The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
+According to this rule, the predators are best described by two clusters (indicated in blue).
+However, $f(2)$ is only *slightly* greater than the threshold $f(3) - s(3)$, and $f$ is determined stochastically, so depending on the iteration the rule may choose $k=3$.
+Based on our prior knowledge of three predator guilds (pagophilic, open water, and intermediate), we'll present the results for $k=3$ here and re-run the analysis for $k=2$ in the supplemental.
+
+```{r}
+#| label: optimal-clusters
+
+cluster_fn <- function(x, k) {
+  bray_dist <- function(x) {
+    vegdist(x, method = "bray")
+  }
+  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
+}
+set.seed(139)
+sightings_gap <- suppressWarnings(
+  clusGap(sightings_mtx, 
+                   FUN = cluster_fn,
+                   K.max = 10, 
+                   B = 500,
+                   verbose = FALSE)
+)
+sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
+  mutate(
+    k = seq_along(logW),
+    is_optimum = k == maxSE(gap, SE.sim, method = "Tibs2001SEmax")
+  )
+ktibs <- filter(sightings_gap_df, is_optimum) %>% pull(k)
+tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
+ggplot(sightings_gap_df, aes(x = k, y = gap)) +
+  geom_line() +
+  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
+                      color = is_optimum)) +
+  geom_segment(x = ktibs, xend = ktibs + 1, 
+               y = tibs_thr, yend = tibs_thr,
+               color = "blue",
+               linetype = "dashed") +
+  scale_color_manual(values = c(`TRUE` = "blue",
+                                `FALSE` = "black")) +
+  theme_classic() +
+  theme(legend.position = "none")
+```
+
 Hierarchical clustering of stations by predator community.
-Chose 5 clusters a priori to match Dietrich et al.
-(2021).
 Note: `r nrow(zoop_sf)` stations were included in the zooplankton clustering, but only `r length(sightings_clust$labels)` stations were used for predator clustering.
@@
 # Cut predators into five clusters
-clust <- cutree(sightings_clust, 5)
+clust <- cutree(sightings_clust, 3)
 g <- split(names(clust), clust)
@@
 Upon visual inspection of the mosaic plot, I don't perceive any clear correlations between the zooplankton and predator cluster assignments.
-I tested the strength of the relationship between the two cluster assignments through simulation of the Rand index (RI).
-RI is a measure of similarity between two clusterings ranging from 0 (no agreement) to 1 (total agreement).
-The histogram below shows the distribution of RI values when the cluster assignments were randomly shuffled 100,000 times.
-The vertical red line is the observed RI value.
-
-```{r}
-#| label: randindex
-
-set.seed(1311)
-
-expected_ri <- replicate(
-  100000,
-  aricode::RI(
-    sample(both_clust$zoop_clust, nrow(both_clust)),
-    sample(both_clust$pred_clust, nrow(both_clust))
-  )
-)
-
-observed_ri <- aricode::RI(both_clust$zoop_clust, both_clust$pred_clust)
-
-ri_pval <- sum(expected_ri > observed_ri) / 100000
-
-ggplot(tibble(`Rand index` = expected_ri), aes(`Rand index`)) +
-  geom_histogram(binwidth = 0.001) +
-  geom_vline(xintercept = observed_ri, color = "red") +
-  theme_classic()
-
-```
-
-Based on this simulation, the correlation between the zooplankton and predator assignments *is* significant (p = `r as.character(ri_pval)`).
-But I haven't figured out how to assess effect size yet, which I imagine is small.
+We're interpreting this as evidence predator communities are structured by the physical environment, specifically sea ice, rather than the prey communities.
 
@@
   pivot_wider(names_from = species,
-              values_from = count,
+              values_from = count_norm,
               values_fill = 0) %>% 
@@
 
-According to the 25% cutoff, clusters 1 and 4 lack indicator species.
-In agreement, every species' highest indicator value falls in clusters 2, 3, or 5.
-This suggests to me 5 clusters is a poor choice for the predators and we should use fewer.
-I'll calculate the gap statistic for the hierarchical cluster analysis to quantitatively choose a data-informed number of clusters and repeat this entire analysis.
-For now, continuing with 5 clusters.
+### Cluster descriptions
 
-I don't know enough about the Antarctic Peninsula megafauna to say whether the indicator species cluster together logically.
-The fulmarine/*Daption* petrels are split across Clusters 2 and 3.
-Cluster 2 lumps Cape Petrel and Antarctic Fulmar with the prion-like Blue Petrel.
-Cluster 3 combines the fulmarine Antarctic and Snow Petrels with the Kerguelen Petrel and fur seals.
-Cluster 5 seems to have the most ice-associated species (phocids and Adelie penguins).
+* Cluster 1 = marginal ice zone (SNPT)
+* Cluster 2 = open ocean (ANFU, CAPT, BLPT)
+* Cluster 3 = pagophilic (FUSE, ADPN, CRSE)
 
@@
 
-What geospatial patterns do we see in the predator clusters?
-Cluster 5 is mostly found in the Bransfield Strait, which makes sense for the Adelie Penguin-crabeater seal-leopard seal complex.
-Cluster 2 (Antarctic Fulmar, Cape Petrel, and Blue Petrel) tends to show up oceanward of the South Shetland Islands.
-Cluster 3 (fur seal, snow petrel, Kerguelen petrel, and Antarctic Petrel) seems the most variable; found both inside and outside the Strait depending on the year.
+What geospatial patterns do we see in the predator communities?
+The marginal ice zone community was mostly observed just north of the South Shetland islands, occasionally farther north beyond Elephant Island.
+The open ocean community occupied the northeastern portion of the study area in 2012, 2014, and 2015.
+In 2013 it contracted westward and in 2016 it was entirely absent.
+The pagophilic predator community dominated the Bransfield Strait itself, expanding north towards Elephant Island in 2013 and 2016.
 
@@
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-  coord_ant(map_lim) +
-  theme(legend.position = "bottom")
-```
-
-## Optimal number of clusters
-
-We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters ($k$).
-The gap statistic ($f$) is a goodness of clustering measure.
-The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
-According to this rule, the predators are best described by two clusters (indicated in blue).
-However, $f(2)$ is only slightly greater than $f(3)-s(3)$ and, because $s$ is estimated by bootstrapping, sometimes re-running the gap statistic estimate yields an optimum at $k=3$ instead.
-An alternative rule suggests choosing the smallest $k$ such $f(k) \geq f_{max} - s_{max}$, where $f_{max}$ and $s_{max}$ are $f$ and $s$ at the first local maximum, respectively.
-In the figure below, $f_{max} - s_{max}$ is represented by the dashed red line, and the smallest $k$ greater than that threshold is $k=3$ (red).
-
-```{r}
-#| label: optimal-clusters
-
-cluster_fn <- function(x, k) {
-  bray_dist <- function(x) {
-    vegdist(x, method = "bray")
-  }
-  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
-}
-set.seed(931)
-sightings_gap <- suppressWarnings(
-  clusGap(sightings_mtx, 
-                   FUN = cluster_fn,
-                   K.max = 10, 
-                   B = 500,
-                   verbose = FALSE)
-)
-sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
-  mutate(
-    k = seq_along(logW),
-    is_optimum = case_when(
-      k == maxSE(gap, SE.sim, method = "Tibs2001SEmax") ~ "tibsSEmax",
-      k == maxSE(gap, SE.sim, method = "firstSEmax") ~ "firstSEmax",
-      TRUE ~ "none"
-  ))
-kfirstmax <- with(
-  sightings_gap_df,
-  maxSE(gap, SE.sim, method = "firstmax")
-)
-ktibs <- with(sightings_gap_df, k[is_optimum == "tibsSEmax"])
-firstSEmax_thr <- with(sightings_gap_df, gap[kfirstmax] - SE.sim[kfirstmax])
-tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
-ggplot(sightings_gap_df, aes(x = k, y = gap)) +
-  geom_line() +
-  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
-                      color = is_optimum)) +
-  geom_segment(x = 3, xend = 4, 
-               y = firstSEmax_thr, yend = firstSEmax_thr,
-               color = "red",
-               linetype = "dashed") +
-  geom_segment(x = ktibs, xend = ktibs + 1, 
-               y = tibs_thr, yend = tibs_thr,
-               color = "blue",
-               linetype = "dashed") +
-  scale_color_manual(values = c(tibsSEmax = "blue",
-                                firstSEmax = "red",
-                                none = "black")) +
-  theme_classic() +
-  theme(legend.position = "none")
-```
-
-## 3-cluster analysis
-
-Repeat the indicator species analysis and maps with three clusters.
-
-```{r}
-#| label: predator_dendro2
-
-# Cut predators into 3 clusters
-clust <- cutree(sightings_clust, 3)
-g <- split(names(clust), clust)
-p <- ggtree(sightings_clust, hang = -1)
-pred_clust_mrca <- sapply(g, function(n) MRCA(p, n))
-
-p %>% 
-  groupClade(pred_clust_mrca, group_name = "Predator cluster") + 
-  aes(color = `Predator cluster`) +
-  layout_dendrogram() +
-  theme_dendrogram()
-```
-
-```{r}
-#| label: indicator_species2
-
-species_wide <- pivot_wider(predators_abundant,
-                            names_from = species,
-                            values_from = count,
-                            values_fill = 0) %>% 
-  column_to_rownames("amlr.station")
-cluster_indval <- labdsv::indval(species_wide, clust)
-indval_mtx <- cluster_indval$indval
-indval_fmt <- apply(indval_mtx, 2, function(x) {
-  ord <- order(x, decreasing = TRUE)
-  result <- ifelse(x >= 0.25, 
-         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
-         NA) %>% 
-    `[`(ord) %>% 
-    na.omit() %>% 
-    paste(collapse = "<br>")
-  ifelse(result == "", "None", result)
-})
-as_tibble(t(indval_fmt)) %>% 
-  knitr::kable(format = "html", escape = FALSE) %>% 
-  kableExtra::kable_styling()
-```
-
-```{r}
-#| label: cluster_maps2
-
-pred_clust_sf <- tibble(amlr.station = names(clust),
-                        pred_clust = factor(clust)) %>% 
-  left_join(select(zoop_sf,
-                   amlr.station,
-                   geometry,
-                   Year),
-            by = "amlr.station") %>% 
-  st_as_sf() %>% 
-  st_transform(ant_proj())
-
-map_lim <- st_bbox(pred_clust_sf) %>% 
-  project_bbox() %>% 
-  expand_bbox(factor = 1.2)
-
-ant_basemap(map_lim) +
-  geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
-  facet_wrap(~ Year) +
-  scale_x_continuous(breaks = c(-60, -55)) +
-  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-  scale_color_brewer(palette = "Dark2") +
   coord_ant(map_lim) +
@@
 ```
+
+## Environmental loadings
+
+I compared the drivers of community structure two ways: physics (hydrology and ice coverage) and zooplankton (community assignments). The result is kind of a mess! Here's what makes jumps out at me:
+
+* Predator cluster 1 (marginal ice) is associated with higher phaeopigment. An ice-edge effect?
+* Predator cluster 2 (open ocean) is associated with higher temperatures and chl a. Associated with macrozooplankton communities 2a and 2b (small euphausiids, amphipods, etc).
+* Predator cluster 3 (pagophilic) is associated with higher salinity and deeper mixed layer. Associated with macrozooplankton communities 3a and 3b (big euphausiids, Metridia, etc).
+* No clear relationship between predator community and ice categories.
+
+```{r}
+#| label: env-loadings
+#| eval: false
+
+station_ice <- predators_stations %>% 
+  group_by(amlr.station) %>% 
+  summarize(ice_cat = ice_cat[1]) %>% 
+  as_tibble() %>% 
+  select(-geometry)
+
+# Evironmental variables
+station_env <- zoop_sf %>% 
+  select(amlr.station, zuml_m, avg.temp, avg.salinity, Integ.chla.100m, 
+         Integ.phae.100m, TOD_2levels_civil, Year, 
+         zoop = `Winter Cluster factor`) %>% 
+  as_tibble() %>% 
+  left_join(station_ice, by = "amlr.station") %>% 
+  mutate(zoop = fct_collapse(zoop,
+                             `2` = c("2a", "2b"),
+                             `3` = c("3a", "3b")),
+         Year = factor(Year)) %>% 
+  slice(map_int(rownames(sightings_mtx),
+                ~ which(.x == amlr.station))) %>% 
+  select(-amlr.station, -geometry)
+
+# Fit multiple regression model
+pred_envfit <- envfit(sightings_nmds, 
+                      station_env, 
+                      permutations = 999, 
+                      choices = 1:3,
+                      na.rm = TRUE)
+
+# Extract envfit vectors and factors
+envfit_vecs <- scores(pred_envfit, 
+                      display = "vectors",
+                      choices = 1:3) %>% 
+  as_tibble(rownames = "envvar") %>% 
+  mutate(across(starts_with("NMDS"),
+                ~ .x * ordiArrowMul(pred_envfit)))
+envfit_facs <- scores(pred_envfit, 
+                      display = "factors",
+                      choices = 1:3) %>% 
+  as_tibble(rownames = "envvar") %>% 
+  mutate(across(starts_with("NMDS"),
+                ~ .x * ordiArrowMul(pred_envfit)))
+
+# Hydrology plot
+ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
+               data = envfit_vecs,
+               linewidth = 1,
+               color = "grey50",
+               alpha = 0.75) +
+  geom_text(aes(x = NMDS1, y = NMDS2, label = envvar),
+            data = envfit_vecs) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+# Ice plot
+ice_facs <- filter(envfit_facs, str_detect(envvar, "ice_cat"))
+ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(aes(x = NMDS1, y = NMDS2),
+             data = ice_facs,
+             size = 2) +
+  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
+                           data = ice_facs) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_classic()
+
+# Macrozooplankton plot
+zoop_facs <- filter(envfit_facs, str_detect(envvar, "zoop"))
+ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(aes(x = NMDS1, y = NMDS2),
+             data = zoop_facs,
+             size = 2) +
+  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
+                           data = zoop_facs) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_classic()
+```

### 2023-05-18 reports/bibliography.qmd
nmds plots finished

---
title: "bibliography"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-05-18 R/cluster.R

@@ -53,7 +53,7 @@ filter_sightings <- function(sightings, stations, station_thr) {
   sightings_agg <- sightings %>%
     as_tibble() %>%
     group_by(amlr.station, species) %>%
-    summarize(count = sum(count), .groups = "drop")
+    summarize(count_norm = sum(count_norm), .groups = "drop")
   retain <- sightings_agg %>%
     group_by(species) %>%
     summarize(stations_present = n(),
@@ -68,7 +68,7 @@ sightings_to_matrix <- function(sightings) {
     as_tibble() %>%
     pivot_wider(id_cols = amlr.station,
                 names_from = species,
-                values_from = count,
+                values_from = count_norm,
                 values_fill = 0) %>%
     mutate(across(-amlr.station, ~ log(.x + 1)))
   sightings_mtx <- as.matrix(select(sightings_wide, -amlr.station))

### 2023-05-18 R/predators.R

@@ -2,6 +2,7 @@ aggregate_predators <- function(predators) {
   predators %>%
     filter(species != "NULL") %>%
     mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
-    group_by(year, interval, species, lon_mean, lat_mean) %>%
-    summarize(count = sum(count_species), .groups = "drop")
+    group_by(year, interval, species, lon_mean, lat_mean, nmi) %>%
+    summarize(count = sum(count_species), .groups = "drop") %>%
+    mutate(count_norm = count / nmi)
 }

### 2023-05-18 R/sp.R

@@ -43,7 +43,7 @@ ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
 
 # By species and year
 rasterize_counts <- function(counts_sf,
-                             count_col = "count",
+                             count_col = "count_norm",
                              species_col = "species",
                              year_col = "year",
                              layer_name = count_col,

### 2023-05-18 _targets.R

@@ -49,7 +49,7 @@ list(
                mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
                group_by(cruise, year, interval, lon_mean, lat_mean) %>%
                summarize(nmi = sum(nmi), .groups = "drop") %>%
-               mutate(species = "survey", count = 1)),
+               mutate(species = "survey", count_norm = 1)),
   tar_target(effort_sf,
              latlon_to_sf(effort, coords = c("lon_mean", "lat_mean"))),
   # Zooplankton

### 2023-05-19 reports/04_pred_clus.qmd
Make cluster maps' color scheme match

@@
 #| label: env-loadings
-#| eval: false
 

### 2023-05-31 reports/04_pred_clus.qmd
Updated predator clusters from raw abundance to ind nm^-1

@@
 library(ggtree)
+library(patchwork)
 library(sf)
@@
   group_by(amlr.station, species) %>% 
-  summarize(count_norm = sum(count_norm), .groups = "drop") %>%
+  summarize(count_station = sum(count),
+            .groups = "drop") %>% 
   group_by(species) %>%
-  summarize(stations_present = n(),
-            station_frac = stations_present / total_stations,
-            .groups = "drop") %>%
+  summarize(total_count = sum(count_station),
+            stations_present = n(),
+            station_frac = stations_present / total_stations) %>%
   arrange(desc(station_frac))
@@
 pred_frac %>% 
-  set_names(c("Species", "Stations", "Fraction")) %>% 
+  set_names(c("Species", "Ind", "Stations", "Fraction")) %>% 
   mutate(Fraction = frac_fn(Fraction)) %>% 
@@
 
-* Cluster 1 = marginal ice zone (SNPT)
-* Cluster 2 = open ocean (ANFU, CAPT, BLPT)
-* Cluster 3 = pagophilic (FUSE, ADPN, CRSE)
+* Cluster 1 = open ocean? (ANFU)
+* Cluster 2 = marginal ice zone (SNPT, ANPT, FUSE)
+* Cluster 3 = pagophilic (ADPN, CRSE, LESE)
 
@@
 
-pred_clust_sf <- tibble(amlr.station = names(clust),
-                        pred_clust = factor(clust)) %>% 
+pred_clust_sf <- tibble(
+  amlr.station = names(clust),
+  pred_clust = factor(clust,
+                      labels = c("Open", "Marginal", "Pagophilic"))
+) %>% 
   left_join(select(zoop_sf,
@@
   rename_with(~ paste0("N", .x), starts_with("MDS")) %>% 
-  right_join(tibble(amlr.station = names(clust),
-                    pred_clust = factor(clust)),
+  right_join(pred_clust_sf %>% 
+               as_tibble() %>% 
+               select(amlr.station, pred_clust),
              by = "amlr.station")
@@
 
-* Predator cluster 1 (marginal ice) is associated with higher phaeopigment. An ice-edge effect?
-* Predator cluster 2 (open ocean) is associated with higher temperatures and chl a. Associated with macrozooplankton communities 2a and 2b (small euphausiids, amphipods, etc).
-* Predator cluster 3 (pagophilic) is associated with higher salinity and deeper mixed layer. Associated with macrozooplankton communities 3a and 3b (big euphausiids, Metridia, etc).
-* No clear relationship between predator community and ice categories.
+* Predator cluster associations with ice are unclear. May change after digging into Christian's ice data?
+* Both predator cluster 1 (open-ish) and 2 (marginal ice zone) are associated with higher temperatures and macrozooplankton cluster 2 (small euphausiids, amphipods, etc)
+* Predator cluster 3 (pagophilic) is associated with higher salinity, deeper mixed layer, and macrozooplankton cluster 3 (big euphausiids, Metridia, etc)
+* Macrozooplankton cluster 1 (salps etc) evenly distributed across predator clusters
 
@@
 # Hydrology plot
-ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+hydro_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
   geom_point(aes(color = pred_clust), alpha = 0.8) +
@@
                color = "grey50",
-               alpha = 0.75) +
+               alpha = 0.6) +
   geom_text(aes(x = NMDS1, y = NMDS2, label = envvar),
@@
   scale_color_brewer(palette = "Dark2") +
-  theme_classic() +
-  theme(legend.position = "bottom")
+  theme_classic()
 
@@
 ice_facs <- filter(envfit_facs, str_detect(envvar, "ice_cat"))
-ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust), alpha = 0.8) +
+ice_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.6) +
   geom_point(aes(x = NMDS1, y = NMDS2),
@@
 zoop_facs <- filter(envfit_facs, str_detect(envvar, "zoop"))
-ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust), alpha = 0.8) +
+zoop_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.6) +
   geom_point(aes(x = NMDS1, y = NMDS2),
@@
   theme_classic()
+
+layout <- "
+AAAA
+BBCC
+"
+hydro_plot + ice_plot + zoop_plot + 
+  plot_layout(design = layout, guides = "collect")
 ```

### 2023-06-01 reports/03_combined.qmd
Incorporate updated ice data into station and cluster analyses

@@
   mutate(n_obs = ice_intervals,
-         n_mode = round(ice_cat_freq * n_obs, digits = 0)) %>% 
-  group_by(n_obs, n_mode, ice_cat) %>% 
+         n_mode = round(ice_type_freq * n_obs, digits = 0)) %>% 
+  group_by(n_obs, n_mode, ice_type) %>% 
   summarize(Stations = n(), .groups = "drop") %>% 
@@
   coord_fixed() +
-  facet_wrap(~ ice_cat, ncol = 2) +
+  facet_wrap(~ ice_type, ncol = 2) +
   theme_classic()
@@
 ant_basemap(map_lim) +
-  geom_sf(aes(color = ice_cat), 
-          predators_stations,
+  geom_sf(aes(color = ice_type), 
+          filter(predators_stations, year != 2014),
           size = 0.5) +

### 2023-06-01 reports/04_pred_clus.qmd
Incorporate updated ice data into station and cluster analyses

@@
 The above mosaic plot shows the distribution of stations by predator and zooplankton clusters.
-For example, the bottom left rectangle indicates 9 stations were assigned to zooplankton cluster 1 and predator cluster 1.
+For example, the bottom left rectangle indicates 34 stations were assigned to zooplankton cluster 1 and predator cluster 1.
 Rectangles color-coded by zooplankton cluster using the color palette from Dietrich et al.
@@
 
-Upon visual inspection of the mosaic plot, I don't perceive any clear correlations between the zooplankton and predator cluster assignments.
-We're interpreting this as evidence predator communities are structured by the physical environment, specifically sea ice, rather than the prey communities.
+Upon visual inspection of the mosaic plot, there appears to be some general correlation.
+Predator cluster 1 associates somewhat indiscriminately with zooplankton clusters 2a-3b.
+Predator cluster 2 associates strongly with zooplankton cluster 2a, as well as with 2b to a lesser extent.
+Predator cluster 3 associates with zooplankton clusters 3a and 3b.
+Zooplankton cluster 1 seems to be the "background".
+
+For context, zooplankton cluster 1 was mostly salps, clusters 2a and 2b were small krill and amphipods (mostly offshore), and clusters 3a and 3b were large krill and *Metridia* (mostly in the Bransfield Strait).
 
@@
 
-I compared the drivers of community structure two ways: physics (hydrology and ice coverage) and zooplankton (community assignments). The result is kind of a mess! Here's what makes jumps out at me:
+I compared the drivers of community structure two ways: physics (hydrology and ice coverage) and zooplankton (community assignments). Here's what makes jumps out at me:
 
-* Predator cluster associations with ice are unclear. May change after digging into Christian's ice data?
-* Both predator cluster 1 (open-ish) and 2 (marginal ice zone) are associated with higher temperatures and macrozooplankton cluster 2 (small euphausiids, amphipods, etc)
+* Predator cluster associations with ice kind of track with what we expected.
+The pagophilic predator cluster correlates with first-year ice, open ocean predators with open ice, marginal ice predators with thin ice.
+Weirdly, multi-year ice falls among the open ocean predators, but only two stations were classified as multi-year ice so it's probably a small sample size thing.
+* Predator cluster 1 (open water) and 2 (marginal ice zone) are associated with higher temperatures and zooplankton cluster 2 (small euphausiids, amphipods, etc).
+They differ in the open water predator cluster was found in areas with greater phaeopigment and the marginal ice cluster in areas with greater chlorophyll-a.
 * Predator cluster 3 (pagophilic) is associated with higher salinity, deeper mixed layer, and macrozooplankton cluster 3 (big euphausiids, Metridia, etc)
@@
   group_by(amlr.station) %>%
-  summarize(ice_cat = ice_cat[1]) %>% 
+
+  summarize(ice_type = ice_type[1]) %>% 
   as_tibble() %>% 
@@
 # Ice plot
-ice_facs <- filter(envfit_facs, str_detect(envvar, "ice_cat"))
+ice_facs <- filter(envfit_facs, str_detect(envvar, "ice_type"))
 ice_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +

### 2023-06-01 R/cluster.R

@@ -26,8 +26,8 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     stations_group <- stations[stations$Year == sightings_key$year, ]
     # Find nearest station within year
     nearest_station <- st_join(sightings_group,
-                                   select(stations_group, amlr.station),
-                                   join = st_nearest_feature)
+                               select(stations_group, amlr.station, date.tow),
+                               join = st_nearest_feature)
     # Get station coordinates
     station_coords <- stations_group %>%
       slice(map_int(nearest_station$amlr.station,
@@ -35,8 +35,8 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     # Calculate distance from predator sighting to nearest station
     nearest_station %>%
       mutate(dist_to_station = st_distance(nearest_station,
-                                               station_coords,
-                                               by_element = TRUE))
+                                           station_coords,
+                                           by_element = TRUE))
   }
 
   # Assign sightings across years
@@ -53,7 +53,7 @@ filter_sightings <- function(sightings, stations, station_thr) {
   sightings_agg <- sightings %>%
     as_tibble() %>%
     group_by(amlr.station, species) %>%
-    summarize(count_norm = sum(count_norm), .groups = "drop")
+    summarize(count_norm = sum(count) / survey_nmi[1], .groups = "drop")
   retain <- sightings_agg %>%
     group_by(species) %>%
     summarize(stations_present = n(),

### 2023-06-01 R/ice.R

@@ -5,34 +5,68 @@ mfv <- function(x) {
   names(x_tbl)[which.max(x_tbl)]
 }
 
-# Adds three columns to stations data frame:
-# ice_cat [fct] OPEN, THIN, FY, MY
-# ice_intervals [int] count of intervals with ice obs associated with station
-# ice_cat_freq [dbl] fraction of intervals assoc w station with modal ice cat
-aggregate_ice <- function(stations, underway, ice_cat) {
-  # Aggregates underway ice observations by nearest AMLR station
-  # Uses mode as aggregation function i.e. mfv()
-  underway_ice <- as_tibble(stations) %>%
-    left_join(transmute(underway,
-                        year = as.numeric(substr(cruise, 5, 8)),
-                        interval,
-                        ice_type_desc),
-              by = c("year", "interval"),
-              multiple = "all") %>%
-    filter(ice_type_desc != "NULL") %>%
-    left_join(ice_cat, by = "ice_type_desc") %>%
-    # Ice observations were unique to intervals, but intervals are
-    # repeated if there were multiple predator sightings
-    group_by(amlr.station,
-             interval) %>%
-    summarize(interval_ice_cat = first(ice_cat),
-              .groups = "drop_last") %>%
-    summarize(ice_cat = mfv(interval_ice_cat),
-              ice_intervals = n(),
-              ice_cat_freq = sum(ice_cat == interval_ice_cat) / n()) %>%
-    mutate(ice_cat = factor(ice_cat,
-                            levels = c("OPEN", "THIN", "FY", "MY")))
+# Clean ice categories
+# Converts e.g., FY/MY to FY and collapses SLUSH and BRASH into THIN
+clean_ice <- function(ice_types) {
+  str_extract(ice_types, "[^/]+") %>%
+    factor(levels = c("OPEN",
+                      "SLUSH",
+                      "BRASH",
+                      "THIN",
+                      "FY",
+                      "MY",
+                      "NULL")) %>%
+    fct_collapse(THIN = "SLUSH",
+                 THIN = "BRASH")
+}
 
-  stations %>%
-    left_join(underway_ice, by = "amlr.station")
+# Adds three columns to stations data frame:
+# ice_type [fct] OPEN, THIN, FY, MY
+# ice_intervals [int] count of intervals with ice obs associated with station
+# ice_type_freq [dbl] fraction of intervals assoc w station with modal ice type
+# ice_coverage [dbl] mean of ice coverage code
+aggregate_ice <- function(stations, ice_raw) {
[…]

### 2023-06-01 _targets.R

@@ -5,7 +5,7 @@ library(tidyverse)
 
 # Set target options:
 tar_option_set(
-  packages = c("sf", "tidyverse"),
+  packages = c("readxl", "sf", "tidyverse"),
   format = "rds"
 )
 
@@ -39,7 +39,7 @@ list(
     here(wamlr_dir, "WAMLR-2012-2016-Underway-Predator-Sightings.csv"),
     format = "file"
   ),
-  tar_target(predators, readr::read_csv(predators_file)),
+  tar_target(predators, read_csv(predators_file)),
   tar_target(predators_agg, aggregate_predators(predators)),
   tar_target(predators_sf,
              latlon_to_sf(predators_agg, coords = c("lon_mean", "lat_mean"))),
@@ -65,11 +65,25 @@ list(
     zoop_long,
     pivot_longer(zoop, -(1:53), names_to = "taxa", values_to = "abundance")
   ),
+  # Ice observations
+  tar_target(
+    ice_file,
+    here("data", "ALl_raw_bird_pred_obs_12_16_winter.xlsx"),
+    format = "file"
+  ),
+  tar_target(
+    ice_raw,
+    read_excel(ice_file) %>%
+      transmute(year_ice = Year,
+                lat = Latitude,
+                lon = Longitude,
+                ice_code = as.numeric(IceCode),
+                ice_type = clean_ice(IceType),
+                date_ice = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
+      filter(ice_type != "NULL") %>%
+      latlon_to_sf(coords = c("lon", "lat"))
+  ),
   # Aggregate predators and ice observations
-  tar_target(ice_cat_file,
-             here("data", "ICECAT.csv"),
-             format = "file"),
-  tar_target(ice_cat, read_csv(ice_cat_file)),
   tar_target(
     station_effort,
     assign_sightings(effort_sf, zoop_sf, max_dist_km = 15) %>%
@@ -79,7 +93,7 @@ list(
   tar_target(
     predators_stations,
     assign_sightings(predators_sf, zoop_sf, max_dist_km = 15) %>%
-      aggregate_ice(predators, ice_cat) %>%
+      aggregate_ice(ice_raw) %>%
       left_join(station_effort %>%
                   as_tibble() %>%
                   select(amlr.station, survey_nmi),

### 2023-06-07 _targets.R
Include 2014 ice data

@@ -68,19 +68,21 @@ list(
   # Ice observations
   tar_target(
     ice_file,
-    here("data", "ALl_raw_bird_pred_obs_12_16_winter.xlsx"),
+    here("data", "ice", "ice.csv"),
     format = "file"
   ),
   tar_target(
     ice_raw,
-    read_excel(ice_file) %>%
+    read_csv(ice_file, na = c("", "NULL")) %>%
       transmute(year_ice = Year,
                 lat = Latitude,
                 lon = Longitude,
                 ice_code = as.numeric(IceCode),
                 ice_type = clean_ice(IceType),
                 date_ice = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
-      filter(ice_type != "NULL") %>%
+      filter(ice_type != "NULL",
+             # 1 ice_code is "85"
+             ice_code <= 10) %>%
       latlon_to_sf(coords = c("lon", "lat"))
   ),
   # Aggregate predators and ice observations

### 2023-06-07 reports/04_pred_clus.qmd
Include 2014 ice data
updated targets meta after fixing ice data

@@
 station_ice <- predators_stations %>% 
-  group_by(amlr.station) %>%
-
-  summarize(ice_type = ice_type[1]) %>% 
-  as_tibble() %>% 
-  select(-geometry)
+  distinct(amlr.station, ice_type, ice_coverage) %>% 
+  as_tibble()
 
@@
 AAAA
+AAAA
 BBCC

### 2023-06-27 reports/01_predators.qmd
Update species maps

@@
 
-## Antarctic Peninsula winter predators
+## Counts
 
-### Counts
-
-The five most abundant species (in ind nmi^-1).
+The five most abundant species (in ind nmi\^-1).
 
@@
 
-### Maps
+## Maps
 
 Survey effort. 21 out of the 19404 intervals were recorded as nmi \> 1000 (as opposed to \~1 nmi for the rest). Those intervals excluded from the survey effort.
+
+### Climatologies
+
+```{r}
+#| label: species_climatologies
+
+library(terra)
+# Mkae "top five" predator sightings spatially aware
+topfive_sf <- predators_agg %>% 
+  semi_join(top_five, by = "species") %>% 
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+# Create a template raster of the study area
+effort_ext <- effort_sf %>% 
+  st_transform(ant_proj()) %>% 
+  ext()
+rast_template <- rast(crs = as.character(ant_proj())[[1]],
+                      extent = effort_ext,
+                      resolution = 30e3)
+# Rasterize effort by year
+effort_rast <- effort_sf %>% 
+  st_transform(ant_proj()) %>% 
+  rasterize(rast_template,
+            field = "nmi",
+            by = "year",
+            fun = sum)
+# Rasterize the predator sightings by year and species
+topfive_rast <- topfive_sf %>% 
+  st_transform(ant_proj()) %>% 
+  mutate(year_species = interaction(year, species)) %>% 
+  rasterize(rast_template, 
+            field = "count",
+            by = "year_species",
+            fun = sum)
+# Normalize counts by effort
+topfive_norm <- topfive_rast
+for (i in seq(nlyr(topfive_rast))) {
+  lyr_name <- names(topfive_rast)[i]
+  lyr_year <- substr(lyr_name, 1, 4)
+  effort_year <- effort_rast[lyr_year] %>% 
+    subst(0, NA)
+  topfive_norm[[i]] <- topfive_norm[[i]] / effort_year
+}
+# Average across years
+topfive_climatology <- unique(top_five$species) %>% 
+  map(function(sp) {
+    sp_rast <- subset(topfive_norm, 
+                      which(str_detect(names(topfive_norm), sp)))
+    mean(sp_rast, na.rm = TRUE)
+  }) %>% 
+  set_names(unique(top_five$species)) %>% 
+  rast()
+# Classify raster (otherwise CRSE throw it all off)
+topfive_clim_class <- topfive_climatology %>% 
+  classify(rcl = c(0, 1, 5, 10, 50))
+
+# Project bounding box
+map_bbox <- ext(-63, -54, -64, -59) %>% 
+  project(from = "+proj=longlat +datum=WGS84",
+          to = as.character(ant_proj())[[1]])
+
+# Make the map! 
+ant_basemap() +
+  tidyterra::geom_spatraster(data = topfive_clim_class) +
+  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  scale_fill_brewer(palette = "RdPu", 
+                    na.value = NA, 
+                    direction = 1,
+                    na.translate = FALSE) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+  labs(fill = "Ind nmi^-1") +
+  facet_wrap(~lyr) +
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  theme(legend.position = "top")
+
+```
+
+### Survey effort
 
@@
 
-effort_stars <- effort_sf %>% 
-  filter(nmi < 1000) %>% 
-  rasterize_counts(count_col = "nmi",
-                   res = 2.5e4)
-map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
-                       crs = "EPSG:4326") %>% 
-  project_bbox() %>% 
-  expand_bbox(factor = 1.2)
-ant_basemap(map_lim) + 
-  stars::geom_stars(aes(fill = nmi), data = effort_stars) +
+map_bbox <- ext(effort_rast)
+
+ant_basemap(map_bbox) +
+  tidyterra::geom_spatraster(data = effort_rast) +
   scale_fill_distiller(palette = "PuRd", 
@@
                        na.value = NA) +
-  scale_x_continuous(breaks = seq(-65, -55, by = 5)) +
-  scale_y_continuous(breaks = seq(-65, -58, by = 2)) +
-  coord_ant(map_lim) +
-  facet_wrap(vars(year)) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -57, by = 2)) +
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  labs(fill = "nmi") +
+  facet_wrap(~lyr) +
   expand_limits(fill = 0)
@@
 
-Distribution maps of the five most abundant species in each survey year, aggregated to 25km bins.
+### Yearly distributions
+
+Distribution maps of the five most abundant species in each survey year, aggregated to 30km bins.
 
@@
 
-topfive_sf <- predators_agg %>% 
-  semi_join(top_five, by = "species") %>% 
-  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+map_bbox <- ext(-63, -54, -64, -59) %>% 
+  project(from = "+proj=longlat +datum=WGS84",
+          to = as.character(ant_proj())[[1]])
 
-topfive_stars <- rasterize_counts(topfive_sf, res = 2.5e4)
+topfive_class <- topfive_rast %>% 
+  classify(rcl = c(0, 1, 5, 10, 50, 100, 1000, Inf))
 
-map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
-                       crs = "EPSG:4326") %>% 
-  project_bbox() %>% 
-  expand_bbox(factor = 1.1)
+topfive_class_df <- as.data.frame(topfive_class, xy = TRUE) %>% 
+  pivot_longer(-c("x", "y"), 
+               names_to = "name", 
+               values_to = "count") %>% 
+  separate_wider_delim(name, 
+                       delim = ".", 
+                       names = c("year", "species")) %>% 
+  drop_na(count)
 
-ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = log(count_norm + 1)), 
-                    data = topfive_stars) +
-  scale_fill_distiller(palette = "PuRd",
+ant_basemap() +
+  geom_raster(aes(x, y, fill = count), topfive_class_df) +
+  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  scale_fill_brewer(palette = "RdPu", 
+                    na.value = NA, 
                     direction = 1,
-                       na.value = NA) +
-  scale_x_continuous(breaks = c(-65, -60, -55)) +
-  scale_y_continuous(breaks = c(-66, -64, -62, -60)) +
-  coord_ant(map_lim) +
+                    na.translate = FALSE) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+  labs(fill = "Ind nmi^-1") +
   facet_grid(species ~ year) +
-  expand_limits(fill = 0) +
-  theme(axis.text.x = element_text(size = 7))
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  theme(legend.position = "right",
+        text = element_text(size = 8))
 ```
@@
 
-soi_stars <- rasterize_counts(soi_sf, res = 2.5e4)
+soi_rast <- soi_sf %>% 
+  st_transform(ant_proj()) %>% 
+  mutate(year_species = interaction(year, species)) %>% 
+  rasterize(rast_template, 
+            field = "count",
+            by = "year_species",
+            fun = sum)
 
-map_lim <- st_bbox(soi_sf) %>% 
-  project_bbox() %>% 
-  expand_bbox(factor = 1.2)
+map_bbox <- ext(-63, -54, -64, -59) %>% 
+  project(from = "+proj=longlat +datum=WGS84",
+          to = as.character(ant_proj())[[1]])
 
-ant_basemap(map_lim) +
-  stars::geom_stars(aes(fill = log(count_norm + 1)), data = soi_stars) +
-  scale_fill_distiller(palette = "PuRd",
+soi_class <- soi_rast %>% 
+  classify(rcl = c(0, 1, 5, 10, 50, 100, 1000))
+
+soi_class_df <- as.data.frame(soi_class, xy = TRUE) %>% 
+  pivot_longer(-c("x", "y"), 
+               names_to = "name", 
+               values_to = "count") %>% 
+  separate_wider_delim(name, 
+                       delim = ".", 
+                       names = c("year", "species")) %>% 
+  drop_na(count)
+
+ant_basemap() +
+  geom_raster(aes(x, y, fill = count), soi_class_df) +
+  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  scale_fill_brewer(palette = "RdPu", 
+                    na.value = NA, 
                     direction = 1,
-                       na.value = NA) +
-  scale_x_continuous(breaks = c(-65, -60, -55)) +
-  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
+                    na.translate = FALSE) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+  labs(fill = "Ind nmi^-1") +
   facet_grid(species ~ year) +
-  expand_limits(fill = 0) +
-  coord_ant(map_lim) +
-  theme(axis.text.x = element_text(size = 7))
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  theme(legend.position = "right",
+        text = element_text(size = 8))
 ```

### 2023-06-27 reports/04_pred_clus.qmd
Update species maps

@@
 * Cluster 3 = pagophilic (ADPN, CRSE, LESE)
+
+### Species within clusters
+
+```{r}
+#| label: clusters_species
+predators_abundant %>% left_join(
+  tibble(
+    amlr.station = names(clust),
+    pred_clust = factor(clust,
+                        labels = c("Open", "Marginal", "Pagophilic"))
+  ), by = "amlr.station") %>% 
+  complete(amlr.station, pred_clust, species, 
+           fill = list(count_norm = 0)) %>% 
+  group_by(species, pred_clust) %>% 
+  summarize(mean_abundance = mean(count_norm), .groups = "drop") %>%
+  mutate(species = fct_reorder(species, 
+                               mean_abundance, 
+                               .fun = sum,
+                               .desc = TRUE)) %>% 
+  ggplot(aes(species, mean_abundance)) +
+  geom_col() +
+  labs(y = "Mean abundance (ind nmi^-1)") +
+  facet_wrap(~ pred_clust, ncol = 1) +
+  theme_classic(base_size = 14) +
+  theme(axis.title.x = element_blank())
+```
 

### 2023-06-27 _targets.R

@@ -46,6 +46,7 @@ list(
   # Effort
   tar_target(effort,
              predators %>%
+               filter(nmi < 1000) %>%
                mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
                group_by(cruise, year, interval, lon_mean, lat_mean) %>%
                summarize(nmi = sum(nmi), .groups = "drop") %>%

### 2023-06-28 reports/04_pred_clus.qmd
Figure showing distribution of ice coverage by predator cluster

@@
 
-* Cluster 1 = open ocean? (ANFU)
-* Cluster 2 = marginal ice zone (SNPT, ANPT, FUSE)
-* Cluster 3 = pagophilic (ADPN, CRSE, LESE)
+-   Cluster 1 = open ocean? (ANFU)
+-   Cluster 2 = marginal ice zone (SNPT, ANPT, FUSE)
+-   Cluster 3 = pagophilic (ADPN, CRSE, LESE)
 
@@
 
-* Predator cluster associations with ice kind of track with what we expected.
+-   Predator cluster associations with ice kind of track with what we expected.
 The pagophilic predator cluster correlates with first-year ice, open ocean predators with open ice, marginal ice predators with thin ice.
 Weirdly, multi-year ice falls among the open ocean predators, but only two stations were classified as multi-year ice so it's probably a small sample size thing.
-* Predator cluster 1 (open water) and 2 (marginal ice zone) are associated with higher temperatures and zooplankton cluster 2 (small euphausiids, amphipods, etc).
+-   Predator cluster 1 (open water) and 2 (marginal ice zone) are associated with higher temperatures and zooplankton cluster 2 (small euphausiids, amphipods, etc).
 They differ in the open water predator cluster was found in areas with greater phaeopigment and the marginal ice cluster in areas with greater chlorophyll-a.
-* Predator cluster 3 (pagophilic) is associated with higher salinity, deeper mixed layer, and macrozooplankton cluster 3 (big euphausiids, Metridia, etc)
-* Macrozooplankton cluster 1 (salps etc) evenly distributed across predator clusters
+-   Predator cluster 3 (pagophilic) is associated with higher salinity, deeper mixed layer, and macrozooplankton cluster 3 (big euphausiids, Metridia, etc)
+-   Macrozooplankton cluster 1 (salps etc) evenly distributed across predator clusters
 
@@
 ```
+
+## Clusters by ice coverage
+
+I asserted the clusters corresponded to open water, marginal ice, and pack ice communities. Let's verify that.
+
+```{r}
+library(ggridges)
+
+pred_clust_ice <- tibble(
+  amlr.station = names(clust),
+  pred_clust = factor(clust,
+                      labels = c("Open", "Marginal", "Pagophilic"))
+) %>% 
+  left_join(station_ice,
+            by = "amlr.station") %>% 
+  drop_na(ice_coverage)
+
+ggplot(pred_clust_ice, aes(x = ice_coverage / 10, fill = after_stat(x))) + 
+  geom_histogram(bins = 20, color = "grey30") +
+  scale_x_continuous(labels = scales::percent) +
+  scale_fill_distiller(palette = "Blues", guide = NULL) +
+  facet_grid(rows = vars(pred_clust)) +
+  labs(x = "Ice coverage",
+       y = "Station count") +
+  theme_minimal()
+
+```

### 2023-07-02 reports/04_pred_clus.qmd
Begin looking at sea ice 65% contour

@@
 ```{r}
-library(ggridges)
-
 pred_clust_ice <- tibble(

### 2023-07-02 scratch/ice_conc.R

@@ -1,0 +1,34 @@
+library(ncdf4)
+library(terra)
+library(tidyterra)
+library(tidyverse)
+
+ext(-67, -50, -68, -55) %>%
+  project(from = "epsg:4326",
+          to = "epsg:3412")
+
+ice_nc <- nc_open("~/Downloads/nsidcG02202v4shmday_df31_2bad_59c8_U1688308936040.nc")
+ice_conc_arr <- ncvar_get(ice_nc, "cdr_seaice_conc_monthly") %>%
+  # NETcdf stores values reversed along x-axis
+  apply(3, pracma::fliplr, simplify = FALSE) %>%
+  simplify2array()
+ice_conc_ext <- ext(
+  min(ncvar_get(ice_nc, "xgrid")),
+  max(ncvar_get(ice_nc, "xgrid")),
+  min(ncvar_get(ice_nc, "ygrid")),
+  max(ncvar_get(ice_nc, "ygrid"))
+)
+ice_conc <- rast(
+  ice_conc_arr,
+  crs = "epsg:3412",
+  ice_conc_ext
+)
+
+plot(ice_conc)
+
+ggplot() +
+  geom_spatraster(data = ice_conc) +
+  geom_spatraster_contour(data = ice_conc, breaks = 0.65, color = "white") +
+  facet_wrap(~lyr)
+
+

### 2023-08-02 reports/04_pred_clus.qmd
Fixed massive count normalization error!

@@
 library(targets)
+library(tidyterra)
 library(tidyverse)
@@
 tar_load(c("predators_stations", "predators_abundant", 
+           "seaice_conc_df", "seaice_contours",
            "sightings_clust", "sightings_mtx", "sightings_dist", 
-           "zoop_sf"), 
+           "zoop_sf", "ant_sf"), 
          store = here::here("_targets"))
@@
 source(here::here("R", "zoop.R"))
+seaice_contours <- terra::unwrap(seaice_contours)
 ```
@@
 The gap statistic ($f$) is a goodness of clustering measure.
-The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
-According to this rule, the predators are best described by two clusters (indicated in blue).
-However, $f(2)$ is only *slightly* greater than the threshold $f(3) - s(3)$, and $f$ is determined stochastically, so depending on the iteration the rule may choose $k=3$.
-Based on our prior knowledge of three predator guilds (pagophilic, open water, and intermediate), we'll present the results for $k=3$ here and re-run the analysis for $k=2$ in the supplemental.
+The authors recommended choosing $k$ at the shoulder.
+Heuristically, the shoulder is the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
+According to this rule, the predators are best described by one cluster (indicated in blue).
+However, there's a clear jump from $k=2$ to $k=3$, so based on our prior knowledge of three predator guilds (pagophilic, open water, and intermediate), we'll present the results for $k=3$ here.
 
@@
 
-# Cut predators into five clusters
+# Cut predators into three clusters
 clust <- cutree(sightings_clust, 3)
@@
 species_wide <- predators_abundant %>% 
+  select(amlr.station, species, count_nmi) %>% 
   pivot_wider(names_from = species,
-              values_from = count_norm,
+              values_from = count_nmi,
               values_fill = 0) %>% 
@@
   complete(amlr.station, pred_clust, species, 
-           fill = list(count_norm = 0)) %>% 
+           fill = list(count_nmi = 0)) %>% 
   group_by(species, pred_clust) %>% 
-  summarize(mean_abundance = mean(count_norm), .groups = "drop") %>%
+  summarize(mean_abundance = mean(count_nmi), .groups = "drop") %>%
   mutate(species = fct_reorder(species, 
@@
 What geospatial patterns do we see in the predator communities?
-The marginal ice zone community was mostly observed just north of the South Shetland islands, occasionally farther north beyond Elephant Island.
-The open ocean community occupied the northeastern portion of the study area in 2012, 2014, and 2015.
-In 2013 it contracted westward and in 2016 it was entirely absent.
-The pagophilic predator community dominated the Bransfield Strait itself, expanding north towards Elephant Island in 2013 and 2016.
+The pagophilic predator community (purple) was largely found within the Bransfield Strait and the open ocean predator community (green) further offshore (with the exception of 2015, when it moved more into the Strait).
+The marginal ice zone predator community (orange) was geographically dynamic, its distribution shifting in space substantially between years.
+Blue line indicates 65% ice concentration contour during the month of August in each year.
 
@@
   geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
+  geom_sf(data = seaice_contours, color = "skyblue") +
   facet_wrap(~ Year) +
@@
   scale_color_brewer(palette = "Dark2") +
+  scale_fill_gradient(lim = c(0, 1), na.value = "transparent") +
   coord_ant(map_lim) +
@@
 hydro_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(aes(color = pred_clust)) +
   geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
@@
 ice_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust), alpha = 0.6) +
+  geom_point(aes(color = pred_clust)) +
   geom_point(aes(x = NMDS1, y = NMDS2),
@@
 zoop_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust), alpha = 0.6) +
+  geom_point(aes(color = pred_clust)) +
   geom_point(aes(x = NMDS1, y = NMDS2),

### 2023-08-02 R/cluster.R

@@ -50,17 +50,23 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
 
 filter_sightings <- function(sightings, stations, station_thr) {
   total_stations <- n_distinct(stations$amlr.station)
-  sightings_agg <- sightings %>%
+  retain <- sightings %>%
+    group_by(species) %>%
+    summarize(stations_present = n()) %>%
+    mutate(station_frac = stations_present / n()) %>%
+    filter(station_frac >= station_thr)
+  semi_join(sightings, retain, by = "species")
+}
+
+normalize_counts <- function(sightings, effort, stations) {
+  sightings %>%
     as_tibble() %>%
     group_by(amlr.station, species) %>%
-    summarize(count_norm = sum(count) / survey_nmi[1], .groups = "drop")
-  retain <- sightings_agg %>%
-    group_by(species) %>%
-    summarize(stations_present = n(),
-              station_frac = stations_present / total_stations,
-              .groups = "drop") %>%
-    filter(station_frac >= station_thr)
-  semi_join(sightings_agg, retain, by = "species")
+    summarize(count = sum(count), .groups = "drop") %>%
+    left_join(select(as_tibble(effort), amlr.station, survey_nmi),
+              by = "amlr.station") %>%
+    mutate(count_nmi = count / survey_nmi) %>%
+    left_join(select(stations, amlr.station, Year), by = "amlr.station")
 }
 
 sightings_to_matrix <- function(sightings) {
@@ -68,7 +74,7 @@ sightings_to_matrix <- function(sightings) {
     as_tibble() %>%
     pivot_wider(id_cols = amlr.station,
                 names_from = species,
-                values_from = count_norm,
+                values_from = count_nmi,
                 values_fill = 0) %>%
     mutate(across(-amlr.station, ~ log(.x + 1)))
   sightings_mtx <- as.matrix(select(sightings_wide, -amlr.station))

### 2023-08-02 R/ice.R

@@ -20,15 +20,15 @@ clean_ice <- function(ice_types) {
                  THIN = "BRASH")
 }
 
-# Adds three columns to stations data frame:
+# Aggregates ice data:
 # ice_type [fct] OPEN, THIN, FY, MY
 # ice_intervals [int] count of intervals with ice obs associated with station
 # ice_type_freq [dbl] fraction of intervals assoc w station with modal ice type
 # ice_coverage [dbl] mean of ice coverage code
 aggregate_ice <- function(stations, ice_raw) {
   stations2 <- stations %>%
-    transmute(amlr_station = amlr.station,
-              year_station = year,
+    transmute(amlr.station,
+              year_station = Year,
               date_station = as.Date(date.tow))
 
   # Spatially join stations to raw ice observations
@@ -50,12 +50,12 @@ aggregate_ice <- function(stations, ice_raw) {
     }) %>%
     ungroup() %>%
     filter(station_distance <= units::as_units(15, "km"),
-           abs(station_lag) <= as.difftime(1, units = "days")) %>%
+           abs(station_lag) <= as.difftime(3, units = "days")) %>%
     st_as_sf(sf_column_name = "geometry")
 
   # Aggregate ice observations by station
-  ice_aggregate <- ice_stations %>%
-    group_by(amlr_station, year = year_ice) %>%
+  ice_stations %>%
+    group_by(amlr.station, year = year_ice) %>%
     summarize(modal_ice_type = mfv(ice_type),
               ice_intervals = n(),
               ice_type_freq = sum(ice_type == modal_ice_type),
@@ -66,7 +66,49 @@ aggregate_ice <- function(stations, ice_raw) {
                                                         "FY",
                                                         "MY"))) %>%
     as_tibble() %>%
-    select(amlr_station, ice_type, ice_intervals, ice_type_freq, ice_coverage)
+    select(amlr.station, ice_type, ice_intervals, ice_type_freq, ice_coverage)
+}
 
-  left_join(stations, ice_aggregate, by = c(amlr.station = "amlr_station"))
+collect_seaice <- function(seaice_dir) {
+  seaice <- seaice_dir %>%
+    dir(full.names = TRUE) %>%
+    map(rast) %>%
+    rast()
+  set.names(seaice, as.character(2012:2016))
+
+  proj_ext <- ext(-67, -50, -68, -55) %>%
[…]

### 2023-08-02 _targets.R

@@ -5,7 +5,7 @@ library(tidyverse)
 
 # Set target options:
 tar_option_set(
-  packages = c("readxl", "sf", "tidyverse"),
+  packages = c("readxl", "sf", "terra", "tidyterra", "tidyverse"),
   format = "rds"
 )
 
@@ -86,6 +86,14 @@ list(
              ice_code <= 10) %>%
       latlon_to_sf(coords = c("lon", "lat"))
   ),
+  # Satellite sea ice data
+  tar_target(
+    seaice_dir,
+    here("data", "ice", "Aug sea ice concentration"),
+    format = "file"
+  ),
+  tar_target(seaice_contours, seaice2contour(seaice_dir, 0.65)),
+  tar_target(seaice_conc_df, seaice2df(seaice_dir)),
   # Aggregate predators and ice observations
   tar_target(
     station_effort,
@@ -94,13 +102,14 @@ list(
       summarize(survey_nmi = sum(nmi), .groups = "drop")
   ),
   tar_target(
+    stations_ice,
+    aggregate_ice(zoop_sf, ice_raw)
+  ),
+  tar_target(
     predators_stations,
     assign_sightings(predators_sf, zoop_sf, max_dist_km = 15) %>%
-      aggregate_ice(ice_raw) %>%
-      left_join(station_effort %>%
-                  as_tibble() %>%
-                  select(amlr.station, survey_nmi),
-                by = "amlr.station")
+      normalize_counts(station_effort, zoop_sf) %>%
+      left_join(stations_ice, by = "amlr.station")
   ),
   tar_target(
     predators_abundant,

### 2023-08-02 scratch/clustrelfreq.R

@@ -25,3 +25,48 @@ pred_prey %>%
              total = "both") %>%
   as_flextable()
 
+
+# Multinomial regression --------------------------------------------------
+
+library(nnet)
+pred_prey2 <- pred_prey %>%
+  mutate(zoop_clust2 = substr(zoop_clust, 1, 1),
+         pred_clust = relevel(pred_clust, ref = "Open"))
+test <- multinom(pred_clust ~ 0 + zoop_clust + ice_coverage,
+                 data = pred_prey2)
+summary(test)
+z <- summary(test)$coefficients/summary(test)$standard.errors
+z
+
+test_pred <- expand_grid(
+  zoop_clust = c("1", "2a", "2b", "3a", "3b"),
+  ice_coverage = seq(0, 1, by = 0.1)
+) %>%
+  mutate(class_probs = as.data.frame(predict(test,
+                                             newdata = .,
+                                             type = "probs"))) %>%
+  unpack(class_probs) %>%
+  pivot_longer(c(Open, Marginal, Pagophilic),
+               names_to = "pred_clust",
+               values_to = "probs")
+pred_freq <- pred_prey %>%
+  group_by(pred_clust) %>%
+  summarize(n = n()) %>%
+  mutate(frac = n / sum(n))
+ggplot(test_pred,
+       aes(ice_coverage, probs, color = pred_clust)) +
+  geom_segment(aes(x = 0, y = frac, xend = 1, yend = frac, color = pred_clust),
+               pred_freq,
+               linetype = "dashed") +
+  geom_line() +
+  facet_grid(cols = vars(zoop_clust)) +
+  theme_classic()
+
+# Is "open" right? --------------------------------------------------------
+
+predators_stations %>%
+  semi_join(filter(pred_prey, pred_clust == "Open"),
+            by = "amlr.station") %>%
+  as.data.frame() %>%
+  select(amlr.station, species, count_norm) %>% View()
+  pivot_wider(names_from = "species", values_from = "count_norm")

### 2023-08-10 reports/04_pred_clus.qmd
Include all predators in cluster analysis (ie remove 5% threshold)

@@
 
-Following the methods of Dietrich et al.
-(2021), clustering analysis is restricted to species sighted near 5% or more of stations.
-This avoids granting undue leverage to rare species.
-`r n_distinct(predators_abundant$species)` species retained (i.e., species above red line).
+Dietrich et al.
+(2021) restricted cluster analysis to taxa observed at 5% or more of stations, which was necessary to avoid giving rare species too much leverage.
+Predators are far less diverse, so this analysis retains all species except uncertain IDs (e.g.
+UNSE = unknown seal).
+`r n_distinct(predators_abundant$species)` species included in analysis.
 
@@
 total_stations <- n_distinct(zoop_sf$amlr.station)
-pred_frac <- predators_stations %>%
+pred_frac <- predators_abundant %>%
   as_tibble() %>% 
@@
   arrange(desc(station_frac))
-cutoff_row <- sum(pred_frac$station_frac >= 0.05)
 
@@
   kableExtra::kable_styling("striped",
-                            full_width = FALSE) %>% 
-  kableExtra::row_spec(cutoff_row, 
-                       extra_css = "border-bottom: 2px dashed red")
+                            full_width = FALSE)
 
@@
 Heuristically, the shoulder is the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
-According to this rule, the predators are best described by one cluster (indicated in blue).
-However, there's a clear jump from $k=2$ to $k=3$, so based on our prior knowledge of three predator guilds (pagophilic, open water, and intermediate), we'll present the results for $k=3$ here.
+According to this rule, the predators are best described by one cluster.
+However, there's a clear jump from $k=2$ to $k=3$, followed by a leveling off.
+Therefore, based on our prior knowledge of three predator guilds (pagophilic, open water, and marginal ice), we'll present the results for $k=3$ here.
 
@@
 
+TODO: add text about how to interpret values.
+
 ```{r}
@@
 
-### Cluster descriptions
+### Cluster interpretation
 
--   Cluster 1 = open ocean? (ANFU)
--   Cluster 2 = marginal ice zone (SNPT, ANPT, FUSE)
--   Cluster 3 = pagophilic (ADPN, CRSE, LESE)
+-   Cluster 1 = open water
+-   Cluster 2 = marginal ice zone
+-   Cluster 3 = pagophilic
+
+Indicator value $d_{ic}$ of species $i$ in cluster $c$ is the product of relative fidelity, $f_{ic}$, and relative abundance, $a_{ic}$. See documentation of `labdsv::indval()`.
+
+Relative fidelity $f_{ic}$ of species $i$ in cluster $c$ is:
+
+$$
+f_{ic} = \frac{\sum_{j \in c}p_{ij}}{n_c}
+$$
+
+Relative abundance $a_{ic}$ is:
+
+$$
+a_{ic}=\frac{\sum_{j \in c}x_{ij}/n_c}{\sum_{k=1}^K (\sum_{j \in k}x_{ij}/n_k)}
+$$
+
+A high indicator value, e.g., $d_{ic}\approx1$, means species $i$ is present in nearly all sites in cluster $c$ and nearly absent from sites not in $c$.
+The greatest $d_{ic}$ observed here was Adèlie penguins in cluster 3 at 0.74.
+ADPN were observed in 45 of the 60 cluster 3 sites, for $f_{ic}=0.75$.
+The average abundance of ADPN in cluster 1 was 4.04 ind nm\^-1, compared to 0.113 and 0.166 in clusters 2 and 3, respectively.
+Yields $a_{ic}=0.935$.
+
+Note that differences in indicator values between species reflect *relative* changes in frequency/abundance.
+For example, in cluster 3 (pagophilic predators), Antarctic fur seals were found at more stations ($f_{ic}=0.783$) and in higher abundances (mean 2.22 ind nm\^-1) than crabeater seals ($f_{ic}=0.650$, mean 1.73 ind nm\^-1), yet they have the same indicator value $d_{ic}=0.63$.
+That's because Antarctic fur seals were still somewhat abundant in another cluster whereas crabeater seals were concentrated in cluster 3.
+Specifically, mean Antarctic fur seal abundance was 0.441 ind nm\^-1 in cluster 2 (marginal ice zone), 19.9% of their cluster 1 abundance.
+Crabeater seals dropped from 1.73 ind nm\^-1 to 0.133 ind nm\^-1, or 7.7% of cluster 1 abundance.
+
+Comparing the highest indicator values between clusters, the pagophilic community has the most distinct collection of predators, followed by the marginal ice zone and open water communities.
+Three species had $d_{ic} \gt 0.6$ in the pagophilic cluster: Adèlie penguins, Antarctic fur seals, and crabeater seals.
+All three of these species were both more prevalent and more abundant in the pagophilic cluster than either of the other two.
+In contrast, the open water community had no species with $d_{ic} \gt 0.35$.
+The three most prevalent species in the open water cluster were snow petrels, Antarctic petrels, and southern giant petrels, all of which were more abundant in the marginal ice zone cluster.
+Meaning, the pagophilic community was distinct because of specific species (Adèlie penguins, Antarctic fur seals, and crabeater seals), while the open water community was characterized by more variable species compositions.
+The marginal ice zone community fell in between.
+Its highest indicator value species (snow petrel, Antarctic petrel, and southern giant petrel) were widespread across communities, so their relative abundances in the marginal ice zone cluster were lower ($a_{ic} \in [0.400 - 0.652]$) than the three highest indicator value species were in the pagophilic cluster ($a_{ic} \in [0.810 - 0.971]$).
+
+```{r}
+#| label: indval_interpret
+
+# Adelie penguin x cluster 3 worked example
+predators_abundant2 <- predators_abundant %>% 
+  mutate(cluster = map_int(amlr.station, \(x) clust[x]))
+adpn <- filter(predators_abundant2, species == "ADPN")
+
+pic_adpn <- sum(adpn$cluster == 3)
+nc3 <- sum(clust == 3)
+fic_adpn <- pic_adpn / nc3
+
+abu_adpn <- adpn %>% 
+  group_by(cluster) %>% 
+  summarize(abu = sum(count_nmi) / n())
+aic_adpn <- abu_adpn$abu[3] / sum(abu_adpn$abu)
+
+# ANFU x cluster 1
+anfu <- filter(predators_abundant2, species == "ANFU")
+
+pic_anfu <- sum(anfu$cluster == 1)
+nc1 <- sum(clust == 1)
+fic_anfu <- pic_anfu / nc1
+
+abu_anfu <- anfu %>% 
+  group_by(cluster) %>% 
+  summarize(abu = sum(count_nmi) / n())
+aic_anfu <- abu_anfu$abu[1] / sum(abu_anfu$abu)
+
+```
 
@@
 
-## Cluster maps
+## Cluster distribution
 
-What geospatial patterns do we see in the predator communities?
-The pagophilic predator community (purple) was largely found within the Bransfield Strait and the open ocean predator community (green) further offshore (with the exception of 2015, when it moved more into the Strait).
-The marginal ice zone predator community (orange) was geographically dynamic, its distribution shifting in space substantially between years.
-Blue line indicates 65% ice concentration contour during the month of August in each year.
+How frequent are the different clusters?
+The marginal ice zone predator community was observed most often, followed by the open water then the pagophilic communities.
+The pagophilic community was observed most frequently in 2016.
+Sea ice extent was less in 2016 than 2013 and 2015, so the relative frequency of the pagophilic community in 2016 is initially surprising.
+However, it may be explained by the greater sea ice in 2013 and 2015 limiting survey access to the Bransfield Strait.
 
 ```{r}
-#| label: cluster_maps
+#| label: cluster_freq
 
-pred_clust_sf <- tibble(
+station_ice <- predators_abundant %>%
+  distinct(amlr.station, ice_coverage)
+
+pred_clust <- tibble(
   amlr.station = names(clust),
@@
             by = "amlr.station") %>% 
+  left_join(station_ice, by = "amlr.station")
+
+pred_clust %>% 
+  mutate(Year = factor(Year)) %>% 
+  crosstable::crosstable("Year", by = "pred_clust", total = "col")
+
+```
+
+What geospatial patterns do we see in the predator communities?
+The pagophilic predator community (purple) was largely found within the Bransfield Strait.
+The marginal ice zone community was mostly distributed around the South Shetland Islands, however in the years 2013 and 2015 (when sea ice extent was greatest) this community was more frequent further offshore.
+The open water community was found furthest off shore (north in 2012 and 2014, west in 2016), with the exception of 2015 when it moved south to the entrance of the Bransfield Strait, even moving within the Strait.
+Is this the result of polynyas?
+Blue line indicates 65% ice concentration contour during the month of August in each year.
+Notice the limited coverage of the Bransfield Strait in 2012 and 2013, and the absence of eastern stations in 2016.
+
+```{r}
+#| label: cluster_maps
+
+pred_clust_sf <- pred_clust %>% 
   st_as_sf() %>% 
@@
 
+icy_open <- filter(pred_clust_sf,
+                   pred_clust == "Open",
+                   ice_coverage >= 7.5)
+
 ant_basemap(map_lim) +
-  geom_sf(aes(color = pred_clust), pred_clust_sf, size = 2, shape = 15) +
+  geom_sf(aes(color = pred_clust), 
+          pred_clust_sf, 
+          size = 2, 
+          shape = 15) +
   geom_sf(data = seaice_contours, color = "skyblue") +
@@
   left_join(station_ice, by = "amlr.station") %>% 
-  mutate(zoop = fct_collapse(zoop,
-                             `2` = c("2a", "2b"),
-                             `3` = c("3a", "3b")),
-         Year = factor(Year)) %>% 
+  mutate(Year = factor(Year)) %>% 
   slice(map_int(rownames(sightings_mtx),
@@
                alpha = 0.6) +
-  geom_text(aes(x = NMDS1, y = NMDS2, label = envvar),
+  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
                            data = envfit_vecs) +
@@
 
+Summaries of environmental conditions by cluster.
+
+```{r}
+#| label: env_summ
+
+env <- tibble(
+  amlr.station = names(clust),
+  pred_clust = factor(clust,
+                      labels = c("Open", "Marginal", "Pagophilic"))
+) %>% 
+  left_join(
+    select(as_tibble(zoop_sf),
+           amlr.station, avg.salinity, Integ.chla.100m),
+    by = "amlr.station") %>% 
+  left_join(select(station_ice, amlr.station, ice_coverage),
+            by = "amlr.station") %>% 
+  pivot_longer(avg.salinity:ice_coverage, 
+               names_to = "envvar",
+               values_to = "envval")
+env %>% 
+  group_by(envvar) %>% 
+  summarize(envaov = list(aov(envval ~ pred_clust))) %>% 
+  mutate(Fval = map_dbl(envaov, \(aov) summary(aov)[[1]]$`F value`[1]),
+         pval = map_dbl(envaov, \(aov) summary(aov)[[1]]$`Pr(>F)`[1]))
+env %>% 
+  group_by(pred_clust, envvar) %>% 
+  summarize(across(envval, list(mean = partial(mean, na.rm = TRUE),
+                                sd = partial(sd, na.rm = TRUE))),
+            .groups = "drop")
+```
+
 ## Clusters by ice coverage
@@
 ```
+
+## Predator *x* prey clusters
+
+Predator/prey cluster associations were statistically significant.
+The most positive residuals were between pagophilic predators x zooplankton 3b (4.07), pagophilic predators x zooplankton 3a (2.68), marginal ice zone predators x zooplankton 2a (2.33), and open water predators x zooplankton 2b (1.82).
+Zooplankton 1 was not
+
+```{r}
+pred_prey <- as_tibble(pred_clust) %>% 
+  select(amlr.station, pred_clust) %>% 
+  left_join(select(as_tibble(zoop_sf), 
+                   amlr.station, 
+                   prey_clust = `Winter Cluster factor`),
+            by = "amlr.station")
+pred_prey_chisq <- with(pred_prey, chisq.test(pred_clust, prey_clust))
+pred_prey_chisq
+pred_prey_chisq$residuals
+
+```
+
+##

### 2023-08-10 R/cluster.R

@@ -48,13 +48,14 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     st_as_sf()
 }
 
-filter_sightings <- function(sightings, stations, station_thr) {
+filter_sightings <- function(sightings, stations) {
+  unk_sp <- c("UNSE", "UNPR", "UGPT", "UNPN")
   total_stations <- n_distinct(stations$amlr.station)
   retain <- sightings %>%
     group_by(species) %>%
     summarize(stations_present = n()) %>%
-    mutate(station_frac = stations_present / n()) %>%
-    filter(station_frac >= station_thr)
+    mutate(station_frac = stations_present / total_stations) %>%
+    filter(!species %in% unk_sp)
   semi_join(sightings, retain, by = "species")
 }
 

### 2023-08-10 _targets.R

@@ -113,7 +113,7 @@ list(
   ),
   tar_target(
     predators_abundant,
-    filter_sightings(predators_stations, zoop_sf, station_thr = 0.05)
+    filter_sightings(predators_stations, zoop_sf)
   ),
   # Clustering
   tar_target(sightings_mtx, sightings_to_matrix(predators_abundant)),

### 2023-08-10 scratch/clustrelfreq.R

@@ -1,9 +1,9 @@
 library(sf)
 library(tidyverse)
 
-tar_load(c("predators_stations", "sightings_clust", "zoop"))
+tar_load(c("predators_abundant", "sightings_clust", "zoop"))
 clust <- cutree(sightings_clust, 3)
-station_ice <- predators_stations %>%
+station_ice <- predators_abundant %>%
   distinct(amlr.station, ice_type, ice_coverage) %>%
   as_tibble()
 pred_prey <- tibble(
@@ -18,9 +18,10 @@ pred_prey <- tibble(
             by = "amlr.station") %>%
   left_join(station_ice, by = "amlr.station")
 
+library(crosstable)
+
 pred_prey %>%
-  mutate(zoop_clust2 = substr(zoop_clust, 1, 1)) %>%
-  crosstable(c(zoop_clust2, ice_coverage),
+  crosstable(c(zoop_clust, ice_coverage),
              by = pred_clust,
              total = "both") %>%
   as_flextable()
@@ -30,8 +31,7 @@ pred_prey %>%
 
 library(nnet)
 pred_prey2 <- pred_prey %>%
-  mutate(zoop_clust2 = substr(zoop_clust, 1, 1),
-         pred_clust = relevel(pred_clust, ref = "Open"))
+  mutate(pred_clust = relevel(pred_clust, ref = "Open"))
 test <- multinom(pred_clust ~ 0 + zoop_clust + ice_coverage,
                  data = pred_prey2)
 summary(test)
@@ -62,11 +62,11 @@ ggplot(test_pred,
   facet_grid(cols = vars(zoop_clust)) +
   theme_classic()
 
-# Is "open" right? --------------------------------------------------------
 
-predators_stations %>%
-  semi_join(filter(pred_prey, pred_clust == "Open"),
+# Icy open water? ---------------------------------------------------------
+
+predators_abundant %>%
+  left_join(select(pred_prey, amlr.station, pred_clust),
             by = "amlr.station") %>%
-  as.data.frame() %>%
-  select(amlr.station, species, count_norm) %>% View()
-  pivot_wider(names_from = "species", values_from = "count_norm")
+  filter(pred_clust == "Open",
+         ice_coverage >= 7.5) %>% View()

### 2023-08-10 reports/04_pred_clus.qmd
refactoring cluster report

@@
 
-tar_load(c("predators_stations", "predators_abundant", 
-           "seaice_conc_df", "seaice_contours",
-           "sightings_clust", "sightings_mtx", "sightings_dist", 
+tar_load(
+  c("predators_clust", "stations_clust", "sightings_cut",
+    "seaice_conc_df", "seaice_contours", "stations_ice",
+    starts_with("sightings"),
+    starts_with("nmds"),
     "zoop_sf", "ant_sf"), 
-         store = here::here("_targets"))
+  store = here::here("_targets")
+)
 source(here::here("R", "sp.R"))
@@
 UNSE = unknown seal).
-`r n_distinct(predators_abundant$species)` species included in analysis.
+`r n_distinct(predators_clust$species)` species included in analysis.
 
@@
 total_stations <- n_distinct(zoop_sf$amlr.station)
-pred_frac <- predators_abundant %>%
+pred_frac <- predators_clust %>%
   as_tibble() %>% 
@@
 The gap statistic ($f$) is a goodness of clustering measure.
-The authors recommended choosing $k$ at the shoulder.
+The authors recommended choosing $k$ at the shoulder of the $f \sim k$ curve.
 Heuristically, the shoulder is the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
 According to this rule, the predators are best described by one cluster.
-However, there's a clear jump from $k=2$ to $k=3$, followed by a leveling off.
+However, there's a clear jump from $k=2$ to $k=3$ (red), followed by a leveling off.
 Therefore, based on our prior knowledge of three predator guilds (pagophilic, open water, and marginal ice), we'll present the results for $k=3$ here.
@@
 
-cluster_fn <- function(x, k) {
-  bray_dist <- function(x) {
-    vegdist(x, method = "bray")
-  }
-  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
-}
-set.seed(139)
-sightings_gap <- suppressWarnings(
-  clusGap(sightings_mtx, 
-                   FUN = cluster_fn,
-                   K.max = 10, 
-                   B = 500,
-                   verbose = FALSE)
-)
-sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
-  mutate(
-    k = seq_along(logW),
-    is_optimum = k == maxSE(gap, SE.sim, method = "Tibs2001SEmax")
-  )
-ktibs <- filter(sightings_gap_df, is_optimum) %>% pull(k)
-tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
-ggplot(sightings_gap_df, aes(x = k, y = gap)) +
+ktibs <- with(sightings_gap, k[is_optimum])
+tibs_thr <- with(sightings_gap, gap[ktibs + 1] - SE.sim[ktibs + 1])
+ggplot(sightings_gap, aes(x = k, y = gap)) +
   geom_line() +
@@
                       color = is_optimum)) +
+  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim),
+                  filter(sightings_gap, k == 3),
+                  color = "red") +
   geom_segment(x = ktibs, xend = ktibs + 1, 
@@
 
-# Cut predators into three clusters
-clust <- cutree(sightings_clust, 3)
-g <- split(names(clust), clust)
+g <- split(names(sightings_cut), sightings_cut)
 p <- ggtree(sightings_clust, hang = -1)
@@
 
-How do the predator clusters compare to the zooplankton clusters?
-
-```{r}
-#| label: predator_mosaic
-#| warning: false
-
-# Join with zooplankton clusters
-both_clust <- zoop_sf %>% 
-  transmute(amlr.station, 
-            zoop_clust = factor(`Winter Cluster factor`)) %>% 
-  right_join(tibble(amlr.station = names(clust),
-                    pred_clust = factor(clust)),
-             by = "amlr.station")
-
-# Mosaic plot
-cluster_pal = c(`1` = "#96B687",
-                `2a` = "#9CB2C8",
-                `2b` = "#CAB3CC",
-                `3a` = "#D29A97",
-                `3b` = "#E8AD5C")
-ggplot(both_clust) +
-  geom_mosaic(aes(x = product(zoop_clust, pred_clust), 
-                  fill = zoop_clust)) +
-  geom_mosaic_text(aes(x = product(zoop_clust, pred_clust),
-                       label = after_stat(.wt)),
-                   as.label = TRUE) +
-  scale_fill_manual(values = cluster_pal) +
-  labs(x = "Predator cluster",
-       y = "Zooplankton cluster",
-       fill = "Zooplankton cluster") +
-  theme_classic()
-```
-
-The above mosaic plot shows the distribution of stations by predator and zooplankton clusters.
-For example, the bottom left rectangle indicates 34 stations were assigned to zooplankton cluster 1 and predator cluster 1.
-Rectangles color-coded by zooplankton cluster using the color palette from Dietrich et al.
-2021 Fig 3a.
-
-Upon visual inspection of the mosaic plot, there appears to be some general correlation.
-Predator cluster 1 associates somewhat indiscriminately with zooplankton clusters 2a-3b.
-Predator cluster 2 associates strongly with zooplankton cluster 2a, as well as with 2b to a lesser extent.
-Predator cluster 3 associates with zooplankton clusters 3a and 3b.
-Zooplankton cluster 1 seems to be the "background".
-
-For context, zooplankton cluster 1 was mostly salps, clusters 2a and 2b were small krill and amphipods (mostly offshore), and clusters 3a and 3b were large krill and *Metridia* (mostly in the Bransfield Strait).
-
 ## Indicator species
@@
 
-TODO: add text about how to interpret values.
-
 ```{r}
@@
 
-species_wide <- predators_abundant %>% 
-  select(amlr.station, species, count_nmi) %>% 
-  pivot_wider(names_from = species,
-              values_from = count_nmi,
-              values_fill = 0) %>% 
-  column_to_rownames("amlr.station")
-cluster_indval <- labdsv::indval(species_wide, clust)
-indval_mtx <- cluster_indval$indval
+indval_mtx <- sightings_indval$indval
 indval_fmt <- apply(indval_mtx, 2, function(x) {
@@
 # Adelie penguin x cluster 3 worked example
-predators_abundant2 <- predators_abundant %>% 
-  mutate(cluster = map_int(amlr.station, \(x) clust[x]))
-adpn <- filter(predators_abundant2, species == "ADPN")
+adpn <- filter(predators_clust, species == "ADPN")
 
-pic_adpn <- sum(adpn$cluster == 3)
-nc3 <- sum(clust == 3)
+pic_adpn <- sum(adpn$pred_clust == "Pagophilic")
+nc3 <- sum(sightings_cut == 3)
 fic_adpn <- pic_adpn / nc3
@@
 abu_adpn <- adpn %>% 
-  group_by(cluster) %>% 
+  group_by(pred_clust) %>% 
   summarize(abu = sum(count_nmi) / n())
@@
 # ANFU x cluster 1
-anfu <- filter(predators_abundant2, species == "ANFU")
+anfu <- filter(predators_clust, species == "ANFU")
 
-pic_anfu <- sum(anfu$cluster == 1)
-nc1 <- sum(clust == 1)
+pic_anfu <- sum(anfu$pred_clust == "Open")
+nc1 <- sum(sightings_cut == 1)
 fic_anfu <- pic_anfu / nc1
@@
 abu_anfu <- anfu %>% 
-  group_by(cluster) %>% 
-  summarize(abu = sum(count_nmi) / n())
+  group_by(pred_clust) %>% 
+  summarize(abu = sum(log(count_nmi + 1)) / n())
 aic_anfu <- abu_anfu$abu[1] / sum(abu_anfu$abu)
 
-```
-
-### Species within clusters
-
-```{r}
-#| label: clusters_species
-predators_abundant %>% left_join(
-  tibble(
-    amlr.station = names(clust),
-    pred_clust = factor(clust,
-                        labels = c("Open", "Marginal", "Pagophilic"))
-  ), by = "amlr.station") %>% 
-  complete(amlr.station, pred_clust, species, 
-           fill = list(count_nmi = 0)) %>% 
-  group_by(species, pred_clust) %>% 
-  summarize(mean_abundance = mean(count_nmi), .groups = "drop") %>%
-  mutate(species = fct_reorder(species, 
-                               mean_abundance, 
-                               .fun = sum,
-                               .desc = TRUE)) %>% 
-  ggplot(aes(species, mean_abundance)) +
-  geom_col() +
-  labs(y = "Mean abundance (ind nmi^-1)") +
-  facet_wrap(~ pred_clust, ncol = 1) +
-  theme_classic(base_size = 14) +
-  theme(axis.title.x = element_blank())
 ```
@@
 
-station_ice <- predators_abundant %>%
-  distinct(amlr.station, ice_coverage)
-
-pred_clust <- tibble(
-  amlr.station = names(clust),
-  pred_clust = factor(clust,
-                      labels = c("Open", "Marginal", "Pagophilic"))
-) %>% 
-  left_join(select(zoop_sf,
-                   amlr.station,
-                   geometry,
-                   Year),
-            by = "amlr.station") %>% 
-  left_join(station_ice, by = "amlr.station")
-
-pred_clust %>% 
+stations_clust %>% 
   mutate(Year = factor(Year)) %>% 
-  crosstable::crosstable("Year", by = "pred_clust", total = "col")
+  crosstable::crosstable("Year", by = "pred_clust", total = "col") %>% 
+  crosstable::as_flextable()
 
@@
 
-pred_clust_sf <- pred_clust %>% 
-  st_as_sf() %>% 
-  st_transform(ant_proj())
-
-map_lim <- st_bbox(pred_clust_sf) %>% 
+map_lim <- st_bbox(stations_clust) %>% 
   project_bbox() %>% 
@@
 
-icy_open <- filter(pred_clust_sf,
+icy_open <- filter(stations_clust,
                    pred_clust == "Open",
@@
   geom_sf(aes(color = pred_clust), 
-          pred_clust_sf, 
+          stations_clust, 
           size = 2, 
           shape = 15) +
+  # geom_sf(data = filter(stations_clust, 
+  #                       amlr.station %in% c("AMLR2015_W1111",
+  #                                           "AMLR2015_W1113")),
+  #         color = "yellow") +
   geom_sf(data = seaice_contours, color = "skyblue") +
@@
 
-nmds_stress <- tibble(
-  k = 1:6,
-  stress = map_dbl(k, function(k) {
-    metaMDS(
-      sightings_mtx, 
-      distance = "bray",
-      k = k,
-      trymax = 100,
-      trace = FALSE
-    )$stress
-  })
-)
-
 ggplot(nmds_stress, aes(k, stress)) +
@@
 
-sightings_nmds <- metaMDS(
-  sightings_mtx, 
-  distance = "bray",
-  k = 3,
-  trymax = 100,
-  trace = FALSE
-)
-
-nmds_isoreg <- isoreg(sightings_nmds$diss, sightings_nmds$dist)
-
-nmds_shepard <- tibble(
-  diss = sightings_nmds$diss,
-  dist = sightings_nmds$dist,
-  mono = nmds_isoreg$yf
-)
-
 nmds_spear <- cor(nmds_shepard$diss,
@@
 nmds_label <- str_glue(
-  "Non-metric fit, R^2 = {round(1 - sightings_nmds$stress^2, 3)}\n",
+  "Non-metric fit, R^2 = {round(1 - nmds_sightings$stress^2, 3)}\n",
   "Linear fit, R^2 = {round(nmds_spear^2, 3)}\n",
-  "k = {sightings_nmds$ndim}\n",
-  "Stress = {round(sightings_nmds$stress, 3)}"
+  "k = {nmds_sightings$ndim}\n",
+  "Stress = {round(nmds_sightings$stress, 3)}"
 )
@@
 #| label: loading
-
-nmds_df <- sightings_nmds$points %>% 
-  as_tibble(rownames = "amlr.station") %>% 
-  rename_with(~ paste0("N", .x), starts_with("MDS")) %>% 
-  right_join(pred_clust_sf %>% 
-               as_tibble() %>% 
-               select(amlr.station, pred_clust),
-             by = "amlr.station")
 
@@
 
+```
+
+```{r}
+#| label: env-loadings-old
+#| eval: false
+
 station_ice <- predators_stations %>% 
@@
   as_tibble() %>% 
-  left_join(station_ice, by = "amlr.station") %>% 
+  left_join(stations_ice, by = "amlr.station") %>% 
   mutate(Year = factor(Year)) %>% 
@@
 # Fit multiple regression model
-pred_envfit <- envfit(sightings_nmds, 
+pred_envfit <- envfit(nmds_sightings, 
                       station_env, 
@@
 #| label: env_summ
+#| eval: false
 
@@
     by = "amlr.station") %>% 
-  left_join(select(station_ice, amlr.station, ice_coverage),
+  left_join(select(stations_ice, amlr.station, ice_coverage),
             by = "amlr.station") %>% 
@@
 ```{r}
+#| eval: false
 pred_clust_ice <- tibble(
@@
 ) %>% 
-  left_join(station_ice,
+  left_join(stations_ice,
             by = "amlr.station") %>% 
@@
 ```{r}
+#| eval: false
 pred_prey <- as_tibble(pred_clust) %>% 

### 2023-08-10 R/cluster.R

@@ -48,17 +48,7 @@ assign_sightings <- function(sightings, stations, max_dist_km) {
     st_as_sf()
 }
 
-filter_sightings <- function(sightings, stations) {
-  unk_sp <- c("UNSE", "UNPR", "UGPT", "UNPN")
-  total_stations <- n_distinct(stations$amlr.station)
-  retain <- sightings %>%
-    group_by(species) %>%
-    summarize(stations_present = n()) %>%
-    mutate(station_frac = stations_present / total_stations) %>%
-    filter(!species %in% unk_sp)
-  semi_join(sightings, retain, by = "species")
-}
-
+# Normalize species counts by survey effort
 normalize_counts <- function(sightings, effort, stations) {
   sightings %>%
     as_tibble() %>%
@@ -70,6 +60,7 @@ normalize_counts <- function(sightings, effort, stations) {
     left_join(select(stations, amlr.station, Year), by = "amlr.station")
 }
 
+# Convert sightings data frame to matrix
 sightings_to_matrix <- function(sightings) {
   sightings_wide <- sightings %>%
     as_tibble() %>%
@@ -81,4 +72,39 @@ sightings_to_matrix <- function(sightings) {
   sightings_mtx <- as.matrix(select(sightings_wide, -amlr.station))
   row.names(sightings_mtx) <- sightings_wide$amlr.station
   sightings_mtx
+}
+
+# Generate gap statistic curve for choosing k
+cluster_gap <- function(mtx, max_k, seed = 139) {
+  cluster_fn <- function(x, k) {
+    list(cluster = stats::cutree(hclust(vegan::vegdist(x, method = "bray"),
+                                        method = "ward.D2"),
+                                 k = k))
+  }
+  set.seed(seed)
+  gap <- suppressWarnings(
+    cluster::clusGap(mtx,
+                     FUN = cluster_fn,
+                     K.max = 10,
+                     B = 500,
+                     verbose = FALSE)
+  )
+  as_tibble(gap$Tab) %>%
+    mutate(
+      k = seq_along(logW),
+      is_optimum = k == cluster::maxSE(gap, SE.sim, method = "Tibs2001SEmax")
+    )
+}
+
+# Create a predator cluster factor from a vector of stations and the cut tree
+station_to_cluster <- function(station, clust) {
+  clust_int <- clust[station]
+  factor(clust_int, levels = 1:3, labels = c("Open", "Marginal", "Pagophilic"))
+}
+
[…]

### 2023-08-10 R/nmds.R

@@ -1,0 +1,47 @@
+# Calculate NMDS stress across numbers of axes
+stress <- function(mtx, k) {
+  tibble(
+    k = k,
+    stress = map_dbl(k, \(k) {
+      vegan::metaMDS(
+        mtx,
+        distance = "bray",
+        k = k,
+        trymax = 100,
+        trace = FALSE
+      )$stress
+    })
+  )
+}
+
+# Create shepard diagram data
+shepard <- function(nmds) {
+  nmds_isoreg <- isoreg(nmds$diss, nmds$dist)
+  tibble(
+    diss = nmds$diss,
+    dist = nmds$dist,
+    mono = nmds_isoreg$yf
+  )
+}
+
+# Convert NMDS object to a data frame, including cluster and environmental data
+nmds_to_df <- function(nmds, clust, env) {
+  nmds$points %>%
+    as_tibble(rownames = "amlr.station") %>%
+    rename_with(\(axis) paste0("N", axis), starts_with("MDS")) %>%
+    mutate(pred_clust = station_to_cluster(amlr.station, clust)) %>%
+    left_join(env, by = "amlr.station")
+}
+
+# Combine ice and station environmental data, filtering down to stations used
+# in this analysis
+env_data <- function(stations, ice, predators) {
+  left_join(
+    select(stations, amlr.station, zuml_m, avg.temp, avg.salinity,
+           Integ.chla.100m, Integ.phae.100m, TOD_2levels_civil, Year,
+           zoop_clust = `Winter Cluster factor`),
+    select(ice, amlr.station, ice_type, ice_coverage),
+    by = "amlr.station"
+  ) %>%
+    semi_join(predators, by = "amlr.station")
+}

### 2023-08-10 _targets.R

@@ -109,16 +109,45 @@ list(
     predators_stations,
     assign_sightings(predators_sf, zoop_sf, max_dist_km = 15) %>%
       normalize_counts(station_effort, zoop_sf) %>%
-      left_join(stations_ice, by = "amlr.station")
-  ),
-  tar_target(
-    predators_abundant,
-    filter_sightings(predators_stations, zoop_sf)
+      left_join(stations_ice, by = "amlr.station") %>%
+      filter(!species %in% c("UNSE", "UNPR", "UGPT", "UNPN"))
   ),
   # Clustering
-  tar_target(sightings_mtx, sightings_to_matrix(predators_abundant)),
+  tar_target(sightings_mtx, sightings_to_matrix(predators_stations)),
   tar_target(sightings_dist, vegan::vegdist(sightings_mtx, method = "bray")),
   tar_target(sightings_clust, hclust(sightings_dist, method = "ward.D2")),
+  tar_target(sightings_gap, cluster_gap(sightings_mtx, max_k = 10, seed = 139)),
+  tar_target(sightings_cut, cutree(sightings_clust, k = 3)),
+  tar_target(sightings_indval, labdsv::indval(sightings_mtx, sightings_cut)),
+  tar_target(
+    predators_clust,
+    mutate(predators_stations,
+           pred_clust = station_to_cluster(amlr.station, sightings_cut))
+  ),
+  tar_target(
+    stations_clust,
+    zoop_sf %>%
+      left_join(stations_ice, by = "amlr.station") %>%
+      mutate(pred_clust = station_to_cluster(amlr.station, sightings_cut)) %>%
+      semi_join(predators_clust, by = "amlr.station")
+  ),
+  # NMDS
+  tar_target(nmds_stress, stress(sightings_mtx, k = 1:6)),
+  tar_target(nmds_sightings,
+             vegan::metaMDS(
+               sightings_mtx,
+               distance = "bray",
+               k = 3,
+               trymax = 100,
+               trace = FALSE)),
+  tar_target(nmds_shepard, shepard(nmds_sightings)),
+  tar_target(nmds_env, env_data(zoop, stations_ice, predators_stations)),
+  tar_target(nmds_df, nmds_to_df(nmds_sightings, sightings_cut, nmds_env)),
+  tar_target(nmds_envfit, vegan::envfit(nmds_sightings,
+                                        select(nmds_env, -amlr.station),
+                                        permutations = 999,
+                                        choices = 1:3,
[…]

### 2023-08-17 reports/01_predators.qmd
Major overhaul of associating predator sighitngs with stations

@@
   group_by(species) %>% 
-  summarize(total_count = sum(count_norm)) %>% 
+  summarize(total_count = sum(count)) %>% 
   arrange(desc(total_count)) %>%
@@
   group_by(year, species) %>% 
-  summarize(total = sum(count_norm), .groups = "drop") %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
   mutate(species = fct_reorder(species, total, .fun = sum),

### 2023-08-17 reports/03_combined.qmd
Major overhaul of associating predator sighitngs with stations

--- 2023-03-26 reports/03_combined.qmd
@@
   scale_color_fermenter(palette = "RdYlBu",
-                        breaks = c(1, 2, 3, 5, 8)) +
+                        breaks = c(10, 20, 100, 200)) +
   coord_fixed() +
-  facet_wrap(~ ice_cat, ncol = 2) +
+  facet_wrap(~ ice_type, ncol = 2) +
   theme_classic()
@@
 ant_basemap(map_lim) +
-  geom_sf(aes(color = ice_cat), 
-          predators_stations,
+  geom_sf(aes(color = ice_type), 
+          filter(predators_stations, year != 2014),
           size = 0.5) +
+  scale_color_brewer("Ice category", palette = "YlGnBu") +
+  guides(color = guide_legend(override.aes = list(size = 5))) +
   facet_wrap(~ year) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "left")
 

### 2023-08-17 reports/04_pred_clus.qmd
Major overhaul of associating predator sighitngs with stations

@@
 
-library(cluster)
-library(ggmosaic)
 library(ggtree)
@@
 
-tar_load(
-  c("predators_clust", "stations_clust", "sightings_cut",
-    "seaice_conc_df", "seaice_contours", "stations_ice",
-    starts_with("sightings"),
-    starts_with("nmds"),
-    "zoop_sf", "ant_sf"), 
-  store = here::here("_targets")
-)
+tar_load_everything(store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 Heuristically, the shoulder is the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$.
-According to this rule, the predators are best described by one cluster.
-However, there's a clear jump from $k=2$ to $k=3$ (red), followed by a leveling off.
-Therefore, based on our prior knowledge of three predator guilds (pagophilic, open water, and marginal ice), we'll present the results for $k=3$ here.
+According to this rule, the predators are best described by three clusters.
 
@@
 
-Which species best represent each cluster? Using Dufrene-Legendre indicator species analysis. Apply indicator value cutoff at 25%.
+Which species best represent each cluster?
+Using Dufrene-Legendre indicator species analysis.
+Indicator values, $d$, presented for each cluster for $d\gt0.25$.
+$d$ is the product of relative frequency (fraction of sites present within cluster) and relative abundance (fraction of abundance found within cluster).
 
@@
 
--   Cluster 1 = open water
--   Cluster 2 = marginal ice zone
--   Cluster 3 = pagophilic
+-   Cluster 1 = uncertain
+-   Cluster 2 = pagophilic
+-   Cluster 3 = open water
 
-Indicator value $d_{ic}$ of species $i$ in cluster $c$ is the product of relative fidelity, $f_{ic}$, and relative abundance, $a_{ic}$. See documentation of `labdsv::indval()`.
+Relative to cluster 3, cluster 2 was composed of more ice associated species.
+The strongest indicator for cluster 2 was the Antarctic fur seal, which was sighted at 98.6% of cluster 2 sites, for 87% of the species' total abundance.
+Pagophilic predators such as crabeater seals, leopard seals, and Adélie penguins were also indicator species for this cluster.
+Another pagophilic predator, the minke whale, was too rare overall to be significant indicator species for any cluster, but was sighted most frequently in cluster 2 (12.2% of sites versus 5.2% and 4.4% of clusters 1 and 3, respectively) and 80.0% of its abundance was in cluster 2.
 
-Relative fidelity $f_{ic}$ of species $i$ in cluster $c$ is:
+Cluster 3 describes a more open water guild.
+Southern fulmars and cape petrels were significant indicator species for this cluster.
+Blue petrels, an open water predator, were most frequently observed in cluster 3 (36.3% of sites, relative to 6.9% and 10.8% of sites in clusters 1 and 2) in greater numbers than elsewhere (66.0% of total abundance).
 
-$$
-f_{ic} = \frac{\sum_{j \in c}p_{ij}}{n_c}
-$$
+Snow petrels, Antarctic petrels, and southern giant petrels were significant indicator species for both clusters 2 and 3.
+These species were the most widespread predators overall, sighted at 59.1%, 53.0%, and 44.9% of stations.
+Snow petrels were also the most abundant predator in the survey, accounting for 34.1% of total individuals observed.
 
-Relative abundance $a_{ic}$ is:
-
-$$
-a_{ic}=\frac{\sum_{j \in c}x_{ij}/n_c}{\sum_{k=1}^K (\sum_{j \in k}x_{ij}/n_k)}
-$$
-
-A high indicator value, e.g., $d_{ic}\approx1$, means species $i$ is present in nearly all sites in cluster $c$ and nearly absent from sites not in $c$.
-The greatest $d_{ic}$ observed here was Adèlie penguins in cluster 3 at 0.74.
-ADPN were observed in 45 of the 60 cluster 3 sites, for $f_{ic}=0.75$.
-The average abundance of ADPN in cluster 1 was 4.04 ind nm\^-1, compared to 0.113 and 0.166 in clusters 2 and 3, respectively.
-Yields $a_{ic}=0.935$.
-
-Note that differences in indicator values between species reflect *relative* changes in frequency/abundance.
-For example, in cluster 3 (pagophilic predators), Antarctic fur seals were found at more stations ($f_{ic}=0.783$) and in higher abundances (mean 2.22 ind nm\^-1) than crabeater seals ($f_{ic}=0.650$, mean 1.73 ind nm\^-1), yet they have the same indicator value $d_{ic}=0.63$.
-That's because Antarctic fur seals were still somewhat abundant in another cluster whereas crabeater seals were concentrated in cluster 3.
-Specifically, mean Antarctic fur seal abundance was 0.441 ind nm\^-1 in cluster 2 (marginal ice zone), 19.9% of their cluster 1 abundance.
-Crabeater seals dropped from 1.73 ind nm\^-1 to 0.133 ind nm\^-1, or 7.7% of cluster 1 abundance.
-
-Comparing the highest indicator values between clusters, the pagophilic community has the most distinct collection of predators, followed by the marginal ice zone and open water communities.
-Three species had $d_{ic} \gt 0.6$ in the pagophilic cluster: Adèlie penguins, Antarctic fur seals, and crabeater seals.
-All three of these species were both more prevalent and more abundant in the pagophilic cluster than either of the other two.
-In contrast, the open water community had no species with $d_{ic} \gt 0.35$.
-The three most prevalent species in the open water cluster were snow petrels, Antarctic petrels, and southern giant petrels, all of which were more abundant in the marginal ice zone cluster.
-Meaning, the pagophilic community was distinct because of specific species (Adèlie penguins, Antarctic fur seals, and crabeater seals), while the open water community was characterized by more variable species compositions.
-The marginal ice zone community fell in between.
-Its highest indicator value species (snow petrel, Antarctic petrel, and southern giant petrel) were widespread across communities, so their relative abundances in the marginal ice zone cluster were lower ($a_{ic} \in [0.400 - 0.652]$) than the three highest indicator value species were in the pagophilic cluster ($a_{ic} \in [0.810 - 0.971]$).
-
-```{r}
-#| label: indval_interpret
-
-# Adelie penguin x cluster 3 worked example
-adpn <- filter(predators_clust, species == "ADPN")
-
-pic_adpn <- sum(adpn$pred_clust == "Pagophilic")
-nc3 <- sum(sightings_cut == 3)
-fic_adpn <- pic_adpn / nc3
-
-abu_adpn <- adpn %>% 
-  group_by(pred_clust) %>% 
-  summarize(abu = sum(count_nmi) / n())
-aic_adpn <- abu_adpn$abu[3] / sum(abu_adpn$abu)
-
-# ANFU x cluster 1
-anfu <- filter(predators_clust, species == "ANFU")
-
-pic_anfu <- sum(anfu$pred_clust == "Open")
-nc1 <- sum(sightings_cut == 1)
-fic_anfu <- pic_anfu / nc1
-
-abu_anfu <- anfu %>% 
-  group_by(pred_clust) %>% 
-  summarize(abu = sum(log(count_nmi + 1)) / n())
-aic_anfu <- abu_anfu$abu[1] / sum(abu_anfu$abu)
-
-```
+No species were identified as significant indicators for cluster 1.
+Species observed at \>25% of cluster 1 sites include Adélie penguins, crabeater seals, and Antarctic fur seals, suggesting a pack ice community.
+In terms of relative abundance, 59.8% of Weddell seals, 43.6% of Adélie penguins, and 20.2% of crabeater seals were sighted in cluster 1.
+If this cluster represents the pack ice community, the lack of significant indicator species may be attributable to difficulty surveying the pack ice?
 
@@
 
-How frequent are the different clusters?
-The marginal ice zone predator community was observed most often, followed by the open water then the pagophilic communities.
-The pagophilic community was observed most frequently in 2016.
-Sea ice extent was less in 2016 than 2013 and 2015, so the relative frequency of the pagophilic community in 2016 is initially surprising.
-However, it may be explained by the greater sea ice in 2013 and 2015 limiting survey access to the Bransfield Strait.
+How frequent are the different clusters and where were they found?
+The open water community was observed most often, followed by the marginal ice zone then the uncertain communities.
+The relative frequency of the open water community was greatest in 2012, when survey effort in the Bransfield Strait was most limited, and least in 2016.
+Generally, the open water community occupied the offshore regions, but in 2014 and 2015 it moved onto the shelf into the triangle between Elephant, Joinville, and George Islands.
+
+The marginal ice zone community occupied the shallower, western part of the Bransfield Strait, moving offshore of the Shetland Islands in 2014 and north of Elephant Island in 2016.
+
+The uncertain community was most commonly observed in the south eastern part of the survey region, near Joinville Island.
+A particularly coherent group of sites in this community formed in 2013 between Joinville Island and the South Shetlands.
 
@@
 
-What geospatial patterns do we see in the predator communities?
-The pagophilic predator community (purple) was largely found within the Bransfield Strait.
-The marginal ice zone community was mostly distributed around the South Shetland Islands, however in the years 2013 and 2015 (when sea ice extent was greatest) this community was more frequent further offshore.
-The open water community was found furthest off shore (north in 2012 and 2014, west in 2016), with the exception of 2015 when it moved south to the entrance of the Bransfield Strait, even moving within the Strait.
-Is this the result of polynyas?
-Blue line indicates 65% ice concentration contour during the month of August in each year.
-Notice the limited coverage of the Bransfield Strait in 2012 and 2013, and the absence of eastern stations in 2016.
-
 ```{r}
 #| label: cluster_maps
+#| warning: false
 
@@
 
-icy_open <- filter(stations_clust,
-                   pred_clust == "Open",
-                   ice_coverage >= 7.5)
-
 ant_basemap(map_lim) +
-  geom_sf(aes(color = pred_clust), 
+  # Sea ice
+  geom_tile(aes(x, y, fill = seaice_conc), 
+            filter(seaice_conc_df, seaice_conc >= 0.65)) +
+  scale_fill_distiller(lim = c(0.65, 1),
+                       palette = "Blues",
+                       na.value = "transparent") +
+  geom_sf(data = seaice_contours, color = "darkblue") +
+  # Predator clusters
+  ggnewscale::new_scale_fill() +
+  geom_sf(aes(fill = pred_clust), 
           stations_clust, 
           size = 2, 
-          shape = 15) +
-  # geom_sf(data = filter(stations_clust, 
-  #                       amlr.station %in% c("AMLR2015_W1111",
-  #                                           "AMLR2015_W1113")),
-  #         color = "yellow") +
-  geom_sf(data = seaice_contours, color = "skyblue") +
+          shape = 22) +
   facet_wrap(~ Year) +
@@
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-  scale_color_brewer(palette = "Dark2") +
-  scale_fill_gradient(lim = c(0, 1), na.value = "transparent") +
+  scale_fill_brewer(palette = "Dark2") +
+  guides(fill = guide_legend(override.aes = list(size = 3))) +
   coord_ant(map_lim) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
 ```
@@
 
-How many NMDS axes should we use? Scree plot
+### NMDS diagnostics
+
+How many NMDS axes should we use? Scree plot indicates 3 axes is acceptable.
 
@@
 
-3 axes for NMDS looks acceptable. How does ordination distance compare to dissimilarity (Shepard diagram)?
+Shepard diagram demonstrates the ordination distance and observed dissimilarity are acceptably correlated.
 
@@
 
-NMDS plots, color-coded by predator cluster.
+### Environmental loading
+
+Only two environmental variables were significantly correlated with the NMDS axes: sampling year and ice coverage, with $r^2$ values of 0.061 and 0.053, respectively.
+Compare this to the macrozooplankton analysis, where 8 environmental values were significantly correlated with the NMDS axes and $r^2$ values were as high as 0.36 (chl *a* in the upper 100 m) and 0.703 (upper mixed layer salinity).
+
+```{r}
+#| label: envfit
+
+nmds_envfit
+
+```
+
+NMDS plots, color-coded by predator cluster with year and ice coverage loadings.
 
@@
 
-ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust)) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic()
-
-ggplot(nmds_df, aes(NMDS1, NMDS3)) +
-  geom_point(aes(color = pred_clust)) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic()
-
-ggplot(nmds_df, aes(NMDS2, NMDS3)) +
-  geom_point(aes(color = pred_clust)) +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic()
-```
-
-## Environmental loadings
-
-I compared the drivers of community structure two ways: physics (hydrology and ice coverage) and zooplankton (community assignments). Here's what makes jumps out at me:
-
--   Predator cluster associations with ice kind of track with what we expected.
-The pagophilic predator cluster correlates with first-year ice, open ocean predators with open ice, marginal ice predators with thin ice.
-Weirdly, multi-year ice falls among the open ocean predators, but only two stations were classified as multi-year ice so it's probably a small sample size thing.
--   Predator cluster 1 (open water) and 2 (marginal ice zone) are associated with higher temperatures and zooplankton cluster 2 (small euphausiids, amphipods, etc).
-They differ in the open water predator cluster was found in areas with greater phaeopigment and the marginal ice cluster in areas with greater chlorophyll-a.
--   Predator cluster 3 (pagophilic) is associated with higher salinity, deeper mixed layer, and macrozooplankton cluster 3 (big euphausiids, Metridia, etc)
--   Macrozooplankton cluster 1 (salps etc) evenly distributed across predator clusters
-
-```{r}
-#| label: env-loadings
-
-```
-
-```{r}
-#| label: env-loadings-old
-#| eval: false
-
-station_ice <- predators_stations %>% 
-  distinct(amlr.station, ice_type, ice_coverage) %>% 
-  as_tibble()
-
-# Evironmental variables
-station_env <- zoop_sf %>% 
-  select(amlr.station, zuml_m, avg.temp, avg.salinity, Integ.chla.100m, 
-         Integ.phae.100m, TOD_2levels_civil, Year, 
-         zoop = `Winter Cluster factor`) %>% 
-  as_tibble() %>% 
-  left_join(stations_ice, by = "amlr.station") %>% 
-  mutate(Year = factor(Year)) %>% 
-  slice(map_int(rownames(sightings_mtx),
-                ~ which(.x == amlr.station))) %>% 
-  select(-amlr.station, -geometry)
-
-# Fit multiple regression model
-pred_envfit <- envfit(nmds_sightings, 
-                      station_env, 
-                      permutations = 999, 
-                      choices = 1:3,
-                      na.rm = TRUE)
-
-# Extract envfit vectors and factors
-envfit_vecs <- scores(pred_envfit, 
+ice_vector <- scores(nmds_envfit, 
                      display = "vectors",
@@
   mutate(across(starts_with("NMDS"),
-                ~ .x * ordiArrowMul(pred_envfit)))
-envfit_facs <- scores(pred_envfit, 
+                ~ .x * ordiArrowMul(nmds_envfit))) %>% 
+  filter(envvar == "ice_coverage")
+
+year_centroids <- scores(nmds_envfit, 
                      display = "factors",
@@
   as_tibble(rownames = "envvar") %>% 
+  filter(str_starts(envvar, "Year")) %>% 
   mutate(across(starts_with("NMDS"),
-                ~ .x * ordiArrowMul(pred_envfit)))
+                ~ .x * ordiArrowMul(nmds_envfit)),
+         Year = substr(envvar, 5, 8))
 
-# Hydrology plot
-hydro_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust)) +
-  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
-               data = envfit_vecs,
-               linewidth = 1,
-               color = "grey50",
-               alpha = 0.6) +
-  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
-                           data = envfit_vecs) +
+ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(data = year_centroids, 
+             shape = 18,
+             size = 2.5,
+             color = "black",
+             alpha = 0.85) +
+  ggrepel::geom_text_repel(aes(label = Year),
+                           year_centroids) +
+  geom_segment(aes(xend = NMDS1, yend = NMDS2),
+               ice_vector, 
+               x = 0, y = 0,
+               size = 1.5,
+               color = "black",
+               alpha = 0.65,
+               arrow = arrow(angle = 20, length = unit(15, "points"))) +
+  ggrepel::geom_text_repel(data = ice_vector,
+                           label = "Ice coverage") +
+  labs(color = "Predator cluster") +
   scale_color_brewer(palette = "Dark2") +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "bottom")
 
-# Ice plot
-ice_facs <- filter(envfit_facs, str_detect(envvar, "ice_type"))
-ice_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust)) +
-  geom_point(aes(x = NMDS1, y = NMDS2),
-             data = ice_facs,
-             size = 2) +
-  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
-                           data = ice_facs) +
+ggplot(nmds_df, aes(NMDS1, NMDS3)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(data = year_centroids, 
+             shape = 18,
+             size = 2.5,
+             color = "black",
+             alpha = 0.85) +
+  ggrepel::geom_text_repel(aes(label = Year),
+                           year_centroids) +
+  geom_segment(aes(xend = NMDS1, yend = NMDS3),
+               ice_vector, 
+               x = 0, y = 0,
+               size = 1.5,
+               color = "black",
+               alpha = 0.65,
+               arrow = arrow(angle = 20, length = unit(15, "points"))) +
+  ggrepel::geom_text_repel(data = ice_vector,
+                           label = "Ice coverage") +
+  labs(color = "Predator cluster") +
   scale_color_brewer(palette = "Dark2") +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "bottom")
 
-# Macrozooplankton plot
-zoop_facs <- filter(envfit_facs, str_detect(envvar, "zoop"))
-zoop_plot <- ggplot(nmds_df, aes(NMDS1, NMDS2)) +
-  geom_point(aes(color = pred_clust)) +
-  geom_point(aes(x = NMDS1, y = NMDS2),
-             data = zoop_facs,
-             size = 2) +
-  ggrepel::geom_text_repel(aes(x = NMDS1, y = NMDS2, label = envvar),
-                           data = zoop_facs) +
+ggplot(nmds_df, aes(NMDS2, NMDS3)) +
+  geom_point(aes(color = pred_clust), alpha = 0.8) +
+  geom_point(data = year_centroids, 
+             shape = 18,
+             size = 2.5,
+             color = "black",
+             alpha = 0.85) +
+  ggrepel::geom_text_repel(aes(label = Year),
+                           year_centroids) +
+  geom_segment(aes(xend = NMDS2, yend = NMDS3),
+               ice_vector, 
+               x = 0, y = 0,
+               size = 1.5,
+               color = "black",
+               alpha = 0.65,
+               arrow = arrow(angle = 20, length = unit(15, "points"))) +
+  ggrepel::geom_text_repel(data = ice_vector,
+                           label = "Ice coverage") +
+  labs(color = "Predator cluster") +
   scale_color_brewer(palette = "Dark2") +
-  theme_classic()
-
-layout <- "
-AAAA
-AAAA
-BBCC
-"
-hydro_plot + ice_plot + zoop_plot + 
-  plot_layout(design = layout, guides = "collect")
-```
-
-Summaries of environmental conditions by cluster.
-
-```{r}
-#| label: env_summ
-#| eval: false
-
-env <- tibble(
-  amlr.station = names(clust),
-  pred_clust = factor(clust,
-                      labels = c("Open", "Marginal", "Pagophilic"))
-) %>% 
-  left_join(
-    select(as_tibble(zoop_sf),
-           amlr.station, avg.salinity, Integ.chla.100m),
-    by = "amlr.station") %>% 
-  left_join(select(stations_ice, amlr.station, ice_coverage),
-            by = "amlr.station") %>% 
-  pivot_longer(avg.salinity:ice_coverage, 
-               names_to = "envvar",
-               values_to = "envval")
-env %>% 
-  group_by(envvar) %>% 
-  summarize(envaov = list(aov(envval ~ pred_clust))) %>% 
-  mutate(Fval = map_dbl(envaov, \(aov) summary(aov)[[1]]$`F value`[1]),
-         pval = map_dbl(envaov, \(aov) summary(aov)[[1]]$`Pr(>F)`[1]))
-env %>% 
-  group_by(pred_clust, envvar) %>% 
-  summarize(across(envval, list(mean = partial(mean, na.rm = TRUE),
-                                sd = partial(sd, na.rm = TRUE))),
-            .groups = "drop")
+  theme_classic() +
+  theme(legend.position = "bottom")
 ```
@@
 
-I asserted the clusters corresponded to open water, marginal ice, and pack ice communities. Let's verify that.
+Median ice coverage increased from the open water cluster (median 33.0%) to the marginal ice (59.6%) and uncertain (70.6%) clusters.
+This supports the labels of "open water" and "marginal ice" for clusters 3 and 2, respectively.
+Also provides evidence for the "pack ice" cluster.
 
 ```{r}
-#| eval: false
-pred_clust_ice <- tibble(
-  amlr.station = names(clust),
-  pred_clust = factor(clust,
-                      labels = c("Open", "Marginal", "Pagophilic"))
-) %>% 
-  left_join(stations_ice,
-            by = "amlr.station") %>% 
-  drop_na(ice_coverage)
+#| label: ice_clusters
 
-ggplot(pred_clust_ice, aes(x = ice_coverage / 10, fill = after_stat(x))) + 
+median_coverge <- stations_clust %>% 
+  drop_na(ice_coverage) %>% 
+  group_by(pred_clust) %>% 
+  summarize(median_ice = median(ice_coverage / 10, na.rm = TRUE))
+
+ggplot(stations_clust, 
+       aes(x = ice_coverage / 10, fill = after_stat(x))) + 
   geom_histogram(bins = 20, color = "grey30") +
+  geom_vline(aes(xintercept = median_ice),
+             median_coverge,
+             color = "red", linetype = "dashed") +
   scale_x_continuous(labels = scales::percent) +
@@
   labs(x = "Ice coverage",
-       y = "Station count") +
+       y = "# sites") +
   theme_minimal()
@@
 
-Predator/prey cluster associations were statistically significant.
-The most positive residuals were between pagophilic predators x zooplankton 3b (4.07), pagophilic predators x zooplankton 3a (2.68), marginal ice zone predators x zooplankton 2a (2.33), and open water predators x zooplankton 2b (1.82).
-Zooplankton 1 was not
+Predator/prey cluster associations were statistically significant overall ($\chi = 53.6, p<0.001$) as were 5 pairwise associations after applying Bonferroni correction.
+The marginal ice and uncertain predator communities were positively associated with macrozooplankton communities 3b (large krill, including *E.
+superba*) and 3a (an extremely diverse assemblage associated with cold, high salinity water), respectively.
+The open water predator community was positively associated with cluster 2b (*Thysanoessa macrura*) and negatively associated with 3a and 3b.
+Macrozooplankton cluster 1 (a ubiquitous cluster indicated by *Salpa* and *Clione*) was not significantly associated with any predator cluster.
+Neither was macrozooplankton cluster 2a (indicated by important prey species such as *E.
+frigida*, *E.
+triacantha*, myctophid larvae, and *Themisto gaudichaudii*), but it associated most often with the open water predator cluster.
 
 ```{r}
-#| eval: false
-pred_prey <- as_tibble(pred_clust) %>% 
-  select(amlr.station, pred_clust) %>% 
-  left_join(select(as_tibble(zoop_sf), 
-                   amlr.station, 
-                   prey_clust = `Winter Cluster factor`),
-            by = "amlr.station")
-pred_prey_chisq <- with(pred_prey, chisq.test(pred_clust, prey_clust))
+#| label: pred_prey
+
+pred_prey_chisq <- with(stations_clust, 
+                        chisq.test(pred_clust, 
+                                   `Winter Cluster factor`,
+                                   simulate.p.value = TRUE))
 pred_prey_chisq
 pred_prey_chisq$residuals
+pred_prey_table <- with(stations_clust, table(pred_clust, 
+                                              `Winter Cluster factor`))
+chisq.posthoc.test::chisq.posthoc.test(pred_prey_table, 
+                                       simulate.p.value = TRUE) %>% 
+  pivot_longer(-c(Dimension, Value), 
+               names_to = "zoop_clust", 
+               values_to = "value") %>% 
+  pivot_wider(names_from = "Value") %>% 
+  arrange(`p values`) %>% 
+  filter(`p values` < 0.05)
 

### 2023-08-17 R/cluster.R

@@ -13,62 +13,18 @@
 # highest agglomerative coefficient
 # ...
 # Indicator values were calculated for each taxa cluster combination based on
-# formulas in Dufrene and Leg- endre (1997) and implemented in the R package
+# formulas in Dufrene and Legendre (1997) and implemented in the R package
 # labdsv
-
-assign_sightings <- function(sightings, stations, max_dist_km) {
-  stopifnot(inherits(sightings, "sf"),
-            inherits(stations, "sf"))
-
-  # Assign sightings to stations within year
-  assign_group <- function(sightings_group, sightings_key) {
-    # Note difference in column title Year vs year
-    stations_group <- stations[stations$Year == sightings_key$year, ]
-    # Find nearest station within year
-    nearest_station <- st_join(sightings_group,
-                               select(stations_group, amlr.station, date.tow),
-                               join = st_nearest_feature)
-    # Get station coordinates
-    station_coords <- stations_group %>%
-      slice(map_int(nearest_station$amlr.station,
-                    ~ which(stations_group$amlr.station == .x)))
-    # Calculate distance from predator sighting to nearest station
-    nearest_station %>%
-      mutate(dist_to_station = st_distance(nearest_station,
-                                           station_coords,
-                                           by_element = TRUE))
-  }
-
-  # Assign sightings across years
-  sightings %>%
-    group_by(year) %>%
-    group_modify(assign_group) %>%
-    ungroup() %>%
-    filter(dist_to_station <= units::as_units(max_dist_km, "km")) %>%
-    st_as_sf()
-}
-
-# Normalize species counts by survey effort
-normalize_counts <- function(sightings, effort, stations) {
-  sightings %>%
-    as_tibble() %>%
-    group_by(amlr.station, species) %>%
-    summarize(count = sum(count), .groups = "drop") %>%
-    left_join(select(as_tibble(effort), amlr.station, survey_nmi),
-              by = "amlr.station") %>%
-    mutate(count_nmi = count / survey_nmi) %>%
-    left_join(select(stations, amlr.station, Year), by = "amlr.station")
-}
 
 # Convert sightings data frame to matrix
 sightings_to_matrix <- function(sightings) {
   sightings_wide <- sightings %>%
     as_tibble() %>%
+    mutate(count_nmi_log = log1p(count_nmi)) %>%
     pivot_wider(id_cols = amlr.station,
[…]

### 2023-08-17 R/ice.R

@@ -79,14 +79,15 @@ collect_seaice <- function(seaice_dir) {
   proj_ext <- ext(-67, -50, -68, -55) %>%
     project(from = "epsg:4326",
             to = crs(seaice))
-  crop(seaice, proj_ext)
+  wrap(crop(seaice, proj_ext))
 }
 
-seaice2contour <- function(seaice_dir, contour_level) {
-  seaice_rast <- collect_seaice(seaice_dir)
+seaice2contour <- function(seaice_rast, contour_level) {
+  seaice_rast <- unwrap(seaice_rast)
   map(
     seq(nlyr(seaice_rast)),
-    ~ as.contour(seaice_rast[[.x]], levels = contour_level * 1000)
+    ~ as.contour(seaice_rast[[.x]],
+                 levels = contour_level * 1000)
   ) %>%
     vect() %>%
     mutate(Year = 2012:2016,
@@ -94,8 +95,8 @@ seaice2contour <- function(seaice_dir, contour_level) {
     wrap()
 }
 
-seaice2df <- function(seaice_dir) {
-  seaice_rast <- collect_seaice(seaice_dir)
+seaice2df <- function(seaice_rast) {
+  seaice_rast <- unwrap(seaice_rast)
   proj_ext <- ext(-67, -50, -68, -55) %>%
     project(from = "epsg:4326",
             to = crs(seaice_rast))

### 2023-08-17 R/nmds.R

@@ -1,11 +1,10 @@
 # Calculate NMDS stress across numbers of axes
-stress <- function(mtx, k) {
+stress <- function(dist, k) {
   tibble(
     k = k,
     stress = map_dbl(k, \(k) {
       vegan::metaMDS(
-        mtx,
-        distance = "bray",
+        dist,
         k = k,
         trymax = 100,
         trace = FALSE
@@ -43,5 +42,6 @@ env_data <- function(stations, ice, predators) {
     select(ice, amlr.station, ice_type, ice_coverage),
     by = "amlr.station"
   ) %>%
-    semi_join(predators, by = "amlr.station")
+    semi_join(predators, by = "amlr.station") %>%
+    mutate(Year = factor(Year))
 }

### 2023-08-17 R/predators.R

@@ -1,8 +1,58 @@
 aggregate_predators <- function(predators) {
   predators %>%
     filter(species != "NULL") %>%
-    mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
-    group_by(year, interval, species, lon_mean, lat_mean, nmi) %>%
-    summarize(count = sum(count_species), .groups = "drop") %>%
-    mutate(count_norm = count / nmi)
+    mutate(UTC_start = lubridate::mdy_hm(UTC_start),
+           year = lubridate::year(UTC_start)) %>%
+    group_by(year, interval, species, UTC_start, lon_mean, lat_mean, nmi) %>%
+    summarize(count = sum(count_species), .groups = "drop")
+}
+
+filter_species <- function(sightings, station_thr) {
+  total_stations <- n_distinct(sightings$amlr.station)
+  sightings %>%
+    group_by(species) %>%
+    summarize(n_stations = n_distinct(amlr.station)) %>%
+    filter(n_stations >= station_thr * total_stations,
+           !str_starts(species, "UN")) %>%
+    semi_join(x = sightings, y = ., by = "species")
+}
+
+assign_sightings <- function(sightings, stations, max_dist_km, max_days) {
+  stopifnot(inherits(sightings, "sf"),
+            inherits(stations, "sf"))
+
+  # Assign sightings to stations within year
+  assign_group <- function(sightings_group, sightings_key) {
+    station_buffers <- stations %>%
+      filter(Year == sightings_key$year) %>%
+      st_transform(ant_proj()) %>%
+      st_buffer(max_dist_km * 1000) %>%
+      transmute(
+        amlr.station,
+        tow.UTC = date.tow + (start.time.UTC - as.POSIXct("1899-12-31", tz = "UTC"))
+      )
+    sightings_group %>%
+      st_transform(ant_proj()) %>%
+      st_join(station_buffers, left = FALSE) %>%
+      mutate(lag_days = abs(as.numeric(UTC_start - tow.UTC, units = "secs")) / 3600 / 24) %>%
+      filter(lag_days <= max_days)
+  }
+
+  # Assign sightings across years
+  sightings %>%
+    group_by(year) %>%
+    group_modify(assign_group) %>%
+    ungroup() %>%
+    st_as_sf()
+}
+
+# Normalize species counts by survey effort
+normalize_counts <- function(sightings, effort) {
+  sightings %>%
+    as_tibble() %>%
[…]

### 2023-08-17 R/sp.R

@@ -43,7 +43,7 @@ ant_raster_template <- function(limits = ant_lims(), res = 5e4) {
 
 # By species and year
 rasterize_counts <- function(counts_sf,
-                             count_col = "count_norm",
+                             count_col = "count",
                              species_col = "species",
                              year_col = "year",
                              layer_name = count_col,

### 2023-08-17 _targets.R

@@ -44,13 +44,16 @@ list(
   tar_target(predators_sf,
              latlon_to_sf(predators_agg, coords = c("lon_mean", "lat_mean"))),
   # Effort
-  tar_target(effort,
-             predators %>%
-               filter(nmi < 1000) %>%
-               mutate(year = lubridate::year(lubridate::mdy_hm(UTC_start))) %>%
-               group_by(cruise, year, interval, lon_mean, lat_mean) %>%
-               summarize(nmi = sum(nmi), .groups = "drop") %>%
-               mutate(species = "survey", count_norm = 1)),
+  tar_target(
+    effort,
+    predators %>%
+      distinct(cruise, processID, interval, UTC_start, UTC_end, nmi,
+               lon_mean, lat_mean) %>%
+      filter(nmi < 1000) %>%
+      mutate(UTC_start = lubridate::mdy_hm(UTC_start),
+             year = lubridate::year(UTC_start),
+             species = "survey")
+  ),
   tar_target(effort_sf,
              latlon_to_sf(effort, coords = c("lon_mean", "lat_mean"))),
   # Zooplankton
@@ -92,12 +95,13 @@ list(
     here("data", "ice", "Aug sea ice concentration"),
     format = "file"
   ),
-  tar_target(seaice_contours, seaice2contour(seaice_dir, 0.65)),
-  tar_target(seaice_conc_df, seaice2df(seaice_dir)),
-  # Aggregate predators and ice observations
+  tar_target(seaice_rast, collect_seaice(seaice_dir)),
+  tar_target(seaice_contours, seaice2contour(seaice_rast, 0.65)),
+  tar_target(seaice_conc_df, seaice2df(seaice_rast)),
+  # Aggregate effort and ice observations
   tar_target(
     station_effort,
-    assign_sightings(effort_sf, zoop_sf, max_dist_km = 15) %>%
+    assign_sightings(effort_sf, zoop_sf, max_dist_km = 15, max_days = 1) %>%
       group_by(year, amlr.station) %>%
       summarize(survey_nmi = sum(nmi), .groups = "drop")
   ),
@@ -107,10 +111,9 @@ list(
   ),
   tar_target(
     predators_stations,
-    assign_sightings(predators_sf, zoop_sf, max_dist_km = 15) %>%
-      normalize_counts(station_effort, zoop_sf) %>%
-      left_join(stations_ice, by = "amlr.station") %>%
-      filter(!species %in% c("UNSE", "UNPR", "UGPT", "UNPN"))
[…]

### 2023-08-17 scratch/samplingeffort.R

@@ -1,0 +1,50 @@
+dat <- predators_clust %>%
+  group_by(amlr.station) %>%
+  summarize(n_sp = n_distinct(species),
+            effort = survey_nmi[1],
+            pred_clust = pred_clust[1])
+
+ggplot(dat, aes(effort, n_sp, color = pred_clust)) +
+  geom_point(alpha = 0.8, position = position_jitter(width = 2, height = 0.2)) +
+  scale_color_brewer(palette = "Dark2") +
+  labs(x = "Survey effort (nmi)", y = "# species") +
+  theme_classic() +
+  theme(legend.position = c(1, 0),
+        legend.justification = c(1.1, -0.1),
+        legend.title = element_blank())
+
+ggplot(dat, aes(pred_clust, effort, fill = pred_clust)) +
+  geom_boxplot() +
+  scale_fill_brewer(palette = "Dark2") +
+  labs(y = "Survey effort (nmi)") +
+  theme_classic() +
+  theme(axis.title.x = element_blank(),
+        legend.position = c(0, 1),
+        legend.justification = c(-0.1, 1),
+        legend.title = element_blank())
+
+dat %>%
+  group_by(pred_clust) %>%
+  summarize(median(effort))
+
+# dissimilarity ~ difference in survey effort
+bc_dist <- as.matrix(sightings_dist) %>%
+  as_tibble(rownames = "amlr.station1") %>%
+  pivot_longer(-amlr.station1, names_to = "amlr.station2", values_to = "dist")
+dat2 <- expand_grid(
+  select(dat,
+         amlr.station1 = amlr.station,
+         effort1 = effort),
+  select(dat,
+         amlr.station2 = amlr.station,
+         effort2 = effort)
+) %>%
+  mutate(diff_effort = abs(effort1 - effort2)) %>%
+  left_join(bc_dist, by = c("amlr.station1", "amlr.station2")) %>%
+  filter(amlr.station1 != amlr.station2)
+ggplot(dat2, aes(diff_effort, dist)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm", formula = y ~ x) +
+  labs(x = "Effort difference (nmi)",
+       y = "Bray-Curtis dissimilarity") +
+  theme_classic()

### 2023-08-18 reports/04_pred_clus.qmd
Cleaning up figures and tables

@@
 library(ggtree)
+library(gridExtra)
 library(patchwork)
@@
 tar_load_everything(store = here::here("_targets"))
+source(here::here("R", "predators.R"))
 source(here::here("R", "sp.R"))
@@
   ord <- order(x, decreasing = TRUE)
+  species <- code_to_common(rownames(indval_mtx))
   result <- ifelse(x >= 0.25, 
-         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
+         str_glue("{species} ({round(x, 2)})"),
          NA) %>% 
@@
 
--   Cluster 1 = uncertain
--   Cluster 2 = pagophilic
+-   Cluster 1 = pack ice
+-   Cluster 2 = marginal ice
 -   Cluster 3 = open water
@@
 How frequent are the different clusters and where were they found?
-The open water community was observed most often, followed by the marginal ice zone then the uncertain communities.
+The open water community was observed most often, followed by the marginal ice zone then the pack ice communities.
 The relative frequency of the open water community was greatest in 2012, when survey effort in the Bransfield Strait was most limited, and least in 2016.
@@
 
-The uncertain community was most commonly observed in the south eastern part of the survey region, near Joinville Island.
+The pack ice community was most commonly observed in the south eastern part of the survey region, near Joinville Island.
 A particularly coherent group of sites in this community formed in 2013 between Joinville Island and the South Shetlands.
@@
 
+# Indicator species grob
+indval_grob <- apply(sightings_indval$indval, 2, function(col) {
+  # Extract indicator values >0.25 from each cluster's column and format with
+  # species name
+  col <- sort(col[col > 0.25], decreasing = TRUE)
+  species <- code_to_common(names(col))
+  if (length(col) == 0) {
+    "None"
+  } else {
+    str_glue("{species} ({round(col, 2)})")
+  }
+}) %>% 
+  # Create a 
+  map(\(x) {
+    nrow = max(map_dbl(., length))
+    c(x, rep("", nrow - length(x)))
+  }) %>% 
+  as_tibble() %>% 
+  set_names(levels(stations_clust$pred_clust)) %>% 
+  tableGrob(
+    rows = NULL,
+    theme = ttheme_minimal(base_size = 6,
+                           padding = unit(c(1, 1), "mm"),
+                           core = list(fg_params = list(hjust = 0,
+                                                        x = 0)),
+                           colhead = list(fg_params = list(hjust = 0,
+                                                           x = 0)))
+  )
+
 ant_basemap(map_lim) +
@@
   geom_tile(aes(x, y, fill = seaice_conc), 
-            filter(seaice_conc_df, seaice_conc >= 0.65)) +
+            filter(seaice_conc_df, between(seaice_conc, 0.65, 1)),
+            alpha = 0.8) +
   scale_fill_distiller(lim = c(0.65, 1),
@@
                        na.value = "transparent") +
-  geom_sf(data = seaice_contours, color = "darkblue") +
+  # geom_sf(data = seaice_contours, color = "darkblue") +
   # Predator clusters
@@
   theme(legend.position = "bottom",
-        legend.title = element_blank())
+        legend.title = element_blank()) +
+  inset_element(indval_grob, 
+                0.67, 0, 1, 0.5, 
+                align_to = "panel")
 ```
@@
 #| label: loading
-
-ice_vector <- scores(nmds_envfit, 
-                     display = "vectors",
-                     choices = 1:3) %>% 
-  as_tibble(rownames = "envvar") %>% 
-  mutate(across(starts_with("NMDS"),
-                ~ .x * ordiArrowMul(nmds_envfit))) %>% 
-  filter(envvar == "ice_coverage")
+#| fig-width: 9
+#| fig-height: 5
 
@@
 
-ggplot(nmds_df, aes(NMDS1, NMDS2)) +
+nmds_plot <- function(axis1, axis2) {
+  ice_ordisurf <- ordisurf(
+    nmds_sightings ~ nmds_env$ice_coverage,
+    choices = c(axis1, axis2),
+    plot = FALSE
+  )
+
+  x <- paste0("NMDS", axis1)
+  y <- paste0("NMDS", axis2)
+
+  ice_df <- expand_grid(
+    axis1 = ice_ordisurf$grid$x,
+    axis2 = ice_ordisurf$grid$y,
+  ) %>% 
+    cbind(as.numeric(ice_ordisurf$grid$z)) %>% 
+    set_names(c(x, y, "ice_coverage")) %>% 
+    mutate(ice_coverage = ice_coverage / 10) %>% 
+    drop_na(ice_coverage)
+
+  ggplot(nmds_df, aes(.data[[x]], .data[[y]])) +
+    stat_contour(aes(z = .data$ice_coverage, 
+                     color = after_stat(level)),
+                 ice_df) +
+    scale_color_distiller(
+      "Ice coverage",
+      palette = "Blues",
+      labels = scales::percent,
+      limits = c(0.25, 0.75),
+      breaks = seq(0.25, 0.75, by = 0.1),
+      guide = guide_colorbar(barwidth = unit(2, "in"),
+                             direction = "horizontal")
+    ) +
+    ggnewscale::new_scale_color() +
     geom_point(aes(color = pred_clust), alpha = 0.8) +
@@
                              year_centroids) +
-  geom_segment(aes(xend = NMDS1, yend = NMDS2),
-               ice_vector, 
-               x = 0, y = 0,
-               size = 1.5,
-               color = "black",
-               alpha = 0.65,
-               arrow = arrow(angle = 20, length = unit(15, "points"))) +
-  ggrepel::geom_text_repel(data = ice_vector,
-                           label = "Ice coverage") +
+    scale_color_brewer(
+      palette = "Dark2",
+      guide = guide_legend(override.aes = list(size = 2),
+                           direction = "horizontal",
+                           order = 1)
+    ) +
     labs(color = "Predator cluster") +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic() +
-  theme(legend.position = "bottom")
+    theme_classic()
+}
 
-ggplot(nmds_df, aes(NMDS1, NMDS3)) +
-  geom_point(aes(color = pred_clust), alpha = 0.8) +
-  geom_point(data = year_centroids, 
-             shape = 18,
-             size = 2.5,
-             color = "black",
-             alpha = 0.85) +
-  ggrepel::geom_text_repel(aes(label = Year),
-                           year_centroids) +
-  geom_segment(aes(xend = NMDS1, yend = NMDS3),
-               ice_vector, 
-               x = 0, y = 0,
-               size = 1.5,
-               color = "black",
-               alpha = 0.65,
-               arrow = arrow(angle = 20, length = unit(15, "points"))) +
-  ggrepel::geom_text_repel(data = ice_vector,
-                           label = "Ice coverage") +
-  labs(color = "Predator cluster") +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-ggplot(nmds_df, aes(NMDS2, NMDS3)) +
-  geom_point(aes(color = pred_clust), alpha = 0.8) +
-  geom_point(data = year_centroids, 
-             shape = 18,
-             size = 2.5,
-             color = "black",
-             alpha = 0.85) +
-  ggrepel::geom_text_repel(aes(label = Year),
-                           year_centroids) +
-  geom_segment(aes(xend = NMDS2, yend = NMDS3),
-               ice_vector, 
-               x = 0, y = 0,
-               size = 1.5,
-               color = "black",
-               alpha = 0.65,
-               arrow = arrow(angle = 20, length = unit(15, "points"))) +
-  ggrepel::geom_text_repel(data = ice_vector,
-                           label = "Ice coverage") +
-  labs(color = "Predator cluster") +
-  scale_color_brewer(palette = "Dark2") +
-  theme_classic() +
-  theme(legend.position = "bottom")
+map2(c(1, 1, 2), c(2, 3, 3), nmds_plot) %>% 
+  reduce(`|`) +
+  plot_annotation(tag_levels = 'A') +
+  plot_layout(guides = "collect") &
+    theme(legend.position = "bottom",
+          legend.direction = "horizontal",
+          plot.tag = element_text(face = "bold"))
 ```
@@
 
-Median ice coverage increased from the open water cluster (median 33.0%) to the marginal ice (59.6%) and uncertain (70.6%) clusters.
+Median ice coverage increased from the open water cluster (median 33.0%) to the marginal ice (59.6%) and pack ice (70.6%) clusters.
 This supports the labels of "open water" and "marginal ice" for clusters 3 and 2, respectively.
@@
 Predator/prey cluster associations were statistically significant overall ($\chi = 53.6, p<0.001$) as were 5 pairwise associations after applying Bonferroni correction.
-The marginal ice and uncertain predator communities were positively associated with macrozooplankton communities 3b (large krill, including *E.
+The marginal ice and pack ice predator communities were positively associated with macrozooplankton communities 3b (large krill, including *E.
 superba*) and 3a (an extremely diverse assemblage associated with cold, high salinity water), respectively.
@@
 
-##
+```{r}
+#| label: contingency
+
+pred_prey_tbl <- with(stations_clust, 
+                      table(pred_clust, `Winter Cluster factor`))
+pred_prey_chisq$expected
+pred_prey_chisq$observed
+
+expand_grid(`Predator cluster` = levels(stations_clust$pred_clust),
+            `Zooplankton cluster` = unique(stations_clust$`Winter Cluster factor`)) %>% 
+  mutate(observed = map2_int(`Predator cluster`, 
+                             `Zooplankton cluster`,
+                             \(i, j) pred_prey_chisq$observed[i, j]),
+         expected = map2_dbl(`Predator cluster`, 
+                             `Zooplankton cluster`,
+                             \(i, j) pred_prey_chisq$expected[i, j])) %>% 
+  mutate(label = str_glue("{observed} ({round(expected, 1)})")) %>% 
+  pivot_wider(id_cols = `Predator cluster`,
+              names_from = `Zooplankton cluster`, 
+              values_from = label) %>% 
+  knitr::kable()
+```

### 2023-08-18 R/cluster.R

@@ -55,7 +55,7 @@ cluster_gap <- function(mtx, max_k, seed = 139) {
 # Create a predator cluster factor from a vector of stations and the cut tree
 station_to_cluster <- function(station, clust) {
   clust_int <- clust[station]
-  factor(clust_int, levels = 1:3, labels = c("Uncertain", "Marginal ice", "Open water"))
+  factor(clust_int, levels = 1:3, labels = c("Pack ice", "Marginal ice", "Open water"))
 }
 
 # Determine if a species is rare (<5% of stations)

### 2023-08-18 R/predators.R

@@ -56,3 +56,28 @@ normalize_counts <- function(sightings, effort) {
               by = "amlr.station") %>%
     mutate(count_nmi = count / survey_nmi)
 }
+
+code_to_common <- function(species_code) {
+  c(ADPN	= "Adélie penguin",
+    ANFU	= "Southern fulmar",
+    ANPT	= "Antarctic petrel",
+    ANSH	= "Antarctic shag",
+    ANTE	= "Antarctic tern",
+    BLPT	= "Blue petrel",
+    CAPT	= "Cape petrel",
+    CRSE	= "Crabeater seal",
+    ELSE	= "Elephant seal",
+    EMPN	= "Emperor penguin",
+    FUSE	= "Antarctic fur seal",
+    GEPN	= "Gentoo penguin",
+    KEGU	= "Kelp gull",
+    KIWH	= "Killer whale",
+    LESE	= "Leopard seal",
+    MIWH	= "Minke whale",
+    PFSB	= "Pale-faced sheathbill",
+    ROSE	= "Ross seal",
+    SBWH	= "Bottlenose whale",
+    SGPT	= "Southern giant petrel",
+    SNPT	= "Snow petrel",
+    WESE	= "Weddell seal")[species_code]
+}

### 2023-10-25 reports/01_predators.qmd
Clean up dist maps, add cluster env associations

@@
 library(targets)
+library(terra)
 library(tidyverse)
@@
 
-The five most abundant species (in ind nmi\^-1).
+The six most abundant species (in ind nmi\^-1).
 
@@
 
-top_five <- predators_agg %>% 
+top_six <- predators_agg %>% 
   group_by(species) %>% 
@@
   arrange(desc(total_count)) %>%
-  slice(1:5)
+  slice(1:6) %>% 
+  mutate(common_name = c("Snow petrel", 
+                         "Antarctic fur seal",
+                         "Adélie penguin",
+                         "Crabeater seal",
+                         "Antarctic petrel",
+                         "Southern fulmar"))
 
-top_five
+top_six
 ```
@@
 
-top_five_plot_data <- predators_agg %>% 
-  semi_join(top_five, by = "species") %>% 
+top_six_plot_data <- predators_agg %>% 
+  semi_join(top_six, by = "species") %>% 
   group_by(year, species) %>% 
@@
          year = factor(year, levels = 2016:2012))
-year_labels <- top_five_plot_data %>% 
+year_labels <- top_six_plot_data %>% 
   filter(species == "SNPT") %>% 
@@
 
-ggplot(top_five_plot_data, aes(species, total, group = year)) +
+ggplot(top_six_plot_data, aes(species, total, group = year)) +
   geom_col(color = "black", fill = "grey80") +
@@
 
-library(terra)
-# Mkae "top five" predator sightings spatially aware
-topfive_sf <- predators_agg %>% 
-  semi_join(top_five, by = "species") %>% 
+# Make "top six" predator sightings spatially aware
+topsix_sf <- predators_agg %>% 
+  semi_join(top_six, by = "species") %>% 
+  left_join(select(top_six, species, common_name), by = "species") %>% 
   latlon_to_sf(coords = c("lon_mean", "lat_mean"))
@@
 # Rasterize the predator sightings by year and species
-topfive_rast <- topfive_sf %>% 
+topsix_rast <- topsix_sf %>% 
   st_transform(ant_proj()) %>% 
-  mutate(year_species = interaction(year, species)) %>% 
+  mutate(year_commonname = interaction(year, common_name)) %>% 
   rasterize(rast_template, 
             field = "count",
-            by = "year_species",
+            by = "year_commonname",
             fun = sum)
 # Normalize counts by effort
-topfive_norm <- topfive_rast
-for (i in seq(nlyr(topfive_rast))) {
-  lyr_name <- names(topfive_rast)[i]
+topsix_norm <- topsix_rast
+for (i in seq(nlyr(topsix_rast))) {
+  lyr_name <- names(topsix_rast)[i]
   lyr_year <- substr(lyr_name, 1, 4)
@@
     subst(0, NA)
-  topfive_norm[[i]] <- topfive_norm[[i]] / effort_year
+  topsix_norm[[i]] <- topsix_norm[[i]] / effort_year
 }
 # Average across years
-topfive_climatology <- unique(top_five$species) %>% 
+topsix_climatology <- unique(top_six$common_name) %>% 
   map(function(sp) {
-    sp_rast <- subset(topfive_norm, 
-                      which(str_detect(names(topfive_norm), sp)))
+    sp_rast <- subset(topsix_norm, 
+                      which(str_detect(names(topsix_norm), sp)))
     mean(sp_rast, na.rm = TRUE)
   }) %>% 
-  set_names(unique(top_five$species)) %>% 
+  set_names(unique(top_six$common_name)) %>% 
   rast()
 # Classify raster (otherwise CRSE throw it all off)
-topfive_clim_class <- topfive_climatology %>% 
+topsix_clim_class <- topsix_climatology %>% 
   classify(rcl = c(0, 1, 5, 10, 50))
@@
 ant_basemap() +
-  tidyterra::geom_spatraster(data = topfive_clim_class) +
+  tidyterra::geom_spatraster(data = topsix_clim_class) +
   geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
@@
 
-Distribution maps of the five most abundant species in each survey year, aggregated to 30km bins.
+Distribution maps of the six most abundant species in each survey year, aggregated to 30km bins.
 
 ```{r}
-#| label: topfive_map
+#| label: topsix_map
 #| warning: false
@@
 
-topfive_class <- topfive_rast %>% 
+topsix_class <- topsix_rast %>% 
   classify(rcl = c(0, 1, 5, 10, 50, 100, 1000, Inf))
 
-topfive_class_df <- as.data.frame(topfive_class, xy = TRUE) %>% 
+topsix_class_df <- as.data.frame(topsix_class, xy = TRUE) %>% 
   pivot_longer(-c("x", "y"), 
@@
 ant_basemap() +
-  geom_raster(aes(x, y, fill = count), topfive_class_df) +
+  geom_raster(aes(x, y, fill = count), topsix_class_df) +
   geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
@@
 
-Distribution maps of five species of interest in each survey year, also aggregated to 25km bins.
+Distribution maps of six species of interest in each survey year, also aggregated to 25km bins.
 
@@
 ```
+
+Distribution maps of all species in all years, for supplemental.
+
+```{r}
+#| label: pred_map
+#| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-height: 9
+#| fig-width: 7.5
+#| fig-dpi: 600
+
+pred_sf <- predators_agg %>% 
+  filter(!str_starts(species, "U")) %>% 
+  mutate(common_name = code_to_common(species)) %>% 
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+
+pred_rast <- pred_sf %>% 
+  st_transform(ant_proj()) %>% 
+  mutate(year_commonname = interaction(year, common_name)) %>% 
+  rasterize(rast_template, 
+            field = "count",
+            by = "year_commonname",
+            fun = sum)
+
+# Normalize counts by effort
+pred_norm <- pred_rast
+for (i in seq(nlyr(pred_rast))) {
+  lyr_name <- names(pred_rast)[i]
+  lyr_year <- substr(lyr_name, 1, 4)
+  effort_year <- effort_rast[lyr_year] %>% 
+    subst(0, NA)
+  pred_norm[[i]] <- pred_norm[[i]] / effort_year
+}
+# Average across years
+pred_climatology <- unique(code_to_common(pred_sf$species)) %>% 
+  map(function(sp) {
+    sp_rast <- subset(pred_norm, 
+                      which(str_detect(names(pred_norm), sp)))
+    mean(sp_rast, na.rm = TRUE)
+  }) %>% 
+  set_names(unique(pred_sf$common_name)) %>% 
+  rast()
+
+map_bbox <- ext(-63, -54, -64, -59) %>% 
+  project(from = "+proj=longlat +datum=WGS84",
+          to = as.character(ant_proj())[[1]])
+
+pred_class <- pred_climatology %>% 
+  classify(rcl = c(0, 1, 5, 10, 50, 100, 1000))
+
+pred_class_df <- as.data.frame(pred_class, xy = TRUE) %>% 
+  pivot_longer(-c("x", "y"), 
+               names_to = "name", 
+               values_to = "count") %>% 
+  drop_na(count) %>% 
+  mutate(name = map_chr(name, \(n) paste(strwrap(n, 15), collapse = "\n")))
+
+p <- ant_basemap() +
+  geom_raster(aes(x, y, fill = count), pred_class_df) +
+  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  scale_fill_brewer(palette = "RdPu", 
+                    na.value = NA, 
+                    direction = 1,
+                    na.translate = FALSE) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+  labs(fill = "Ind nmi^-1") +
+  facet_wrap(~ name, ncol = 6) +
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  theme(legend.position = "top",
+        text = element_text(size = 8))
+
+ggsave("figs/s1.jpg", p, width = 6, height = 8, units = "in", dpi = 600)
+```

### 2023-10-25 reports/04_pred_clus.qmd
Clean up dist maps, add cluster env associations

@@
 -   Cluster 1 = pack ice
--   Cluster 2 = marginal ice
--   Cluster 3 = open water
+-   Cluster 2 = open water
+-   Cluster 3 = marginal ice
 
@@
 
-# Indicator species grob
-indval_grob <- apply(sightings_indval$indval, 2, function(col) {
-  # Extract indicator values >0.25 from each cluster's column and format with
-  # species name
-  col <- sort(col[col > 0.25], decreasing = TRUE)
-  species <- code_to_common(names(col))
-  if (length(col) == 0) {
-    "None"
-  } else {
-    str_glue("{species} ({round(col, 2)})")
-  }
-}) %>% 
-  # Create a 
-  map(\(x) {
-    nrow = max(map_dbl(., length))
-    c(x, rep("", nrow - length(x)))
-  }) %>% 
-  as_tibble() %>% 
-  set_names(levels(stations_clust$pred_clust)) %>% 
-  tableGrob(
-    rows = NULL,
-    theme = ttheme_minimal(base_size = 6,
-                           padding = unit(c(1, 1), "mm"),
-                           core = list(fg_params = list(hjust = 0,
-                                                        x = 0)),
-                           colhead = list(fg_params = list(hjust = 0,
-                                                           x = 0)))
-  )
-
 ant_basemap(map_lim) +
@@
                        na.value = "transparent") +
-  # geom_sf(data = seaice_contours, color = "darkblue") +
   # Predator clusters
@@
   theme(legend.position = "bottom",
-        legend.title = element_blank()) +
-  inset_element(indval_grob, 
-                0.67, 0, 1, 0.5, 
-                align_to = "panel")
+        legend.title = element_blank())
+```
+
+```{r}
+#| label: cluster_maps2
+#| warning: false
+
+library(eks)
+cluster_kde <- stations_clust %>% 
+  group_by(pred_clust) %>% 
+  group_map(\(rows, keys) st_kde(rows)$sf %>% 
+              filter(contlabel %in% c(50, 95)) %>% 
+              mutate(pred_clust = keys$pred_clust)) %>% 
+  reduce(sf:::rbind.sf)
+
+ant_basemap(map_lim) +
+  # Predator clusters
+  ggnewscale::new_scale_fill() +
+  geom_sf(aes(color = pred_clust,
+              linetype = contlabel), 
+          cluster_kde, 
+          fill = NA,
+          linewidth = 1) +
+  facet_wrap(~ pred_clust) +
+  scale_x_continuous(breaks = c(-60, -55)) +
+  scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+  scale_color_brewer(palette = "Dark2") +
+  scale_linetype_manual(values = c(2, 1), guide = "none") +
+  coord_ant(map_lim) +
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
 ```
@@
 
+## Environmental conditions by cluster
+
+```{r}
+#| label: env_clust
+
+kw <- map(c("zuml_m", "avg.temp", "avg.salinity", 
+            "Integ.chla.100m", "Integ.phae.100m"),
+          \(v) {
+            kw <- kruskal.test(stations_clust[[v]], stations_clust$pred_clust)
+            data.frame(var = v, p = kw$p.value * 5)
+          }) %>% 
+  list_rbind()
+
+stations_clust %>% 
+  as.data.frame() %>% 
+  group_by(pred_clust) %>%
+  summarize(
+    across(c(zuml_m, avg.temp, avg.salinity, Integ.chla.100m, Integ.phae.100m),
+           list(mean = partial(mean, na.rm = TRUE), 
+                q1 = \(x) quantile(x, 0.25, na.rm = TRUE), 
+                q3 = \(x) quantile(x, 0.75, na.rm = TRUE)),
+           .names = "{.col}={.fn}")
+  ) %>% 
+  pivot_longer(-pred_clust) %>% 
+  separate_wider_delim(name, "=", names = c("var", "fn")) %>% 
+  pivot_wider(names_from = "fn", values_from = "value") %>% 
+  mutate(across(mean:q3, \(x) formatC(signif(x, digits=3), digits = 3, format = "fg", flag = "#"))) %>% 
+  transmute(pred_clust, var, value = str_glue("{mean} ({q1} - {q3})")) %>% 
+  pivot_wider(names_from = pred_clust, values_from = "value") %>% 
+  left_join(kw, by = "var") %>% 
+  mutate(var = ifelse(p <= 0.05, paste0(var, " *"), var)) %>% 
+  select(-p) %>% 
+  knitr::kable()
+
+```
+
 ## Clusters by ice coverage
 
-Median ice coverage increased from the open water cluster (median 33.0%) to the marginal ice (59.6%) and pack ice (70.6%) clusters.
+Median ice coverage increased from the open water cluster (median 34.1%) to the marginal ice (56.0%) and pack ice (74.9%) clusters.
 This supports the labels of "open water" and "marginal ice" for clusters 3 and 2, respectively.

### 2023-10-25 scratch/poster_figures.R

@@ -26,16 +26,14 @@ p <- ant_basemap(map_lim) +
           cluster_kde,
           fill = NA,
           linewidth = 1) +
-  facet_wrap(~ pred_clust) +
+  facet_grid(cols = vars(pred_clust)) +
   scale_x_continuous(breaks = c(-60, -55)) +
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
   scale_color_brewer(palette = "Dark2") +
   scale_linetype_manual(values = c(2, 1), guide = "none") +
   coord_ant(map_lim) +
-  theme(legend.position = "bottom",
-        legend.title = element_blank(),
-        strip.text = element_text(size = 14, face = "bold"),
-        legend.text = element_text(size = 14, face = "bold"))
+  theme(legend.position = "none",
+        strip.text = element_text(size = 14, face = "bold"))
 
 # Generate the ggplot2 plot grob
 g <- grid.force(ggplotGrob(p))
@@ -56,17 +54,17 @@ guide_names <- str_subset(
 txt_colors <- RColorBrewer::brewer.pal(3, "Dark2")
 for (i in 1:3) {
   g <- editGrob(grob = g, gPath = strip_names[i], gp = gpar(col = txt_colors[i]))
-  g <- editGrob(grob = g, gPath = guide_names[i], gp = gpar(col = txt_colors[i]))
 }
 # Draw the edited plot
 grid.newpage()
 grid.draw(g)
 
-# 14.31x7.15
-scale_factor <- 0.40
+# 14.31x4.5
+scale_factor <- 0.60
 ggsave("figs/poster/community_distributions.jpg",
+       g,
        width = 14.31 * scale_factor,
-       height = 7.15 * scale_factor,
+       height = 4.5 * scale_factor,
        units = "in",
        dpi = 600)
 
@@ -78,7 +76,7 @@ median_coverge <- stations_clust %>%
   group_by(pred_clust) %>%
   summarize(median_ice = median(ice_coverage / 10, na.rm = TRUE))
 
-ggplot(stations_clust,
+p <- ggplot(stations_clust,
        aes(x = ice_coverage / 10, fill = after_stat(x))) +
   geom_histogram(bins = 20, color = "grey30") +
   geom_vline(aes(xintercept = median_ice),
@@ -92,11 +90,33 @@ ggplot(stations_clust,
   theme_minimal(base_size = 14) +
[…]

### 2023-10-26 reports/03_combined.qmd
Add supplemental table 1

@@
 
-tar_load(c("predators_agg", "predators_stations", "zoop", "ant_sf"), 
-         store = here::here("_targets"))
+tar_load_everything(store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 
-Each station was associated with the modal ice type (multi-year, first-year, thin, or open) from the underway observations, which were only associated with predator sightings.
-**Seems weird that no stations shook out as multi-year ice, right?** This figure shows how frequent the modal ice type was at each station.
+```{r}
+ice_qa <- zoop_sf %>% 
+  right_join(stations_ice, by = "amlr.station") %>%
+  distinct(amlr.station, ice_type, ice_intervals, ice_type_freq) %>%
+  mutate(n_obs = ice_intervals,
+         n_mode = round(ice_type_freq, digits = 0))
+```
+
+Each station was associated with the modal ice type (multi-year, first-year, thin, or open) from the underway observations.
+This figure shows how frequent the modal ice type was at each station.
 The two grey lines represent 1:1 and 2:1.
@@
 Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station.
-Most points on the figure include multiple stations, indicated by the color.
-For example, in the THIN ice category (bottom), 8 stations (red point) had 4 intervals with ice category observations where all 4 were THIN (x,y = 4,4).
+
+The frequency of the modal ice type was 90% or greater for 57.8% of stations. The modal ice type was \<50% frequency at just 10 stations (\<5% of total stations).
 
@@
 
-predators_stations %>% 
-  mutate(n_obs = ice_intervals,
-         n_mode = round(ice_type_freq * n_obs, digits = 0)) %>% 
-  group_by(n_obs, n_mode, ice_type) %>% 
-  summarize(Stations = n(), .groups = "drop") %>% 
-  ggplot(aes(n_obs, n_mode, color = Stations)) +
+ggplot(ice_qa, aes(n_obs, n_mode)) +
   geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
-  geom_point() +
+  geom_point(alpha = 0.5) +
   labs(x = "Count of intervals with ice category observation",
        y = "Count of intervals with most frequent ice category") +
-  scale_color_fermenter(palette = "RdYlBu",
-                        breaks = c(10, 20, 100, 200)) +
   coord_fixed() +
@@
 
+ice_sf <- station_effort %>% 
+  select(amlr.station, year) %>% 
+  right_join(stations_ice, by = "amlr.station")
+
 ant_basemap(map_lim) +
   geom_sf(aes(color = ice_type), 
-          filter(predators_stations, year != 2014),
+          ice_sf,
           size = 0.5) +
@@
 ```
+
+```{r}
+ice_type_grand_total_row <- stations_ice %>% 
+  mutate(YEAR = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  count(ice_type) %>% 
+  pivot_wider(names_from = "ice_type", 
+              values_from = "n", 
+              values_fill = 0) %>% 
+  mutate(TOTAL = OPEN + THIN + FY + MY,
+         across(-TOTAL, \(x) x / TOTAL),
+         YEAR = "Total")
+
+stations_ice %>% 
+  mutate(YEAR = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  count(YEAR, ice_type) %>% 
+  pivot_wider(names_from = "ice_type", 
+              values_from = "n", 
+              values_fill = 0) %>% 
+  mutate(TOTAL = OPEN + THIN + FY + MY,
+         across(c(OPEN, THIN, FY, MY), \(x) x / TOTAL)) %>% 
+  rbind(ice_type_grand_total_row) %>% 
+  mutate(across(-c(TOTAL, YEAR), scales::percent)) %>% 
+  knitr::kable()
+
+stations_ice %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  count(year, ice_type) %>% 
+  group_by(year) %>% 
+  mutate(frac = n / sum(n)) %>% 
+  ungroup() %>% 
+  mutate(ice_type = fct_rev(ice_type)) %>% 
+  ggplot(aes(year, frac, fill = ice_type)) +
+  geom_col() +
+  scale_fill_brewer(palette = "Blues") +
+  scale_y_continuous("Sea ice composition", labels = scales::percent) +
+  theme_minimal() +
+  theme(axis.title.x = element_blank(),
+        legend.position = "bottom",
+        legend.title = element_blank())
+stations_ice %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  group_by(year, ice_type) %>% 
+  summarize(mean_cov = mean(ice_coverage) / 10,
+            n = n(),
+            .groups = "drop_last") %>% 
+  mutate(n = sum(n),
+         se_cov = sqrt(mean_cov * (1 - mean_cov) / n)) %>% 
+  ungroup() %>% 
+  mutate(ice_type = fct_rev(ice_type)) %>% 
+  ggplot(aes(year, mean_cov, fill = ice_type)) +
+  geom_errorbar(aes(ymin = mean_cov - se_cov, ymax = mean_cov + se_cov),
+                width = 0.2,
+                color = "grey50",
+                position = "dodge") +
+  geom_point(size = 2, 
+             shape = 21, 
+             color = "grey50", 
+             position = position_dodge(width = 0.2)) +
+  scale_fill_brewer(palette = "Blues") +
+  scale_y_continuous(limits = c(NA, 1),
+                     labels = scales::percent) +
+  theme_minimal() +
+  theme()
+
+stations_ice %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  group_by(year) %>% 
+  summarize(mean = mean(ice_coverage),
+            sd = sd(ice_coverage),
+            .groups = "drop") %>% 
+  mutate(mean_sd = str_glue("{format(mean, digits = 2)} \U00B1 {format(sd, digits = 2)}")) %>% 
+  knitr::kable()
+```
+
+## Effort
+
+### Number of stations
+
+Total of `r n_distinct(predators_stations$amlr.station)` stations. Breakdown by year:
+
+```{r}
+predators_stations %>% 
+  distinct(amlr.station) %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  count(year)
+```
+
+### Line transects
+
+```{r}
+survey_nmi <- predators_stations %>% 
+  distinct(amlr.station, survey_nmi) %>% 
+  pull(survey_nmi)
+mean_nmi <- format(mean(survey_nmi), digits = 3) 
+sd_nmi <- format(sd(survey_nmi), digits = 3) 
+```
+
+`r mean_nmi` $\pm$ `r sd_nmi` nm of survey effort associated with each station.
+
+### Ice and effort combined
+
+```{r}
+percent <- partial(scales::percent, accuracy = 0.1)
+
+ice_type_lookup = c(
+  OPEN = "Open water",
+  THIN = "Thin ice",
+  FY = "First-year ice",
+  MY = "Multi-year ice"
+)
+
+ice_table <- stations_ice %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  group_by(year, ice_type) %>% 
+  summarize(n_year_type = n(),
+            ice_mean = mean(ice_coverage / 10),
+            .groups = "drop_last") %>% 
+  mutate(n_year = sum(n_year_type),
+         ice_se = sqrt(ice_mean * (1 - ice_mean) / n_year)) %>% 
+  ungroup() %>% 
+  transmute(year,
+            ice_type = ice_type_lookup[ice_type],
+            ice_val = str_glue("{n_year_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
+  pivot_wider(names_from = ice_type, 
+              values_from = ice_val, 
+              values_fill = "0") %>% 
+  rbind(
+    stations_ice %>% 
+      mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+      group_by(ice_type) %>% 
+      summarize(n_type = n(),
+                ice_mean = mean(ice_coverage / 10)) %>% 
+      mutate(ice_se = sqrt(ice_mean * (1 - ice_mean) / sum(n_type)),
+             ice_val = str_glue("{n_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
+      transmute(year = "Total", ice_type = ice_type_lookup[ice_type], ice_val = str_glue("{n_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
+      pivot_wider(names_from = ice_type, 
+              values_from = ice_val)
+  )
+
+effort_table <- predators_stations %>% 
+  distinct(amlr.station, survey_nmi) %>% 
+  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+  group_by(year) %>% 
+  summarize(`Stations sampled` = n(),
+            mean_nmi = mean(survey_nmi),
+            sd_nmi = sd(survey_nmi),
+            `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean_nmi)} \U00B1 {sprintf('%0.1f', sd_nmi)}")) %>% 
+  select(year, `Stations sampled`, `Effort per station (nmi)`) %>% 
+  rbind(
+    predators_stations %>% 
+      distinct(amlr.station, survey_nmi) %>% 
+      mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
+      summarize(
+        year = "Total",
+        `Stations sampled` = n(),
+        `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean(survey_nmi))} \U00B1 {sprintf('%0.1f', sd(survey_nmi))}")
+      )
+  )
+
+effort_table %>% 
+  left_join(ice_table, by = "year") %>% 
+  knitr::kable() %>% 
+  kableExtra::add_header_above(c(" ", "Survey effort" = 2, "Ice conditions" = 4))
+```

### 2023-10-26 R/predators.R
figs s1 and s2 hacking

@@ -86,6 +86,6 @@ code_to_common <- function(species_code) {
     SBWH = "Bottlenose whale",
     SGPT = "Southern giant petrel",
     SNPT = "Snow petrel",
-    SPSK = "SPSK WHAT",
+    SPSK = "South polar skua",
     WESE = "Weddell seal")[species_code]
 }

### 2023-10-26 reports/01_predators.qmd
figs s1 and s2

@@
 
-tar_load(c("predators_agg", "ant_sf", "effort_sf"), 
-         store = here::here("_targets"))
+tar_load_everything(store = here::here("_targets"))
 source(here::here("R", "sp.R"))
+source(here::here("R", "predators.R"))
 ```
@@
 
-Distribution maps of all species in all years, for supplemental.
+Distribution maps of all species in all years, for supplemental. Organized by mammals and birds
 
 ```{r}
-#| label: pred_map
-#| warning: false
-#| column: screen
-#| out-width: 100%
-#| fig-height: 9
-#| fig-width: 7.5
-#| fig-dpi: 600
-
 pred_sf <- predators_agg %>% 
@@
   drop_na(count) %>% 
-  mutate(name = map_chr(name, \(n) paste(strwrap(n, 15), collapse = "\n")))
+  mutate(name = map_chr(name, \(n) paste(strwrap(n, 20), collapse = "\n")))
+
+```
+
+```{r}
+#| label: bird_climatologies
+#| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-height: 9
+#| fig-width: 6.5
+#| fig-dpi: 600
+
+bird_class_df <- pred_class_df %>% 
+  semi_join(tibble(name = c(
+    "Blue petrel",
+    "Antarctic shag",
+    "Cape petrel",
+    "Kelp gull",
+    "Antarctic petrel",
+    "Southern giant\npetrel",
+    "Adélie penguin",
+    "Kerguelen petrel",
+    "Gentoo penguin",
+    "Antarctic tern",
+    "Northern giant\npetrel",
+    "Pale-faced\nsheathbill",
+    "Chinstrap penguin",
+    "Southern fulmar",
+    "Antarctic prion",
+    "Common diving\npetrel",
+    "South polar skua",
+    "Snow petrel",
+    "Emperor penguin",
+    "Black-browed\nalbatross"
+  )), by = "name")
 
 p <- ant_basemap() +
-  geom_raster(aes(x, y, fill = count), pred_class_df) +
-  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  geom_raster(aes(x, y, fill = count), bird_class_df) +
+  geom_sf(data = ant_sf) +
   scale_fill_brewer(palette = "RdPu", 
@@
   labs(fill = "Ind nmi^-1") +
-  facet_wrap(~ name, ncol = 6) +
+  facet_wrap(~ name, ncol = 5) +
   coord_sf(xlim = map_bbox[1:2],
@@
 
-ggsave("figs/s1.jpg", p, width = 6, height = 8, units = "in", dpi = 600)
+ggsave(here::here("figs/s1.jpg"), 
+       p, 
+       width = 6.5, 
+       height = 9, 
+       units = "in", 
+       dpi = 600)
+
+p
+
 ```
+
+```{r}
+#| label: mammal_climatologies
+#| warning: false
+#| column: screen
+#| out-width: 100%
+#| fig-height: 9
+#| fig-width: 6.5
+#| fig-dpi: 600
+
+mammal_class_df <- pred_class_df %>% 
+  anti_join(bird_class_df, by = "name")
+
+p <- ant_basemap() +
+  geom_raster(aes(x, y, fill = count), mammal_class_df) +
+  geom_sf(data = ant_sf) +
+  scale_fill_brewer(palette = "RdPu", 
+                    na.value = NA, 
+                    direction = 1,
+                    na.translate = FALSE) +
+  scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+  scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+  labs(fill = "Ind nmi^-1") +
+  facet_wrap(~ name, ncol = 3) +
+  coord_sf(xlim = map_bbox[1:2],
+           ylim = map_bbox[3:4],
+           crs = ant_proj()) +
+  theme(legend.position = "top",
+        text = element_text(size = 8))
+
+ggsave(here::here("figs/s2.jpg"), 
+       p, 
+       width = 6.5, 
+       height = 9, 
+       units = "in", 
+       dpi = 600)
+
+p
+
+```

### 2023-11-14 README.Rmd
figs s1 and s2 hacking
Learning forking and pulling!
Gotcha!
HEHE
This is Andrea

@@
 
-Characterizing the marine winter predator community (seabirds and marine mammals) around the northern Antarctic Peninsula (including the Bransfield Strait and South Shetland Islands) using observations collected during the 2012-2016 NOAA/NSF austral winter cruises on the RV Ice Breaker Nathaniel B.
-Palmer.
+![Im a silly person](images/max-czapanskiy_0.jpg)
 
-*Work in progress!!*
-
-## Code
-
-All code under development for **antwincomm** can be found in this repository.
-The analysis pipeline [uses `targets`](https://books.ropensci.org/targets/) to manage workflow caching and dependencies.
-Better documentation to follow (promise).
-This repository will be archived on Zenodo or somewhere similar at publication.
-
-## Data
-
-The **antwincomm** data are too large for GitHub, so they're stored in the [antwincomm Open Science Framework project](https://osf.io/hwnvy/).
-Currently private, will be made public no later than time of publication.
-For access to the OSF project, please contact Max Czapanskiy (max [dot] czapanskiy [at] noaa [dot] gov).
-
-For the code to work, you must download data/WAMLR from the OSF project into data/WAMLR in your local copy of the repository.
-You can do that manually or by running the following lines of code.
-Until the project is made public, though, you'll first have to [authenticate `osfr`](https://docs.ropensci.org/osfr/articles/auth.html).
-
-```{r eval=FALSE}
-dir.create(here::here("data"))
-osf_wamlr <- osfr::osf_retrieve_file("6407bf6228e5c503cc9376b3")
-osfr::osf_download(osf_wamlr, path = here::here("data"), recurse = TRUE)
-```
-
-## Outputs
-
-See the [project website](https://flukeandfeather.github/io/antwincomm) for a list of reports documenting the analysis pipeline.
+##

### 2023-11-29 reports/01_predators.qmd
Clean up targets dependencies

@@
 library(terra)
+library(tidyterra)
 library(tidyverse)
 
-tar_load_everything(store = here::here("_targets"))
+tar_load(c("ant_sf", "effort_sf", "predators_agg"),
+         store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 
-The six most abundant species (in ind nmi\^-1).
+The six most abundant species (in ind km\^-1).
 
@@
   st_transform(ant_proj()) %>% 
-  rasterize(rast_template,
-            field = "nmi",
-            by = "year",
-            fun = sum)
+  rasterize_by("km", "year", sum, rast_template)
+
 # Rasterize the predator sightings by year and species
@@
   st_transform(ant_proj()) %>% 
-  mutate(year_commonname = interaction(year, common_name)) %>% 
-  rasterize(rast_template, 
-            field = "count",
-            by = "year_commonname",
-            fun = sum)
+  mutate(year_commonname = as.character(interaction(year, common_name))) %>% 
+  rasterize_by("count", "year_commonname", sum, rast_template)
+
 # Normalize counts by effort
@@
 ant_basemap() +
-  tidyterra::geom_spatraster(data = topsix_clim_class) +
-  geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
+  geom_spatraster(data = topsix_clim_class) +
+  geom_sf(data = ant_sf) +
   scale_fill_brewer(palette = "RdPu", 
@@
   scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
-  labs(fill = "Ind nmi^-1") +
+  labs(fill = expression("Ind km"^-1)) +
   facet_wrap(~lyr) +
@@
 
-ant_basemap(map_bbox) +
-  tidyterra::geom_spatraster(data = effort_rast) +
+ant_basemap() +
+  geom_spatraster(data = effort_rast) +
   scale_fill_distiller(palette = "PuRd", 
@@
            crs = ant_proj()) +
-  labs(fill = "nmi") +
+  labs(fill = "km") +
   facet_wrap(~lyr) +
@@
 
-topsix_class_df <- as.data.frame(topsix_class, xy = TRUE) %>% 
+topsix_class_df <- topsix_class %>% 
+  as.data.frame(xy = TRUE) %>% 
   pivot_longer(-c("x", "y"), 
@@
   scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
-  labs(fill = "Ind nmi^-1") +
+  labs(fill = expression("Ind km"^-1)) +
   facet_grid(species ~ year) +
@@
   st_transform(ant_proj()) %>% 
-  mutate(year_species = interaction(year, species)) %>% 
-  rasterize(rast_template, 
-            field = "count",
-            by = "year_species",
-            fun = sum)
+  mutate(year_species = as.character(interaction(year, species))) %>% 
+  rasterize_by("count", "year_species", sum, rast_template)
 
@@
 
-soi_class_df <- as.data.frame(soi_class, xy = TRUE) %>% 
+soi_class_df <- soi_class %>% 
+  as.data.frame(xy = TRUE) %>% 
   pivot_longer(-c("x", "y"), 
@@
   st_transform(ant_proj()) %>% 
-  mutate(year_commonname = interaction(year, common_name)) %>% 
-  rasterize(rast_template, 
-            field = "count",
-            by = "year_commonname",
-            fun = sum)
+  mutate(year_commonname = as.character(interaction(year, common_name))) %>% 
+  rasterize_by("count", "year_commonname", sum, rast_template)
 
@@
   drop_na(count) %>% 
-  mutate(name = map_chr(name, \(n) paste(strwrap(n, 20), collapse = "\n")))
+  mutate(name = map_chr(name, \(n) paste(strwrap(n, 20), collapse = "\n"))) %>% 
+  # one sighting each, maps not useful
+  filter(!name %in% c("Antarctic prion", "Humpback whale"))
 
@@
   scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
-  labs(fill = "Ind nmi^-1") +
+  labs(fill = expression("Ind km"^-1)) +
   facet_wrap(~ name, ncol = 5) +
@@
   scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
-  labs(fill = "Ind nmi^-1") +
+  labs(fill = expression("Ind km"^-1)) +
   facet_wrap(~ name, ncol = 3) +

### 2023-11-29 reports/02_stations.qmd
Clean up targets dependencies

@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   geom_sf(aes(color = time.of.day), zoop_sf, size = 0.2) +
@@
   }
-  ant_basemap(map_lim) +
+  ant_basemap() +
     geom_sf(aes(color = .data[[var]]), zoop_trans, size = 2, shape = 15) +

### 2023-11-29 reports/03_combined.qmd
Clean up targets dependencies

@@
 
+library(lubridate)
 library(sf)
@@
 
-tar_load_everything(store = here::here("_targets"))
+tar_load(c("effort_sf", "predators_agg", "predators_stations", 
+           "station_effort", "stations_ice", "zoop", "zoop_sf"),
+         store = here::here("_targets"))
 source(here::here("R", "sp.R"))
@@
 ## Sightings and stations
+
+Visual description of line transect and point station aggregation, for supplemental.
+
+```{r}
+#| label: fig-aggregation
+#| fig-cap: "Biophysical observations (e.g., macrozooplankton net tows and CTD casts) collected at stations (yellow point) were associated with visual surveys for predators collected along line transects.
+All visual surveys conducted within 15 km (shaded circle) and 3 days were aggregated with station observations to form sites.
+In this example, the station was sampled on August 15, 2016 and visual surveys were condected within 15 km of the station twice: on August 15 (red) and on August 23 (blue).
+Only the visual surveys collected on August 15 were associated with this site."
+#| warning: false
+
+points_to_line <- function(pts) {
+  pts %>% 
+    group_by(segmentid) %>% 
+    summarize(begin = paste0(first(processID), first(interval)),
+              end = paste0(last(processID), last(interval)),
+              lbl = format(first(UTC_start), "%b %d %Y"),
+              do_union = FALSE) %>% 
+    mutate(segmentid = factor(segmentid, labels = lbl)) %>% 
+    st_cast("LINESTRING")
+}
+
+tow_origin <- as.POSIXct("1899-12-31", tz = "UTC")
+station_prj <- slice(zoop_sf, 25) %>% 
+  st_transform(ant_proj()) %>% 
+  mutate(tow.UTC = date.tow + (start.time.UTC - tow_origin),
+         lbl = format(tow.UTC, "%b %d"))
+station_buffer <- st_buffer(station_prj, 15e3)
+station_bbox <- st_bbox(station_buffer) %>% 
+  expand_bbox(1.5)
+transect_prj <- effort_sf %>% 
+  filter(between(UTC_start, 
+                 station_prj$tow.UTC - days(3),
+                 station_prj$tow.UTC + days(3))) %>% 
+  st_transform(ant_proj()) %>% 
+  st_intersection(station_buffer) %>% 
+  mutate(lag = UTC_start - lag(UTC_start, default = UTC_start[1]),
+         segmentid = factor(cumsum(lag > hours(6))))
+transect_line <- transect_prj %>% 
+  points_to_line()
+transect_prj2 <- effort_sf %>% 
+  filter(year == station_prj$Year) %>% 
+  st_transform(ant_proj()) %>% 
+  st_crop(station_bbox) %>% 
+  mutate(lag = UTC_start - lag(UTC_start, default = UTC_start[1]),
+         segmentid = factor(cumsum(lag > hours(6))))
+transect_line2 <- transect_prj2 %>% 
+  points_to_line()
+
+transect_endpoints <- transect_prj2 %>% 
+  filter(paste0(processID, interval) %in% c(transect_line2$begin, 
+                                            transect_line2$end))
+
+ant_basemap() +
+  geom_sf(data = station_buffer,
+          fill = "grey50",
+          alpha = 0.1,
+          linetype = "dashed") +
+  geom_sf(aes(color = segmentid),
+          transect_line2,
+          linetype = "dotted",
+          linewidth = 1) +
+  geom_sf(aes(color = segmentid),
+          transect_line,
+          linewidth = 1) +
+  geom_sf(data = station_prj,
+          shape = 23,
+          size = 4,
+          fill = "gold") +
+  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
+  coord_ant(station_bbox) +
+  theme(legend.position = c(0.03, 0.98),
+        legend.justification = c(0, 1),
+        legend.title = element_blank())
+
+```
 
@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   stars::geom_stars(aes(fill = count), data = topfive_stars) +
@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   stars::geom_stars(aes(fill = count), data = soi_stars) +
@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   geom_sf(aes(color = ice_type), 

### 2023-11-29 reports/04_pred_clus.qmd
Clean up targets dependencies

@@
 
-tar_load_everything(store = here::here("_targets"))
+tar_load(c("nmds_df", "nmds_env", "nmds_envfit", "nmds_shepard", "nmds_sightings", 
+           "nmds_stress", "predators_clust", "seaice_conc_df", "seaice_contours", 
+           "sightings_clust", "sightings_cut", "sightings_gap", "sightings_indval", 
+           "stations_clust", "zoop_sf"), 
+         store = here::here("_targets"))
 source(here::here("R", "predators.R"))
@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   # Sea ice
@@
 
-ant_basemap(map_lim) +
+ant_basemap() +
   # Predator clusters

### 2023-11-29 R/predators.R

@@ -54,7 +54,8 @@ normalize_counts <- function(sightings, effort) {
     summarize(count = sum(count), .groups = "drop") %>%
     right_join(select(as_tibble(effort), amlr.station, survey_nmi),
                by = "amlr.station") %>%
-    mutate(count_nmi = count / survey_nmi)
+    mutate(count_nmi = count / survey_nmi,
+           count_km = nmi_to_km(count_nmi))
 }
 
 code_to_common <- function(species_code) {

### 2023-11-29 R/sp.R

@@ -83,7 +83,7 @@ latlon_to_sf <- function(df, coords = c("x", "y")) {
                crs = sp::CRS("+proj=longlat +datum=WGS84"))
 }
 
-ant_basemap <- function(map_lim = NULL) {
+ant_basemap <- function() {
   ggplot() +
     geom_sf(data = tar_read("ant_sf", store = here::here("_targets"))) +
     theme_minimal() +
@@ -118,5 +118,25 @@ expand_bbox <- function(x, factor) {
   result["xmax"] <- x0 + xrng * factor / 2
   result["ymin"] <- y0 - yrng * factor / 2
   result["ymax"] <- y0 + yrng * factor / 2
+  result
+}
+
+nmi_to_km <- function(nmi) nmi * 1.852
+
+rasterize_by <- function(x_sf, field, by, fun, template) {
+  result <- x_sf %>%
+    vect() %>%
+    rasterize(template,
+              field = field,
+              by = by,
+              fun = fun)
+  result_names <- names(result) %>%
+    str_extract("c[(]([^,]+)", 1) %>%
+    str_replace_all("\\\"", "")
+  names(result) <- ifelse(
+    !is.na(result_names),
+    result_names,
+    names(result)
+  )
   result
 }

### 2023-11-29 reports/05_synthesis/ui.R

@@ -1,32 +1,0 @@
-source(here::here("reports", "05_synthesis", "util.R"))
-
-ui <- fillPage(
-
-  fillRow(
-    # Left column: map and env histogram
-    fillCol(
-      girafeOutput("community_map", height = "100%"),
-      div(
-        selectInput("env_var",
-                    "Environmental variable",
-                    choices = list("Salinity" = "avg.salinity",
-                                   "Ice coverage" = "ice_coverage",
-                                   "Int. chl a" = "Integ.chla.100m"),
-                    selected = "Salinity"),
-        plotOutput("env_hist", height = "150px")
-      ),
-      flex = c(1, NA)
-    ),
-
-    # Right column: tree, station table, sightings table
-    fillCol(
-      plotOutput("cluster_dendro", height = "100%"),
-      DT::dataTableOutput("station_table"),
-      DT::dataTableOutput("sightings_table"),
-      flex = c(3, 2, 3)
-    ),
-
-    flex = c(2, 1)
-  )
-
-)

### 2023-11-29 scratch/poster_figures.R

@@ -18,7 +18,7 @@ cluster_kde <- stations_clust %>%
               mutate(pred_clust = keys$pred_clust)) %>%
   reduce(sf:::rbind.sf)
 
-p <- ant_basemap(map_lim) +
+p <- ant_basemap() +
   # Predator clusters
   ggnewscale::new_scale_fill() +
   geom_sf(aes(color = pred_clust,

### 2023-11-30 reports/03_combined.qmd
Finishing edits for November revisions

@@
 ```{r}
-survey_nmi <- predators_stations %>% 
-  distinct(amlr.station, survey_nmi) %>% 
-  pull(survey_nmi)
-mean_nmi <- format(mean(survey_nmi), digits = 3) 
-sd_nmi <- format(sd(survey_nmi), digits = 3) 
+survey_km <- predators_stations %>% 
+  distinct(amlr.station, survey_km) %>% 
+  pull(survey_km)
+mean_km <- format(mean(survey_km), digits = 3) 
+sd_km <- format(sd(survey_km), digits = 3) 
 ```
 
-`r mean_nmi` $\pm$ `r sd_nmi` nm of survey effort associated with each station.
+`r mean_km` $\pm$ `r sd_km` km of survey effort associated with each station.
 
@@
 effort_table <- predators_stations %>%
-  distinct(amlr.station, survey_nmi) %>% 
+  distinct(amlr.station, survey_km) %>% 
   mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
@@
   summarize(`Stations sampled` = n(),
-            mean_nmi = mean(survey_nmi),
-            sd_nmi = sd(survey_nmi),
-            `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean_nmi)} \U00B1 {sprintf('%0.1f', sd_nmi)}")) %>% 
-  select(year, `Stations sampled`, `Effort per station (nmi)`) %>% 
+            mean_km = mean(survey_km),
+            sd_km = sd(survey_km),
+            `Effort per station (km)` = str_glue("{sprintf('%0.1f', mean_km)} \U00B1 {sprintf('%0.1f', sd_km)}")) %>% 
+  select(year, `Stations sampled`, `Effort per station (km)`) %>% 
   rbind(
     predators_stations %>% 
-      distinct(amlr.station, survey_nmi) %>% 
+      distinct(amlr.station, survey_km) %>% 
       mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
@@
         `Stations sampled` = n(),
-        `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean(survey_nmi))} \U00B1 {sprintf('%0.1f', sd(survey_nmi))}")
+        `Effort per station (km)` = str_glue("{sprintf('%0.1f', mean(survey_km))} \U00B1 {sprintf('%0.1f', sd(survey_km))}")
       )

### 2023-11-30 reports/04_pred_clus.qmd
Finishing edits for November revisions

@@
 
-tar_load(c("nmds_df", "nmds_env", "nmds_envfit", "nmds_shepard", "nmds_sightings", 
-           "nmds_stress", "predators_clust", "seaice_conc_df", "seaice_contours", 
-           "sightings_clust", "sightings_cut", "sightings_gap", "sightings_indval", 
-           "stations_clust", "zoop_sf"), 
+tar_load(c("nmds_df", "nmds_env", "nmds_envfit", "nmds_shepard", 
+           "nmds_sightings", "nmds_stress", "predators_clust", 
+           "seaice_conc_df", "seaice_contours", "sightings_clust", 
+           "sightings_cut", "sightings_gap", "sightings_indval", 
+           "station_effort", "stations_clust", "zoop_sf"), 
          store = here::here("_targets"))
@@
 
-ktibs <- with(sightings_gap, k[is_optimum])
+ktibs <- 3
 tibs_thr <- with(sightings_gap, gap[ktibs + 1] - SE.sim[ktibs + 1])
@@
   geom_line() +
-  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
-                      color = is_optimum)) +
+  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim)) +
   geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim),
@@
                linetype = "dashed") +
+  labs(x = expression("Clusters" ~~ (italic(k))),
+       y = expression("Gap statistic" ~~ (italic(f)))) +
   scale_color_manual(values = c(`TRUE` = "blue",
@@
 
-Relative to cluster 3, cluster 2 was composed of more ice associated species.
-The strongest indicator for cluster 2 was the Antarctic fur seal, which was sighted at 98.6% of cluster 2 sites, for 87% of the species' total abundance.
-Pagophilic predators such as crabeater seals, leopard seals, and Adélie penguins were also indicator species for this cluster.
-Another pagophilic predator, the minke whale, was too rare overall to be significant indicator species for any cluster, but was sighted most frequently in cluster 2 (12.2% of sites versus 5.2% and 4.4% of clusters 1 and 3, respectively) and 80.0% of its abundance was in cluster 2.
-
-Cluster 3 describes a more open water guild.
-Southern fulmars and cape petrels were significant indicator species for this cluster.
-Blue petrels, an open water predator, were most frequently observed in cluster 3 (36.3% of sites, relative to 6.9% and 10.8% of sites in clusters 1 and 2) in greater numbers than elsewhere (66.0% of total abundance).
-
-Snow petrels, Antarctic petrels, and southern giant petrels were significant indicator species for both clusters 2 and 3.
-These species were the most widespread predators overall, sighted at 59.1%, 53.0%, and 44.9% of stations.
-Snow petrels were also the most abundant predator in the survey, accounting for 34.1% of total individuals observed.
-
-No species were identified as significant indicators for cluster 1.
-Species observed at \>25% of cluster 1 sites include Adélie penguins, crabeater seals, and Antarctic fur seals, suggesting a pack ice community.
-In terms of relative abundance, 59.8% of Weddell seals, 43.6% of Adélie penguins, and 20.2% of crabeater seals were sighted in cluster 1.
-If this cluster represents the pack ice community, the lack of significant indicator species may be attributable to difficulty surveying the pack ice?
+```{r}
+#| label: abu_by_clust
+predators_clust %>% 
+  select(amlr.station, species, count) %>% 
+  complete(species, amlr.station, fill = list(count = 0)) %>% 
+  left_join(stations_clust %>% 
+              as_tibble() %>% 
+              select(amlr.station, pred_clust),
+            by = "amlr.station") %>% 
+  rbind(mutate(., pred_clust = "All clusters")) %>% 
+  left_join(station_effort %>% 
+              as_tibble() %>% 
+              transmute(amlr.station, survey_km = nmi_to_km(survey_nmi)),
+            by = "amlr.station") %>% 
+  group_by(species, pred_clust) %>% 
+  summarize(density_km = sum(count) / sum(survey_km),
+            freq = sum(count > 0) / n(),
+            density_freq = str_glue("{sprintf('%0.3f', density_km)} [{scales::percent(freq, accuracy = 0.1)}]"),
+            .groups = "drop") %>% 
+  select(-density_km, -freq) %>% 
+  pivot_wider(names_from = pred_clust, values_from = density_freq) %>%
+  arrange(desc(`All clusters`)) %>% 
+  knitr::kable()
+```
 
@@
 
-nmds_envfit
+vector_tbl <- with(nmds_envfit$vectors, 
+                   cbind(arrows, 
+                         r2 = r, 
+                         p = pvals)) %>% 
+  as_tibble(rownames = "var") %>% 
+  mutate(type = "Vector",
+         across(NMDS1:p, \(x) sprintf("%0.3f", x)))
+
+factor_tbl <- with(nmds_envfit$factors, 
+                   cbind(centroids, 
+                         r2 = r[var.id],
+                         p = pvals[var.id])) %>% 
+  as_tibble(rownames = "var") %>% 
+  mutate(type = "Factor",
+         across(NMDS1:p, \(x) sprintf("%0.3f", x)))
+
+rbind(vector_tbl, factor_tbl) %>% 
+  relocate(type) %>% 
+  knitr::kable()
 

### 2023-11-30 R/predators.R

@@ -2,8 +2,9 @@ aggregate_predators <- function(predators) {
   predators %>%
     filter(species != "NULL") %>%
     mutate(UTC_start = lubridate::mdy_hm(UTC_start),
-           year = lubridate::year(UTC_start)) %>%
-    group_by(year, interval, species, UTC_start, lon_mean, lat_mean, nmi) %>%
+           year = lubridate::year(UTC_start),
+           km = nmi_to_km(nmi)) %>%
+    group_by(year, interval, species, UTC_start, lon_mean, lat_mean, nmi, km) %>%
     summarize(count = sum(count_species), .groups = "drop")
 }
 
@@ -52,7 +53,7 @@ normalize_counts <- function(sightings, effort) {
     as_tibble() %>%
     group_by(amlr.station, species) %>%
     summarize(count = sum(count), .groups = "drop") %>%
-    right_join(select(as_tibble(effort), amlr.station, survey_nmi),
+    right_join(select(as_tibble(effort), amlr.station, survey_nmi, survey_km),
                by = "amlr.station") %>%
     mutate(count_nmi = count / survey_nmi,
            count_km = nmi_to_km(count_nmi))

### 2024-05-30 reports/01_predators.qmd
Final commit before Zenodo preparations

@@
 # Make the map! 
-ant_basemap() +
+sublabels <- tibble(lyr = names(topsix_clim_class),
+                    x = -62.25, y = -59.25) %>% 
+  arrange(lyr) %>% 
+  mutate(labels = LETTERS[1:6]) %>% 
+  st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+p <- ant_basemap() +
   geom_spatraster(data = topsix_clim_class) +
   geom_sf(data = ant_sf) +
+  geom_sf_text(aes(label = labels), sublabels, fontface = "bold") +
   scale_fill_brewer(palette = "RdPu", 
@@
            crs = ant_proj()) +
-  theme(legend.position = "top")
+  theme(legend.position = "top",
+        panel.background = element_rect(color = "grey20"),
+        strip.text = element_blank())
+ggsave(here::here("figs/Figure 2.eps"), p, 
+       width = 169, height = 121, units = "mm")
+p
 

### 2024-05-30 reports/04_pred_clus.qmd
Final commit before Zenodo preparations

@@
 
-ant_basemap() +
+sublabels <- tibble(Year = 2012:2016,
+                    x = -63, y = -60,
+                    labels = LETTERS[1:5]) %>% 
+  st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+p <- ant_basemap() +
   # Sea ice
@@
             alpha = 0.8) +
-  scale_fill_distiller(lim = c(0.65, 1),
+  scale_fill_distiller("Ice concentration",
+                       lim = c(0.65, 1),
+                       labels = scales::percent,
                        palette = "Blues",
-                       na.value = "transparent") +
+                       na.value = "transparent",
+                       guide = guide_colorbar(title.position = "top")) +
   # Predator clusters
@@
   facet_wrap(~ Year) +
+  # Labels
+  geom_sf_text(aes(label = labels), sublabels, fontface = "bold", vjust = 0) +
   scale_x_continuous(breaks = c(-60, -55)) +
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-  scale_fill_brewer(palette = "Dark2") +
-  guides(fill = guide_legend(override.aes = list(size = 3))) +
+  scale_fill_brewer("Predator cluster", palette = "Dark2",
+                    guide = guide_legend(override.aes = list(size = 3),
+                                         title.position = "top")) +
   coord_ant(map_lim) +
   theme(legend.position = "bottom",
-        legend.title = element_blank())
+        strip.text = element_blank(),
+        panel.background = element_rect())
+
+ggsave(here::here("figs/Figure 4.eps"), p, 
+       width = 169, height = 121, units = "mm")
+p
 ```
@@
 
-ant_basemap() +
+sublabels <- tibble(pred_clust = factor(levels(cluster_kde$pred_clust)),
+                    x = -63, y = -60,
+                    labels = LETTERS[1:3]) %>% 
+  st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+p <- ant_basemap() +
   # Predator clusters
@@
           linewidth = 1) +
+  geom_sf_text(aes(label = labels), sublabels, fontface = "bold", vjust = 0) +
   facet_wrap(~ pred_clust) +
@@
   scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
-  scale_color_brewer(palette = "Dark2") +
+  scale_color_brewer("Predator cluster", palette = "Dark2") +
   scale_linetype_manual(values = c(2, 1), guide = "none") +
@@
   theme(legend.position = "bottom",
-        legend.title = element_blank())
+        strip.text = element_blank(),
+        panel.background = element_rect())
+
+ggsave(here::here("figs/Figure 5.eps"), p, 
+       width = 169, height = 57, units = "mm")
+p
 ```
@@
     ggnewscale::new_scale_color() +
-    geom_point(aes(color = pred_clust), alpha = 0.8) +
+    geom_point(aes(color = pred_clust), size = 1, alpha = 0.8) +
     geom_point(data = year_centroids, 
@@
     labs(color = "Predator cluster") +
-    theme_classic()
+    theme_classic(base_size = 10)
 }
 
-map2(c(1, 1, 2), c(2, 3, 3), nmds_plot) %>% 
+p <- map2(c(1, 1, 2), c(2, 3, 3), nmds_plot) %>% 
   reduce(`|`) +
@@
     theme(legend.position = "bottom",
-          legend.direction = "horizontal",
+          legend.box = "vertical",
+          legend.box.margin = margin(t = -3),
+          legend.spacing.y = unit(-5, "pt"),
           plot.tag = element_text(face = "bold"))
+ggsave(here::here("figs/Figure 6.eps"), p, 
+       width = 169, height = 100, units = "mm")
+p
+
 ```
@@
 
-ggplot(stations_clust, 
+sublabels <- tibble(
+  pred_clust = factor(levels(stations_clust$pred_clust)),
+  labels = LETTERS[1:3],
+  x = 1, y = 15
+)
+p <- ggplot(stations_clust, 
        aes(x = ice_coverage / 10, fill = after_stat(x))) + 
@@
              color = "red", linetype = "dashed") +
+  geom_text(aes(x = x, y = y, label = labels), sublabels, 
+            hjust = 1, vjust = 1,
+            fontface = "bold") +
   scale_x_continuous(labels = scales::percent) +
@@
        y = "# sites") +
-  theme_minimal()
+  theme_classic() +
+  theme(strip.text = element_blank())
+
+ggsave(here::here("figs/Figure 3.eps"), p, 
+       width = 169, height = 121, units = "mm")
+p
 

### 2024-05-30 R/sp.R

@@ -140,3 +140,11 @@ rasterize_by <- function(x_sf, field, by, fun, template) {
   )
   result
 }
+
+make_overview_map <- function() {
+  map_bbox <- st_bbox(c(xmin = -64, xmax = -53, ymin = -64.5, ymax = -58.5),
+                      crs = "EPSG:4326") %>%
+    project_bbox()
+  ant_basemap() +
+    coord_ant(map_bbox)
+}

### 2024-05-30 _targets.R

@@ -153,6 +153,8 @@ list(
                                         permutations = 999,
                                         choices = 1:3,
                                         na.rm = TRUE)),
+  # Overview map (figure 1)
+  tar_target(overview_map, make_overview_map()),
   # Reports
   tar_quarto(reports, here("reports"))
 )

### 2024-05-30 README.Rmd
max loves bagels
yooo
Add readme

@@
 
-```{r, include = FALSE}
+```{r, echo = FALSE}
 knitr::opts_chunk$set(
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  fig.path = "README-"
 )
+# Please put your title here to include it in the file below.
+Title <- "Sea-ice and macrozooplankton distribution as determinants of top predator community structure in Antarctic winter"
 ```
@@
 
-<!-- badges: start -->
+[![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.11397472.svg)](http://dx.doi.org/10.5281/zenodo.11397472)
 
-<!-- badges: end -->
+This repository contains the code for our paper:
 
-![Im a silly person](images/max-czapanskiy_0.jpg)
+> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024). _`r Title`_. Marine Ecology Progress Series <https://doi.org/xxx/xxx>
 
-##
+### How to cite
+
+Please cite this compendium as:
+
+> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024).
+(`r format(Sys.Date(), "%Y")`).
+_Compendium of R code for `r Title`_.
+Accessed `r format(Sys.Date(), "%d %b %Y")`.
+Online at <https://doi.org/10.5281/zenodo.11397472>
+
+## Contents
+
+The **analysis** directory contains:
+
+  - [:file\_folder: paper](/analysis/paper): Quarto source document
+    for manuscript. Includes code to reproduce the figures and tables
+    generated by the analysis. It also has a rendered version,
+    `paper.docx`, suitable for reading (the code is replaced by figures
+    and tables in this file)
+  - [:file\_folder: figures](/analysis/figures): Plots and other
+    illustrations
+  - [:file\_folder:
+    supplementary-materials](/analysis/supplementary-materials):
+    Supplementary materials including notes and other documents
+    prepared and collected during the analysis.
+
+## How to download and run locally
+
+This research compendium has been developed using the statistical programming
+language R. To work with the compendium, you will need
+installed on your computer the [R software](https://cloud.r-project.org/)
+itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).
+
+You can download the compendium as a zip from from this URL:
+[main.zip](/archive/main.zip). After unzipping:
+- open the `.Rproj` file in RStudio
+- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
+[DESCRIPTION](/DESCRIPTION) file).
+- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console
+
+### Licenses
+
+**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
+
+**Code :** See the [DESCRIPTION](DESCRIPTION) file

### 2024-05-30 analysis/paper/paper.qmd
begin writing paper: intro

---
title: "Sea-ice and macrozooplankton distribution as determinants of top predator community structure in Antarctic winter"
author:
  - Max F. Czapanskiy:
      orcid: 0000-0002-6302-905X
      institute:
        - ims
        - fed
        - esd1
  - Jarrod A. Santora:
      correspondence: "yes"
      email: jarrod.santora@noaa.gov
      institute: 
        - fed
        - dam
  - Kimberly S. Dietrich:
      institute: 
        - ims
        - fed
  - Megan A. Cimino:
      orcid: 0000-0002-1715-2903
      institute:
        - ims
        - esd1
  - Elliott L. Hazen:
      institute: esd1
  - Christian S. Reiss:
      institute: esd2
  - Richard R. Veit:
      institute: cuny
institute:
  - ims:
      name: Institute of Marine Sciences, University of California Santa Cruz
      address: Santa Cruz, California 95060, USA
  - fed: 
      name: Fisheries Ecology Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: Santa Cruz, California 95060, USA
  - esd1:
      name: Ecosystem Science Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: Monterey, California 93940, USA
  - dam:
      name: Department of Applied Math, University of California Santa Cruz
      address: Santa Cruz, California 95060, USA
  - esd2:
      name: Ecosystem Science Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: La Jolla, California 92037, USA
  - cuny:
      name: City University of New York, College of Staten Island, Biology Department
      address: Staten Island, New York 10314, USA
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/marine-ecology-progress-series.csl" # Insert path for the bib-style
abstract: |
  The Antarctic Peninsula marine ecosystem is highly productive, with large 
  populations of commercially and ecologically important species including 
  Antarctic krill (*Euphausia superba*), Adelié Adélie penguins (*Pygoscelis 
  adeliae*), and crabeater seals (*Lobodon carcinophagus*). The ecology of the 
  peninsula is rapidly changing due to accelerating climate change and fishing 
  pressure. Systematic ecosystem surveys have focused on austral spring and 
  summer, leaving an information gap on winter ecosystem dynamics. Using data 
  from 5 consecutive ecosystem surveys, we quantified the composition and 
  distribution of winter predator communities and investigated the physical and 
  biological influences on community structure. Seabirds and marine mammals 
  clustered into 3 communities: an ice-associated community represented by 
  Adélie penguins and crabeater seals; a diverse marginal ice zone community
  dominated by fur seals and several species of seabirds including 3 petrels, 
  kelp gulls (*Larus dominicanus*), and Antarctic terns (*Sterna vittata*); and 
  an open water community consisting of Southern fulmars (*Fulmarus 
  glacialoides*) and 4 species of petrels. These communities were distributed 
  along an environmental gradient ranging from ice-covered, cold, saline waters 
  to ice-free, warmer, and fresher water with greater chlorophyll 
  concentrations. Predator communities were also associated with different 
  communities of macrozooplankton: ice-associated predators with an extremely 
  diverse assemblage of typically mesopelagic zooplankton; marginal ice zone 
  predators with a community of large-bodied euphausiids community (*E. superba*
  , *E. crystallorophias*); and open water predators with a community of small-
  bodied euphausiids (*Thysanoessa macrura*). Our synthesis of integrated winter 
  predator and macrozooplankton communities relative to sea-ice concentration 
  provides reference points for future ecosystem assessments within this rapidly 
  changing region.
keywords: |
  Biogeography; Krill; Marine mammal; Polar regions; Sea ice; Seabird; 
  Zooplankton
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

<!-- The actual document text starts here: -->

# Introduction

The marine ecosystem surrounding the Antarctic Peninsula is highly productive and rapidly changing [@clarkeClimateChangeMarine2007; @montes-hugoRecentChangesPhytoplankton2009]. Integrated assessments and syntheses across trophic levels will help us better understand variability of the ecosystem and potential impacts of climate change, such as effects of phenological shifts and carry-over effects between seasons. Variability in the spring biomass and distribution of Antarctic phytoplankton, abundance and diversity of zooplankton, and population dynamics of upper trophic level species (e.g. reproductive success and recruitment) are controlled, in part, by physical and biogeochemical processes during the preceding winter [@hinkeDailyActivityMinimum2011; @sabaWinterSpringControls2014; @meyerWinterPackiceZone2017]. However, systematic ecosystem surveys have focused on austral spring and summer due to more favorable weather. Thus, there is a paucity of information on winter ecosystem dynamics (i.e. diversity and distribution of communities). Since climate-related effects will likely accelerate and percolate through the ecosystem, it is important to document what these effects are, including anthropogenic stressors.

Winter-time climate change near the northern Antarctic Peninsula and fisheries pressure are increasing faster than in other seasons. Atmospheric temperatures around the Antarctic Peninsula increased faster than anywhere else in the southern hemisphere over the last 50 yr, with the most rapid warming occurring during austral winter [@jonesSixtyYearsWidespread2019; @ducklowMarinePelagicEcosystem2022]. Warmer winters, combined with stronger winds, have reduced the duration of sea-ice coverage in the region by an average of \>90 d yr^–1^ [@ducklow2013west], leading to reduced biodiversity and biomass at the base of the food web, including phytoplankton, microzooplankton, and larval fish [@linDeclinePlanktonDiversity2021; @corsoClimateDrivesLongterm2022]. Further, recent record low winter sea-ice coverage in the Antarctic Peninsula suggests that the rate of warming and overall resilience of the marine ecosystem may have been underestimated [@purichRecordLowAntarctic2023]. These winter ecosystem changes have negatively impacted predator populations, including those of penguins [@trivelpiece2011variability] and fur seals [@krauseRapidPopulationCollapse2022; @forcadaNinetyYearsChange2023] that had been recovering. Concurrently, the annual commercial catch of Antarctic krill Euphausia superba increased 4-fold from 2010 to 2020 [@kawaguchi2020krill]. As winter sea-ice extent and concentration decrease, the winter fishery activities have expanded farther south and have become more spatially concentrated [@nicol2016fishery]. This spatial shift has resulted in greater overlap between fisheries and predators at local scales [@hinkeIdentifyingRiskConcurrent2017]. The increasing overlap between commercial fisheries, and foraging seals and penguins, brings into question whether negative impact on predators will increase or decrease [@wattersLongtermObservationsAntarctica2020]. Effective ecosystem-based management of the Antarctic Peninsula is thus hindered by limited data on winter-time community structure and foraging interactions among species [@bestleyMarineEcosystemAssessment2020].

The winter communities of seabirds and marine mammals in the northern Antarctic Peninsula is a mix of overwintering and seasonal members [@santoraEnvironmentalDeterminantsTop2014]. These species include a variety of ice-obligate, ice-avoidant, and intermediate species. Ice-obligate (pagophilic) species consist primarily of swimmers, such as crabeater seals *Lobodon carcinophaga*, Adélie penguins *Pygoscelis adeliae*, and leopard seals *Hydrurga leptonyx*. Most of these either follow the ice edge north as it advances in winter [@thiebotAdeliePenguinsExtensive2019] or remain in the pack ice [@burnsWinterHabitatUse2004] all year. Pagophilic predators exploit prey resources associated with pack ice and rely on ice floes to haul out for rest, avoid predators, and reproduce [@southwellTimingPuppingPackice2003; @ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017]. The ice-avoidant (open water) species include several fulmarine petrels, such as southern fulmars *Fulmarus glacialoides*, cape petrels *Daption capense*, and blue petrels *Halobaena caerulea*. This guild is abundant in the region during the austral summer [@santoraSpatiotemporalPersistenceTop2013; @warwick-evansMultiscaleAssessmentDistribution2021], but many individuals migrate farther north in winter [@delordSpeciesSpecificForaging2016]. These seabirds have broader wing morphology adapted for efficient flight in strong, consistent winds [@ainleySpeciesHabitatRelationshipsAntarctic1993]. In contrast, intermediate between these 2 guilds, snow petrels *Pagodroma nivea* and Antarctic petrels *Thalassoica antarctica* have narrower wings and forage in areas of marginal ice coverage [@santoraEnvironmentalDeterminantsTop2014]. During winter, they are largely confined to the marginal ice zone [@delordAntarcticPetrelsIce2020; @delordSpeciesSpecificForaging2016].

At fine spatial scales (\<10 km), the structure of the top predator community of the Antarctic Peninsula in winter is driven by both physical (sea ice coverage) and biological (prey abundance and distribution) factors [@fraserIceEdgesSeabird1986; @reissOverwinterHabitatSelection2017]. The mesoscale (10s to 100s km) distribution of macrozooplankton communities is more geographically consistent in winter than summer, independent of ice conditions [@dietrich2021winter], whereas the fine-scale distribution of seabird communities dynamically responds to ice coverage in all seasons [@ainleySpeciesHabitatRelationshipsAntarctic1993]. Five winter communities of macrozooplankton have previously been identified in this region [@dietrich2021winter]. Two of these communities contain smaller euphausiids (*Thysanoessa macrura*, *E. frigida*, *E. triacantha*) and amphipods (*Primno*, *Themisto*) and occupy the region offshore of the shelf, north and west of the South Shetland Islands. Within the Bransfield Strait, a community dominated by large euphausiids (*E. superba*, *E. crystallorophias*) occurs in the western part of the strait, while a highly diverse assemblage including chaetognaths, *Calanoides acutus*, and siphonophores occurs in the eastern strait. A fifth macrozooplankton community is ubiquitous within this region and contains pelagic tunicates (*Salpa thompsoni*) and gastropods (*Clione limacina*). If the distribution of prey communities influences structure of predator communities, then we would expect associations between the open water seabird guild and the off-shelf macrozooplankton communities (called 2a and 2b by @dietrich2021winter), as well as the pagophilic penguin/pinniped predator community with the Bransfield Strait macrozooplankton communities (called 3a and 3b by @dietrich2021winter). Conversely, if the physical factors influencing predator community distribution at the fine-scale also drive mesoscale patterns, then predator communities would be associated directly with ice coverage (and perhaps hydrology) rather than macrozooplankton communities.

Here, we investigated the ecological determinants of top predator community structure in the Antarctic Peninsula during winter (August to September) using data from 5 consecutive ecosystem surveys in 2012–2016. The study period coincided with several high and low sea-ice years, and captured natural variability in the extent and distribution of the winter marginal ice zone habitat within the northern Antarctic Peninsula marine ecosystem [@santoraEnvironmentalDeterminantsTop2014; @reissOverwinterHabitatSelection2017; @dietrich2021winter]. Antarctic sea ice reached a record high in 2014, before the current regime of rapid declines [@parkinson40yRecordReveals2019], and an extreme El Niño event in 2015/2016 had substantial impacts on phytoplankton and zooplankton communities [@costaLargeDiatomBloom2021; @walshExtremeNinoSouthern2023]. We clustered top predator observations into distinct communities following the methods used for macrozooplankton communities from the same surveys [@dietrich2021winter]. We quantified associations between the top predator communities and (1) physical environmental factors and (2) macrozooplankton communities to test the hypothesis that community structure at lower trophic levels, combined with sea-ice extent, determines community structure at higher trophic levels. Our winter ecosystem observations have potential implications for future climate and fishery impact assessments that emphasize relationships between the physical environment, macrozooplankton distribution, and top predator community structure.

# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### 2024-05-30 analysis/scripts/01_download_data.R

@@ -1,0 +1,8 @@
+# Download dataset zip from USAP Data Center
+download.file(
+  "https://www.usap-dc.org/dataset/usap-dc/601795/2024-05-28T19:16:39.2Z/AntPeninsula_winter_marine_data.zip",
+  "analysis/data/raw_data/AntPeninsula_winter_marine_data.zip"
+)
+
+# If necessary to do manually, visit https://doi.org/10.15784/601795 and
+# download AntPeninsula_winter_marine_data.zip to analysis/data/raw_data.

### 2024-05-30 analysis/paper/paper.qmd
paper: methods

@@
 
-# Background
+# Materials and Methods
 
-# Methods
+## Survey area data
+
+### Survey design
+
+We surveyed the South Shetland Island and Elephant Island regions during the austral winters (August to September) of 2012–2016 as part of the US Antarctic Marine Living Resources (AMLR) program aboard the RV Ice Breaker (RVIB) ‘Nathaniel B.
+Palmer’ (Fig.
+1).
+The main objectives of the survey were to map hydrographic conditions, study the distribution and abundance of fish, krill, and other zooplankton through acoustic methods and net hauls, and describe the spatial distribution of seabirds and marine mammals.
+We followed methods consistent with the long-term austral summer surveys conducted by the US AMLR program [@reissOverwinterHabitatSelection2017].
+We sampled between 33 and 108 stations per year, dependent on ice conditions.
+Stations were spaced approximately 25–30 km apart, and the full survey area covered an area of approximately 120000 km^2^.
+
+### Visual surveys
+
+Sighting surveys for seabirds and marine mammals were conducted from the bridge of the ship, following established methodologies described in previous studies [@santoraEnvironmentalDeterminantsTop2014].
+Two observers collected observations simultaneously during daylight hours.
+One observer recorded seabirds, while the other scanned for marine mammals in the water and on the ice using binoculars.
+Observers used 10 × 50 binoculars and image-stabilized 20 × 60 binoculars.
+The pilot house/bridge of the RVIB ‘Nathaniel B.
+Palmer’ is 54 ft (\~16.5 m) above sea level.
+Observations only stopped when the ship could not move forward though the sea ice, and no observations were made during station sampling.
+
+During each survey, we continuously monitored and classified sea-ice conditions according to the international sea ice protocol (ASPeCt) [@worbyPartObservationalTechnique1999; @santoraSpatiotemporalPersistenceTop2013; @reissOverwinterHabitatSelection2017].
+This involved estimating the percentage of ice cover on a scale of 0 to 10, where 0 is open water and 10 is complete ice coverage (fully consolidated).
+We also identified different types of ice, including first-year, multi-year, and thin gray ice.
+An automatic logger recorded the ice conditions every 10 s, updated by the observer as conditions changed.
+
+### Hydrology
+
+We measured hydrology at each survey station.
+Conductivity and temperature in the water column were measured using a Sea-Bird Inc.
+SBE-9/11+ CTD profiler (Table 1).
+We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations (Table 1).
+For full details, see @dietrich2021winter.
+
+## Statistical analysis
+
+We standardized our data to integrate visual surveys (line transects) with oceanographic sampling stations (discrete points).
+First, we aggregated all transect observations (relative abundance counts of seabirds and mammals, and ice conditions) into 1.85 km bins.
+We then associated each sampling station with all transect intervals conducted within 15 km and 3 d of the station to form ‘sites’.
+We excluded intervals farther in space or time than 15 km and 3 d from a sample station from the community analysis.
+After associating transect intervals and sample stations in sites, we aggregated the relative predator abundance and ice conditions of each site.
+We normalized relative predator abundance by sampling effort (ind.
+km^–1^), then applied a pseudo-log transformation (i.e.
+log~10~(*x* + 1)) to minimize the influence of highly abundant taxa.
+The analysis of macrozooplankton in these surveys [@dietrich2021winter] excluded taxa observed at fewer than 5% of sample stations; the seabird and marine mammal assemblage is much less diverse than macrozooplankton, so we did not exclude any species.
+We aggregated ice conditions across intervals within a site as the most frequent ice type (open, thin, first-year, or multi-year sea ice) and mean percent coverage.
+
+We assessed the community structure from relative abundance data using cluster and ordination analyses.
+First, we calculated a Bray-Curtis dissimilarity matrix using the R package 'vegan' [@oksanenVeganCommunityEcology2022] to estimate dissimilarities in species composition between sites.
+We then generated hierarchical clusters using Ward's minimum variance linkage method, implemented in the R package 'cluster' [@maechlerClusterClusterAnalysis2022], applied to the dissimilarity matrix.
+We determined the optimal number of clusters (*k*) to retain from the hierarchy using the gap statistic [@tibshiraniEstimatingNumberClusters2001].
+The gap statistic compares the within-cluster variation of different values of *k* to their expected values under a null reference distribution.
+The gap statistic, *f*, usually increases with *k*, but exhibits a shoulder at the optimal number of clusters.
+A recommended heuristic for finding the shoulder is selecting the smallest *k* such that the gap statistic for *k*, *f*(*k*), is greater than the gap statistic for *k + 1* less the standard error, *s* [@eq-gap-stat].
+
+$$
+f(k)>f(k+1)-s(k+1)
+$$ {#eq-gap-stat}
+
+We described each predator cluster with a smaller number of indicator species.
+Using the R package 'labdsv', we calculated indicator values, *d*, which indicate the relative affinity of a species to a particular cluster [@dufrene1997species; @robertsLabdsvOrdinationMultivariate2019].
+The indicator value for species *i* in cluster *c*, *d~ic~*, is the product of relative frequency, *f~ic~*, and relative abundance, *a~ic~*, where *f~ic~* is the fraction of sites in *c* where *i* was present and *a~ic~* is the ratio of per-site abundance of *i* in *c* to the overall per-site abundance of *i*.
+*d* ranges between 0 and 1, where values close to 1 indicate a species that was ubiquitous within the cluster and highly abundant relative to other clusters.
+Indicator values ≥0.25 were retained as significant indicators.
+
+To explore the relationship between community structure and the environment, we reduced the dimensionality of the community data using non-metric multidimensional scaling (NMDS).
+For this purpose, we used the same Bray-Curtis dissimilarity matrix from the cluster analysis.
+We determined the number of NMDS axes to retain by examining a scree plot, which plots overall stress of the fit against the number of axes, and assessed quality of fit using a Shepard diagram, which plots the dissimilarity distances in the reduced space against the original dissimilarity distances.
+We chose the smallest number of axes with stress \<0.2 and examined the correlation in the Shepard diagram for goodness of fit.
+
+We used linear regression to quantify the relationship between the retained NMDS axes and the environmental, biological, and temporal variables.
+We performed this analysis with the `envfit()` function in the 'vegan' package, which treats environmental variables as the response variables and NMDS ordination scores as the predictors.
+The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables (Table 1).
+Temporal variables included time of day of the net tow, to account for diel patterns in macrozooplankton distribution, as well as year, to account for interannual variability.
+The regression coefficients obtained from the analysis represent the direction and magnitude of the maximal change in community structure associated with each environmental variable.
+We tested for significant associations between predator and macrozooplankton communities using a chi-squared test and assessed the statistical significance of individual predator–prey community associations via post hoc analysis of the Pearson residuals.
 

### 2024-05-31 analysis/paper/paper.qmd
Results and discussion

@@
 
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
-
-<!-- With the following code you can access and display values from the yml header above. -->
-
 Keywords: `r rmarkdown::metadata$keywords`
-
-<!-- The actual document text starts here: -->
 
@@
 
-<!-- Here's some example analysis code: -->
+## Visual surveys
 
-```{r}
-#| label: get-data
-#| eval: false
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
-```
+During the 5 survey years, we sighted a total of 22 species of seabirds and marine mammals (Table 2).
+The most abundant species were snow petrels, Adélie penguins, and Antarctic fur seals.
+The most widespread species observed were snow petrels, Antarctic petrels, and southern giant petrels.
+The 6 most abundant species (total individuals observed in Table 2) averaged across years were widely distributed across the study region, although densities varied by species (Fig.
+2).
+Distribution maps of all seabird (Fig.
+S1) and marine mammal species (Fig.
+S2) are provided in the Supplement.
+We observed predators in proximity (15 km, 3 d) to 245 sampling stations, ranging from 19 stations in 2012 to 68 stations in 2014.
+Each station was associated with 30.2 ± 10.3 km (mean ± SD) of visual survey effort (Table S1).
 
-```{r}
-#| label: fig-demo-plot
-#| fig-cap: "A plot of random numbers"
-plot(rnorm(10))
-```
+With the exception of 2012, when sampling was most limited, ice conditions observed during visual surveys exhibited more interannual variability in type (multi-year ice, first-year ice, thin ice, open water) than percent coverage.
+In 2012, surveys were restricted to the northernmost part of the survey region, and open water was the most common type (93.3%).
+In the other survey years, the most common ice type was thin ice (2013, 2015, and 2016) or first-year ice (2014) (Table S1).
+Ice coverage was relatively consistent within ice types across years.
+Overall, sites categorized as open water had ice coverage of 6.4 ± 1.6%; thin ice and first-year ice sites had coverage of 55.6 ± 3.2 and 57.0 ± 3.2% (mean ± SE) (Table S1).
 
-@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`
+## Predator clusters
 
-```{r}
-#| label: demo-inline-code
-#| echo: false
-x <- round(pi, 2)
-```
+We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage (Fig.
+3).
+See Fig.
+S3 for the gap statistic curve, which we used to identify the optimal number of clusters.
 
-Here is an example of inline code `r x` in the middle of a sentence.
+The marginal ice community had the greatest number of indicator species (6) and the highest indicator values (Antarctic fur seals: 0.84; snow petrels: 0.48).
+The highest indicator value in the open water cluster belonged to southern fulmars (0.42), closely followed by 4 other species of petrels.
+The marginal ice and open water clusters shared 3 indicator species: snow petrel, Antarctic petrel, and southern giant petrel.
+The pack ice cluster had only 2 indicator species: Adélie penguins (0.37) and crabeater seals (0.30).
+
+Predator densities varied among clusters (Table S2).
+Snow petrels and Adélie penguins had the greatest overall density across clusters (1.665 and 1.153 ind.
+km^–1^, respectively), though their densities were greatest in the marginal ice zone (3.116 and 4.036 ind.
+km^–1^, respectively).
+Other abundant predators in the marginal ice zone included Antarctic fur seals (2.892 ind.
+km^–1^), crabeater seals (0.437 ind.
+km^–1^), Antarctic petrels (0.258 ind.
+km^–1^), and kelp gulls (0.158 ind.
+km^–1^).
+Antarctic fur seals and snow petrels were ubiquitous in the marginal ice cluster (observed at 100.0 and 96.3% of sites, respectively).
+Despite their high densities, Adélie penguins were only observed at 42.6% of sites in the marginal ice cluster (compared to 67.9% of pack ice sites).
+
+Crabeater seals and Adélie penguins were the most abundant predators in the pack ice cluster (1.380 and 0.948 ind.
+km^–1^), followed by snow petrels (0.463 ind.
+km^–1^), Antarctic petrels (0.177 ind.
+km^–1^), Antarctic fur seals (0.121 ind.
+km^–1^), and kelp gulls (0.108 ind.
+km^–1^).
+Whereas 5 species were observed at 70% or more of the marginal ice sites, only snow petrels (92.5%) were observed at 70% or more of the pack ice sites.
+Predator densities were lower in the open water cluster.
+The most abundant predators, snow petrels (1.554 ind.
+km^–1^) and Antarctic petrels (0.404 ind.
+km^–1^), were also the most prevalent (95.7 and 92.8% of sites, respectively).
+Antarctic fulmars (0.240 ind.
+km^–1^, 59.4% of sites) and blue petrels (0.125 ind.
+km^–1^, 32.6% of sites) were observed in greater densities and at higher frequencies in the open water cluster than other clusters.
+Cape petrels were observed with greater frequency in the open water cluster (47.8% of sites) than the marginal ice (20.4%) and pack ice (13.2%), but their densities were greater in the marginal ice (0.116 ind.
+km^–1^) than the open water (0.084 ind.
+km^–1^) cluster.
+
+The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice (Fig.
+4).
+Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands (Fig.
+5).
+We found predators characteristic of the marginal ice community in the western Bransfield Strait and north of Elephant Island.
+The pack ice community primarily occupied the eastern Bransfield Strait.
+
+## Environmental determinants of community structure
+
+We retained 3 NMDS axes for investigating the relationships between the environment and community structure of predators (stress = 0.135).
+Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; Table 3; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
+The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters (Fig.
+6).
+The marginal ice and open water clusters were well separated by NMDS axis 2.
+
+The predator clusters differed in temperature, salinity, and chl *a*, but not by mixed layer depth or phaeopigment (Table 5).
+Relative to the open water community, the pack ice community was associated with colder, more saline water with lower chl *a* concentrations.
+The marginal ice community occupied intermediate habitats.
+
+The predator clusters were significantly associated with macrozooplankton clusters (χ^*2*^ test, p \< 0.001).
+Five pairwise associations of predators with prey were significant after applying Bonferroni correction (Table 6).
+The pack ice community was positively associated with macrozooplankton community 3a (an extremely diverse assemblage associated with cold, high-salinity water) and negatively associated with cluster 2b (*Thysanoessa macrura*).
+The marginal ice community was positively associated with macrozooplankton community 3b (large krill, including *Euphausia superba*).
+The open water predator community was positively associated with cluster 2b and negatively associated with 3b.
+Macrozooplankton cluster 1 (a ubiquitous cluster indicated by *Salpa* and *Clione*) was not significantly associated with any predator cluster.
+Macrozooplankton cluster 2a (indicated by important prey species such as *E.
+frigida*, *E.
+triacantha*, myctophid larvae, and *Themisto gaudichaudii*) was also not significantly associated with any predator cluster, although it was associated most often with the open water predator cluster (Table 6).
 
@@
 
-# Conclusion
+We analyzed the composition and distribution of predator communities around the northern Antarctic Peninsula in winter to investigate associations among marine predators, their prey, and the physical environment.
+We identified 3 communities of predators, and hypothesized that the spatial organization of these communities is related to the distribution of macrozooplankton communities and a gradient in sea-ice coverage.
+The pagophilic community (e.g.
+Adélie penguins, crabeater seals) consisted of species with diets dominated by Antarctic krill (Table 2) that rely on the pack ice throughout the annual cycle [@ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017].
+Predators in the marginal ice zone (e.g.
+Antarctic fur seals, snow petrels) and in open water (e.g.
+southern fulmars, other petrels) communities also consume Antarctic krill, but their diets are more diverse, with a greater reliance on fish, especially myctophids such as *Electrona antarctica* (Table 2) [@ridoux1989diets; @ainley1992does; @lorentsen1998diet].
+Despite the diet overlap between these 2 communities, they tend to segregate spatially according to ice coverage due to differing foraging strategies.
+Marginal ice zone predators use ambush and pursuit plunging behavior to exploit prey among ice floes, which open water predators have not been observed to do [@ainleySpeciesHabitatRelationshipsAntarctic1993].
+
+The spatial associations between ecological communities and sea-ice conditions largely supported our hypothesis that spatial associations between predator and macrozooplankton communities would fall along a sea-ice coverage gradient.
+Open water predators associated with an offshore zooplankton community, as predicted.
+Two different ice-associated predator communities emerged from our analysis: one that favors marginal sea-ice conditions and the western Bransfield Strait, and another that prefers greater sea-ice coverage and the eastern Bransfield Strait.
+The significant association between marginal ice predators and the *Euphausia superba*/*E.
+crystallorophias* zooplankton community reflects a direct trophic connection between krill predators and their prey.
+However, the mechanism driving the association between pack ice predators and a diverse zooplankton community indicated by chaetognaths, *Calanoides acutus*, siphonophores, and many other taxa (called 3a by @dietrich2021winter) is less clear.
+Many of the indicator taxa for zooplankton community 3a occupy the mesopelagic in open water and marginal ice conditions, but their relative abundances in the epipelagic and surface waters increase when sea-ice coverage is greater [@ainleyAntarcticMesopelagicMicronekton1986; @hopkinsZooplanktonCommunityVicinity1988; @bathmannWinterDistributionOverwintering1993; @floresSeasonalChangesVertical2014].
+Thus, the association between pack ice predators and this macrozooplankton community is likely not a direct trophic relationship, but rather one that reflects shared habitat selection for greater sea-ice coverage.
+
+Two non-indicator species, cape petrels and Antarctic minke whales *Balaenoptera bonaerensis*, are abundant in the northern Antarctic Peninsula during summer, yet neither emerged as an indicator species in our analysis.
+Cape petrels are one of the most common breeding seabirds around the Antarctic Peninsula and form dense hotspots in spring and summer [@huntDistributionAbundanceCommunity1990; @santoraSpatiotemporalPersistenceTop2013; @warwick-evansMultiscaleAssessmentDistribution2021].
+However, in winter, the bulk of the population shifts to lower latitudes as far north as the Patagonian Shelf [@croxallImportancePatagonianShelf2002; @delordSpeciesSpecificForaging2016].
+Consistent with what is known about their habitat associations, cape petrels were more abundant in the open water community than in either of the other 2 predator communities.
+Nonetheless, their overall abundance was sufficiently low that they did not meet the threshold for indicator species.
+Previous surveys of the winter seabird assemblages around the Antarctic Peninsula observed greater abundances of cape, blue, and Kerguelen petrels in the offshore areas west and north of our study area [@whitehouseDistributionAbundanceSeabirds1994; @ribicWaterMassesOcean2011].
+Like cape petrels, Antarctic minke whales are thought to be abundant predators in the northern Antarctic Peninsula during summer; however, their preference for dense sea-ice coverage presents a substantial challenge for abundance estimation [@herrAerialSurveysAntarctic2019; @rischCommonAntarcticMinke2019].
+Although part of the population migrates to lower latitudes in winter [@leeBehaviorSatellitetrackedAntarctic2017; @espositoFirstSightingsAntarctic2021], Antarctic minke whales are acoustically detectable at higher latitudes year-round [@dominelloSeasonalityAntarcticMinke2016; @filunFrozenVersesAntarctic2020].
+The characteristic ‘bio-duck’ minke call is frequently detected during austral winter east of our study region in the Weddell Sea [@filunFrozenVersesAntarctic2020] and further south along the western Antarctic Peninsula [@dominelloSeasonalityAntarcticMinke2016].
+Therefore, it is possible that Antarctic minke whales were more abundant in our study region than our data indicate, distributed in areas inaccessible to research vessels.
+Given the limited sample sizes and biases in our data, we cannot determine the predator community associations of Antarctic minke whales.
+Our analysis included 43 Antarctic minke whales: 30 individuals at 7 marginal ice sites, 8 individuals at 5 open water sites, and 5 individuals at 5 pack ice sites.
+This pattern of relative abundance may indicate that Antarctic minke whales are members of the marginal ice predator community (which was also associated with the prey of these species, large-bodied krill).
+Alternatively, Ainley et al.
+@ainleyCetaceanOccurrencePatterns2007 suggested that Antarctic minke whales are pagophilic members of the pack ice predator community, and their apparent association with marginal ice communities is the result of sampling bias due to habitat inaccessibility.
+
+The winter seabird assemblage that we documented seems to suggest a shift of southerly species northwards to the edge of the marginal ice zone (snow and Antarctic petrels broadly, and Adélie penguins in the eastern region) as well as ubiquitous species, such as kelp gulls and southern giant petrels.
+These latter species opportunistically forage and scavenge within the marginal ice zone.
+Positive associations between these ice-dependent seabirds and pinnipeds likely represent a stable foraging community and reflect concentrations of krill and other macrozooplankton within the marginal ice zone.
+Therefore, increases in sea-ice concentrations towards maximum coverage (no open water present) would likely shift the entire predator community further to the north, as access to open water would be limited to the south.
+
+Our data represent a baseline of biophysical conditions during a period of relatively high sea-ice coverage compared to recent declines [@eayrsRapidDeclineAntarctic2021].
+Our 5 yr winter synthesis of sea ice, hydrography, macrozooplankton, and top predators is unlikely to be replicated in the near future, due to increased costs of shipboard research, reduced availability of research vessels, and prioritized sampling during the Antarctic summer.
+The spatial averages of community composition (i.e.
+predator clusters; Fig.
+5) may be considered essential foraging habitats for overwintering predators.
+The open water assemblage occupies the most temporally variable ocean habitat that we sampled, which changes based on currents and latitudinal position of the marginal ice zone.
+The marginal ice habitat was considerably smaller, with 2 important concentrations: southwest part of the Bransfield Strait, and open water to the north of Elephant Island.
+Antarctic krill are consistently concentrated in the southern region over multiple years [@reissOverwinterHabitatSelection2017], and stable sea-ice concentrations are more likely to occur there due to coastal ocean processes (coastal currents, formation of eddies) and shifts in ice from higher latitudes [@moffatShelfOceanExchange2018].
+We found consistently high concentrations of crabeater seals and fur seals within both marginal ice zone habitats (Figs.
+2 & 4), and the presence of sea ice likely contributes by supporting haul-out habitat for seal thermoregulation and rest between foraging bouts.
+There were more concentrations of fur seals compared to crabeater seals within the marginal ice zone north of Elephant Island.
+We did not always encounter high concentrations of krill in this region, which might explain the reduction in crabeater seals, or that fur seals were targeting other forage taxa, perhaps myctophids [@daneriDietAntarcticFur2005; @politoInvestigatingUseStable2010; @descalzoRelevanceFishSummer2023], that our net hauls were inadequate at catching.
+
+Comparatively, the marginal ice zone habitat north of Elephant Island is likely more hydrographically dynamic and less stable or predictable as suitable foraging habitat for ice-obligate species, compared to the southern Bransfield Strait.
+Within the southern Drake Passage, the southern Antarctic Circumpolar Current and its associated frontal boundary likely influences the rate at which sea ice can form and accumulate within the Elephant Island area [@stammerjohnSeaIceWestern2008].
+Winds and ocean currents play a significant role in how sea ice forms, and ocean atmospheric conditions further to the south are more stable for facilitating sea-ice growth [@massomAntarcticSeaIce2010].
+The pack ice community where we repeatedly encountered dense first- and multi-year ice floes was restricted to the eastern Bransfield Strait and reflects ocean conditions from the higher-latitude Weddell Sea, and where we observed some of the largest concentrations of Adélie penguins and crabeater seals.
+During winter, this essential foraging habitat likely reflects the only available habitat within this region, as areas to the south are mostly completely ice covered with little access to open water for diving predators.
+Therefore, declines in sea ice within this area would likely have significant impacts on the ecology and overwintering survival of these ice-dependent predators.
+
+Species interactions play a key role in the community organization of marine predators [@stachowiczMutualismFacilitationStructure2001; @veitPositiveInteractionsForaging2017].
+Identifying habitat utilization and ecological associations of individual species is crucial for establishing a robust baseline for their needs and formulating conservation strategies.
+However, examining habitat associations within species assemblages may provide further insights into how members of a community interact with each other [@hoffmanEcologySeabirdFeeding1981; @levinIntegratedEcosystemAssessments2009].
+Further, if the presence or abundance of one species is critical to the stability and resilience of the community to environmental change, then the changes in the abundance structure of that community need to be assessed [@stachowiczMutualismFacilitationStructure2001; @dakosCriticalSlowingEarly2014].
+For example, we found a predator community within the marginal ice zone that contained flying (gulls, terns, and petrels) and diving (fur seals) predators.
+If the aerial predators rely on diving predators to drive prey nearer to the surface [@harrisonObservationsMultispeciesSeabird1991; @veitPositiveInteractionsForaging2017], then changes in sea-ice habitat used by fur seals for haul out and rest would have additional indirect effects on the winter feeding ecology of flying seabirds within this region [@evansAssociationsSeabirdsCetaceans1982; @veitAggregationPatternsPelagic1993; @monierChangesPositiveAssociations2020].
+Our analysis of species assemblages within 3 distinct communities (i.e.
+open water, marginal ice, and pack ice) may provide a simplified approach to examining future changes in the entire predator community, as opposed to multiple assessments of individual species’ responses to climate change.
+Therefore, we recommend that future studies aiming to link climate change projections, including changes in winter sea-ice concentration, assess how the 3 communities may shift or contract/expand geographically under different scenarios.
+Importantly, the geographic extent, environmental conditions, and macrozooplankton species associations underlying the 3 communities (Tables 4–6) are provided and can be used to develop predator community habitat models allowing for prediction in years outside of the intensive sampling period.
+Suggested geographic community analyses may include quantifying the habitat size and range (e.g.
+horizontal area and extent) of each community relative to various climate change scenarios.
+This community synthesis should provide helpful data for assessing climate change impacts, as well as winter-related conservation and ecosystem management needs.
+
+There is concern that expansion of the Antarctic krill fishery into austral autumn and winter may impact the overwintering ecology of seabirds and marine mammals [@hinkeIdentifyingRiskConcurrent2017; @wattersLongtermObservationsAntarctica2020; @trathanEcosystemApproachManagement2022].
+Potential impacts include bycatch through fishing gear interactions, entanglement, and ship strikes, and possibly depletion of key prey patches that predators are reliant upon.
+Given the spatial consistency of Antarctic krill concentrations in southern Bransfield Strait during winter [@reissOverwinterHabitatSelection2017; @dietrich2021winter] and the dense concentrations of pinnipeds and seabirds present within the marginal sea-ice zone, increased fishing activity within this location has potential negative consequences to the predator community.
+Further, as global warming exacerbates long-term reductions in winter sea-ice concentration and extent, lower sea-ice years may allow for greater seasonal expansion of the krill fishery, thereby reducing important prey resources and creating potentially more conflicts with predators that are unable to shift their distributions.
+
+# Figures
+
+```{r}
+#| label: fig-top6map
+
+plot(1, 1)
+
+```
 

### 2024-05-31 analysis/paper/paper.qmd
Figure 1

@@
   comment: "#>"
-  fig-path: "../figures/"
   fig-dpi: 600
+knitr:
+  opts_chunk: 
+    dev: # Output figures as .png and .eps 
+      - png
+      - postscript 
+    fig.path: "../figures/"
 filters:
@@
 Keywords: `r rmarkdown::metadata$keywords`
+
+```{r}
+#| label: load
+#| echo: false
+
+# Load this analysis as a package
+devtools::load_all()
+
+```
 
@@
 #| label: fig-top6map
+#| fig-cap: "Geographic distributions of the 6 most abundant species observed during the surveys, averaged across survey years: (A) Adélie penguin, (B) Antarctic fur seal, (C) Antarctic petrel, (D) crabeater seal, (E) snow petrel, (F) southern fulmar.
+Unlike the cluster analysis, these maps aggregate sightings from all visual surveys within the study area.
+The sightings associated with sites are a subset of these (i.e.
+sightings more than 15 km or 3 d from a sampling station are included here)."
+#| out-width: 169mm
+#| out-height: 121mm
 
-plot(1, 1)
+make_fig_top6map(read_obj("topsix_clim_class", rast = TRUE))
 

### 2024-05-31 R/figures.R

@@ -1,0 +1,41 @@
+#' Create a figure with climatological distribution of six most abundant
+#' predators
+#'
+#' As a side effect, saves figure to EPS in figures/
+#'
+#' @return ggplot object
+#' @export
+make_fig_top6map <- function(topsix_clim_class) {
+  ant_sf <- read_obj("ant_sf")
+
+  # Map bounding box, projected
+  map_bbox <- terra::ext(-63, -54, -64, -59) %>%
+    terra::project(from = "+proj=longlat +datum=WGS84",
+                   to = as.character(ant_proj())[[1]])
+
+  # Map sublabels
+  sublabels <- tibble(lyr = names(topsix_clim_class),
+                      x = -62.25, y = -59.25) %>%
+    arrange(lyr) %>%
+    mutate(labels = LETTERS[1:6]) %>%
+    sf:: st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+
+  ant_basemap() +
+    tidyterra::geom_spatraster(data = topsix_clim_class) +
+    geom_sf(data = ant_sf) +
+    geom_sf_text(aes(label = labels), sublabels, fontface = "bold") +
+    scale_fill_brewer(palette = "RdPu",
+                      na.value = NA,
+                      direction = 1,
+                      na.translate = FALSE) +
+    scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+    scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+    labs(fill = expression("Ind km"^-1)) +
+    facet_wrap(~lyr) +
+    coord_sf(xlim = map_bbox[1:2],
+             ylim = map_bbox[3:4],
+             crs = ant_proj()) +
+    theme(legend.position = "top",
+          panel.background = element_rect(color = "grey20"),
+          strip.text = element_blank())
+}

### 2024-05-31 R/predators.R

@@ -1,8 +1,13 @@
+#' Aggregate predator sightings
+#'
+#' @param predators Predator data (from raw CSV file)
+#'
+#' @return Predators aggregated in intervals
+#' @export
 aggregate_predators <- function(predators) {
   predators %>%
     filter(species != "NULL") %>%
-    mutate(UTC_start = lubridate::mdy_hm(UTC_start),
-           year = lubridate::year(UTC_start),
+    mutate(year = lubridate::year(UTC_start),
            km = nmi_to_km(nmi)) %>%
     group_by(year, interval, species, UTC_start, lon_mean, lat_mean, nmi, km) %>%
     summarize(count = sum(count_species), .groups = "drop")
@@ -59,6 +64,12 @@ normalize_counts <- function(sightings, effort) {
            count_km = nmi_to_km(count_nmi))
 }
 
+#' Convert species code to common name
+#'
+#' @param species_code 4-letter species codes (character vector)
+#'
+#' @return Species' common names
+#' @export
 code_to_common <- function(species_code) {
   c(ADPN = "Adélie penguin",
     ANFU = "Southern fulmar",

### 2024-05-31 R/sp.R

@@ -1,33 +1,45 @@
-# Use WGS 84 / Antarctic Polar Stereographic for mapping, recentered on ant_lims()
+#' Use WGS 84 / Antarctic Polar Stereographic for mapping, recentered on ant_lims()
+#'
+#' @return CRS object
+#' @export
 ant_proj <- function() {
   lat_ts <- (ant_lims()["ymin"] + ant_lims()["ymax"]) / 2
   lon_0 <- (ant_lims()["xmin"] + ant_lims()["xmax"]) / 2
-  glue::glue(
-    "+proj=stere +lat_0=-90 +lat_ts={lat_ts} +lon_0={lon_0} +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
-  ) %>%
-  st_crs()
+  sf::st_crs(str_glue("+proj=stere +lat_0=-90 +lat_ts={lat_ts} +lon_0={lon_0} +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"))
 }
 
+#' Spatial limits of study region
+#'
+#' @return A named vector with xmin, xmax, ymin, and ymax values
+#' @export
 ant_lims <- function() {
   c(xmin = -67, xmax = -50, ymin = -68, ymax = -55)
 }
 
-download_ne <- function(ne_dir) {
-  if (!dir.exists(ne_dir)) dir.create(ne_dir, recursive = TRUE)
-  rnaturalearth::ne_download(scale = "large",
-                             type = "land",
-                             category = "physical",
-                             destdir = ne_dir,
-                             returnclass = "sf")
-}
-
+#' Create Antarctic Peninsula sf object using Natural Earth data
+#'
+#' @param ne_dir directory to store Natural Earth data
+#' @param limits spatial limits
+#'
+#' @return Antarctic Peninsula sf object
+#' @export
 create_ant_sf <- function(ne_dir, limits = ant_lims()) {
+  # Download data from Natural Earth, if necessary
+  if (!dir.exists(ne_dir)) dir.create(ne_dir, recursive = TRUE)
+  if (!file.exists(file.path(ne_dir, "ne_10m_land.shp")))
+    rnaturalearth::ne_download(scale = "large",
+                               type = "land",
+                               category = "physical",
+                               destdir = ne_dir,
+                               returnclass = "sf")
+
+  # Load and crop data
   land <- rnaturalearth::ne_load(scale = "large",
                                  type = "land",
                                  category = "physical",
[…]

### 2024-05-31 R/util.R

@@ -1,0 +1,43 @@
+#' Save an object to derived_data/
+#'
+#' @param obj object to be saved
+#' @param obj_name name of object
+#' @param rast whether object is a terra raster. false by default. requires
+#'   extra steps.
+#'
+#' @return nothing, but saves object in analysis/data/derived_data/
+#' @export
+save_obj <- function(obj, obj_name, rast = FALSE) {
+  if (rast) {
+    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.tif"))
+    terra::writeRaster(obj, path, overwrite = TRUE)
+  } else {
+    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.rds"))
+    saveRDS(obj, path)
+  }
+}
+
+#' Read an object from derived_data/
+#'
+#' @param obj_name name of object
+#' @param rast whether object is a terra raster. false by default. requires
+#'   extra steps.
+#'
+#' @return object read from file in analysis/data/derived_data/
+#' @export
+read_obj <- function(obj_name, rast = FALSE) {
+  if (rast) {
+    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.tif"))
+    terra::rast(path)
+  } else {
+    readRDS(here::here(str_glue("analysis/data/derived_data/{obj_name}.rds")))
+  }
+}
+
+# The following allows internal use of some tidyverse packages without using ::
+#' @import dplyr
+#' @import ggplot2
+#' @import purrr
+#' @import readr
+#' @import stringr
+NULL

### 2024-05-31 analysis/scripts/01_process_data.R

@@ -1,0 +1,38 @@
+# Load this research compendium as a package
+# This is functionally equivalent to library(antwincomm)
+devtools::load_all()
+
+# Check raw data have been downloaded
+if (!all(file.exists(file.path(
+  "analysis", "data", "raw_data", "AntPeninsula_winter_marine_data",
+  c("Ice.csv", "PredatorSightings.csv", "StationZooplanktonPhysics.csv")
+)))) {
+  stop("RAW DATA NOT FOUND. See: analysis/data/raw_data/README.md.")
+}
+
+# Predator data
+predators_file <- "analysis/data/raw_data/AntPeninsula_winter_marine_data/PredatorSightings.csv"
+predators <- read_csv(predators_file, show_col_types = FALSE)
+predators_agg <- aggregate_predators(predators)
+save_obj(predators_agg, "predators_agg")
+
+message("Predator data processed")
+
+# Effort data
+effort_sf <- predators %>%
+  distinct(cruise, interval, UTC_start, UTC_end, nmi,
+           lon_mean, lat_mean) %>%
+  filter(nmi < 1000) %>%
+  mutate(year = lubridate::year(UTC_start),
+         species = "survey",
+         km = nmi_to_km(nmi)) %>%
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+save_obj(effort_sf, "effort_sf")
+
+message("Effort data processed")
+
+# Natural Earth data
+ant_sf <- create_ant_sf(ne_dir = "analysis/data/derived_data/ne")
+save_obj(ant_sf, "ant_sf")
+
+message("Natural Earth data processed")

### 2024-05-31 analysis/scripts/02_species_climatologies.R

@@ -1,0 +1,69 @@
+# Load this research compendium as a package
+# This is functionally equivalent to library(antwincomm)
+devtools::load_all()
+
+# Load necessary derived data
+predators_agg <- read_obj("predators_agg")
+effort_sf <- read_obj("effort_sf")
+
+# Find six most abundant species
+top_six <- predators_agg %>%
+  group_by(species) %>%
+  summarize(total_count = sum(count)) %>%
+  arrange(desc(total_count)) %>%
+  slice(1:6) %>%
+  mutate(common_name = code_to_common(species))
+
+# Convert abundant species' data to sf
+topsix_sf <- predators_agg %>%
+  semi_join(top_six, by = "species") %>%
+  left_join(select(top_six, species, common_name), by = "species") %>%
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+
+# Create a template raster of the study area
+effort_ext <- effort_sf %>%
+  sf::st_transform(ant_proj()) %>%
+  terra::ext()
+rast_template <- terra::rast(crs = as.character(ant_proj())[[1]],
+                             extent = effort_ext,
+                             resolution = 30e3)
+
+# Rasterize effort by year
+effort_rast <- effort_sf %>%
+  sf::st_transform(ant_proj()) %>%
+  rasterize_by("km", "year", sum, rast_template)
+
+# Rasterize the predator sightings by year and species
+topsix_rast <- topsix_sf %>%
+  sf::st_transform(ant_proj()) %>%
+  mutate(year_commonname = as.character(interaction(year, common_name))) %>%
+  rasterize_by("count", "year_commonname", sum, rast_template)
+
+# Normalize counts by effort
+topsix_norm <- topsix_rast
+for (i in seq(terra::nlyr(topsix_rast))) {
+  lyr_name <- names(topsix_rast)[i]
+  lyr_year <- substr(lyr_name, 1, 4)
+  effort_year <- effort_rast[lyr_year] %>%
+    terra::subst(0, NA)
+  topsix_norm[[i]] <- topsix_norm[[i]] / effort_year
+}
+
+# Average across years
+topsix_climatology <- unique(top_six$common_name) %>%
+  map(function(sp) {
+    sp_rast <- terra::subset(topsix_norm,
+                             which(str_detect(names(topsix_norm), sp)))
+    terra::mean(sp_rast, na.rm = TRUE)
+  }) %>%
[…]

### 2024-06-06 analysis/paper/paper.qmd
Figure 3

@@
 
-make_fig_top6map(read_obj("topsix_clim_class", rast = TRUE))
+make_fig_top6map(read_obj("topsix_clim_class", type = "rast"))
 
+```
+
+```{r}
+#| label: fig-seaice
+#| fig-cap: "Distribution of sea ice coverage among sites, by predator cluster: (A) open water, (B) marginal ice, (C) pack ice. Dashed red line: median ice coverage."
+#| out-width: 169mm
+#| out-height: 121mm
+
+make_fig_seaiceclust(read_obj("stations_clust"))
 ```

### 2024-06-06 R/figures.R

@@ -1,8 +1,6 @@
 #' Create a figure with climatological distribution of six most abundant
 #' predators
 #'
-#' As a side effect, saves figure to EPS in figures/
-#'
 #' @return ggplot object
 #' @export
 make_fig_top6map <- function(topsix_clim_class) {
@@ -38,4 +36,38 @@ make_fig_top6map <- function(topsix_clim_class) {
     theme(legend.position = "top",
           panel.background = element_rect(color = "grey20"),
           strip.text = element_blank())
+}
+
+#' Create a figure with distribution of sea ice by predator cluster
+#'
+#' @return ggplot object
+#' @export
+make_fig_seaiceclust <- function(stations_clust) {
+  median_coverge <- stations_clust %>%
+    drop_na(ice_coverage) %>%
+    group_by(pred_clust) %>%
+    summarize(median_ice = median(ice_coverage / 10, na.rm = TRUE))
+
+  sublabels <- tibble(
+    pred_clust = factor(levels(stations_clust$pred_clust)),
+    labels = LETTERS[1:3],
+    x = 1, y = 15
+  )
+
+  ggplot(stations_clust,
+              aes(x = ice_coverage / 10, fill = after_stat(x))) +
+    geom_histogram(bins = 20, color = "grey30") +
+    geom_vline(aes(xintercept = median_ice),
+               median_coverge,
+               color = "red", linetype = "dashed") +
+    geom_text(aes(x = x, y = y, label = labels), sublabels,
+              hjust = 1, vjust = 1,
+              fontface = "bold") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_fill_distiller(palette = "Blues", guide = NULL) +
+    facet_grid(rows = vars(pred_clust)) +
+    labs(x = "Ice coverage",
+         y = "# sites") +
+    theme_classic() +
+    theme(strip.text = element_blank())
 }

### 2024-06-06 R/ice.R

@@ -5,8 +5,14 @@ mfv <- function(x) {
   names(x_tbl)[which.max(x_tbl)]
 }
 
-# Clean ice categories
-# Converts e.g., FY/MY to FY and collapses SLUSH and BRASH into THIN
+#' Clean ice categories
+#'
+#' Converts e.g., FY/MY to FY and collapses SLUSH and BRASH into THIN
+#'
+#' @param ice_types Character vector of ice types
+#'
+#' @return Cleaned up ice types
+#' @export
 clean_ice <- function(ice_types) {
   str_extract(ice_types, "[^/]+") %>%
     factor(levels = c("OPEN",
@@ -29,18 +35,18 @@ aggregate_ice <- function(stations, ice_raw) {
   stations2 <- stations %>%
     transmute(amlr.station,
               year_station = Year,
-              date_station = as.Date(date.tow))
+              date_station = as.Date(start.time.UTC))
 
   # Spatially join stations to raw ice observations
   ice_stations <- ice_raw %>%
     group_by(year_ice) %>%
     group_modify(function(ice_by_year, ice_keys) {
       stations_by_year <- filter(stations2, year_station == ice_keys$year_ice)
-      nearest_stations_idx <- st_nearest_feature(ice_by_year, stations_by_year)
+      nearest_stations_idx <- sf::st_nearest_feature(ice_by_year, stations_by_year)
       nearest_stations <- stations_by_year[nearest_stations_idx, ]
-      station_distance <- st_distance(ice_by_year,
-                                      nearest_stations,
-                                      by_element = TRUE)
+      station_distance <- sf::st_distance(ice_by_year,
+                                          nearest_stations,
+                                          by_element = TRUE)
       station_lag <- ice_by_year$date_ice - nearest_stations$date_station
       stations_no_sf <- select(as_tibble(nearest_stations), -geometry)
       ice_by_year %>%
@@ -51,7 +57,7 @@ aggregate_ice <- function(stations, ice_raw) {
     ungroup() %>%
     filter(station_distance <= units::as_units(15, "km"),
            abs(station_lag) <= as.difftime(3, units = "days")) %>%
-    st_as_sf(sf_column_name = "geometry")
+    sf::st_as_sf(sf_column_name = "geometry")
 
   # Aggregate ice observations by station
   ice_stations %>%
[…]

### 2024-06-06 R/predators.R

@@ -23,6 +23,15 @@ filter_species <- function(sightings, station_thr) {
     semi_join(x = sightings, y = ., by = "species")
 }
 
+#' Assign sightings to nearest station
+#'
+#' @param sightings predator sighting data
+#' @param stations station data
+#' @param max_dist_km maximum spatial tolerance (km)
+#' @param max_days maximum temporal tolerance (days)
+#'
+#' @return predator data associated with nearest station
+#' @export
 assign_sightings <- function(sightings, stations, max_dist_km, max_days) {
   stopifnot(inherits(sightings, "sf"),
             inherits(stations, "sf"))
@@ -31,16 +40,16 @@ assign_sightings <- function(sightings, stations, max_dist_km, max_days) {
   assign_group <- function(sightings_group, sightings_key) {
     station_buffers <- stations %>%
       filter(Year == sightings_key$year) %>%
-      st_transform(ant_proj()) %>%
-      st_buffer(max_dist_km * 1000) %>%
-      transmute(
+      sf::st_transform(ant_proj()) %>%
+      sf::st_buffer(max_dist_km * 1000) %>%
+      select(
         amlr.station,
-        tow.UTC = date.tow + (start.time.UTC - as.POSIXct("1899-12-31", tz = "UTC"))
+        start.time.UTC
       )
     sightings_group %>%
-      st_transform(ant_proj()) %>%
-      st_join(station_buffers, left = FALSE) %>%
-      mutate(lag_days = abs(as.numeric(UTC_start - tow.UTC, units = "secs")) / 3600 / 24) %>%
+      sf::st_transform(ant_proj()) %>%
+      sf::st_join(station_buffers, left = FALSE) %>%
+      mutate(lag_days = abs(as.numeric(UTC_start - start.time.UTC, units = "secs")) / 3600 / 24) %>%
       filter(lag_days <= max_days)
   }
 
@@ -49,7 +58,7 @@ assign_sightings <- function(sightings, stations, max_dist_km, max_days) {
     group_by(year) %>%
     group_modify(assign_group) %>%
     ungroup() %>%
-    st_as_sf()
+    sf::st_as_sf()
 }
 
 # Normalize species counts by survey effort
@@ -57,7 +66,7 @@ normalize_counts <- function(sightings, effort) {
   sightings %>%
     as_tibble() %>%
     group_by(amlr.station, species) %>%
-    summarize(count = sum(count), .groups = "drop") %>%
[…]

### 2024-06-06 R/util.R

@@ -2,42 +2,59 @@
 #'
 #' @param obj object to be saved
 #' @param obj_name name of object
-#' @param rast whether object is a terra raster. false by default. requires
-#'   extra steps.
+#' @param type whether object is a basic R object, raster, or vector. Raster and
+#'   vector data require additional processing.
 #'
 #' @return nothing, but saves object in analysis/data/derived_data/
 #' @export
-save_obj <- function(obj, obj_name, rast = FALSE) {
-  if (rast) {
-    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.tif"))
-    terra::writeRaster(obj, path, overwrite = TRUE)
-  } else {
-    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.rds"))
-    saveRDS(obj, path)
-  }
+save_obj <- function(obj, obj_name, type = c("r", "rast", "vect")) {
+  type = match.arg(type)
+  switch (type,
+    r = {
+      path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.rds"))
+      saveRDS(obj, path)
+    },
+    rast = {
+      path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.tif"))
+      terra::writeRaster(obj, path, overwrite = TRUE)
+    },
+    vect = {
+      path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.shp"))
+      terra::writeVector(obj, path, overwrite = TRUE)
+    }
+  )
 }
 
 #' Read an object from derived_data/
 #'
 #' @param obj_name name of object
-#' @param rast whether object is a terra raster. false by default. requires
-#'   extra steps.
+#' @param type whether object is a basic R object, raster, or vector. Raster and
+#'   vector data require additional processing.
 #'
 #' @return object read from file in analysis/data/derived_data/
 #' @export
-read_obj <- function(obj_name, rast = FALSE) {
-  if (rast) {
-    path <- here::here(str_glue("analysis/data/derived_data/{obj_name}.tif"))
-    terra::rast(path)
-  } else {
-    readRDS(here::here(str_glue("analysis/data/derived_data/{obj_name}.rds")))
-  }
+read_obj <- function(obj_name, type = c("r", "rast", "vect")) {
+  type = match.arg(type)
+  switch (
+    type,
+    r = readRDS(here::here(str_glue("analysis/data/derived_data/{obj_name}.rds"))),
+    rast = {
[…]

### 2024-06-06 R/zoop.R

@@ -1,14 +1,19 @@
+#' Read station (zooplankton) data
+#'
+#' @param zoop_path path to StationZooplanktonPhysics.csv
+#'
+#' @return data frame with station-level data
+#' @export
 read_zoop <- function(zoop_path) {
-  readxl::read_excel(zoop_path, sheet = 1) %>%
-    mutate(time.of.day = factor(time.of.day,
-                                levels = c("D", "T", "N"),
-                                labels = c("Day", "Twilight", "Night")),
-           across(c(zuml_m,
-                    starts_with("avg"),
-                    starts_with("Integ"),
-                    Net.ice.percent),
-                  parse_number)) %>%
-    select(-57) %>%
+  sun_angle <- function(t, x, y) {
+    pmap_dbl(list(t, x, y), \(.t, .x, .y) oce::sunAngle(.t, .x, .y)$altitude)
+  }
+
+  read_csv(zoop_path, show_col_types = FALSE) %>%
+    mutate(sun_angle = sun_angle(start.time.UTC, dec.longitude, dec.latitude),
+           time.of.day = cut(sun_angle,
+                             c(-Inf, -18, 0, Inf),
+                             labels = c("Night", "Twilight", "Day"))) %>%
     filter(leg == "W")
 }
 

### 2024-06-06 analysis/scripts/01_process_data.R

@@ -10,15 +10,19 @@ if (!all(file.exists(file.path(
   stop("RAW DATA NOT FOUND. See: analysis/data/raw_data/README.md.")
 }
 
-# Predator data
+# Predator data -----------------------------------------------------------
 predators_file <- "analysis/data/raw_data/AntPeninsula_winter_marine_data/PredatorSightings.csv"
-predators <- read_csv(predators_file, show_col_types = FALSE)
+predators <- read_csv(predators_file, show_col_types = FALSE) %>%
+  mutate(year = lubridate::year(UTC_start))
+predators_sf <- latlon_to_sf(predators, coords = c("lon_mean", "lat_mean"))
 predators_agg <- aggregate_predators(predators)
+save_obj(predators_sf, "predators_sf")
 save_obj(predators_agg, "predators_agg")
 
 message("Predator data processed")
 
-# Effort data
+
+# Effort data -------------------------------------------------------------
 effort_sf <- predators %>%
   distinct(cruise, interval, UTC_start, UTC_end, nmi,
            lon_mean, lat_mean) %>%
@@ -31,8 +35,65 @@ save_obj(effort_sf, "effort_sf")
 
 message("Effort data processed")
 
-# Natural Earth data
+
+# Zooplankton data --------------------------------------------------------
+zoop_file <- "analysis/data/raw_data/AntPeninsula_winter_marine_data/StationZooplanktonPhysics.csv"
+zoop <- read_zoop(zoop_file)
+zoop_sf <- latlon_to_sf(zoop, coords = c("dec.longitude", "dec.latitude"))
+save_obj(zoop_sf, "zoop_sf")
+
+message("Zooplankton data processed")
+
+
+# Ice data ----------------------------------------------------------------
+
+ice_file <- "analysis/data/raw_data/AntPeninsula_winter_marine_data/Ice.csv"
+ice_raw <- read_csv(ice_file, na = c("", "NULL")) %>%
+  transmute(year_ice = Year,
+            lat = Latitude,
+            lon = Longitude,
+            ice_code = as.numeric(IceCode),
+            ice_type = clean_ice(IceType),
+            date_ice = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
+  filter(ice_type != "NULL",
+         # 1 ice_code is "85"
+         ice_code <= 10) %>%
+  latlon_to_sf(coords = c("lon", "lat"))
+save_obj(ice_raw, "ice_raw")
+
[…]

### 2024-06-06 analysis/scripts/02_species_climatologies.R

@@ -64,6 +64,6 @@ topsix_clim_class <- topsix_climatology %>%
   terra::classify(rcl = c(0, 1, 5, 10, 50))
 
 # Save to derived data
-save_obj(topsix_clim_class, "topsix_clim_class", rast = TRUE)
+save_obj(topsix_clim_class, "topsix_clim_class", type = "rast")
 
 message("Species climatologies generated")

### 2024-06-06 analysis/scripts/03_cluster_analysis.R

@@ -1,0 +1,33 @@
+# Load this research compendium as a package
+# This is functionally equivalent to library(antwincomm)
+devtools::load_all()
+
+# Load necessary derived data
+predators_stations <- read_obj("predators_stations")
+effort_sf <- read_obj("effort_sf")
+zoop_sf <- read_obj("zoop_sf")
+stations_ice <- read_obj("stations_ice")
+
+# Cluster sites by predator sightings
+sightings_mtx <- sightings_to_matrix(predators_stations)
+sightings_dist <- vegan::vegdist(sightings_mtx, method = "bray")
+sightings_clust <- hclust(sightings_dist, method = "ward.D2")
+sightings_gap <- cluster_gap(sightings_mtx, max_k = 10, seed = 139)
+sightings_cut <- cutree(sightings_clust, k = 3)
+
+# Indicator values
+sightings_indval <- labdsv::indval(sightings_mtx, sightings_cut)
+
+# Associate predators and sites with predator clusters
+predators_clust <- mutate(
+  predators_stations,
+  pred_clust = station_to_cluster(amlr.station, sightings_cut)
+)
+
+stations_clust <- zoop_sf %>%
+  left_join(stations_ice, by = "amlr.station") %>%
+  mutate(pred_clust = station_to_cluster(amlr.station, sightings_cut)) %>%
+  semi_join(predators_clust, by = "amlr.station")
+
+# Save derived objects
+save_obj(stations_clust, "stations_clust")

### 2024-06-06 analysis/paper/paper.qmd
Figures 1 and 4

@@
 
+![Overview of study area.
+Circles indicate the locations of the US Antarctic Marine Living Resources (AMLR) station grid.
+Solid and dashed gray lines represent the 1000 and 2500 m isobaths, respectively.
+Not all stations were sampled every year.
+Inset indicates the location of the study area (red shading) relative to South America and the Antarctic Peninsula (BS: Bellingshausen Sea; WS: Weddell Sea).
+Bathymetry from IBCSO v2 [@dorschelInternationalBathymetricChart2022]; land from Natural Earth (www.naturalearthdata.com).](images/fig-overviewmap.png){#fig-overviewmap width="116mm"}
+
 ```{r}
@@
 
+```{r}
+#| label: fig-predclustannual
+#| fig-cap: "Geographic distribution of predator communities in each survey year (A–E: 2012–2016, respectively).
+Areas of extensive sea-ice coverage are indicated by blue gradient, where satellite-derived mean sea-ice coverage in the month of August in each year exceeded 65%."
+#| out-width: 169mm
+#| out-height: 121mm
+
+make_fig_predclustannual(
+  read_obj("stations_clust"), 
+  read_obj("seaice_conc_df")
+)
+
+```
+
 # Acknowledgements

### 2024-06-06 R/figures.R

@@ -71,3 +71,48 @@ make_fig_seaiceclust <- function(stations_clust) {
     theme_classic() +
     theme(strip.text = element_blank())
 }
+
+#' Create a figure with geographic distribution of predator clusters by year
+#'
+#' @return ggplot object
+#' @export
+make_fig_predclustannual <- function(stations_clust, seaice_conc_df) {
+  map_lim <- sf::st_bbox(stations_clust) %>%
+    project_bbox() %>%
+    expand_bbox(factor = 1.2)
+
+  sublabels <- tibble(Year = 2012:2016,
+                      x = -63, y = -60,
+                      labels = LETTERS[1:5]) %>%
+    sf::st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+
+  ant_basemap() +
+    # Sea ice
+    geom_tile(aes(x, y, fill = seaice_conc),
+              filter(seaice_conc_df, between(seaice_conc, 0.65, 1)),
+              alpha = 0.8) +
+    scale_fill_distiller("Ice concentration",
+                         lim = c(0.65, 1),
+                         labels = scales::percent,
+                         palette = "Blues",
+                         na.value = "transparent",
+                         guide = guide_colorbar(title.position = "top")) +
+    # Predator clusters
+    ggnewscale::new_scale_fill() +
+    geom_sf(aes(fill = pred_clust),
+            stations_clust,
+            size = 2,
+            shape = 22) +
+    facet_wrap(~ Year) +
+    # Labels
+    geom_sf_text(aes(label = labels), sublabels, fontface = "bold", vjust = 0) +
+    scale_x_continuous(breaks = c(-60, -55)) +
+    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+    scale_fill_brewer("Predator cluster", palette = "Dark2",
+                      guide = guide_legend(override.aes = list(size = 3),
+                                           title.position = "top")) +
+    coord_ant(map_lim) +
+    theme(legend.position = "bottom",
+          strip.text = element_blank(),
+          panel.background = element_rect())
+}

### 2024-06-06 R/sp.R

@@ -125,13 +125,26 @@ coord_ant <- function(map_lim = NULL) {
            crs = ant_proj())
 }
 
+#' Project a bounding box
+#'
+#' @param x sf::bbox
+#'
+#' @return sf::bbox x projected from lat/lon into Antarctic projection (see
+#'   ant_proj())
+#' @export
 project_bbox <- function(x) {
   x %>%
-    st_as_sfc(crs = "EPSG:4326") %>%
-    st_transform(ant_proj()) %>%
-    st_bbox()
+    sf::st_as_sfc(crs = "EPSG:4326") %>%
+    sf::st_transform(ant_proj()) %>%
+    sf::st_bbox()
 }
 
+#' Expand a bounding box
+#'
+#' @param x sf::bbox
+#'
+#' @return sf::bbox x expanded by factor
+#' @export
 expand_bbox <- function(x, factor) {
   result <- x
   xrng <- x["xmax"] - x["xmin"]

### 2024-06-06 analysis/paper/paper.qmd
figure 5

@@
 sightings more than 15 km or 3 d from a sampling station are included here)."
+#| fig-width: 6.65
+#| fig-height: 4.76
 #| out-width: 169mm
@@
 #| fig-cap: "Distribution of sea ice coverage among sites, by predator cluster: (A) open water, (B) marginal ice, (C) pack ice. Dashed red line: median ice coverage."
+#| fig-width: 6.65
+#| fig-height: 4.76
 #| out-width: 169mm
@@
 Areas of extensive sea-ice coverage are indicated by blue gradient, where satellite-derived mean sea-ice coverage in the month of August in each year exceeded 65%."
+#| fig-width: 6.65
+#| fig-height: 4.76
 #| out-width: 169mm
@@
 )
+
+```
+
+```{r}
+#| label: fig-predclustkde
+#| fig-cap: "Geographic distribution of predator communities across years.
+Solid and dashed lines indicate the 50 and 95% contours of the kernel density estimate for each cluster, respectively, pooling sites from all years."
+#| fig-width: 6.65
+#| fig-height: 2.24
+#| out-width: 169mm
+#| out-height: 57mm
+
+make_fig_predclustkde(read_obj("stations_clust"))
 

### 2024-06-06 R/figures.R

@@ -116,3 +116,45 @@ make_fig_predclustannual <- function(stations_clust, seaice_conc_df) {
           strip.text = element_blank(),
           panel.background = element_rect())
 }
+
+#' Create a figure with geographic distribution of predator clusters across
+#' years
+#'
+#' @return ggplot object
+#' @export
+make_fig_predclustkde <- function(stations_clust) {
+  map_lim <- sf::st_bbox(stations_clust) %>%
+    project_bbox() %>%
+    expand_bbox(factor = 1.2)
+
+  cluster_kde <- stations_clust %>%
+    group_by(pred_clust) %>%
+    group_map(\(rows, keys) eks::st_kde(rows)$sf %>%
+                filter(contlabel %in% c(50, 95)) %>%
+                mutate(pred_clust = keys$pred_clust)) %>%
+    reduce(sf:::rbind.sf)
+
+  sublabels <- tibble(pred_clust = factor(levels(cluster_kde$pred_clust)),
+                      x = -63, y = -60,
+                      labels = LETTERS[1:3]) %>%
+    sf::st_as_sf(crs = "EPSG:4326", coords = c("x", "y"))
+
+  ant_basemap() +
+    # Predator clusters
+    ggnewscale::new_scale_fill() +
+    geom_sf(aes(color = pred_clust,
+                linetype = contlabel),
+            cluster_kde,
+            fill = NA,
+            linewidth = 1) +
+    geom_sf_text(aes(label = labels), sublabels, fontface = "bold", vjust = 0) +
+    facet_wrap(~ pred_clust) +
+    scale_x_continuous(breaks = c(-60, -55)) +
+    scale_y_continuous(breaks = c(-63, -62, -61, -60)) +
+    scale_color_brewer("Predator cluster", palette = "Dark2") +
+    scale_linetype_manual(values = c(2, 1), guide = "none") +
+    coord_ant(map_lim) +
+    theme(legend.position = "bottom",
+          strip.text = element_blank(),
+          panel.background = element_rect())
+}

### 2024-06-06 analysis/paper/paper.qmd
Figure 6

@@
 
-make_fig_predclustannual(
-  read_obj("stations_clust"), 
-  read_obj("seaice_conc_df")
-)
+make_fig_predclustannual(read_obj("stations_clust"), 
+                         read_obj("seaice_conc_df"))
 
@@
 
+```{r}
+#| label: fig-nmds
+#| fig-cap: "Distribution of sites and environmental loadings in non-metric multidimensional scaling (NMDS) ordination: (A) first and second NMDS axes, (B) first and third NMDS axes, (C) second and third NMDS axes.
+Black diamonds indicate centroids of survey years, and contours represent the fitted surface of sea-ice coverage.
+Sites are color-coded by predator cluster assignment."
+#| fig-width: 6.65
+#| fig-height: 3.94
+#| out-width: 169mm
+#| out-height: 100mm
+
+make_fig_nmds(read_obj("nmds_envfit"),
+              read_obj("nmds_df"),
+              read_obj("nmds_sightings"),
+              read_obj("nmds_env"))
+```
+
 # Acknowledgements

### 2024-06-06 R/figures.R

@@ -158,3 +158,80 @@ make_fig_predclustkde <- function(stations_clust) {
           strip.text = element_blank(),
           panel.background = element_rect())
 }
+
+#' Create a figure with NMDS loadings
+#' years
+#'
+#' @return ggplot object
+#' @export
+make_fig_nmds <- function(nmds_envfit, nmds_df, nmds_sightings, nmds_env) {
+  year_centroids <- vegan::scores(nmds_envfit,
+                                  display = "factors",
+                                  choices = 1:3) %>%
+    as_tibble(rownames = "envvar") %>%
+    filter(str_starts(envvar, "Year")) %>%
+    mutate(across(starts_with("NMDS"),
+                  ~ .x * vegan::ordiArrowMul(nmds_envfit)),
+           Year = substr(envvar, 5, 8))
+
+  nmds_plot <- function(axis1, axis2) {
+    ice_ordisurf <- vegan::ordisurf(
+      nmds_sightings ~ nmds_env$ice_coverage,
+      choices = c(axis1, axis2),
+      plot = FALSE
+    )
+
+    x <- paste0("NMDS", axis1)
+    y <- paste0("NMDS", axis2)
+
+    ice_df <- expand_grid(
+      axis1 = ice_ordisurf$grid$x,
+      axis2 = ice_ordisurf$grid$y,
+    ) %>%
+      cbind(as.numeric(ice_ordisurf$grid$z)) %>%
+      set_names(c(x, y, "ice_coverage")) %>%
+      mutate(ice_coverage = ice_coverage / 10) %>%
+      drop_na(ice_coverage)
+
+    ggplot(nmds_df, aes(.data[[x]], .data[[y]])) +
+      stat_contour(aes(z = .data$ice_coverage,
+                       color = after_stat(level)),
+                   ice_df) +
+      scale_color_distiller(
+        "Ice coverage",
+        palette = "Blues",
+        labels = scales::percent,
+        limits = c(0.25, 0.75),
+        breaks = seq(0.25, 0.75, by = 0.1),
+        guide = guide_colorbar(barwidth = unit(2, "in"),
+                               direction = "horizontal")
+      ) +
+      ggnewscale::new_scale_color() +
+      geom_point(aes(color = pred_clust), size = 1, alpha = 0.8) +
+      geom_point(data = year_centroids,
+                 shape = 18,
+                 size = 2.5,
+                 color = "black",
+                 alpha = 0.85) +
+      ggrepel::geom_text_repel(aes(label = Year),
[…]

### 2024-06-06 R/nmds.R

@@ -1,4 +1,10 @@
-# Calculate NMDS stress across numbers of axes
+#' Calculate NMDS stress across numbers of axes
+#'
+#' @param dist distance matrix
+#' @param k number of axes (vectorized, e.g., 1:6 is ok)
+#'
+#' @return tibble with k and stress
+#' @export
 stress <- function(dist, k) {
   tibble(
     k = k,
@@ -13,7 +19,13 @@ stress <- function(dist, k) {
   )
 }
 
-# Create shepard diagram data
+#' Create shepard diagram data
+#'
+#' @param nmds nmds object (see vegan::metaMDS)
+#'
+#' @return tibble with diss, dist, and monotonically increasing nonparametric
+#'   fit
+#' @export
 shepard <- function(nmds) {
   nmds_isoreg <- isoreg(nmds$diss, nmds$dist)
   tibble(
@@ -23,7 +35,14 @@ shepard <- function(nmds) {
   )
 }
 
-# Convert NMDS object to a data frame, including cluster and environmental data
+#' Convert NMDS object to a data frame, including cluster and environmental data
+#'
+#' @param nmds NMDS object
+#' @param clust Cluster data
+#' @param env Environmental data
+#'
+#' @return NMDS object as data frame
+#' @export
 nmds_to_df <- function(nmds, clust, env) {
   nmds$points %>%
     as_tibble(rownames = "amlr.station") %>%
@@ -32,12 +51,21 @@ nmds_to_df <- function(nmds, clust, env) {
     left_join(env, by = "amlr.station")
 }
 
-# Combine ice and station environmental data, filtering down to stations used
-# in this analysis
+#' Combine ice and station environmental data
+#'
+#' Filters to stations used in this analysis
+#'
+#' @param stations Zooplankton station data
+#' @param ice Ice data
+#' @param predators Predator data
+#'
+#' @return Environmental data for NMDS loadings
+#' @export
 env_data <- function(stations, ice, predators) {
+  stations <- as_tibble(stations)
   left_join(
     select(stations, amlr.station, zuml_m, avg.temp, avg.salinity,
-           Integ.chla.100m, Integ.phae.100m, TOD_2levels_civil, Year,
+           Integ.chla.100m, Integ.phae.100m, time.of.day, Year,
            zoop_clust = `Winter Cluster factor`),
     select(ice, amlr.station, ice_type, ice_coverage),
     by = "amlr.station"

### 2024-06-06 analysis/scripts/03_cluster_analysis.R

@@ -30,4 +30,7 @@ stations_clust <- zoop_sf %>%
   semi_join(predators_clust, by = "amlr.station")
 
 # Save derived objects
+save_obj(sightings_mtx, "sightings_mtx")
+save_obj(sightings_dist, "sightings_dist")
 save_obj(stations_clust, "stations_clust")
+save_obj(sightings_cut, "sightings_cut")

### 2024-06-06 analysis/scripts/04_nmds.R

@@ -1,0 +1,32 @@
+# Load this research compendium as a package
+# This is functionally equivalent to library(antwincomm)
+devtools::load_all()
+
+# Load necessary derived data
+sightings_mtx <- read_obj("sightings_mtx")
+sightings_cut <- read_obj("sightings_cut")
+sightings_dist <- read_obj("sightings_dist")
+zoop_sf <- read_obj("zoop_sf")
+stations_ice <- read_obj("stations_ice")
+predators_stations <- read_obj("predators_stations")
+
+# NMDS
+nmds_stress <- stress(sightings_mtx, k = 1:6)
+nmds_sightings <- vegan::metaMDS(sightings_dist,
+                                 k = 3,
+                                 trymax = 100,
+                                 trace = FALSE)
+nmds_shepard <- shepard(nmds_sightings)
+nmds_env <- env_data(zoop_sf, stations_ice, predators_stations)
+nmds_df <- nmds_to_df(nmds_sightings, sightings_cut, nmds_env)
+nmds_envfit <- vegan::envfit(nmds_sightings,
+                             select(nmds_env, -amlr.station),
+                             permutations = 999,
+                             choices = 1:3,
+                             na.rm = TRUE)
+
+# Save derived data
+save_obj(nmds_sightings, "nmds_sightings")
+save_obj(nmds_env, "nmds_env")
+save_obj(nmds_df, "nmds_df")
+save_obj(nmds_envfit, "nmds_envfit")

### 2024-06-06 analysis/paper/paper.qmd
figure cross-references in main text

@@
 We surveyed the South Shetland Island and Elephant Island regions during the austral winters (August to September) of 2012–2016 as part of the US Antarctic Marine Living Resources (AMLR) program aboard the RV Ice Breaker (RVIB) ‘Nathaniel B.
-Palmer’ (Fig.
-1).
+Palmer’ [@fig-overviewmap].
 The main objectives of the survey were to map hydrographic conditions, study the distribution and abundance of fish, krill, and other zooplankton through acoustic methods and net hauls, and describe the spatial distribution of seabirds and marine mammals.
@@
 The most widespread species observed were snow petrels, Antarctic petrels, and southern giant petrels.
-The 6 most abundant species (total individuals observed in Table 2) averaged across years were widely distributed across the study region, although densities varied by species (Fig.
-2).
+The 6 most abundant species (total individuals observed in Table 2) averaged across years were widely distributed across the study region, although densities varied by species [@fig-top6map].
 Distribution maps of all seabird (Fig.
@@
 
-We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage (Fig.
-3).
+We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage [@fig-seaice].
 See Fig.
@@
 
-The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice (Fig.
-4).
-Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands (Fig.
-5).
+The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice [@fig-predclustannual].
+Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands [@fig-predclustkde].
 We found predators characteristic of the marginal ice community in the western Bransfield Strait and north of Elephant Island.
@@
 Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; Table 3; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
-The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters (Fig.
-6).
+The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters [@fig-nmds].
 The marginal ice and open water clusters were well separated by NMDS axis 2.
@@
 The spatial averages of community composition (i.e.
-predator clusters; Fig.
-5) may be considered essential foraging habitats for overwintering predators.
+predator clusters; @fig-predclustkde) may be considered essential foraging habitats for overwintering predators.
 The open water assemblage occupies the most temporally variable ocean habitat that we sampled, which changes based on currents and latitudinal position of the marginal ice zone.
@@
 Antarctic krill are consistently concentrated in the southern region over multiple years [@reissOverwinterHabitatSelection2017], and stable sea-ice concentrations are more likely to occur there due to coastal ocean processes (coastal currents, formation of eddies) and shifts in ice from higher latitudes [@moffatShelfOceanExchange2018].
-We found consistently high concentrations of crabeater seals and fur seals within both marginal ice zone habitats (Figs.
-2 & 4), and the presence of sea ice likely contributes by supporting haul-out habitat for seal thermoregulation and rest between foraging bouts.
+We found consistently high concentrations of crabeater seals and fur seals within both marginal ice zone habitats [@fig-top6map; @fig-predclustannual], and the presence of sea ice likely contributes by supporting haul-out habitat for seal thermoregulation and rest between foraging bouts.
 There were more concentrations of fur seals compared to crabeater seals within the marginal ice zone north of Elephant Island.

### 2024-06-06 analysis/paper/paper.qmd
add table 1; clean up word template

@@
 execute:
-  echo: true
+  echo: false
   warning: false
@@
 Conductivity and temperature in the water column were measured using a Sea-Bird Inc.
-SBE-9/11+ CTD profiler (Table 1).
-We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations (Table 1).
+SBE-9/11+ CTD profiler [@tbl-envvar].
+We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations [@tbl-envvar].
 For full details, see @dietrich2021winter.
@@
 We performed this analysis with the `envfit()` function in the 'vegan' package, which treats environmental variables as the response variables and NMDS ordination scores as the predictors.
-The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables (Table 1).
+The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables [@tbl-envvar].
 Temporal variables included time of day of the net tow, to account for diel patterns in macrozooplankton distribution, as well as year, to account for interannual variability.
@@
 
+# Acknowledgements
+
+We are grateful for the diligent work by the field personnel who made these surveys possible, including the captains, crew, and support personnel of the RVIB ‘Nathaniel B.
+Palmer’.
+The surveys were conducted by a partnership of NOAA Fisheries Antarctic Ecosystem Research Division and the NSF US Antarctic Program (NSF Office of Polar Programs project 2011285).
+We thank Dr.
+Jack Conroy for his contributions and feedback on matters related to zooplankton communities.
+J.A.S.
+thanks M.
+P.
+Force and A.
+Borker for their dedicated support during the visual surveys
+
+\newpage
+
+# Tables
+
+<!-- Note: cross-referencing a table breaks column alignment. See https://github.com/quarto-dev/quarto-cli/issues/8970 -->
+
+| Variable                   | Description and units                                                                                                                                                                                                                                                                                                 | Collection location |
+|:-----------------|:------------------------------------|:-----------------|
+| UML depth                  | Depth at which the density differed by 0.05 kg m^−3^ from the average density of the upper 10 m of the water column [@mitchell1991] \[meters\]                                                                                                                                                                        | Station             |
+| Temperature                | Average UML temperature \[°C\]                                                                                                                                                                                                                                                                                        | Station             |
+| Salinity                   | Average UML salinity \[PSU\]                                                                                                                                                                                                                                                                                          | Station             |
+| Chl *a*                    | Average integrated chl *a* to 100 m from bottles \[mg m^-2^\]                                                                                                                                                                                                                                                         | Station             |
+| Phaeopigment               | Average integrated phaeopigment to 100 m from bottles \[mg m^-2^\]                                                                                                                                                                                                                                                    | Station             |
+| Ice type                   | Factor with 4 levels: open, thin, first-year, multi-year                                                                                                                                                                                                                                                              | Transect            |
+| Ice coverage               | Percent cover                                                                                                                                                                                                                                                                                                         | Transect            |
+| Macrozooplankton community | Factor with 5 levels corresponding to the clusters identified by @dietrich2021winter: 1 (*Salpa* and *Clione*), 2a (small *Euphausia*, myctophid larvae, and amphipods), 2b (*Thysanoessa*), 3a (extremely diverse, including siphonophores, chaetognaths, copepods), and 3b (*E.
+superba* and *E.
+crystallorophias*) | Station             |
+
+: Environmental variables collected during survey.
+Hydrographic variables (upper mixed layer [UML] depth, temperature, salinity, chlorophyll a [chl a], and phaeopigment) were collected at sampling stations.
+Ice variables (type and coverage) were collected along transects and associated with the nearest station.
+Macrozooplankton communities were classified based on net tows collected at the sampling stations (see @dietrich2021winter).
+{#tbl-envvar} {tbl-colwidths="[20,60,20]"}
+
+\newpage
+
 # Figures
@@
 Bathymetry from IBCSO v2 [@dorschelInternationalBathymetricChart2022]; land from Natural Earth (www.naturalearthdata.com).](images/fig-overviewmap.png){#fig-overviewmap width="116mm"}
+
+\newpage
 
@@
 
+\newpage
+
 ```{r}
@@
 ```
+
+\newpage
 
@@
 
-# Acknowledgements
-
-<!-- The following line inserts a page break  -->
-
 \newpage

### 2024-06-07 analysis/paper/paper.qmd
Table 2

@@
 #| label: load
-#| echo: false
 
@@
 We surveyed the South Shetland Island and Elephant Island regions during the austral winters (August to September) of 2012–2016 as part of the US Antarctic Marine Living Resources (AMLR) program aboard the RV Ice Breaker (RVIB) ‘Nathaniel B.
-Palmer’ [@fig-overviewmap].
+Palmer’ ([@fig-overviewmap]).
 The main objectives of the survey were to map hydrographic conditions, study the distribution and abundance of fish, krill, and other zooplankton through acoustic methods and net hauls, and describe the spatial distribution of seabirds and marine mammals.
@@
 
-During the 5 survey years, we sighted a total of 22 species of seabirds and marine mammals (Table 2).
+During the 5 survey years, we sighted a total of 22 species of seabirds and marine mammals ([@tbl-predators]).
 The most abundant species were snow petrels, Adélie penguins, and Antarctic fur seals.
 The most widespread species observed were snow petrels, Antarctic petrels, and southern giant petrels.
-The 6 most abundant species (total individuals observed in Table 2) averaged across years were widely distributed across the study region, although densities varied by species [@fig-top6map].
+The 6 most abundant species (total individuals observed in @tbl-predators) averaged across years were widely distributed across the study region, although densities varied by species ([@fig-top6map]).
 Distribution maps of all seabird (Fig.
@@
 
-We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage [@fig-seaice].
+We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage ([@fig-seaice]).
 See Fig.
@@
 
-The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice [@fig-predclustannual].
-Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands [@fig-predclustkde].
+The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice ([@fig-predclustannual]).
+Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands ([@fig-predclustkde]).
 We found predators characteristic of the marginal ice community in the western Bransfield Strait and north of Elephant Island.
@@
 Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; Table 3; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
-The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters [@fig-nmds].
+The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters ([@fig-nmds]).
 The marginal ice and open water clusters were well separated by NMDS axis 2.
@@
 The pagophilic community (e.g.
-Adélie penguins, crabeater seals) consisted of species with diets dominated by Antarctic krill (Table 2) that rely on the pack ice throughout the annual cycle [@ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017].
+Adélie penguins, crabeater seals) consisted of species with diets dominated by Antarctic krill (\[\@tbl-predators\]) that rely on the pack ice throughout the annual cycle [@ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017].
 Predators in the marginal ice zone (e.g.
 Antarctic fur seals, snow petrels) and in open water (e.g.
-southern fulmars, other petrels) communities also consume Antarctic krill, but their diets are more diverse, with a greater reliance on fish, especially myctophids such as *Electrona antarctica* (Table 2) [@ridoux1989diets; @ainley1992does; @lorentsen1998diet].
+southern fulmars, other petrels) communities also consume Antarctic krill, but their diets are more diverse, with a greater reliance on fish, especially myctophids such as *Electrona antarctica* (\[\@tbl-predators\]) [@ridoux1989diets; @ainley1992does; @lorentsen1998diet].
 Despite the diet overlap between these 2 communities, they tend to segregate spatially according to ice coverage due to differing foraging strategies.
@@
 Antarctic krill are consistently concentrated in the southern region over multiple years [@reissOverwinterHabitatSelection2017], and stable sea-ice concentrations are more likely to occur there due to coastal ocean processes (coastal currents, formation of eddies) and shifts in ice from higher latitudes [@moffatShelfOceanExchange2018].
-We found consistently high concentrations of crabeater seals and fur seals within both marginal ice zone habitats [@fig-top6map; @fig-predclustannual], and the presence of sea ice likely contributes by supporting haul-out habitat for seal thermoregulation and rest between foraging bouts.
+We found consistently high concentrations of crabeater seals and fur seals within both marginal ice zone habitats ([@fig-top6map; @fig-predclustannual]), and the presence of sea ice likely contributes by supporting haul-out habitat for seal thermoregulation and rest between foraging bouts.
 There were more concentrations of fur seals compared to crabeater seals within the marginal ice zone north of Elephant Island.
@@
 {#tbl-envvar} {tbl-colwidths="[20,60,20]"}
+
+\newpage
+
+```{r}
+#| label: tbl-predators
+#| tbl-cap: "Predator species observed during surveys.
+Ind: total individuals observed, Sites: number of sites where species was present, Freq: percentage of sites where species was present.
+Typical prey and foraging behavior are presented for context."
+#| tbl-colwidths: [18, 7, 7, 11, 17, 16, 24]
+
+format_percent <- scales::label_percent(0.1)
+
+predator_context <- tribble(
+  ~species, ~`Typical prey`, ~`Foraging behavior`,     ~Reference,
+  "SNPT",   "Myctophids",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordSpeciesSpecificForaging2016",
+  "ANPT",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordAntarcticPetrelsIce2020",
+  "SGPT",   "Juvenile pinnipeds and seabirds, carrion",    "Scavenging, surface feeding (<5 m)", "@hunterFoodFeedingEcology1983",
+  "FUSE",   "Krill, myctophids",    "Diving (~15-50 m)", "@stanilandForagingBehaviourTwo2011",
+  "KEGU",   "Limpets, krill",    "Surface feeding (<5 m)", "@silvaDoesAccessHighquality2001",
+  "ANFU",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
+  "CAPT",   "Krill, myctophids",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
+  "ADPN",   "Krill",    "Diving (~15-50 m)", "@ciminoClimatedrivenSympatryMay2016; @juaresDietAdeliePenguins2018",
+  "CRSE",   "Krill",    "Diving (~50-100 m)", "@burnsWinterHabitatUse2004; @huckstadtDietSpecialistChanging2012",
+  "BLPT",   "Krill, myctophids, amphipods",    "Surface feeding (<5 m)", "@princeFoodFeedingEcology1980",
+  "LESE",   "Seals, penguins, fish, rkill",    "Diving (~50-100 m)", "@walkerSeasonalOccurrenceDiet1998; @casauxDietLeopardSeal2009; @krauseNovelForagingStrategies2015",
+  "ANTE",   "Krill, myctophids",    "Surface feeding (<5 m)", "@croxallFoodFeedingEcology1980; @ainley1992does",
+  "PFSB",   "Carrion",    "Scavenging", "@faveroForagingEcologyPaleFaced1996",
+  "MIWH",   "Krill",    "Diving (~5-100 m)", "@friedlaenderFeedingRatesUnderice2014",
+  "WESE",   "Fish, squid",    "Diving (~100-350 m)", "@testaOverwinterMovementsDiving1994; @lakeRegionalTemporalFinescale2003",
+  "GEPN",   "Krill, fish",    "Diving (~25-100 m)", "@croxallFoodFeedingEcology1980; @ciminoClimatedrivenSympatryMay2016",
+  "ELSE",   "Myctophids, squid",    "Diving (~200-800 m)", "@guinetSouthernElephantSeal2014; @daneriFeedingHabitsSouthern2015",
+  "KIWH",   "Whales, seals, toothfish, penguins",    "Complex and diverse", "@pitmanThreeFormsKiller2003",
+  "EMPN",   "Krill, silverfish",    "Diving (~50-300 m)", "@kirkwoodForagingEcologyFemale1997; @rodaryBenthicDivingMale2000",
+  "SBWH",   "Squid",    "Diving (>500 m)", "@macleodReviewDataDiets2003",
+  "ANSH",   "Demersal fish",    "Diving (<50 m)", "@casauxShagsAntarcticaTheir2006",
+  "ROSE",   "Myctophids, squid",    "Diving (~100-300 m)", "@blixRossSealOmmatophoca2007; @southwellReviewDataAbundance2012"
+)
+
+n_sites <- nrow(read_obj("zoop_sf"))
+
+read_obj("predators_stations") %>% 
+  mutate(common = code_to_common(species),
+         scientific = code_to_scientific(species),
+         Species = str_glue("{common} _{scientific}_")) %>% 
+  group_by(Species, species) %>% 
+  summarize(Ind = sum(count),
+            Sites = n(),
+            `Freq (%)` = format_percent(Sites / n_sites[1]),
+            .groups = "drop") %>% 
+  left_join(predator_context, by = "species") %>% 
+  select(-species) %>% 
+  arrange(desc(Sites)) %>% 
+  knitr::kable()
+
+```
 

### 2024-06-07 R/predators.R

@@ -111,3 +111,34 @@ code_to_common <- function(species_code) {
     SPSK = "South polar skua",
     WESE = "Weddell seal")[species_code]
 }
+
+#' Convert species code to scientific name
+#'
+#' @param species_code 4-letter species codes (character vector)
+#'
+#' @return Species' scientific names
+#' @export
+code_to_scientific <- function(species_code) {
+  c(ADPN = "Pygoscelis adeliae",
+    ANFU = "Fulmarus glacialoides",
+    ANPT = "Thalassoica antarctica",
+    ANSH = "Leucocarbo bransfieldensis",
+    ANTE = "Sterna vittata",
+    BLPT = "Halobaena caerulea ",
+    CAPT = "Daption capense",
+    CRSE = "Lobodon carcinophagus",
+    ELSE = "Mirounga leonina",
+    EMPN = "Aptenodytes forsteri",
+    FUSE = "Arctocephalus gazella",
+    GEPN = "Pygoscelis papua",
+    KEGU = "Larus dominicanus",
+    KIWH = "Orcinus orca",
+    LESE = "Hydrurga leptonyx",
+    MIWH = "Balaenoptera bonaerensis",
+    PFSB = "Chionis albus",
+    ROSE = "Ommatophoca rossii",
+    SBWH = "Hyperoodon planifrons",
+    SGPT = "Macronectes giganteus",
+    SNPT = "Pagodroma nivea",
+    WESE = "Leptonychotes weddellii")[species_code]
+}

### 2024-06-07 analysis/paper/paper.qmd
refactor table 2 code

@@
 
-format_percent <- scales::label_percent(0.1)
-
-predator_context <- tribble(
-  ~species, ~`Typical prey`, ~`Foraging behavior`,     ~Reference,
-  "SNPT",   "Myctophids",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordSpeciesSpecificForaging2016",
-  "ANPT",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordAntarcticPetrelsIce2020",
-  "SGPT",   "Juvenile pinnipeds and seabirds, carrion",    "Scavenging, surface feeding (<5 m)", "@hunterFoodFeedingEcology1983",
-  "FUSE",   "Krill, myctophids",    "Diving (~15-50 m)", "@stanilandForagingBehaviourTwo2011",
-  "KEGU",   "Limpets, krill",    "Surface feeding (<5 m)", "@silvaDoesAccessHighquality2001",
-  "ANFU",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
-  "CAPT",   "Krill, myctophids",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
-  "ADPN",   "Krill",    "Diving (~15-50 m)", "@ciminoClimatedrivenSympatryMay2016; @juaresDietAdeliePenguins2018",
-  "CRSE",   "Krill",    "Diving (~50-100 m)", "@burnsWinterHabitatUse2004; @huckstadtDietSpecialistChanging2012",
-  "BLPT",   "Krill, myctophids, amphipods",    "Surface feeding (<5 m)", "@princeFoodFeedingEcology1980",
-  "LESE",   "Seals, penguins, fish, rkill",    "Diving (~50-100 m)", "@walkerSeasonalOccurrenceDiet1998; @casauxDietLeopardSeal2009; @krauseNovelForagingStrategies2015",
-  "ANTE",   "Krill, myctophids",    "Surface feeding (<5 m)", "@croxallFoodFeedingEcology1980; @ainley1992does",
-  "PFSB",   "Carrion",    "Scavenging", "@faveroForagingEcologyPaleFaced1996",
-  "MIWH",   "Krill",    "Diving (~5-100 m)", "@friedlaenderFeedingRatesUnderice2014",
-  "WESE",   "Fish, squid",    "Diving (~100-350 m)", "@testaOverwinterMovementsDiving1994; @lakeRegionalTemporalFinescale2003",
-  "GEPN",   "Krill, fish",    "Diving (~25-100 m)", "@croxallFoodFeedingEcology1980; @ciminoClimatedrivenSympatryMay2016",
-  "ELSE",   "Myctophids, squid",    "Diving (~200-800 m)", "@guinetSouthernElephantSeal2014; @daneriFeedingHabitsSouthern2015",
-  "KIWH",   "Whales, seals, toothfish, penguins",    "Complex and diverse", "@pitmanThreeFormsKiller2003",
-  "EMPN",   "Krill, silverfish",    "Diving (~50-300 m)", "@kirkwoodForagingEcologyFemale1997; @rodaryBenthicDivingMale2000",
-  "SBWH",   "Squid",    "Diving (>500 m)", "@macleodReviewDataDiets2003",
-  "ANSH",   "Demersal fish",    "Diving (<50 m)", "@casauxShagsAntarcticaTheir2006",
-  "ROSE",   "Myctophids, squid",    "Diving (~100-300 m)", "@blixRossSealOmmatophoca2007; @southwellReviewDataAbundance2012"
-)
-
-n_sites <- nrow(read_obj("zoop_sf"))
-
-read_obj("predators_stations") %>% 
-  mutate(common = code_to_common(species),
-         scientific = code_to_scientific(species),
-         Species = str_glue("{common} _{scientific}_")) %>% 
-  group_by(Species, species) %>% 
-  summarize(Ind = sum(count),
-            Sites = n(),
-            `Freq (%)` = format_percent(Sites / n_sites[1]),
-            .groups = "drop") %>% 
-  left_join(predator_context, by = "species") %>% 
-  select(-species) %>% 
-  arrange(desc(Sites)) %>% 
-  knitr::kable()
+make_predatortbl(read_obj("predators_stations"), read_obj("zoop_sf"))
 

### 2024-06-07 R/tables.R

@@ -1,0 +1,52 @@
+#' Make a table with predator occurrence and additional details
+#'
+#' @param predators predator data aggregated to stations
+#' @param stations station-level data
+#'
+#' @return knitr::knitr_kable
+#' @export
+make_predatortbl <- function(predators, stations) {
+  format_percent <- scales::label_percent(0.1)
+
+  predator_context <- tribble(
+    ~species, ~`Typical prey`, ~`Foraging behavior`,     ~Reference,
+    "SNPT",   "Myctophids",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordSpeciesSpecificForaging2016",
+    "ANPT",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@spearMorphologicalDifferencesRelative1998; @delordAntarcticPetrelsIce2020",
+    "SGPT",   "Juvenile pinnipeds and seabirds, carrion",    "Scavenging, surface feeding (<5 m)", "@hunterFoodFeedingEcology1983",
+    "FUSE",   "Krill, myctophids",    "Diving (~15-50 m)", "@stanilandForagingBehaviourTwo2011",
+    "KEGU",   "Limpets, krill",    "Surface feeding (<5 m)", "@silvaDoesAccessHighquality2001",
+    "ANFU",   "Myctophids, krill, squid",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
+    "CAPT",   "Krill, myctophids",    "Surface feeding (<5 m)", "@ridoux1989diets; @spearMorphologicalDifferencesRelative1998",
+    "ADPN",   "Krill",    "Diving (~15-50 m)", "@ciminoClimatedrivenSympatryMay2016; @juaresDietAdeliePenguins2018",
+    "CRSE",   "Krill",    "Diving (~50-100 m)", "@burnsWinterHabitatUse2004; @huckstadtDietSpecialistChanging2012",
+    "BLPT",   "Krill, myctophids, amphipods",    "Surface feeding (<5 m)", "@princeFoodFeedingEcology1980",
+    "LESE",   "Seals, penguins, fish, rkill",    "Diving (~50-100 m)", "@walkerSeasonalOccurrenceDiet1998; @casauxDietLeopardSeal2009; @krauseNovelForagingStrategies2015",
[…]

### 2024-06-07 analysis/paper/paper.qmd
Table 3

@@
 
+Note: the repeated cells in the first column of [@tbl-indicators] must be merged manually to produce the formatting seen in the published paper.
+
+```{r}
+#| label: tbl-indicators
+#| tbl-cap: "Indicator species for each predator cluster.
+Indicator values are the product of relative frequency and relative abundance, ranging from 0 to 1.
+Only species with indicator values ≥0.25 are shown for each cluster.
+See Methods for details."
+
+make_indicatortbl(read_obj("sightings_indval"))
+
+```
+
+\newpage
+
 # Figures

### 2024-06-07 R/tables.R

@@ -50,3 +50,27 @@ make_predatortbl <- function(predators, stations) {
     arrange(desc(Sites)) %>%
     knitr::kable()
 }
+
+#' Create indicator species table
+#'
+#' @param indval labdsv::indval Indicator value object
+#'
+#' @return knitr::kable table
+#' @export
+make_indicatortbl <- function(indval) {
+  indval_mtx <- indval$indval
+  cluster_names <- c("Pack ice", "Open water", "Marginal ice")
+
+  # Filter to species >-0.25, format nicely
+  apply(indval_mtx, 2, \(col) {
+    tibble(
+      `Indicator species` = code_to_common(names(col)[col >= 0.25]),
+      `Indicator value` = round(col[col >= 0.25], 2)
+    )
+  }) %>%
+    set_names(cluster_names) %>%
+    bind_rows(.id = "Predator cluster") %>%
+    arrange(`Predator cluster`, desc(`Indicator value`)) %>%
+  # Create table
+    knitr::kable()
+}

### 2024-06-07 analysis/scripts/03_cluster_analysis.R

@@ -34,3 +34,6 @@ save_obj(sightings_mtx, "sightings_mtx")
 save_obj(sightings_dist, "sightings_dist")
 save_obj(stations_clust, "stations_clust")
 save_obj(sightings_cut, "sightings_cut")
+save_obj(sightings_indval, "sightings_indval")
+save_obj(predators_clust, "predators_clust")
+save_obj(stations_clust, "stations_clust")

### 2024-06-07 analysis/paper/paper.qmd
Table 4

@@
 | Variable                   | Description and units                                                                                                                                                                                                                                                                                                 | Collection location |
-|:-----------------|:------------------------------------|:-----------------|
+|:---------|:---------------------------------------------------|:---------|
 | UML depth                  | Depth at which the density differed by 0.05 kg m^−3^ from the average density of the upper 10 m of the water column [@mitchell1991] \[meters\]                                                                                                                                                                        | Station             |
@@
 make_indicatortbl(read_obj("sightings_indval"))
+
+```
+
+\newpage
+
+Note: an additional header row with "No." and "Frequency" columns must be created manually to produce the formatting seen in the published paper.
+
+```{r}
+#| label: tbl-commfreq
+#| tbl-cap: "Relative frequency of predator communities by year. No.: number of sites assigned to each cluster; Frequency: relative frequency of that cluster."
+
+make_commfreqtbl(read_obj("stations_clust"))
 

### 2024-06-07 R/tables.R

@@ -74,3 +74,38 @@ make_indicatortbl <- function(indval) {
   # Create table
     knitr::kable()
 }
+
+#' Create predator community frequency table
+#'
+#' Note: originally I implemented this using flextable output, but there's a bug
+#' involving flextable, docx output, and table captions. See
+#' https://github.com/quarto-dev/quarto-cli/issues/9922.
+#'
+#' @param stations_clust Station-level cluster data
+#'
+#' @return knitr::kable table
+#' @export
+make_commfreqtbl <- function(stations_clust) {
+  # Add a "total" row to the table
+  add_total_row <- function(df) {
+    total_row <- tibble(Year = "Total") %>%
+      cbind(summarize(df, across(-Year, sum)))
+    rbind(df, total_row)
+  }
+
+  # Format number of clusters as number and frequency
+  format_cluster <- function(c, t) {
+    str_glue("{c}\t{round(c / t * 100, 2)}")
+  }
+
+  stations_clust %>%
+    as_tibble() %>%
+    count(Year, pred_clust) %>%
+    pivot_wider(names_from = "pred_clust", values_from = "n") %>%
+    add_total_row() %>%
+    mutate(total = `Open water` + `Marginal ice` + `Pack ice`,
+           across(c("Open water", "Marginal ice", "Pack ice"),
+                  \(c) format_cluster(c, total))) %>%
+    select(-total) %>%
+    knitr::kable()
+}

### 2024-06-07 analysis/paper/paper.qmd
Table 5

@@
 
+Note: an additional header row with "Mean", "Q1", and "Q3" columns must be created manually to produce the formatting seen in the published paper.
+
+```{r}
+#| label: tbl-clustenv
+#| tbl-cap: "Summary of environmental conditions associated with predator clusters.
+See Table 1 for descriptions of environmental variables.
+Environmental variables that differed significantly (p < 0.05) between clusters are indicated with an asterisk (Kruskal-Wallis test with Bonferroni correction)."
+
+make_clustenvtbl(read_obj("stations_clust"))
+
+```
+
+\newpage
+
 # Figures
@@
 
+# The NMDS figure is slightly stochastic, so setting the seed prevents minor 
+# changes each time we render the paper
+set.seed(1053)
 make_fig_nmds(read_obj("nmds_envfit"),

### 2024-06-07 R/tables.R

@@ -109,3 +109,56 @@ make_commfreqtbl <- function(stations_clust) {
     select(-total) %>%
     knitr::kable()
 }
+
+#' Create cluster x environment table
+#'
+#' @param stations_clust Station-level cluster data
+#'
+#' @return knitr::kable table
+#' @export
+make_clustenvtbl <- function(stations_clust) {
+  kw <- map(c("zuml_m", "avg.temp", "avg.salinity",
+              "Integ.chla.100m", "Integ.phae.100m"),
+            \(v) {
+              kw <- kruskal.test(stations_clust[[v]], stations_clust$pred_clust)
+              tibble(Variable = v, p = kw$p.value * 5)
+            }) %>%
+    list_rbind()
+
+  var_lbls <- c(
+    zuml_m = "UML depth (m)",
+    avg.temp = "Temperature (°C)",
+    avg.salinity = "Salinity (PSU)",
+    Integ.chla.100m = "Chl a (mg m−2)",
+    Integ.phae.100m = "Phaeopigment (mg m−2)"
+  )
+
+  format_val <- function(x) {
+    formatC(signif(x, digits = 3), digits = 3, format = "fg", flag = "#")
+  }
+
+  stations_clust %>%
+    as_tibble() %>%
+    group_by(pred_clust) %>%
+    summarize(
+      across(c(zuml_m, avg.temp, avg.salinity, Integ.chla.100m,
+               Integ.phae.100m),
+             list(mean = partial(mean, na.rm = TRUE),
+                  q1 = \(x) quantile(x, 0.25, na.rm = TRUE),
+                  q3 = \(x) quantile(x, 0.75, na.rm = TRUE)),
+             .names = "{.col}={.fn}")
+    ) %>%
+    pivot_longer(-pred_clust) %>%
+    separate_wider_delim(name, "=", names = c("Variable", "fn")) %>%
+    group_by(pred_clust, Variable) %>%
+    summarize(value = paste(format_val(value), collapse = "\t"),
+              .groups = "drop") %>%
+    pivot_wider(names_from = pred_clust, values_from = value) %>%
+    left_join(kw, by = "Variable") %>%
+    mutate(Variable = ifelse(p <= 0.05,
+                             str_glue("{var_lbls[Variable]} *"),
+                             var_lbls[Variable])) %>%
+    select(-p) %>%
+    slice(c(5, 4, 3, 1, 2)) %>%
+    knitr::kable()
+}

### 2024-06-07 analysis/paper/paper.qmd
Table 6

@@
 
+```{r}
+#| label: tbl-predprey
+#| tbl-cap: "Contingency table of predator and macrozooplankton clusters.
+Macrozooplankton cluster names (1, 2a, 2b, 3a, 3b) from @dietrich2021winter.
+Significant differences (p < 0.05) are indicated in bold with asterisks (post hoc analysis of the Pearson residuals)."
+
+make_predpreytbl(read_obj("stations_clust"))
+
+```
+
+\newpage
+
 # Figures
+
+<!-- Figure 1 was created using QGIS -->
 

### 2024-06-07 R/tables.R

@@ -162,3 +162,52 @@ make_clustenvtbl <- function(stations_clust) {
     slice(c(5, 4, 3, 1, 2)) %>%
     knitr::kable()
 }
+
+#' Create predator-prey contingency table
+#'
+#' @param stations_clust Station-level cluster data
+#'
+#' @return knitr::kable table
+#' @export
+make_predpreytbl <- function(stations_clust) {
+  pred_prey_chisq <- with(stations_clust,
+                          chisq.test(pred_clust,
+                                     `Winter Cluster factor`,
+                                     simulate.p.value = TRUE))
+
+  pred_prey_table <- with(stations_clust,
+                          table(pred_clust,
+                                `Winter Cluster factor`))
+  posthoc <- chisq.posthoc.test::chisq.posthoc.test(
+    pred_prey_table,
+    simulate.p.value = TRUE
+  ) %>%
+    pivot_longer(-c(Dimension, Value),
+                 names_to = "zoop_clust",
+                 values_to = "value") %>%
+    pivot_wider(names_from = "Value") %>%
+    rename("Predator cluster" = Dimension,
+           "Zooplankton cluster" = zoop_clust,
+           p = `p values`) %>%
+    select(-Residuals)
+
+  expand_grid(
+    `Predator cluster` = levels(stations_clust$pred_clust),
+    `Zooplankton cluster` = unique(stations_clust$`Winter Cluster factor`)
+  ) %>%
+    mutate(
+      observed = pred_prey_chisq$observed[cbind(`Predator cluster`,
+                                                `Zooplankton cluster`)],
+      expected = pred_prey_chisq$expected[cbind(`Predator cluster`,
+                                                `Zooplankton cluster`)]
+    ) %>%
+    left_join(posthoc, by = c("Predator cluster", "Zooplankton cluster")) %>%
+    mutate(
+      signif = ifelse(p <= 0.05, "*", ""),
+      label = str_glue("{observed}{signif}\t{round(expected, 1)}{signif}")
+    ) %>%
+    pivot_wider(id_cols = `Predator cluster`,
+                names_from = `Zooplankton cluster`,
+                values_from = label) %>%
+    knitr::kable()
+}

### 2024-06-07 analysis/paper/paper.qmd
fix cross-references

@@
 Conductivity and temperature in the water column were measured using a Sea-Bird Inc.
-SBE-9/11+ CTD profiler [@tbl-envvar].
-We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations [@tbl-envvar].
+SBE-9/11+ CTD profiler ([@tbl-envvar]).
+We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations ([@tbl-envvar]).
 For full details, see @dietrich2021winter.
@@
 The gap statistic, *f*, usually increases with *k*, but exhibits a shoulder at the optimal number of clusters.
-A recommended heuristic for finding the shoulder is selecting the smallest *k* such that the gap statistic for *k*, *f*(*k*), is greater than the gap statistic for *k + 1* less the standard error, *s* [@eq-gap-stat].
+A recommended heuristic for finding the shoulder is selecting the smallest *k* such that the gap statistic for *k*, *f*(*k*), is greater than the gap statistic for *k + 1* less the standard error, *s* ([@eq-gap-stat]).
 
@@
 We performed this analysis with the `envfit()` function in the 'vegan' package, which treats environmental variables as the response variables and NMDS ordination scores as the predictors.
-The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables [@tbl-envvar].
+The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables ([@tbl-envvar]).
 Temporal variables included time of day of the net tow, to account for diel patterns in macrozooplankton distribution, as well as year, to account for interannual variability.
@@
 
-We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species (Table 3) and relationships to sea-ice coverage ([@fig-seaice]).
+We retained 3 predator clusters from the hierarchical analysis, which we refer to as the ‘pack ice’, ‘marginal ice’, and ‘open water’ communities based on their indicator species ([@tbl-indicators]) and relationships to sea-ice coverage ([@fig-seaice]).
 See Fig.
@@
 
-The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities (Table 4), with variability between years related to the spatial distribution of sea ice ([@fig-predclustannual]).
+The open water community was observed most frequently (56.3% of sites), followed by the marginal ice (22.0%) and pack ice (21.6%) communities ([@tbl-commfreq]), with variability between years related to the spatial distribution of sea ice ([@fig-predclustannual]).
 Geographically, in general, the open water community occupied the northern offshore regions and the gap between Elephant Island and the rest of the Shetland Islands ([@fig-predclustkde]).
@@
 We retained 3 NMDS axes for investigating the relationships between the environment and community structure of predators (stress = 0.135).
-Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; Table 3; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
+Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; [@tbl-indicators]; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
 The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters ([@fig-nmds]).
@@
 
-The predator clusters differed in temperature, salinity, and chl *a*, but not by mixed layer depth or phaeopigment (Table 5).
+The predator clusters differed in temperature, salinity, and chl *a*, but not by mixed layer depth or phaeopigment ([@tbl-clustenv]).
 Relative to the open water community, the pack ice community was associated with colder, more saline water with lower chl *a* concentrations.
@@
 The predator clusters were significantly associated with macrozooplankton clusters (χ^*2*^ test, p \< 0.001).
-Five pairwise associations of predators with prey were significant after applying Bonferroni correction (Table 6).
+Five pairwise associations of predators with prey were significant after applying Bonferroni correction ([@tbl-predprey]).
 The pack ice community was positively associated with macrozooplankton community 3a (an extremely diverse assemblage associated with cold, high-salinity water) and negatively associated with cluster 2b (*Thysanoessa macrura*).
@@
 frigida*, *E.
-triacantha*, myctophid larvae, and *Themisto gaudichaudii*) was also not significantly associated with any predator cluster, although it was associated most often with the open water predator cluster (Table 6).
+triacantha*, myctophid larvae, and *Themisto gaudichaudii*) was also not significantly associated with any predator cluster, although it was associated most often with the open water predator cluster ([@tbl-predprey]).
 
@@
 The pagophilic community (e.g.
-Adélie penguins, crabeater seals) consisted of species with diets dominated by Antarctic krill (\[\@tbl-predators\]) that rely on the pack ice throughout the annual cycle [@ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017].
+Adélie penguins, crabeater seals) consisted of species with diets dominated by Antarctic krill ([@tbl-predators]) that rely on the pack ice throughout the annual cycle [@ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017].
 Predators in the marginal ice zone (e.g.
 Antarctic fur seals, snow petrels) and in open water (e.g.
-southern fulmars, other petrels) communities also consume Antarctic krill, but their diets are more diverse, with a greater reliance on fish, especially myctophids such as *Electrona antarctica* (\[\@tbl-predators\]) [@ridoux1989diets; @ainley1992does; @lorentsen1998diet].
+southern fulmars, other petrels) communities also consume Antarctic krill, but their diets are more diverse, with a greater reliance on fish, especially myctophids such as *Electrona antarctica* ([@tbl-predators]) [@ridoux1989diets; @ainley1992does; @lorentsen1998diet].
 Despite the diet overlap between these 2 communities, they tend to segregate spatially according to ice coverage due to differing foraging strategies.
@@
 Therefore, we recommend that future studies aiming to link climate change projections, including changes in winter sea-ice concentration, assess how the 3 communities may shift or contract/expand geographically under different scenarios.
-Importantly, the geographic extent, environmental conditions, and macrozooplankton species associations underlying the 3 communities (Tables 4–6) are provided and can be used to develop predator community habitat models allowing for prediction in years outside of the intensive sampling period.
+Importantly, the geographic extent, environmental conditions, and macrozooplankton species associations underlying the 3 communities ([@tbl-commfreq; @tbl-clustenv; @tbl-predprey]) are provided and can be used to develop predator community habitat models allowing for prediction in years outside of the intensive sampling period.
 Suggested geographic community analyses may include quantifying the habitat size and range (e.g.

### 2024-06-13 analysis/paper/paper.qmd
Table S1

@@
   fig-dpi: 600
+crossref:
+  custom:
+    - kind: float
+      key: supptbl
+      reference-prefix: Table S
+      latex-env: supptbl
+      space-before-numbering: false
+      caption-location: top
 knitr:
@@
 We observed predators in proximity (15 km, 3 d) to 245 sampling stations, ranging from 19 stations in 2012 to 68 stations in 2014.
-Each station was associated with 30.2 ± 10.3 km (mean ± SD) of visual survey effort (Table S1).
+Each station was associated with 30.2 ± 10.3 km (mean ± SD) of visual survey effort ([@supptbl-effortice]).
 
@@
 In 2012, surveys were restricted to the northernmost part of the survey region, and open water was the most common type (93.3%).
-In the other survey years, the most common ice type was thin ice (2013, 2015, and 2016) or first-year ice (2014) (Table S1).
+In the other survey years, the most common ice type was thin ice (2013, 2015, and 2016) or first-year ice (2014) ([@supptbl-effortice]).
 Ice coverage was relatively consistent within ice types across years.
-Overall, sites categorized as open water had ice coverage of 6.4 ± 1.6%; thin ice and first-year ice sites had coverage of 55.6 ± 3.2 and 57.0 ± 3.2% (mean ± SE) (Table S1).
+Overall, sites categorized as open water had ice coverage of 6.4 ± 1.6%; thin ice and first-year ice sites had coverage of 55.6 ± 3.2 and 57.0 ± 3.2% (mean ± SE) ([@supptbl-effortice]).
 
@@
 | Variable                   | Description and units                                                                                                                                                                                                                                                                                                 | Collection location |
-|:---------|:---------------------------------------------------|:---------|
+|:----------------|:--------------------------------------|:----------------|
 | UML depth                  | Depth at which the density differed by 0.05 kg m^−3^ from the average density of the upper 10 m of the water column [@mitchell1991] \[meters\]                                                                                                                                                                        | Station             |
@@
 
+# Supplemental material
+
+## Supplemental tables
+
+Note: an additional header row with "Survey effort" and "Ice conditions" must be created manually to produce the formatting seen in the published paper.
+
+::: {#supptbl-effortice}
+```{r}
+make_tbl_effortice(read_obj("stations_ice"), 
+                   read_obj("station_effort"), 
+                   read_obj("predators_stations"))
+
+```
+
+Survey effort and ice conditions by year.
+Under **Survey effort**, **Stations sampled** and **Effort per station** indicate how many stations were sampled in that year and the amount of visual survey effort associated with each station (summarized as mean ± standard deviation).
+Under **Ice conditions**, each column describes a different ice category, indicating the number of stations and the ice coverage (summarized as mean ± standard error).
+Note: some visual surveys did not record ice conditions, so the total number of stations under **Ice conditions** may not equal the **Stations sampled**.
+:::
+
+\newpage
+
 # References

### 2024-06-13 R/supplemental.R

@@ -1,0 +1,67 @@
+#' Make a table summarizing survey effort and ice conditions
+#'
+#' @param stations_ice Station-level ice data
+#' @param station_effort Station-level effort data
+#' @param predators_stations Station-level predator data
+#'
+#' @return knitr::knitr_kable
+#' @export
+make_tbl_effortice <- function(stations_ice, station_effort, predators_stations) {
+  # Effort
+  effort_by_year <- station_effort %>%
+    as_tibble() %>%
+    semi_join(predators_stations, by = "amlr.station") %>%
+    group_by(Year = year) %>%
+    summarize(
+      `Stations sampled` = n(),
+      `Effort per station (km)` = sprintf("%0.1f \U00B1 %0.1f",
+                                          mean(survey_km),
+                                          sd(survey_km))
+    )
+  effort_total <- station_effort %>%
+    as_tibble() %>%
+    semi_join(predators_stations, by = "amlr.station") %>%
+    summarize(
+      `Stations sampled` = n(),
+      `Effort per station (km)` = sprintf("%0.1f \U00B1 %0.1f",
+                                          mean(survey_km),
+                                          sd(survey_km))
+    ) %>%
+    mutate(Year = "Total")
+  effort <- rbind(effort_by_year, effort_total)
+
+  # Ice
+  se <- function(p, n) sqrt(p * (1 - p) / n)
+  ice_lbl <- c(OPEN = "Open water",
+               THIN = "Thin ice",
+               FY = "First-year ice",
+               MY = "Multi-year ice")
+  ice_by_year <- stations_ice %>%
+    semi_join(predators_stations, by = "amlr.station") %>%
+    mutate(ice_coverage = ice_coverage / 10,
+           Year = as.numeric(substr(amlr.station, 5, 8)),
+           ice_type = ice_lbl[ice_type]) %>%
+    group_by(Year, ice_type) %>%
+    summarize(ice = sprintf("%d (%0.1f%% \U00B1 %0.1f%%)",
+                            n(),
+                            mean(ice_coverage) * 100,
+                            se(mean(ice_coverage), n())),
+              .groups = "drop") %>%
+    pivot_wider(names_from = ice_type, values_from = ice, values_fill = "0")
+  ice_total <- stations_ice %>%
+    semi_join(predators_stations, by = "amlr.station") %>%
+    mutate(ice_coverage = ice_coverage / 10,
[…]

### 2024-06-13 analysis/scripts/01_process_data.R

@@ -93,6 +93,7 @@ predators_stations <- predators_sf %>%
   assign_sightings(zoop_sf, max_dist_km = 15, max_days = 3) %>%
   normalize_counts(station_effort) %>%
   filter_species(station_thr = 0.00)
+save_obj(station_effort, "station_effort")
 save_obj(stations_ice, "stations_ice")
 save_obj(predators_stations, "predators_stations")
 

### 2024-06-13 analysis/paper/paper.qmd
table s2

@@
 
-Predator densities varied among clusters (Table S2).
+Predator densities varied among clusters ([@supptbl-densfreq]).
 Snow petrels and Adélie penguins had the greatest overall density across clusters (1.665 and 1.153 ind.
@@
 
+::: {#supptbl-densfreq}
+```{r}
+make_tbl_densfreq(read_obj("predators_clust"), 
+                  read_obj("stations_clust"), 
+                  read_obj("station_effort"))
+
+```
+
+Density (ind km^-1^) and frequency (% sites present) of predators within and across clusters.
+See @tbl-predators for scientific names.
+Note: we removed sites with no observed predators from the cluster analysis, so the frequencies in the **All clusters** column are greater than the frequencies in @tbl-predators (which included sites without predators).
+:::
+
+\newpage
+
 # References

### 2024-06-13 R/supplemental.R

@@ -65,3 +65,65 @@ make_tbl_effortice <- function(stations_ice, station_effort, predators_stations)
   cbind(effort, ice) %>%
     knitr::kable()
 }
+
+make_tbl_densfreq <- function(predators_clust, stations_clust, station_effort) {
+  # Density/frequency by cluster
+  # Effort
+  effort_by_clust <- as_tibble(stations_clust) %>%
+    left_join(station_effort, by = "amlr.station") %>%
+    select(amlr.station, pred_clust, survey_km) %>%
+    group_by(pred_clust) %>%
+    summarize(survey_km = sum(survey_km),
+              n_station = n())
+  # Species
+  species_by_clust <- predators_clust %>%
+    group_by(Species = species, pred_clust) %>%
+    summarize(count = sum(count),
+              n_present = n(),
+              .groups = "drop")
+  # Combined
+  fmt_densfreq <- function(d, f) {
+    case_when(
+      f == 0 ~ "0.000 (0.0%)",
+      f > 0 & d < 0.001 ~ sprintf("<0.001 (%0.1f%%)", f * 100),
+      TRUE ~ sprintf("%0.3f (%0.1f%%)", d, f * 100)
+    )
+  }
+  densfreq_by_clust <- species_by_clust %>%
+    complete(Species, pred_clust, fill = list(count = 0,
+                                              survey_km = 1,
+                                              n_present = 0)) %>%
+    left_join(effort_by_clust, by = "pred_clust") %>%
+    mutate(density = count / survey_km,
+           freq = n_present / n_station,
+           pred = fmt_densfreq(density, freq)) %>%
+    select(Species, pred_clust, pred) %>%
+    pivot_wider(names_from = pred_clust,
+                values_from = pred)
+
+  # Density/frequency overall
+  # Effort
+  effort_total <- summarize(effort_by_clust,
+                            survey_km = sum(survey_km),
+                            n_station = sum(n_station))
+  # Species
+  species_total <- species_by_clust %>%
+    group_by(Species) %>%
+    summarize(count = sum(count),
+              n_present = sum(n_present))
+  # Combined
+  densfreq_total <- species_total %>%
+    cbind(effort_total) %>%
+    mutate(density = count / survey_km,
+           freq = n_present / n_station,
+           `All clusters` = fmt_densfreq(density, freq)) %>%
+    arrange(desc(density)) %>%
+    select(Species, `All clusters`)
+
[…]

### 2024-06-13 analysis/paper/paper.qmd
Table S3

@@
 We retained 3 NMDS axes for investigating the relationships between the environment and community structure of predators (stress = 0.135).
-Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; [@tbl-indicators]; see Table S3 for the coefficient, r^2^, and p-values of the environmental fit regression).
+Two variables, sampling year and ice coverage, were significantly correlated with the ordination scores (p = 0.002 and 0.013, respectively), although the effect size was small (r^2^ = 0.061 and 0.054, respectively; [@tbl-indicators]; see [@supptbl-nmds] for the coefficient, r^2^, and p-values of the environmental fit regression).
 The pack ice cluster occupied a wide range of values on NMDS axis 1, overlapping with the other 2 clusters ([@fig-nmds]).
@@
 
+::: {#supptbl-nmds}
+```{r}
+make_tbl_nmdsdetail(read_obj("nmds_envfit"))
+
+```
+
+Direction cosines along the reduced dimensions (NMDS axes 1-3), goodness of fit ($r^2$), and significance ($p$) of the multiple regression analysis of environmental variables with respect to non-metric multidimensional scaling ordination scores.
+For descriptions of environmental variables, see [@tbl-envvar].
+Statistically significant variables indicated with asterisk.
+:::
+
+\newpage
+
 # References

### 2024-06-13 R/supplemental.R

@@ -66,6 +66,14 @@ make_tbl_effortice <- function(stations_ice, station_effort, predators_stations)
     knitr::kable()
 }
 
+#' Make a table summarizing density and frequency of predators by cluster
+#'
+#' @param predators_clust Predator-level cluster data
+#' @param stations_clust Station-level cluster data
+#' @param station_effort Station-level effort data
+#'
+#' @return knitr::knitr_kable
+#' @export
 make_tbl_densfreq <- function(predators_clust, stations_clust, station_effort) {
   # Density/frequency by cluster
   # Effort
@@ -125,5 +133,81 @@ make_tbl_densfreq <- function(predators_clust, stations_clust, station_effort) {
     left_join(densfreq_by_clust, by = "Species") %>%
     relocate(`All clusters`, .after = last_col()) %>%
     mutate(Species = code_to_common(Species)) %>%
+    knitr::kable()
+}
+
+#' Make a table summarizing NMDS ordinations fit to environmental data
+#'
+#' @param nmds_envfit vegan::envfit
+#'
+#' @return knitr::knitr_kable
+#' @export
+make_tbl_nmdsdetail <- function(nmds_envfit) {
+  vector_lbls <- c(
+    zuml_m = "Mixed layer depth",
+    avg.temp = "Temperature",
+    avg.salinity = "Salinity",
+    Integ.chla.100m = "Chl _a_",
+    Integ.phae.100m = "Phaeopigment",
+    ice_coverage = "Ice coverage"
+  )
+  vector_tbl <- with(nmds_envfit$vectors,
+                     cbind(arrows,
+                           `_r^2^_` = r,
+                           `_p_` = pvals)) %>%
+    as_tibble(rownames = "Variables") %>%
+    mutate(Variables = paste0(vector_lbls[Variables],
+                              ifelse(`_p_` < 0.05, "*", "")),
+           across(-Variables, \(x) sprintf("%0.3f", x)))
+
+#' Title
+#'
+#' @param varid
+#' @param var_name
+#' @param fct_names
+#'
+#' @return
+#' @export
+#'
+#' @examples
+  reformat_factor <- function(varid, var_name, fct_names) {
+    centroids <- nmds_envfit$factors$centroids %>%
+      as_tibble(rownames = "Variables") %>%
+      filter(str_starts(Variables, varid)) %>%
+      mutate(Variables = paste0("\t", fct_names),
[…]

### 2024-06-13 R/zoop.R

@@ -11,14 +11,11 @@ read_zoop <- function(zoop_path) {
 
   read_csv(zoop_path, show_col_types = FALSE) %>%
     mutate(sun_angle = sun_angle(start.time.UTC, dec.longitude, dec.latitude),
-           time.of.day = cut(sun_angle,
-                             c(-Inf, -18, 0, Inf),
-                             labels = c("Night", "Twilight", "Day"))) %>%
+           time.of.day = factor(ifelse(sun_angle > -6, "Day", "Night"))) %>%
     filter(leg == "W")
 }
 
 tod_pal <- function() {
   c(Day = "gold",
-    Twilight = "violet",
     Night = "darkblue")
 }

### 2024-06-13 analysis/scripts/_make_all.R

@@ -1,0 +1,3 @@
+for (script in dir("analysis/scripts", "^[0-9]", full.names = TRUE)) {
+  source(script)
+}

### 2024-06-13 analysis/paper/paper.qmd
Supplemental figures

@@
       caption-location: top
+    - kind: float
+      key: suppfig
+      reference-prefix: Fig. S
+      caption-prefix: Figure S
+      latex-env: suppfig
+      space-before-numbering: false
 knitr:
@@
 The 6 most abundant species (total individuals observed in @tbl-predators) averaged across years were widely distributed across the study region, although densities varied by species ([@fig-top6map]).
-Distribution maps of all seabird (Fig.
-S1) and marine mammal species (Fig.
-S2) are provided in the Supplement.
+Distribution maps of all seabird ([@suppfig-birdclim]) and marine mammal species ([@suppfig-mammclim]) are provided in the Supplement.
 We observed predators in proximity (15 km, 3 d) to 245 sampling stations, ranging from 19 stations in 2012 to 68 stations in 2014.
@@
 | Variable                   | Description and units                                                                                                                                                                                                                                                                                                 | Collection location |
-|:----------------|:--------------------------------------|:----------------|
+|:-----------------|:------------------------------------|:-----------------|
 | UML depth                  | Depth at which the density differed by 0.05 kg m^−3^ from the average density of the upper 10 m of the water column [@mitchell1991] \[meters\]                                                                                                                                                                        | Station             |
@@
 
+## Supplemental figures
+
+::: {#suppfig-birdclim}
+```{r}
+#| label: suppfig-birdclim
+#| fig-height: 7.15
+#| fig-width: 6.5
+make_fig_predclim(read_obj("pred_class_df"),
+                  c("Blue petrel",
+                    "Antarctic shag",
+                    "Cape petrel",
+                    "Kelp gull",
+                    "Antarctic petrel",
+                    "Southern giant\npetrel",
+                    "Adélie penguin",
+                    "Kerguelen petrel",
+                    "Gentoo penguin",
+                    "Antarctic tern",
+                    "Northern giant\npetrel",
+                    "Pale-faced\nsheathbill",
+                    "Chinstrap penguin",
+                    "Southern fulmar",
+                    "Antarctic prion",
+                    "Common diving\npetrel",
+                    "South polar skua",
+                    "Snow petrel",
+                    "Emperor penguin",
+                    "Black-browed\nalbatross"),
+                  nrow = 4)
+
+```
+
+Geographic distributions of seabird species observed during visual surveys, averaged across years.
+:::
+
+\newpage
+
+::: {#suppfig-mammclim}
+```{r}
+#| label: suppfig-mammclim
+#| fig-height: 7.6
+#| fig-width: 6.5
+make_fig_predclim(read_obj("pred_class_df"),
+                  c("Antarctic fur seal",
+                    "Leopard seal",
+                    "Bottlenose whale",
+                    "Crabeater seal",
+                    "Elephant seal",
+                    "Killer whale",
+                    "Weddell seal",
+                    "Ross seal",
+                    "Minke whale"),
+                  nrow = 3)
+
+```
+
+Geographic distributions of marine mammal species observed during visual surveys, averaged across years.
+:::
+
+\newpage
+
 # References

### 2024-06-13 R/supplemental.R

@@ -211,3 +211,35 @@ make_tbl_nmdsdetail <- function(nmds_envfit) {
   rbind(vector_tbl, factor_tbl) %>%
     knitr::kable()
 }
+
+#' Make a species climatology figure
+#'
+#' @param pred_class_df Predator climatologies, with binned densities
+#' @param keep Which species to include
+#'
+#' @return ggplot
+#' @export
+make_fig_predclim <- function(pred_class_df, keep, nrow) {
+  species_df <- filter(pred_class_df, name %in% keep)
+
+  ant_sf <- read_obj("ant_sf")
+  map_bbox <- terra::ext(-63, -54, -64, -59) %>%
+    terra::project(from = "+proj=longlat +datum=WGS84",
+                   to = as.character(ant_proj())[[1]])
+  ant_basemap() +
+    geom_raster(aes(x, y, fill = count), species_df) +
+    geom_sf(data = ant_sf) +
+    scale_fill_brewer(palette = "RdPu",
+                      na.value = NA,
+                      direction = 1,
+                      na.translate = FALSE) +
+    scale_x_continuous(breaks = seq(-64, -54, by = 4)) +
+    scale_y_continuous(breaks = seq(-64, -59, by = 2)) +
+    labs(fill = expression("Ind km"^-1)) +
+    facet_wrap(~ name, nrow = nrow) +
+    coord_sf(xlim = map_bbox[1:2],
+             ylim = map_bbox[3:4],
+             crs = ant_proj()) +
+    theme(legend.position = "top",
+          text = element_text(size = 10))
+}

### 2024-06-13 analysis/scripts/02_species_climatologies.R

@@ -6,6 +6,9 @@ devtools::load_all()
 predators_agg <- read_obj("predators_agg")
 effort_sf <- read_obj("effort_sf")
 
+
+# Top six -----------------------------------------------------------------
+
 # Find six most abundant species
 top_six <- predators_agg %>%
   group_by(species) %>%
@@ -63,7 +66,55 @@ topsix_climatology <- unique(top_six$common_name) %>%
 topsix_clim_class <- topsix_climatology %>%
   terra::classify(rcl = c(0, 1, 5, 10, 50))
 
+
+# All species -------------------------------------------------------------
+
+pred_sf <- predators_agg %>%
+  filter(!str_starts(species, "U")) %>%
+  mutate(common_name = code_to_common(species)) %>%
+  latlon_to_sf(coords = c("lon_mean", "lat_mean"))
+
+pred_rast <- pred_sf %>%
+  sf::st_transform(ant_proj()) %>%
+  mutate(year_commonname = as.character(interaction(year, common_name))) %>%
+  rasterize_by("count", "year_commonname", sum, rast_template)
+
+pred_norm <- pred_rast
+for (i in seq(terra::nlyr(pred_rast))) {
+  lyr_name <- names(pred_rast)[i]
+  lyr_year <- substr(lyr_name, 1, 4)
+  effort_year <- effort_rast[lyr_year] %>%
+    terra::subst(0, NA)
+  pred_norm[[i]] <- pred_norm[[i]] / effort_year
+}
+
+pred_climatology <- unique(pred_sf$common_name) %>%
+  map(function(sp) {
+    sp_rast <- terra::subset(pred_norm,
+                             which(str_detect(names(pred_norm), sp)))
+    terra::mean(sp_rast, na.rm = TRUE)
+  }) %>%
+  set_names(unique(pred_sf$common_name)) %>%
+  terra::rast()
+
+map_bbox <- terra::ext(-63, -54, -64, -59) %>%
+  terra::project(from = "+proj=longlat +datum=WGS84",
+                 to = as.character(ant_proj())[[1]])
+
+pred_class <- pred_climatology %>%
+  terra::classify(rcl = c(0, 1, 5, 10, 50, 100, 1000))
+
+pred_class_df <- as.data.frame(pred_class, xy = TRUE) %>%
+  pivot_longer(-c("x", "y"),
+               names_to = "name",
+               values_to = "count") %>%
+  drop_na(count) %>%
[…]

### 2024-06-13 README.Rmd
Update README

@@
 
-> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024). _`r Title`_. Marine Ecology Progress Series <https://doi.org/xxx/xxx>
+> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024). _`r Title`_. Marine Ecology Progress Series <https://doi.org/10.3354/meps14609>
 
@@
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
+  - [:file\_folder: scripts](/analysis/scripts): R scripts for processing data and performing the analysis
+  - [:file\_folder: data](/analysis/data): Raw and derived data
 
@@
 [DESCRIPTION](/DESCRIPTION) file).
-- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console
+- download the raw data (see [/analysis/data/raw_data/README.md](/analysis/data/raw_data/README.md))
+- run all the data processing and analysis scripts (run [/analysis/scripts/_make_all.R](/analysis/scripts/_make_all.R))
+- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`
 
